### 2023-09-21 metadata/parametadata_control.Rmd
First pass at metadata script

---
title: "Metadata collation"
author: "William Kumler"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggmap)

```

## Read in water column .gof file

```{r download raw data, eval=!file.exists("raw_data/mesoscope.gof")}
sum_url <- "http://scope.soest.hawaii.edu/data/scope2021/paragon1.sum"
download.file(sum_url, destfile = "raw_data/paragon1.sum")

gof_url <- "http://scope.soest.hawaii.edu/FTP/scope/water/paragon1.gof"
download.file(gof_url, destfile = "raw_data/paragon1.gof")

seaflow_url <- "http://scope.soest.hawaii.edu/data/scope2021/SeaFlow_dataset_v1.4.xlsx"
download.file(seaflow_url, destfile = "raw_data/SeaFlow_dataset_v1.4.xlsx", mode = "wb")

# register_google("AIzaSyBVNJV0Myoy3M50IBBDIjXZhUyU8Mnagvs")
# gm <- get_map(location = c(-156.5, 21.8), zoom = 8, maptype = "hybrid")
# saveRDS(object = gm, file = "raw_data/ggmap_of_aloha.rds")

magick::image_read("http://scope.soest.hawaii.edu/data/scope2021/Paragon1CTDchlp.pdf", density = 300) %>%
  magick::image_trim() %>%
  magick::image_write("raw_data/chl_pdf_url.png", format = "png")
```

![](raw_data/chl_pdf_url.png)

```{r From SCOPE FTP}
para_cast_times <- "raw_data/paragon1.sum" %>%
  read_table(skip = 3, col_names = FALSE, col_types = cols(.default = col_character())) %>%
  mutate(mdyhms=paste(X3, X4, X5, X6)) %>%
  mutate(time=as.POSIXct(strptime(mdyhms, "%b %d  %Y %X", tz="UTC"))) %>%
  select(station=X1, cast=X2, time) %>%
  mutate(station=as.numeric(station), cast=as.numeric(cast))

gof_names <- c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
              "sal", "ctd_oxy", "ctd_chl", "theta", "sigma", "oxy_um", "dic", "alk", "ph",
              "po4_um", "no23_um", "sio4_um", "LLN", "LLP", "PC", "PN", "PP",
              "PSi", "chl_ug", "phaeo_ug", "het_abund", 
              "pro_abund", "syn_abund", "euk_abund",
              "q1", "q2", "q3", "q4")
qual_names <- setdiff(gof_names, c(head(gof_names, 6), tail(gof_names, 4)))

# From ftp://ftp.soest.hawaii.edu/hot/water/Readme.water.jgofs
# Quality Indicators:
# Flag  Meaning
# 1    not quality controled
# 2    good data
# 3    suspect (i.e.  questionable) data
# 4    bad data
# 5    missing data
# 9    variable not measured during this cast
raw_gof <- read_table("raw_data/paragon1.gof", skip = 4, col_names = gof_names,
                      col_types = cols(.default = col_double()))
qual_df <- select(raw_gof, matches("^q\\d+$")) %>%
  unite("value", sep = "") %>%
  separate(sep = "", col = "value", into = c(NA, qual_names)) %>%
  mutate(across(everything(), as.numeric))
raw_gof[,qual_names][qual_df!=2] <- NA
para_core <- raw_gof %>% 
  left_join(para_cast_times) %>%
  select(station:rosette, time, lat:oxy_um, chl_ug:phaeo_ug) %>%
  mutate(lon=-lon)
```

```{r visualize}
my_casts <- c(4, 5, 9, 11, 13, 14, 16, 18, 25, 27, 31, 32, 34, 36, 39, 41)

gm <- readRDS("raw_data/ggmap_of_aloha.rds")
stations_toplot <- para_core %>%
  distinct(station, cast, lat, lon) %>%
  mutate(station_cast=paste0("S", station, "C", cast)) %>%
  mutate(plotcol=ifelse(cast%in%my_casts, "red", "black"))
satmap <- ggmap(gm) +
  geom_path(aes(x=lon, y=lat), data=stations_toplot, color="grey80") +
  geom_point(aes(x=lon, y=lat), data = stations_toplot, color="white") +
  scale_color_identity()
statmap <- ggplot(stations_toplot, aes(x=lon, y=lat)) +
  geom_path(color="black") +
  geom_text(aes(x=lon, y=lat, color=plotcol, label=station_cast)) +
  scale_color_identity() +
  theme_bw() +
  coord_fixed()
cowplot::plot_grid(satmap, statmap, nrow=1)

para_core %>%
  mutate(station_cast=paste0("S", station, "C", cast)) %>%
  mutate(plotcol=ifelse(cast%in%my_casts, "red", "black")) %>%
  count(plotcol, station_cast, time) %>%
  ggplot(aes(x=time, y=n, fill=plotcol, label=station_cast)) +
  geom_col() +
  scale_fill_identity() +
  scale_x_datetime(timezone = "HST", date_breaks = "1 day", date_labels="%b %d") +
  theme_bw()

para_core %>%
  pivot_longer(temp:phaeo_ug) %>%
  filter(pressure<260) %>%
  drop_na() %>%
  ggplot(aes(x=value, y=-pressure, group=cast, color=time)) +
  geom_point() +
  facet_wrap(~name, nrow = 2, scales="free_x") +
  scale_color_datetime(timezone = "HST", date_breaks = "1 day", date_labels="%b %d")

seaflowdata <- readxl::read_excel("raw_data/SeaFlow_dataset_v1.4.xlsx")
seaflowdata %>%
  mutate(time=str_remove(time, "\\+.*")) %>%
  mutate(time=strptime(time, format = "%Y-%m-%dT%H:%M:%S")) %>%
  mutate(time=as.POSIXct(time)) %>%
  select(time, starts_with("biomass")) %>%
  pivot_longer(-time) %>%
  filter(!(name=="biomass_synecho" & value>0.2)) %>%
  ggplot() +
  geom_point(aes(x=time, y=value)) +
  facet_wrap(~name, ncol=1, scales="free_y") +
  scale_x_datetime(timezone = "HST", date_breaks = "1 day", date_labels="%b %d")
```

## Create file data from MS samples

```{r}
samp_filedata <- "../mzMLs" %>%
  list.files(pattern = "Smp.*mzML", recursive = TRUE) %>% 
  data.frame(filepath=.) %>%
  mutate(filename=basename(filepath)) %>%
  mutate(shortname=basename(filepath)) %>%
  mutate(shortname=str_remove(shortname, "-wIS|-noIS-neg|-noIS-pos")) %>%
  mutate(shortname=str_remove(shortname, ".*_Smp_")) %>%
  mutate(IS_type=str_extract(filepath, "wIS|noIS")) %>%
  mutate(polarity=dirname(filepath)) %>%
  mutate(amendment=str_extract(shortname, "Nit|Amm|Arg|GMP")) %>%
  mutate(depth=str_extract(shortname, "Deep|Surf")) %>%
  mutate(startime=str_extract(shortname, "Eve|Morn")) %>%
  mutate(timepoint=str_extract(shortname, "T\\d+")) %>%
  mutate(tripl=str_extract(shortname, "[A-C](?=\\.mzML$)"))

station_samp_map <- tribble(
  ~amendment, ~depth, ~startime, ~cast,
  "Nit", "Surf", "Morn", 4,
  "Nit", "Surf", "Eve", 5,
  "Nit", "Deep", "Morn", 9,
  "Nit", "Deep", "Eve", 11,
  "Arg", "Surf", "Morn", 13,
  "Arg", "Surf", "Eve", 14,
  "Arg", "Deep", "Morn", 16,
  "Arg", "Deep", "Eve", 18,
  "GMP", "Surf", "Morn", 25,
  "GMP", "Surf", "Eve", 27,
  "GMP", "Deep", "Morn", 31,
  "GMP", "Deep", "Eve", 32,
  "Amm", "Surf", "Morn", 34,
  "Amm", "Surf", "Eve", 36,
  "Amm", "Deep", "Morn", 39,
  "Amm", "Deep", "Eve", 41
) %>%
  mutate(station=2)

samp_filedata <- samp_filedata %>%
  left_join(station_samp_map)
```

```{r write out}
write_csv(samp_filedata, file = "samp_filedata.csv")
```

### 2023-09-21 metadata/parametadata_control.Rmd
Switching to single project directory

@@
 library(ggmap)
-
 ```
@@
 ```{r}
-samp_filedata <- "../mzMLs" %>%
-  list.files(pattern = "Smp.*mzML", recursive = TRUE) %>% 
+samp_filedata <- "mzMLs" %>%
+  list.files(pattern = "Smp.*mzML", recursive = TRUE, full.names = TRUE) %>% 
   data.frame(filepath=.) %>%
   mutate(filename=basename(filepath)) %>%
-  mutate(shortname=basename(filepath)) %>%
-  mutate(shortname=str_remove(shortname, "-wIS|-noIS-neg|-noIS-pos")) %>%
+  mutate(shortname=str_remove(filename, "-wIS|-noIS-neg|-noIS-pos")) %>%
   mutate(shortname=str_remove(shortname, ".*_Smp_")) %>%
   mutate(IS_type=str_extract(filepath, "wIS|noIS")) %>%
-  mutate(polarity=dirname(filepath)) %>%
+  mutate(polarity=str_extract(filepath, "neg|pos")) %>%
   mutate(amendment=str_extract(shortname, "Nit|Amm|Arg|GMP")) %>%
@@
 ```{r write out}
-write_csv(samp_filedata, file = "samp_filedata.csv")
+write_csv(samp_filedata, file = "metadata/samp_filedata.csv")
 ```

### 2023-09-21 metadata/parametadata_control.R
Adding timestamps to metadata, continuing project switch
Also added purling for easy backjob access

@@ -1,20 +1,11 @@
----
-title: "Metadata collation"
-author: "William Kumler"
-date: "`r Sys.Date()`"
-output: html_document
----
-
-```{r setup, include=FALSE}
+## ----setup, include=FALSE-------------------------------------------------------------------
 knitr::opts_chunk$set(echo = TRUE)
 library(tidyverse)
 library(ggmap)
 library(RaMS)
-```
 
-## Read in water column .gof file
 
-```{r download raw data, eval=!file.exists("metadata/raw_data/paragon1.gof")}
+## ----download raw data, eval=!file.exists("metadata/raw_data/paragon1.gof")-----------------
 sum_url <- "http://scope.soest.hawaii.edu/data/scope2021/paragon1.sum"
 download.file(sum_url, destfile = "metadata/raw_data/paragon1.sum")
 
@@ -31,11 +22,9 @@ download.file(seaflow_url, destfile = "metadata/raw_data/SeaFlow_dataset_v1.4.xl
 magick::image_read("http://scope.soest.hawaii.edu/data/scope2021/Paragon1CTDchlp.pdf", density = 300) %>%
   magick::image_trim() %>%
   magick::image_write("metadata/raw_data/chl_pdf_url.png", format = "png")
-```
 
-![](metadata/raw_data/chl_pdf_url.png)
 
-```{r From SCOPE FTP}
+## ----From SCOPE FTP-------------------------------------------------------------------------
 para_cast_times <- "metadata/raw_data/paragon1.sum" %>%
   read_table(skip = 3, col_names = FALSE, col_types = cols(.default = col_character())) %>%
   mutate(mdyhms=paste(X3, X4, X5, X6)) %>%
@@ -71,9 +60,9 @@ para_core <- raw_gof %>%
   left_join(para_cast_times) %>%
   select(station:rosette, time, lat:oxy_um, chl_ug:phaeo_ug) %>%
   mutate(lon=-lon)
-```
 
-```{r visualize}
+
+## ----visualize------------------------------------------------------------------------------
 my_casts <- c(4, 5, 9, 11, 13, 14, 16, 18, 25, 27, 31, 32, 34, 36, 39, 41)
 
 gm <- readRDS("metadata/raw_data/ggmap_of_aloha.rds")
@@ -124,12 +113,9 @@ seaflowdata %>%
   geom_point(aes(x=time, y=value)) +
   facet_wrap(~name, ncol=1, scales="free_y") +
[…]

### 2023-09-22 metadata/parametadata_control.Rmd
Adding other files to metadata sheet

@@
 library(ggmap)
+library(RaMS)
 ```
@@
 
-```{r download raw data, eval=!file.exists("raw_data/mesoscope.gof")}
+```{r download raw data, eval=!file.exists("metadata/raw_data/paragon1.gof")}
 sum_url <- "http://scope.soest.hawaii.edu/data/scope2021/paragon1.sum"
-download.file(sum_url, destfile = "raw_data/paragon1.sum")
+download.file(sum_url, destfile = "metadata/raw_data/paragon1.sum")
 
 gof_url <- "http://scope.soest.hawaii.edu/FTP/scope/water/paragon1.gof"
-download.file(gof_url, destfile = "raw_data/paragon1.gof")
+download.file(gof_url, destfile = "metadata/raw_data/paragon1.gof")
 
 seaflow_url <- "http://scope.soest.hawaii.edu/data/scope2021/SeaFlow_dataset_v1.4.xlsx"
-download.file(seaflow_url, destfile = "raw_data/SeaFlow_dataset_v1.4.xlsx", mode = "wb")
+download.file(seaflow_url, destfile = "metadata/raw_data/SeaFlow_dataset_v1.4.xlsx", mode = "wb")
 
@@
 # gm <- get_map(location = c(-156.5, 21.8), zoom = 8, maptype = "hybrid")
-# saveRDS(object = gm, file = "raw_data/ggmap_of_aloha.rds")
+# saveRDS(object = gm, file = "metadata/raw_data/ggmap_of_aloha.rds")
 
@@
   magick::image_trim() %>%
-  magick::image_write("raw_data/chl_pdf_url.png", format = "png")
+  magick::image_write("metadata/raw_data/chl_pdf_url.png", format = "png")
 ```
 
-![](raw_data/chl_pdf_url.png)
+![](metadata/raw_data/chl_pdf_url.png)
 
 ```{r From SCOPE FTP}
-para_cast_times <- "raw_data/paragon1.sum" %>%
+para_cast_times <- "metadata/raw_data/paragon1.sum" %>%
   read_table(skip = 3, col_names = FALSE, col_types = cols(.default = col_character())) %>%
@@
 # 9    variable not measured during this cast
-raw_gof <- read_table("raw_data/paragon1.gof", skip = 4, col_names = gof_names,
+raw_gof <- read_table("metadata/raw_data/paragon1.gof", skip = 4, col_names = gof_names,
                       col_types = cols(.default = col_double()))
@@
 
-gm <- readRDS("raw_data/ggmap_of_aloha.rds")
+gm <- readRDS("metadata/raw_data/ggmap_of_aloha.rds")
 stations_toplot <- para_core %>%
@@
 
-seaflowdata <- readxl::read_excel("raw_data/SeaFlow_dataset_v1.4.xlsx")
+seaflowdata <- readxl::read_excel("metadata/raw_data/SeaFlow_dataset_v1.4.xlsx")
 seaflowdata %>%
@@
   mutate(shortname=str_remove(shortname, ".*_Smp_")) %>%
+  mutate(samp_type="Smp") %>%
   mutate(IS_type=str_extract(filepath, "wIS|noIS")) %>%
@@
 
-samp_filedata <- samp_filedata %>%
+other_filedata <- "mzMLs" %>%
+  list.files(recursive = TRUE, full.names = TRUE) %>%
+  str_subset("Smp.*mzML", negate = TRUE) %>%
+  str_subset("MSMS", negate = TRUE) %>%
+  str_subset("TMAO", negate = TRUE) %>%
+  data.frame(filepath=.) %>%
+  mutate(filename=basename(filepath)) %>%
+  mutate(shortname=str_remove(filename, "-wIS|-noIS-neg|-noIS-pos")) %>%
+  mutate(shortname=str_remove(shortname, ".*_Poo_|.*_Std_(4uMStds)?")) %>%
+  mutate(samp_type=str_extract(filename, "Poo|Std")) %>%
+  mutate(IS_type=str_extract(filepath, "wIS|noIS")) %>%
+  mutate(polarity=str_extract(filepath, "neg|pos")) %>%
+  mutate(amendment=str_extract(shortname, "Nit|Amm|Arg|GMP")) %>%
+  fill(IS_type, amendment, .direction = "down")
+
+all_filedata <- samp_filedata %>%
+  bind_rows(other_filedata) %>%
   left_join(station_samp_map)
@@
 
+## Get file timestamps using RaMS
+
+```{r}
+mzml_timestamps <- grabMSdata(all_filedata$filepath, grab_what = "metadata") %>%
+  pluck("metadata") %>%
+  select(filename, timestamp)
+parametadata <- left_join(all_filedata, mzml_timestamps)
+```
+
+## Write out
+
 ```{r write out}
-write_csv(samp_filedata, file = "metadata/samp_filedata.csv")
+write_csv(parametadata, file = "metadata/parametadata.csv")
 ```

### 2023-09-22 metadata/parametadata_control.R

@@ -1,171 +1,0 @@
-## ----setup, include=FALSE-------------------------------------------------------------------
-knitr::opts_chunk$set(echo = TRUE)
-library(tidyverse)
-library(ggmap)
-library(RaMS)
-
-
-## ----download raw data, eval=!file.exists("metadata/raw_data/paragon1.gof")-----------------
-sum_url <- "http://scope.soest.hawaii.edu/data/scope2021/paragon1.sum"
-download.file(sum_url, destfile = "metadata/raw_data/paragon1.sum")
-
-gof_url <- "http://scope.soest.hawaii.edu/FTP/scope/water/paragon1.gof"
-download.file(gof_url, destfile = "metadata/raw_data/paragon1.gof")
-
-seaflow_url <- "http://scope.soest.hawaii.edu/data/scope2021/SeaFlow_dataset_v1.4.xlsx"
-download.file(seaflow_url, destfile = "metadata/raw_data/SeaFlow_dataset_v1.4.xlsx", mode = "wb")
-
-# register_google("AIzaSyBVNJV0Myoy3M50IBBDIjXZhUyU8Mnagvs")
-# gm <- get_map(location = c(-156.5, 21.8), zoom = 8, maptype = "hybrid")
-# saveRDS(object = gm, file = "metadata/raw_data/ggmap_of_aloha.rds")
-
-magick::image_read("http://scope.soest.hawaii.edu/data/scope2021/Paragon1CTDchlp.pdf", density = 300) %>%
-  magick::image_trim() %>%
-  magick::image_write("metadata/raw_data/chl_pdf_url.png", format = "png")
-
-
-## ----From SCOPE FTP-------------------------------------------------------------------------
-para_cast_times <- "metadata/raw_data/paragon1.sum" %>%
-  read_table(skip = 3, col_names = FALSE, col_types = cols(.default = col_character())) %>%
-  mutate(mdyhms=paste(X3, X4, X5, X6)) %>%
-  mutate(time=as.POSIXct(strptime(mdyhms, "%b %d  %Y %X", tz="UTC"))) %>%
-  select(station=X1, cast=X2, time) %>%
-  mutate(station=as.numeric(station), cast=as.numeric(cast))
-
-gof_names <- c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
-              "sal", "ctd_oxy", "ctd_chl", "theta", "sigma", "oxy_um", "dic", "alk", "ph",
[…]

### 2023-09-22 untargeted/paruntargeted_control.Rmd
Beginning to track untargeted folder

---
title: "Untargeted control doc"
author: "William Kumler"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, purl=FALSE)
options(pillar.sigfig=7)
options(readr.num_columns = 0)
options(readr.show_col_types = FALSE)
library(tidyverse)
library(data.table)
library(xcms)
library(RaMS)
library(arrow)

parametadata <- read_csv("metadata/samp_filedata.csv") %>% distinct()

# polarity <- "pos"
polarity <- "neg"
amendment <- "Amm"
# amendment <- "Nit"
# amendment <- "GMP"
# amendment <- "Arg"
IS_type <- "noIS"
# IS_type <- "wIS"

output_folder <- paste(amendment, polarity, IS_type, "output", sep = "_") %>%
  paste0("untargeted/", ., "/")
if(!dir.exists(output_folder))dir.create(output_folder)
output_csv_write <- function(object){
  obj_name <- deparse(substitute(object))
  write_csv(object, file = paste0(output_folder, obj_name, ".csv"))
}

xcms_file_data <- parametadata %>%
  filter(amendment==!!amendment) %>%
  filter(polarity==!!polarity) %>%
  filter(IS_type==!!IS_type) %>%
  mutate(samp_group=paste(amendment, depth, startime, timepoint, sep="-"))
```

```{r xcms}
parallel_param <- SnowParam(
  workers = parallel::detectCores()-1, 
  tasks = nrow(xcms_file_data), 
  progressbar = TRUE
)

register(BPPARAM = SerialParam(progressbar = TRUE))
msnexp <- readMSData(
  files = xcms_file_data$filepath, 
  pdata = new("NAnnotatedDataFrame", xcms_file_data), 
  msLevel. = 1, 
  mode = "onDisk"
)

cwp <- CentWaveParam(
  ppm = 5, 
  peakwidth = c(20, 80), 
  prefilter = c(3, 1e5), 
  snthresh = 0, 
  verboseColumns = TRUE, 
  extendLengthMSW = TRUE, 
  integrate = 2,
  verboseBetaColumns = TRUE
)
msnexp_withpeaks <- findChromPeaks(msnexp, cwp, BPPARAM=parallel_param)

# mid_file <- xcms_file_data %>%
#   arrange(abs(timestamp-mean(timestamp))) %>%
#   select(filename, timestamp) %>%
#   slice(1) %>%
#   pull(filename)
# mid_file_idx <- which(basename(fileNames(msnexp_filled))==mid_file)
# obp <- ObiwarpParam(binSize = 1, response = 1, distFun = "cor_opt", 
#                     centerSample = mid_file_idx)
# msnexp_rtcor <- adjustRtime(msnexp_withpeaks, obp)

register(BPPARAM = SerialParam(progressbar = TRUE))
pdp <- PeakDensityParam(
  sampleGroups = xcms_file_data$samp_group, 
  bw = 12, 
  minFraction = 0, 
  binSize = 0.001, 
  minSamples = 2
)
msnexp_grouped <- groupChromPeaks(msnexp_withpeaks, pdp)

fpp <- FillChromPeaksParam(ppm = 10)
msnexp_filled <- fillChromPeaks(msnexp_grouped, fpp, BPPARAM=parallel_param)

saveRDS(msnexp_filled, paste0(output_folder, "msnexp_filled.rds"))
```

```{r xcms QC, eval=FALSE}
msnexp_filled <- readRDS(paste0(output_folder, "msnexp_filled.rds"))

rt_corrections <- data.frame(initrt=rtime(msnexp_rtcor, adjusted = FALSE),
                             adjrt=rtime(msnexp_rtcor, adjusted = TRUE)) %>%
  rownames_to_column("fileidx") %>%
  mutate(fileidx=as.numeric(str_extract(fileidx, "(?<=F)\\d+"))) %>%
  mutate(filename=basename(fileNames(msnexp_rtcor)[fileidx]))

rt_corrections %>%
  left_join(xcms_file_data) %>%
  ggplot() +
  geom_line(aes(x=adjrt, y=adjrt-initrt, group=filename, color=timestamp))

bet_eic <- dset %>%
  filter(mz%between%pmppm(118.0865, 10)) %>%
  dplyr::collect()
bet_eic %>%
  mutate(rt=round(rt*60, digits=7)) %>%
  left_join(rt_corrections %>% mutate(initrt=round(initrt, digits=7)), 
            by=c("filename", rt="initrt")) %>%
  # filter(is.na(adjrt))
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))
```

```{r making peak_data}
msnexp_filled <- readRDS(paste0(output_folder, "msnexp_filled.rds"))
peak_data_long <- msnexp_filled %>%
  chromPeaks() %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  mutate(peakidx=row_number())
peak_data <- msnexp_filled %>%
  featureDefinitions() %>%
  as.data.frame() %>%
  select(mzmed, rtmed, npeaks, peakidx) %>%
  rownames_to_column("id") %>%
  unnest_longer(peakidx) %>%
  rename_with(~paste0("feat_", .x), .cols = -peakidx) %>%
  dplyr::rename(feature="feat_id") %>%
  left_join(peak_data_long, by = join_by(peakidx)) %>%
  mutate(filename=basename(fileNames(msnexp_filled))[sample])
output_csv_write(peak_data)
```

### 2023-09-22 mzMLs/msconvert.R
Now tracking msconvert.R
For converting .raw files (on Synology drives) to .mzML

@@ -1,0 +1,40 @@
+
+folder_paths <- c(
+  "Z:/1_QEdata/2021/2021_Will_PARAGON/210813_Will_NIT-Expt_PARAGON_HILIC",
+  "Z:/1_QEdata/2021/2021_Will_PARAGON/210916_Will_Amm-Expt_PARAGON_HILIC",
+  "Z:/1_QEdata/2022/220715_Will_GMP-Expt_PARAGON_HILIC",
+  "Z:/1_QEdata/2022/220729_Will_Arg-Expt_PARAGON_HILIC"
+)
+
+sapply(folder_paths, function(path_i){
+  sapply(c("positive", "negative"), function(polarity){
+    if(path_i=="Z:/1_QEdata/2021/2021_Will_PARAGON/210813_Will_NIT-Expt_PARAGON_HILIC" & polarity=="positive"){
+      return(0)
+    }
+    message(paste("Converting", basename(path_i), polarity))
+    short_pol <- gsub("[ia]tive", "", polarity)
+    match_pattern <- paste0("wIS.raw|noIS-", short_pol, ".*.raw|_Std_.*raw")
+    files_to_convert <- list.files(path_i, pattern = match_pattern, full.names = TRUE)
+    mscommand <- paste(
+      "msconvert",
+      paste(files_to_convert, collapse = " "),
+      "--mzML",
+      paste0('-o ', short_pol),
+      '--filter "peakPicking true 1-"',
+      paste('--filter "polarity', polarity, '"')
+    )
+    system(mscommand)
+    
+    msmsmatch_pattern <- paste0("_DDA", short_pol, ".*.raw")
+    msms_to_convert <- list.files(path_i, pattern = msmsmatch_pattern, full.names = TRUE)
+    msmscommand <- paste(
+      "msconvert",
+      paste(msms_to_convert, collapse = " "),
+      "--mzML",
+      paste0('-o ', short_pol, "/MSMS"),
+      '--filter "peakPicking true 1-"',
+      paste('--filter "polarity', polarity, '"')
+    )
+    system(msmscommand)
+  })
+})

### 2023-09-25 metadata/parametadata_control.Rmd
Saving metadata timepoints separately

@@
   bind_rows(other_filedata) %>%
-  left_join(station_samp_map)
+  left_join(station_samp_map) %>%
+  distinct()
 ```
@@
 ```{r}
+if(!file.exists("metadata/raw_data/mzml_timestamps.csv")){
   mzml_timestamps <- grabMSdata(all_filedata$filepath, grab_what = "metadata") %>%
@@
     select(filename, timestamp)
+  write_csv(mzml_timestamps, "metadata/raw_data/mzml_timestamps.csv")
+}
+mzml_timestamps <- read_csv("metadata/raw_data/mzml_timestamps.csv")
 parametadata <- left_join(all_filedata, mzml_timestamps)

### 2023-09-25 metadata/parametadata_control.Rmd
Now only making metadata for mzML files, oops

@@
 other_filedata <- "mzMLs" %>%
-  list.files(recursive = TRUE, full.names = TRUE) %>%
+  list.files(recursive = TRUE, full.names = TRUE, pattern = ".mzML") %>%
   str_subset("Smp.*mzML", negate = TRUE) %>%

### 2023-09-25 metadata/parametadata_control.Rmd
Cleaning up parametadata
Adding static standards sheet

@@
   bind_rows(other_filedata) %>% 
-  left_join(station_samp_map) %>%
-  distinct()
+  left_join(station_samp_map)
 ```
@@
 }
-mzml_timestamps <- read_csv("metadata/raw_data/mzml_timestamps.csv")
-parametadata <- left_join(all_filedata, mzml_timestamps)
+mzml_timestamps <- read_csv("metadata/raw_data/mzml_timestamps.csv") %>%
+  distinct()
+parametadata <- all_filedata %>%
+  left_join(mzml_timestamps, by="filename")
 ```
 
-## Write out
+## Write out clean parametadata file
 
@@
 ```
+
+## Create a clean and static list of standards
+
+```{r stans}
+base_url <- "https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards/"
+mix_stans <- base_url %>%
+  paste0("1ddd97186fa9df1646941aec903dc64a65ef6cfe/Ingalls_Lab_Standards.csv") %>%
+  read_csv() %>%
+  select(compound_name=Compound_Name, mix=HILIC_Mix) %>%
+  drop_na() %>%
+  distinct()
+custom_stans <- base_url %>%
+  paste0("master/data_extra/custom_stans.csv") %>%
+  read_csv() %>%
+  select(compound_name=Compound_Name, mz, rt=RT_minute, z,
+         conc_um=Concentration_uM, compound_type=Compound_Type,
+         formula=Empirical_Formula, date_added=Date_Added, mix=HILIC_Mix) %>%
+  mutate(column="HILIC") %>%
+  mutate(date_added=230101) %>%
+  mutate(compound_name=str_remove(compound_name, "\\?$"))
+
+all_stans <- paste0(base_url, "master/Ingalls_Lab_Standards.csv") %>%
+  read_csv() %>%
+  select(compound_name=Compound_Name, mz, rt=RT_minute, z, column=Column, 
+         conc_um=Concentration_uM, compound_type=Compound_Type,
+         formula=Empirical_Formula, date_added=Date_Added) %>%
+  filter(column=="HILIC") %>%
+  left_join(mix_stans, by="compound_name") %>%
+  bind_rows(custom_stans) %>%
+  mutate(date_added=ifelse(is.na(date_added), 230101, date_added)) %>%
+  mutate(compound_name=ifelse(date_added<210813, compound_name,
+                              paste0(compound_name, "?"))) %>%
+  mutate(polarity=ifelse(z<0, "neg", "pos")) %>%
+  arrange(mz, rt)
+write_csv(all_stans, "metadata/all_stans.csv")
+```

### 2023-09-26 untargeted/paruntargeted_control.Rmd
Adding new parametadata and stans plus masscheck function

@@
 library(data.table)
+library(arrow)
 library(xcms)
 library(RaMS)
-library(arrow)
 
-parametadata <- read_csv("metadata/samp_filedata.csv") %>% distinct()
-
-# polarity <- "pos"
-polarity <- "neg"
+polarity <- "pos"
+# polarity <- "neg"
 amendment <- "Amm"
@@
 # IS_type <- "wIS"
+
+parametadata <- read_csv("metadata/parametadata.csv") %>% 
+  filter(polarity==!!polarity) %>%
+  filter(amendment==!!amendment) %>%
+  filter(IS_type==!!IS_type)
+
+chosen_stans <- read_csv("metadata/all_stans.csv") %>%
+  filter(polarity==!!polarity)
 
@@
 msnexp_filled <- readRDS(paste0(output_folder, "msnexp_filled.rds"))
-
-rt_corrections <- data.frame(initrt=rtime(msnexp_rtcor, adjusted = FALSE),
-                             adjrt=rtime(msnexp_rtcor, adjusted = TRUE)) %>%
-  rownames_to_column("fileidx") %>%
-  mutate(fileidx=as.numeric(str_extract(fileidx, "(?<=F)\\d+"))) %>%
-  mutate(filename=basename(fileNames(msnexp_rtcor)[fileidx]))
-
-rt_corrections %>%
-  left_join(xcms_file_data) %>%
-  ggplot() +
-  geom_line(aes(x=adjrt, y=adjrt-initrt, group=filename, color=timestamp))
-
-bet_eic <- dset %>%
-  filter(mz%between%pmppm(118.0865, 10)) %>%
-  dplyr::collect()
-bet_eic %>%
-  mutate(rt=round(rt*60, digits=7)) %>%
-  left_join(rt_corrections %>% mutate(initrt=round(initrt, digits=7)), 
-            by=c("filename", rt="initrt")) %>%
-  # filter(is.na(adjrt))
-  ggplot() +
-  geom_line(aes(x=rt, y=int, group=filename))
 ```
 
-```{r making peak_data}
+```{r making peak_data and feat_data}
 msnexp_filled <- readRDS(paste0(output_folder, "msnexp_filled.rds"))
@@
 output_csv_write(peak_data)
+
+all_feat_data <- peak_data %>%
+  left_join(parametadata) %>%
+  filter(samp_type=="Smp") %>%
+  group_by(feature) %>%
+  summarize(mzmed=unique(feat_mzmed), rtmed=unique(feat_rtmed)/60, n=n(),
+            med_cor=median(beta_cor, na.rm=TRUE), 
+            med_snr=median(beta_snr, na.rm=TRUE))
+# all_feat_data %>%
+#   # filter(med_cor>0.75) %>%
+#   mutate(n_group=cut(n, breaks = c(0, 2, 10, 40, 60, 62, 100, 130, Inf))) %>%
+#   ggplot(aes(label=feature)) +
+#   geom_point(aes(x=med_cor, y=med_snr, color=n_group))
+
+best_feat_data <- all_feat_data %>%
+  filter(med_cor>0.75) %>%
+  select(-med_cor, -med_snr)
+output_csv_write(best_feat_data)
 ```
+
+```{r diagnostic plots}
+mass_check <- function(mz_i){
+  mass_feats <- best_feat_data %>% 
+    filter(mzmed%between%pmppm(mz_i, 5)) %>%
+    print()
+  chosen_eic <- open_dataset("tmzMLs/pqds") %>%
+    filter(polarity==!!polarity) %>%
+    filter(amendment==!!amendment) %>%
+    filter(mz%between%pmppm(mz_i, 5)) %>%
+    filter(rt%between%c(2, 20)) %>%
+    dplyr::collect()
+  ggtitle_i <- ifelse(is.null(names(mz_i)), mz_i, names(mz_i))
+  chosen_eic %>%
+    left_join(parametadata, by = join_by(filename, samp_type, amendment, polarity)) %>%
+    mutate(samp_type=str_extract(filename, "Mix\\d|Deep|Surf|Poo|Blk")) %>%
+    filter(!str_detect(filename, "Mat")) %>%
+    mutate(plot_facet=ifelse(samp_type%in%c("Mix1", "Mix2"), "Mixes", "Samples")) %>%
+    ggplot() +
+    geom_line(aes(x=rt, y=int, group=filename, color=samp_type)) +
+    facet_wrap(~plot_facet, scales = "free_y", ncol=1) +
+    geom_vline(aes(xintercept=rtmed), color="black", data = mass_feats) +
+    ggtitle(ggtitle_i) +
+    theme_bw()
+}
+mass_check(c("Carn/THB/HMSB"=162.1125))
+mass_check(c("ACH/TMAB"=146.1181))
+mass_check(c("Ala/bAla/Sarc"=90.0555))
+mass_check(c("Arg"=175.119501))
+mass_check(c("Leu/Ile/TMAP"=132.1025))
+mass_check(c("15N Ile"=132.1025+0.997))
+```

### 2023-09-26 metadata/parametadata_control.Rmd
Removing turicine from standards list

@@
   left_join(mix_stans, by="compound_name") %>%
+  filter(compound_name!="Turicine") %>%
   bind_rows(custom_stans) %>%

### 2023-09-26 untargeted/paruntargeted_control.Rmd
Attempt at automating annotation before deciding to switch to manual

@@
 
-```{r making peak_data and feat_data}
+```{r making peak_data and best_feat_data}
 msnexp_filled <- readRDS(paste0(output_folder, "msnexp_filled.rds"))
@@
 
+if(amendment=="Amm"){
+  if(polarity=="pos"){
+    rescue_feats <- c("FT3015", "FT1319", "FT1321")
+  } else if(polarity=="neg"){
+
+  }
+}
+if(amendment=="Nit"){
+  if(polarity=="pos"){
+
+  } else if(polarity=="neg"){
+
+  } else {
+    stop("Unknown polarity")
+  }
+}
+if(amendment=="GMP"){
+  if(polarity=="pos"){
+
+  } else if(polarity=="neg"){
+
+  } else {
+    stop("Unknown polarity")
+  }
+}
+if(amendment=="Arg"){
+  if(polarity=="pos"){
+
+  } else if(polarity=="neg"){
+
+  } else {
+    stop("Unknown polarity")
+  }
+}
+
 best_feat_data <- all_feat_data %>%
-  filter(med_cor>0.75) %>%
+  filter(med_cor>0.75 | feature%in%rescue_feats) %>%
   select(-med_cor, -med_snr)
@@
 ```{r diagnostic plots}
-mass_check <- function(mz_i){
+mass_check <- function(mz_i, all_samps=FALSE){
   mass_feats <- best_feat_data %>% 
@@
     print()
+  if(all_samps){
     chosen_eic <- open_dataset("tmzMLs/pqds") %>%
@@
       dplyr::collect()
+  } else {
+    chosen_eic <- open_dataset("tmzMLs/pqds") %>%
+      filter(polarity==!!polarity) %>%
+      filter(amendment==!!amendment) %>%
+      filter(samp_type=="Poo" | samp_type=="Std") %>%
+      filter(mz%between%pmppm(mz_i, 5)) %>%
+      filter(rt%between%c(2, 20)) %>%
+      dplyr::collect()
+  }
   ggtitle_i <- ifelse(is.null(names(mz_i)), mz_i, names(mz_i))
@@
 mass_check(c("Carn/THB/HMSB"=162.1125))
-mass_check(c("ACH/TMAB"=146.1181))
+mass_check(c("ACH/TMAB"=146.1181), all_samps = TRUE)
 mass_check(c("Ala/bAla/Sarc"=90.0555))
@@
 mass_check(c("15N Ile"=132.1025+0.997))
+mass_check(60.04510)
+mass_check(c("Ectoine"=143.0821))
+mass_check(c("Betonicine"=160.0974))
+mass_check(c("Adenine"=136.0623))
+mass_check(c("Pyroglut"=130.0504))
+mass_check(c("Glutamine"=147.076968))
+mass_check(c("Cytosine"=244.093347))
+mass_check(c("Proline"=116.0712))
 ```
+
+```{r annotate}
+all_stans <- read_csv("metadata/all_stans.csv") %>%
+  filter(polarity==!!polarity)
+
+stan_mix_data <- peak_data %>% 
+  select(feature, filename, into) %>%
+  left_join(parametadata) %>%
+  filter(samp_type=="Std") %>%
+  filter(!str_detect(filename, "Mat")) %>%
+  mutate(mix=str_extract(filename, "Mix\\d")) %>%
+  select(feature, mix, into) %>%
+  pivot_wider(names_from = mix, values_from = into, values_fn = mean)
+
+stan_annotations <- best_feat_data %>%
+  pmap(function(...){
+    row_data <- list(...)
+    all_stans %>%
+      select(compound_name, mz, rt) %>%
+      filter(mz%between%pmppm(row_data$mzmed, 10)) %>%
+      mutate(feature=row_data$feature, mzmed=row_data$mzmed,
+             rtmed=row_data$rtmed, n=row_data$n)
+  }, .progress = TRUE) %>%
+  bind_rows() %>%
+  arrange(mz, rt, rtmed) %>%
+  left_join(stan_mix_data) %>%
+  left_join(all_stans %>% select(compound_name, mix)) %>%
+  mutate(mix_ratio=ifelse(mix=="Mix1", Mix1/Mix2, Mix2/Mix1)) %>%
+  filter(mix_ratio>3 | is.na(mix_ratio)) %>%
+  # select(feature, compound_name)
+  ungroup()
+
+missed_stans <- all_stans %>%
+  anti_join(stan_annotations %>% select(compound_name)) %>%
+  print(n=Inf)
+
+stan_choose <- function(df, compound_name_i, feature_i){
+  anti_df <- df %>%
+    filter(compound_name==compound_name_i) %>%
+    filter(feature!=feature_i)
+  join_vec <- join_by(compound_name, mz, rt, feature, mzmed, rtmed, n, 
+                      Mix1, Mix2, mix, mix_ratio)
+  df %>% anti_join(anti_df, by = join_vec)
+}
+stan_add <- function(df, compound_name_i, feature_i){
+  add_row(df, compound_name=compound_name_i, feature=feature_i)
+}
+stan_remove <- function(df, compound_name_i){
+  df %>% filter(!compound_name%in%compound_name_i)
+}
+if(amendment=="Amm"){
+  if(polarity=="pos"){
+    stan_annotations %>%
+      stan_choose("Sarcosine", "FT0348") %>%
+      stan_choose("beta-Alanine", "FT0347") %>%
+      stan_choose("Cytosine", "FT0769") %>%
+      stan_add("Cytosine from cytidine", "FT0770") %>%
+      stan_choose("L-Proline", "FT0862") %>%
+      stan_remove("5-Oxoproline") %>%
+      stan_choose("L-Leucine", "FT1244") %>%
+      stan_choose("L-Isoleucine", "FT1248") %>%
+      stan_choose("Proline betaine", "FT1572") %>%
+      stan_choose("Betonicine", "FT1842") %>%
+      stan_choose("Carnitine", "FT1857") %>%
+      stan_remove("Dimethylglycine") %>%
+      stan_remove("4-Aminobutyric acid") %>%
+      stan_remove("Creatinine from creatine?") %>%
+      stan_remove("L-Valine") %>%
+      stan_remove(c("L-Threonine", "L-Homoserine")) %>%
+      stan_add("Threonine/Homoserine", "FT0953") %>%
+      # stan_remove("Allopurinol") %>%
+      stan_remove("Guanine from guanosine?") %>%
+      stan_add("L-Glutamic acid", "FT1635") %>%
+      group_by(feature) %>%
+      group_by(compound_name) %>%
+      # filter(n()>1)
+      ungroup() -> stan_annotations
+  } else if(polarity=="neg"){
+
+  } else {
+    stop("Unknown polarity")
+  }
+}
+if(amendment=="Nit"){
+  if(polarity=="pos"){
+
+  } else if(polarity=="neg"){
+
+  } else {
+    stop("Unknown polarity")
+  }
+}
+if(amendment=="GMP"){
+  if(polarity=="pos"){
+
+  } else if(polarity=="neg"){
+
+  } else {
+    stop("Unknown polarity")
+  }
+}
+if(amendment=="Arg"){
+  if(polarity=="pos"){
+
+  } else if(polarity=="neg"){
+
+  } else {
+    stop("Unknown polarity")
+  }
+}
+
+mass_check(c("Cytidine"=244.0933))
+```

### 2023-09-26 untargeted/paruntargeted_control.Rmd
Switching to fully-manual Excel annotation

@@
 
-stan_mix_data <- peak_data %>% 
-  select(feature, filename, into) %>%
-  left_join(parametadata) %>%
-  filter(samp_type=="Std") %>%
+pdf(paste0(output_folder, "anno_eics.pdf"), width = 8, height = 4)
+all_stans %>%
+  filter(compound_type!="Internal Standard") %>%
+  mutate(mz=ifelse(compound_name=="Carnitine", mz[which(compound_name=="Threonine betaine?")], mz)) %>%
+  group_by(mz) %>%
+  summarise(cmpd_lab=paste0(compound_name, " (", mix, ",", rt, ")", collapse = ", ")) %>%
+  # head(6) %>%
+  pwalk(function(...){
+    row_data <- list(...)
+    mass_feats <- best_feat_data %>% filter(mzmed%between%pmppm(row_data$mz, 5))
+    chosen_eic <- open_dataset("tmzMLs/pqds") %>%
+      filter(polarity==!!polarity) %>%
+      filter(amendment==!!amendment) %>%
+      filter(samp_type=="Poo" | samp_type=="Std") %>%
+      filter(mz%between%pmppm(row_data$mz, 5)) %>%
+      filter(rt%between%c(2, 20)) %>%
+      dplyr::collect()
+    gp_plot <- chosen_eic %>%
+      left_join(parametadata, by = join_by(filename, samp_type, amendment, polarity)) %>%
+      mutate(samp_type=str_extract(filename, "Mix\\d|Deep|Surf|Poo|Blk")) %>%
       filter(!str_detect(filename, "Mat")) %>%
-  mutate(mix=str_extract(filename, "Mix\\d")) %>%
-  select(feature, mix, into) %>%
-  pivot_wider(names_from = mix, values_from = into, values_fn = mean)
-
-stan_annotations <- best_feat_data %>%
-  pmap(function(...){
-    row_data <- list(...)
-    all_stans %>%
-      select(compound_name, mz, rt) %>%
-      filter(mz%between%pmppm(row_data$mzmed, 10)) %>%
-      mutate(feature=row_data$feature, mzmed=row_data$mzmed,
-             rtmed=row_data$rtmed, n=row_data$n)
-  }, .progress = TRUE) %>%
-  bind_rows() %>%
-  arrange(mz, rt, rtmed) %>%
-  left_join(stan_mix_data) %>%
-  left_join(all_stans %>% select(compound_name, mix)) %>%
-  mutate(mix_ratio=ifelse(mix=="Mix1", Mix1/Mix2, Mix2/Mix1)) %>%
-  filter(mix_ratio>3 | is.na(mix_ratio)) %>%
-  # select(feature, compound_name)
-  ungroup()
-
-missed_stans <- all_stans %>%
-  anti_join(stan_annotations %>% select(compound_name)) %>%
-  print(n=Inf)
-
-stan_choose <- function(df, compound_name_i, feature_i){
-  anti_df <- df %>%
-    filter(compound_name==compound_name_i) %>%
-    filter(feature!=feature_i)
-  join_vec <- join_by(compound_name, mz, rt, feature, mzmed, rtmed, n, 
-                      Mix1, Mix2, mix, mix_ratio)
-  df %>% anti_join(anti_df, by = join_vec)
-}
-stan_add <- function(df, compound_name_i, feature_i){
-  add_row(df, compound_name=compound_name_i, feature=feature_i)
-}
-stan_remove <- function(df, compound_name_i){
-  df %>% filter(!compound_name%in%compound_name_i)
-}
-if(amendment=="Amm"){
-  if(polarity=="pos"){
-    stan_annotations %>%
-      stan_choose("Sarcosine", "FT0348") %>%
-      stan_choose("beta-Alanine", "FT0347") %>%
-      stan_choose("Cytosine", "FT0769") %>%
-      stan_add("Cytosine from cytidine", "FT0770") %>%
-      stan_choose("L-Proline", "FT0862") %>%
-      stan_remove("5-Oxoproline") %>%
-      stan_choose("L-Leucine", "FT1244") %>%
-      stan_choose("L-Isoleucine", "FT1248") %>%
-      stan_choose("Proline betaine", "FT1572") %>%
-      stan_choose("Betonicine", "FT1842") %>%
-      stan_choose("Carnitine", "FT1857") %>%
-      stan_remove("Dimethylglycine") %>%
-      stan_remove("4-Aminobutyric acid") %>%
-      stan_remove("Creatinine from creatine?") %>%
-      stan_remove("L-Valine") %>%
-      stan_remove(c("L-Threonine", "L-Homoserine")) %>%
-      stan_add("Threonine/Homoserine", "FT0953") %>%
-      # stan_remove("Allopurinol") %>%
-      stan_remove("Guanine from guanosine?") %>%
-      stan_add("L-Glutamic acid", "FT1635") %>%
-      group_by(feature) %>%
-      group_by(compound_name) %>%
-      # filter(n()>1)
-      ungroup() -> stan_annotations
-  } else if(polarity=="neg"){
-
-  } else {
-    stop("Unknown polarity")
-  }
-}
-if(amendment=="Nit"){
-  if(polarity=="pos"){
-
-  } else if(polarity=="neg"){
-
-  } else {
-    stop("Unknown polarity")
-  }
-}
-if(amendment=="GMP"){
-  if(polarity=="pos"){
-
-  } else if(polarity=="neg"){
-
-  } else {
-    stop("Unknown polarity")
-  }
-}
-if(amendment=="Arg"){
-  if(polarity=="pos"){
-
-  } else if(polarity=="neg"){
-
-  } else {
-    stop("Unknown polarity")
-  }
-}
-
-mass_check(c("Cytidine"=244.0933))
+      mutate(plot_facet=ifelse(samp_type%in%c("Mix1", "Mix2"), "Mixes", "Samples")) %>%
+      ggplot() +
+      geom_line(aes(x=rt, y=int, group=filename, color=samp_type)) +
+      facet_wrap(~plot_facet, scales = "free_y", ncol=1) +
+      geom_vline(aes(xintercept=rtmed), color="black", data = mass_feats) +
+      geom_text(aes(x=rtmed, y=Inf, label=feature), 
+                angle=90, hjust=1, vjust=0.5, data = mass_feats) +
+      ggtitle(row_data$cmpd_lab) +
+      scale_x_continuous(breaks = 2:20) +
+      theme_bw() +
+      theme(legend.position = c(0.9, 0.8), legend.title = element_blank(),
+            axis.title = element_blank())
+    print(gp_plot)
+  }, .progress = TRUE)
+dev.off()
 ```

### 2023-09-26 untargeted/paruntargeted_control.Rmd
Beginning manual annotations

@@
 all_stans <- read_csv("metadata/all_stans.csv") %>%
-  filter(polarity==!!polarity)
+  mutate(polarity=factor(polarity, levels=c("pos", "neg"))) %>%
+  arrange(polarity)
+# map(c("Amm", "Nit", "GMP", "Arg"), function(amendment_i){
+#   all_stans %>%
+#       mutate(amendment=amendment_i) %>%
+#       select(amendment, polarity, compound_name, mz, rt, date_added, mix) %>%
+#       write_csv("untargeted/anno_template.csv", append = TRUE)
+# })
+# # Gets copy-pasted into the Excel file manual_annotations.xlsx
 
@@
   group_by(mz) %>%
-  summarise(cmpd_lab=paste0(compound_name, " (", mix, ",", rt, ")", collapse = ", ")) %>%
+  summarise(cmpd_lab=paste0(compound_name, " (", mix, " ,", rt, ")", collapse = ", ")) %>%
   # head(6) %>%

### 2023-09-26 untargeted/paruntargeted_control.Rmd
Dropping internal stans and fixing labels

--- 2023-09-26 untargeted/paruntargeted_control.Rmd
@@
 all_stans <- read_csv("metadata/all_stans.csv") %>%
-  filter(polarity==!!polarity)
+  mutate(polarity=factor(polarity, levels=c("pos", "neg"))) %>%
+  arrange(polarity) %>%
+  filter(compound_type!="Internal Standard")
+# map(c("Amm", "Nit", "GMP", "Arg"), function(amendment_i){
+#   all_stans %>%
+#       mutate(amendment=amendment_i) %>%
+#       select(amendment, polarity, compound_name, mz, rt, date_added, mix) %>%
+#       write_csv("untargeted/anno_template.csv", append = TRUE)
+# })
+# # Gets copy-pasted into the Excel file manual_annotations.xlsx
 
@@
 all_stans %>%
-  filter(compound_type!="Internal Standard") %>%
+  filter(polarity==!!polarity) %>%
   mutate(mz=ifelse(compound_name=="Carnitine", mz[which(compound_name=="Threonine betaine?")], mz)) %>%
   group_by(mz) %>%
-  summarise(cmpd_lab=paste0(compound_name, " (", mix, ",", rt, ")", collapse = ", ")) %>%
-  # head(6) %>%
+  summarise(cmpd_lab=paste0(compound_name, " (", mix, " ,", rt, ")", collapse = ", ")) %>%
+  slice(2) %>%
   pwalk(function(...){
@@
       geom_text(aes(x=rtmed, y=Inf, label=feature), 
-                angle=90, hjust=1, vjust=0.5, data = mass_feats) +
+                angle=90, hjust=1, vjust=1, data = mass_feats) +
       ggtitle(row_data$cmpd_lab) +

### 2023-09-26 arrowscript.R
Now tracking arrowscript and eicscript

@@ -1,0 +1,38 @@
+library(data.table)
+library(RaMS)
+library(arrow)
+
+parametadata <- fread("metadata/parametadata.csv")[
+  IS_type=="noIS",c("filepath", "filename", "samp_type", "amendment", "polarity")
+]
+parametadata[, .N, by=c("samp_type", "amendment", "polarity")]
+
+for(samp_type_i in unique(parametadata$samp_type)){
+  for(amendment_i in unique(parametadata$amendment)){
+    for(polarity_i in unique(parametadata$polarity)){
+      ms_files <- parametadata[samp_type==samp_type_i][amendment==amendment_i][polarity==polarity_i]$filepath
+      # ms_files <- head(ms_files, 3)
+      msdata <- grabMSdata(ms_files, grab_what = "MS1")
+      parq_dir <- paste0("tmzMLs/pqds",
+                         "/samp_type=", samp_type_i, 
+                         "/amendment=", amendment_i, 
+                         "/polarity=", polarity_i)
+      if(!dir.exists(parq_dir))dir.create(parq_dir, recursive = TRUE)
+      parq_name <- paste0(parq_dir, "/part-0.parquet")
+      print(parq_name)
+      write_parquet(msdata$MS1, parq_name)
+    }
+  }
+}
+
+# library(tidyverse)
+# open_dataset("tmzMLs/pqds") %>%
+#   filter(amendment=="Amm") %>%
+#   filter(polarity=="neg") %>%
+#   filter(samp_type=="Smp") %>%
+#   filter(mz%between%pmppm(150.041585, 5)) %>%
+#   filter(rt%between%c(7.5, 9)) %>%
+#   collect() %>%
+#   left_join(parametadata) %>%
+#   ggplot() +
+#   geom_line(aes(x=rt, y=int, group=filename, color=samp_type))

### 2023-09-26 eicscript.R

@@ -1,0 +1,66 @@
+library(tidyverse)
+library(arrow)
+library(RaMS)
+
+IS_type <- "noIS"
+
+all_stans <- read_csv("metadata/all_stans.csv") %>%
+  mutate(polarity=factor(polarity, levels=c("pos", "neg"))) %>%
+  arrange(polarity) %>%
+  filter(compound_type!="Internal Standard")
+
+for(polarity in c("pos", "neg")){
+  for(amendment in c("Amm", "Nit", "GMP", "Arg")){
+    parametadata <- read_csv("metadata/parametadata.csv") %>% 
+      filter(polarity==!!polarity) %>%
+      filter(amendment==!!amendment) %>%
+      filter(IS_type==!!IS_type)
+    output_folder <- paste(amendment, polarity, IS_type, "output", sep = "_") %>%
+      paste0("untargeted/", ., "/")
+    best_feat_data <- read_csv(paste0(output_folder, "peak_data.csv")) %>%
+      left_join(parametadata) %>%
+      filter(samp_type=="Smp") %>%
+      group_by(feature) %>%
+      summarize(mzmed=unique(feat_mzmed), rtmed=unique(feat_rtmed)/60, n=n(),
+                med_cor=median(beta_cor, na.rm=TRUE), 
+                med_snr=median(beta_snr, na.rm=TRUE)) %>%
+      filter(med_cor>0.75) %>%
+      select(-med_cor, -med_snr)
+    
+    pdf(paste0(output_folder, "anno_eics.pdf"), width = 8, height = 4)
+    all_stans %>%
+      filter(polarity==!!polarity) %>%
+      mutate(mz=ifelse(compound_name=="Carnitine", mz[which(compound_name=="Threonine betaine?")], mz)) %>%
+      group_by(mz) %>%
+      summarise(cmpd_lab=paste0(compound_name, " (", mix, " ,", rt, ")", collapse = ", ")) %>%
+      pwalk(function(...){
+        row_data <- list(...)
+        mass_feats <- best_feat_data %>% filter(mzmed%between%pmppm(row_data$mz, 5))
+        chosen_eic <- open_dataset("tmzMLs/pqds") %>%
+          filter(polarity==!!polarity) %>%
+          filter(amendment==!!amendment) %>%
+          filter(samp_type=="Poo" | samp_type=="Std") %>%
+          filter(mz%between%pmppm(row_data$mz, 5)) %>%
+          filter(rt%between%c(2, 20)) %>%
+          dplyr::collect()
[…]

### 2023-09-26 untargeted/paruntargeted_control.Rmd
Finished annotating Amm pos with rescue feats

--- 2023-09-26 untargeted/paruntargeted_control.Rmd
@@
   if(polarity=="pos"){
-    rescue_feats <- c("FT3015", "FT1319", "FT1321")
+    rescue_feats <- c("FT3015", Allopurinol="FT1319", Hypoxanthine="FT1321",
+                      `DMSOP`="FT1688", "Guan from guan"="FT1707", Tyrosine="FT2200",
+                      Phosphocholine="FT2228", Trimethyllysine="FT2278",
+                      Dexpanthenol="FT2426", Cytidine="FT2792",
+                      Mycogly='FT2814', Inosine="FT2960", Guanosine="FT3015")
   } else if(polarity=="neg"){
@@
 all_stans <- read_csv("metadata/all_stans.csv") %>%
-  filter(polarity==!!polarity)
+  mutate(polarity=factor(polarity, levels=c("pos", "neg"))) %>%
+  arrange(polarity) %>%
+  filter(compound_type!="Internal Standard")
+# map(c("Amm", "Nit", "GMP", "Arg"), function(amendment_i){
+#   all_stans %>%
+#       mutate(amendment=amendment_i) %>%
+#       select(amendment, polarity, compound_name, mz, rt, date_added, mix) %>%
+#       write_csv("untargeted/anno_template.csv", append = TRUE)
+# })
+# # Gets copy-pasted into the Excel file manual_annotations.xlsx
+# # Then run eicscript.R to render the EICs for each standard and fill it in
+# # Afterward, use the "rescue_feats" object above to add those that didn't
+# # pass QC for shape or SNR back in
 
-pdf(paste0(output_folder, "anno_eics.pdf"), width = 8, height = 4)
-all_stans %>%
-  filter(compound_type!="Internal Standard") %>%
-  mutate(mz=ifelse(compound_name=="Carnitine", mz[which(compound_name=="Threonine betaine?")], mz)) %>%
-  group_by(mz) %>%
-  summarise(cmpd_lab=paste0(compound_name, " (", mix, ",", rt, ")", collapse = ", ")) %>%
-  # head(6) %>%
-  pwalk(function(...){
-    row_data <- list(...)
-    mass_feats <- best_feat_data %>% filter(mzmed%between%pmppm(row_data$mz, 5))
-    chosen_eic <- open_dataset("tmzMLs/pqds") %>%
-      filter(polarity==!!polarity) %>%
-      filter(amendment==!!amendment) %>%
-      filter(samp_type=="Poo" | samp_type=="Std") %>%
-      filter(mz%between%pmppm(row_data$mz, 5)) %>%
-      filter(rt%between%c(2, 20)) %>%
-      dplyr::collect()
-    gp_plot <- chosen_eic %>%
-      left_join(parametadata, by = join_by(filename, samp_type, amendment, polarity)) %>%
-      mutate(samp_type=str_extract(filename, "Mix\\d|Deep|Surf|Poo|Blk")) %>%
-      filter(!str_detect(filename, "Mat")) %>%
-      mutate(plot_facet=ifelse(samp_type%in%c("Mix1", "Mix2"), "Mixes", "Samples")) %>%
-      ggplot() +
-      geom_line(aes(x=rt, y=int, group=filename, color=samp_type)) +
-      facet_wrap(~plot_facet, scales = "free_y", ncol=1) +
-      geom_vline(aes(xintercept=rtmed), color="black", data = mass_feats) +
-      geom_text(aes(x=rtmed, y=Inf, label=feature), 
-                angle=90, hjust=1, vjust=0.5, data = mass_feats) +
-      ggtitle(row_data$cmpd_lab) +
-      scale_x_continuous(breaks = 2:20) +
-      theme_bw() +
-      theme(legend.position = c(0.9, 0.8), legend.title = element_blank(),
-            axis.title = element_blank())
-    print(gp_plot)
-  }, .progress = TRUE)
-dev.off()
+manual_anno <- readxl::read_excel("untargeted/manual_annotations.xlsx")
+
+manual_anno %>% filter(str_detect(notes, "Unpicked"))
 ```

### 2023-09-27 untargeted/paruntargeted_control.Rmd
Finished Amm neg mode annotations (not many good ones, oof)

@@
   } else if(polarity=="neg"){
-
+    rescue_feats <- c()
   }
@@
 
-```{r diagnostic plots}
+```{r mass_check plots, eval=FALSE}
 mass_check <- function(mz_i, all_samps=FALSE){
@@
 manual_anno %>% filter(str_detect(notes, "Unpicked"))
+
 ```

### 2023-09-27 untargeted/paruntargeted_control.Rmd
Annotated Nit pos, needs rescuing still

@@
 # polarity <- "neg"
-amendment <- "Amm"
-# amendment <- "Nit"
+# amendment <- "Amm"
+amendment <- "Nit"
 # amendment <- "GMP"

### 2023-09-27 untargeted/paruntargeted_control.Rmd
Added rudimentary isotope finder

@@
 manual_anno <- readxl::read_excel("untargeted/manual_annotations.xlsx")
+manual_anno %>% filter(str_detect(notes, "Unpicked"))
+```
 
-manual_anno %>% filter(str_detect(notes, "Unpicked"))
+```{r check anno}
+all_stans <- read_csv("metadata/all_stans.csv") %>%
+  filter(polarity==!!polarity) %>%
+  filter(compound_type!="Internal Standard")
+manual_anno <- readxl::read_excel("untargeted/manual_annotations.xlsx") %>%
+  filter(polarity==!!polarity) %>%
+  filter(amendment==!!amendment)
+best_feat_data <- output_folder %>%
+  paste0("best_feat_data.csv") %>%
+  read_csv()
+mz_rt_check_df <- best_feat_data %>%
+  left_join(manual_anno) %>%
+  select(feature, compound_name, mzmed, rtmed) %>%
+  filter(!is.na(compound_name)) %>%
+  left_join(all_stans) %>%
+  select(feature, compound_name, mzmed, rtmed, mz, rt)
+mz_rt_check_df %>%
+  ggplot(aes(label=compound_name)) +
+  geom_point(aes(x=rtmed, y=rt)) +
+  geom_abline(slope = 1) +
+  coord_fixed()
+mz_rt_check_df %>%
+  ggplot(aes(label=compound_name)) +
+  geom_point(aes(x=mzmed, y=mz-mzmed))
+```
 
+```{r find isotopes (stans only)}
+all_stans <- read_csv("metadata/all_stans.csv") %>%
+  filter(polarity==!!polarity) %>%
+  filter(compound_type!="Internal Standard") %>%
+  select(compound_name, formula)
+manual_anno <- readxl::read_excel("untargeted/manual_annotations.xlsx") %>%
+  filter(polarity==!!polarity) %>%
+  filter(amendment==!!amendment) %>%
+  select(feature, compound_name)
+best_feat_data <- output_folder %>%
+  paste0("best_feat_data.csv") %>%
+  read_csv() %>%
+  left_join(manual_anno) %>%
+  left_join(all_stans)
+peak_rt_bounds <- output_folder %>%
+  paste0("peak_data.csv") %>%
+  read_csv() %>%
+  select(feature, filename, rtmin, rtmax) %>%
+  distinct()
+iso_mzs <- best_feat_data %>%
+  filter(!is.na(formula)) %>%
+  mutate(n_count=str_extract(formula, "N\\d+")) %>%
+  mutate(n_count=as.numeric(str_remove(n_count, "N"))) %>%
+  mutate(n_count=ifelse(str_detect(formula, "N") & is.na(n_count), 1, n_count)) %>%
+  filter(!is.na(n_count)) %>%
+  slice(rep(seq(n()), n_count+1)) %>%
+  group_by(feature) %>%
+  mutate(n_count=1:n()) %>%
+  mutate(labeled_mzmed=mzmed+0.997025*(n_count-1)) %>%
+  select(compound_name, feature, labeled_mzmed, n_count) %>%
+  left_join(peak_rt_bounds, relationship = "many-to-many") %>%
+  left_join(parametadata) %>%
+  filter(samp_type=="Smp") %>%
+  select(compound_name, feature, n_count, labeled_mzmed, filename, rtmin, rtmax)
+
+iso_areas <- iso_mzs %>%
+  split(.$filename) %>% 
+  imap(function(file_peaks, file_name_i){
+    msdata <- open_dataset("tmzMLs/pqds") %>%
+      filter(amendment=="Amm") %>%
+      filter(polarity=="pos") %>%
+      filter(filename==file_name_i) %>%
+      dplyr::collect()
+    file_peaks %>%
+      pmap(function(...){
+        row_data <- data.frame(...)
+        area <- msdata[mz%between%pmppm(row_data$labeled_mzmed, 100)] %>%
+          .[rt%between%c(row_data$rtmin/60, row_data$rtmax/60)] %>%
+          .[,sum(int)]
+        cbind(row_data, area=area)
+      }) %>%
+      bind_rows()
+  }, .progress=TRUE) %>%
+  bind_rows() %>%
+  select(compound_name, filename, area, n_count)
+
+output_csv_write(iso_areas)
+
+iso_areas %>%
+  filter(compound_name=="L-Arginine") %>%
+  left_join(parametadata) %>%
+  filter(!is.na(depth)) %>%
+  mutate(timepoint=factor(timepoint, levels=paste0("T", c(0,1,3,10,26)))) %>%
+  ggplot() +
+  geom_boxplot(aes(x=factor(n_count), y=area, color=timepoint)) +
+  facet_grid(depth~startime, scales="free_y")
 ```

### 2023-09-27 untargeted/paruntargeted_control.Rmd
Finished rescuing Nit neg after some confusion

@@
 
-polarity <- "pos"
-# polarity <- "neg"
+# polarity <- "pos"
+polarity <- "neg"
 # amendment <- "Amm"
@@
   left_join(peak_data_long, by = join_by(peakidx)) %>%
-  mutate(filename=basename(fileNames(msnexp_filled))[sample])
+  mutate(filename=basename(fileNames(msnexp_filled))[sample]) %>%
+  select(feature, feat_mzmed, feat_rtmed, feat_npeaks, filename,
+         mz, mzmin, mzmax, rt, rtmin, rtmax, into, sn, beta_cor, beta_snr)
 output_csv_write(peak_data)
@@
             med_snr=median(beta_snr, na.rm=TRUE))
+output_csv_write(all_feat_data)
 # all_feat_data %>%
@@
 
+# Rescue feats
 if(amendment=="Amm"){
@@
   if(polarity=="pos"){
-
+    rescue_feats <- c()
   } else if(polarity=="neg"){
-
+    rescue_feats <- c(Bisulfate="FT0706", "Itaconic acid"="FT1544", 
+                      "Salicylic"="FT1841")
   } else {
@@
 
-```{r mass_check plots, eval=FALSE}
+```{r rescue feats rendering}
+anno_manu <- readxl::read_excel("untargeted/manual_annotations.xlsx") %>%
+  filter(polarity==!!polarity) %>%
+  filter(amendment==!!amendment)
+all_feat_data <- read_csv(paste0(output_folder, "all_feat_data.csv"))
+anno_manu %>%
+  filter(str_detect(notes, "[Rr]escue"))
+all_feat_data %>%
+  filter(mzmed%between%pmppm(137.0239, 10))
+```
+
+```{r plots from mass check and others, eval=FALSE}
+best_feat_data %>%
+  ggplot() +
+  geom_point(aes(x=rtmed, y=mzmed, color=log10(n))) +
+  scale_color_viridis_c()
+
 mass_check <- function(mz_i, all_samps=FALSE){
@@
     geom_vline(aes(xintercept=rtmed), color="black", data = mass_feats) +
+    scale_x_continuous(breaks = 2:20) +
     ggtitle(ggtitle_i) +
@@
 }
+
+best_feat_data %>% slice_sample(n=1) %>% pull(mzmed) %>% mass_check()
+
 mass_check(c("Carn/THB/HMSB"=162.1125))

### 2023-09-28 untargeted/paruntargeted_control.Rmd
Finished rescuing Nit pos

@@
 
-# polarity <- "pos"
-polarity <- "neg"
+polarity <- "pos"
+# polarity <- "neg"
 # amendment <- "Amm"
@@
   if(polarity=="pos"){
-    rescue_feats <- c()
+    rescue_feats <- c(Creatinine="FT1339", `Homocysteine thiolactone`="FT1444",
+                      Leucine="FT1956", Ornithine="FT1980", Lysine="FT2404",
+                      THB="FT2706", Lysinebetaineish='FT3317', glucogly="FT4428")
+
   } else if(polarity=="neg"){
@@
 all_feat_data %>%
-  filter(mzmed%between%pmppm(137.0239, 10))
+  filter(mzmed%between%pmppm(132.1025, 10))
 ```

### 2023-09-30 metadata/parametadata_control.Rmd
Dropped .mzML from shortname

@@
   mutate(shortname=str_remove(shortname, ".*_Smp_")) %>%
+  mutate(shortname=str_remove(shortname, ".mzML$")) %>%
   mutate(samp_type="Smp") %>%
@@
   mutate(timepoint=str_extract(shortname, "T\\d+")) %>%
-  mutate(tripl=str_extract(shortname, "[A-C](?=\\.mzML$)"))
+  mutate(tripl=str_extract(shortname, "[A-C]$"))
 

### 2023-09-30 untargeted/paruntargeted_control.Rmd
Adding factors to parametadata on first read in

@@
   filter(amendment==!!amendment) %>%
-  filter(IS_type==!!IS_type)
+  filter(IS_type==!!IS_type) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
+  mutate(timepoint=factor(timepoint, levels=paste0("T", c(0,1,3,10,26,73))))
 

### 2023-09-30 untargeted/paruntargeted_control.Rmd
Adding headers, isotope exploration, and multivariate analysis code

@@
   mutate(samp_group=paste(amendment, depth, startime, timepoint, sep="-"))
+
+mz_group <- function(mz_vals, ppm){
+  group_vec <- numeric(length(mz_vals))
+  group_num <- 1L
+  init_vec <- mz_vals
+  names(init_vec) <- seq_along(init_vec)
+  while(length(init_vec)>0){
+    mz_i <- init_vec[1]
+    err <- mz_i*ppm/1000000
+    mz_idxs <- init_vec>mz_i-err & init_vec<mz_i+err
+    group_vec[as.numeric(names(mz_idxs)[mz_idxs])] <- group_num
+    init_vec <- init_vec[!mz_idxs]
+    group_num <- group_num+1L
+  }
+  group_vec
+}
 ```
+
+# XCMS things
 
@@
 
-```{r rescue feats rendering}
+# Annotation
+
+```{r rescue feats rendering, eval=FALSE}
 anno_manu <- readxl::read_excel("untargeted/manual_annotations.xlsx") %>%
@@
 manual_anno <- readxl::read_excel("untargeted/manual_annotations.xlsx")
-manual_anno %>% filter(str_detect(notes, "Unpicked"))
+# manual_anno %>% filter(str_detect(notes, "Unpicked"))
 ```
@@
 ```
+
+# Isotope exploration
 
@@
     msdata <- open_dataset("tmzMLs/pqds") %>%
-      filter(amendment=="Amm") %>%
-      filter(polarity=="pos") %>%
+      filter(amendment==!!amendment) %>%
+      filter(polarity==!!polarity) %>%
       filter(filename==file_name_i) %>%
@@
         row_data <- data.frame(...)
-        area <- msdata[mz%between%pmppm(row_data$labeled_mzmed, 100)] %>%
+        area <- msdata[mz%between%pmppm(row_data$labeled_mzmed, 10)] %>%
           .[rt%between%c(row_data$rtmin/60, row_data$rtmax/60)] %>%
@@
   bind_rows() %>%
-  select(compound_name, filename, area, n_count)
+  mutate(labeled_n_count=n_count-1) %>%
+  select(compound_name, filename, area, labeled_n_count)
 
 output_csv_write(iso_areas)
+```
 
-iso_areas %>%
-  filter(compound_name=="L-Arginine") %>%
+```{r individual isotope EIC}
+single_cmpd_iso_req <- iso_mzs %>%
+  filter(compound_name=="Taurine") %>%
+  distinct(compound_name, n_count, labeled_mzmed) %>%
+  mutate(labeled_mzmed=round(labeled_mzmed, 5)) %>%
+  summarise(req_cmd=paste0("mz%between%pmppm(", labeled_mzmed, ", 10)", collapse="|")) %>%
+  pull(req_cmd)
+full_arrow_req <- paste(
+  'open_dataset("tmzMLs/pqds") %>%',
+  'filter(amendment==!!amendment) %>%',
+  'filter(samp_type=="Smp") %>%',
+  'filter(polarity==!!polarity) %>% ',
+  'filter(', single_cmpd_iso_req, ') %>% ',
+  'dplyr::collect()'
+)
+single_cmpd_iso_eic <- eval(parse(text=full_arrow_req))
+
+iso_mzs %>%
+  distinct(compound_name, feature, n_count, labeled_mzmed) %>%
+  filter(compound_name=="Taurine") %>%
+  pmap(function(...){
+    row_data <- list(...)
+    single_cmpd_iso_eic %>%
+      filter(mz%between%pmppm(row_data$labeled_mzmed, 10)) %>%
+      mutate(n_count=row_data$n_count)
+  }) %>%
+  bind_rows() %>%
+  # filter(rt%between%c(15, 17)) %>%
+  left_join(parametadata) %>%
+  filter(depth=="Surf" & startime=="Morn") %>%
+  ggplot() +
+  geom_line(aes(x=rt, y=int, group=filename, color=timepoint)) +
+  facet_grid(timepoint~n_count)
+```
+
+```{r individual isotope area plots}
+iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv"))
+
+cmpd_name_i <- "Taurine"
+single_cmpd_iso_data <- iso_areas %>%
+  filter(compound_name==cmpd_name_i) %>%
   left_join(parametadata) %>%
   filter(!is.na(depth)) %>%
-  mutate(timepoint=factor(timepoint, levels=paste0("T", c(0,1,3,10,26)))) %>%
+  mutate(labeled_n_count=factor(labeled_n_count))
+
+single_cmpd_iso_data %>%
   ggplot() +
-  geom_boxplot(aes(x=factor(n_count), y=area, color=timepoint)) +
-  facet_grid(depth~startime, scales="free_y")
+  geom_boxplot(aes(x=labeled_n_count, y=area, color=timepoint)) +
+  facet_grid(startime~depth)
+
+single_cmpd_iso_data %>%
+  group_by(depth, startime, labeled_n_count) %>%
+  mutate(area=area/sum(area)) %>%
+  ggplot() +
+  geom_boxplot(aes(x=labeled_n_count, y=area, color=timepoint)) +
+  facet_grid(startime~depth)
+
+single_cmpd_iso_data %>%
+  group_by(depth, startime, labeled_n_count, timepoint) %>%
+  summarize(area=median(area)) %>%
+  group_by(depth, startime, timepoint) %>%
+  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
+  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
+  ggplot() +
+  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
+                  fill=labeled_n_count, group=labeled_n_count)) +
+  facet_grid(startime~depth) +
+  scale_fill_brewer(type = "qual", palette = 3) +
+  theme_bw()
 ```
+
+```{r isotopes for all targ cmpds}
+pdf(paste0(output_folder, "targ_cmpd_mids.pdf"), width = 6, height = 3)
+walk(unique(iso_areas$compound_name), function(cmpd_name_i){
+  gp <- iso_areas %>%
+    filter(compound_name==cmpd_name_i) %>%
+    left_join(parametadata, by="filename") %>%
+    filter(!is.na(depth)) %>%
+    mutate(labeled_n_count=factor(labeled_n_count)) %>%
+    group_by(depth, startime, labeled_n_count, timepoint) %>%
+    summarize(area=median(area), .groups = "drop") %>%
+    group_by(depth, startime, timepoint) %>%
+    mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
+    mutate(upper_bound=lead(lower_bound, default = 1)) %>%
+    ggplot() +
+    geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
+                    fill=labeled_n_count, group=labeled_n_count)) +
+    facet_grid(startime~depth) +
+    scale_fill_brewer(type = "qual", palette = 3) +
+    labs(x="Timepoint", fill="15N count", y="Fraction of peak area in\neach isotopologue") +
+    theme_bw() +
+    ggtitle(cmpd_name_i)
+  plot(gp)
+}, .progress = TRUE)
+dev.off()
+```
+
+# Multivariate approaches
+
+```{r overall analysis}
+iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv"))
+areas_scaled <- iso_areas %>%
+  filter(labeled_n_count>0) %>%
+  left_join(parametadata) %>%
+  filter(!is.na(depth)) %>%
+  mutate(compound_iso=paste0(compound_name, " 15N", labeled_n_count)) %>%
+  arrange(desc(area)) %>%
+  select(shortname, compound_iso, area) %>%
+  group_by(shortname, compound_iso) %>%
+  slice(1) %>%
+  ungroup() %>%
+  complete(shortname, compound_iso, fill = list(area=0)) %>%
+  group_by(compound_iso) %>%
+  # mutate(norm_area=(area-min(area))/(max(area)-min(area))) %>%
+  # mutate(norm_area=rank(area)) %>%
+  mutate(norm_area=scale(area)[,1]) %>%
+  ungroup() %>%
+  select(shortname, compound_iso, norm_area)
+
+areas_scaled %>%
+  pivot_wider(names_from = shortname, values_from = norm_area) %>%
+  column_to_rownames("compound_iso") %>%
+  data.matrix() %>%
+  t() %>%
+  prcomp() %>%
+  pluck("x") %>%
+  as.data.frame() %>%
+  rownames_to_column("shortname") %>%
+  select(shortname, PC1:PC3) %>%
+  left_join(parametadata) %>%
+  arrange(timepoint) %>%
+  mutate(shortname=fct_inorder(shortname)) %>%
+  pivot_longer(starts_with("PC")) %>%
+  filter(name%in%paste0("PC", 1:3)) %>%
+  ggplot() +
+  geom_col(aes(x=shortname, y=value, fill=timepoint)) +
+  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5)) +
+  facet_wrap(~name, ncol=1)
+
+areas_scaled %>%
+  left_join(parametadata, by="shortname") %>%
+  select(compound_iso, shortname, norm_area) %>%
+  group_by(compound_iso) %>%
+  filter(sum(norm_area==0)<n()*0.9) %>%
+  pivot_wider(names_from = "compound_iso", values_from = norm_area) %>%
+  column_to_rownames("shortname") %>%
+  data.matrix()
+
+```
+
+```{r multivar analysis by depth}
+multivar_analysis <- areas_scaled %>%
+  left_join(parametadata, by="shortname") %>%
+  nest(data=-depth) %>%
+  mutate(wide_peaks=map(data, function(metab_df){
+    metab_df %>%
+      select(compound_iso, shortname, norm_area) %>%
+      group_by(compound_iso) %>%
+      filter(sum(norm_area==0)<n()*0.9) %>%
+      pivot_wider(names_from = "compound_iso", values_from = norm_area) %>%
+      column_to_rownames("shortname") %>%
+      data.matrix()
+  })) %>%
+  mutate(nmds_output=map(wide_peaks, function(metab_mat){
+    # env_vec <- pull(parametadata, timepoint, shortname)[rownames(metab_mat)]
+    vegan::metaMDS(metab_mat, k = 2, distance = "manhattan", trace = 0, 
+                   autotransform = FALSE, noshare = FALSE, wascores = FALSE)
+    # vegan::MDSrotate(env_vec)
+  })) %>%
+  mutate(nmds_stress=map_dbl(nmds_output, pluck, "stress")) %>%
+  mutate(perm_output=map(wide_peaks, function(metab_mat){
+    adon_expl <- data.frame(shortname=rownames(metab_mat)) %>%
+      left_join(parametadata %>% select(shortname, timepoint, startime), by="shortname")
+    vegan::adonis2(
+      metab_mat~timepoint+startime, data = adon_expl, by = "margin", 
+      permutations = 999, method = "manhattan")
+  })) %>%
+  mutate(pca_output=map(wide_peaks, prcomp)) %>%
+  mutate(hclust_obj=map(wide_peaks, ~hclust(dist(.x))))
+
+multivar_analysis %>%
+  mutate(nmds_points=map(nmds_output, function(nmds_output_i){
+    nmds_output_i %>% pluck("points") %>% as.data.frame()
+  })) %>%
+  select(depth, nmds_points) %>%
+  unnest_longer(nmds_points, indices_to = "shortname") %>%
+  unnest_wider(nmds_points) %>%
+  left_join(parametadata %>% distinct(shortname, shortname, timepoint, startime)) %>%
+  # mutate(MDS2=ifelse(str_detect(depth,"Deep"), MDS2*-1, MDS2)) %>%
+  ggplot(aes(label=shortname)) +
+  geom_vline(xintercept = 0) +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(x=MDS1, y=MDS2, fill=timepoint), color="black", size=4, pch=21) +
+  facet_wrap(~depth, ncol=1) +
+  theme_bw() +
+  coord_fixed()
+
+multivar_analysis$perm_output
+```

### 2023-09-30 untargeted/paruntargeted_control.Rmd
Adding full dataset PERMANOVA exploration

@@
 # polarity <- "neg"
-# amendment <- "Amm"
-amendment <- "Nit"
+amendment <- "Amm"
+# amendment <- "Nit"
 # amendment <- "GMP"
@@
 areas_scaled <- iso_areas %>%
-  filter(labeled_n_count>0) %>%
+  # filter(labeled_n_count>0) %>%
   left_join(parametadata) %>%
@@
 
-areas_scaled %>%
+areas_mat <- areas_scaled %>%
   left_join(parametadata, by="shortname") %>%
@@
 
+adon_expl <- data.frame(shortname=rownames(areas_mat)) %>%
+  left_join(parametadata %>% select(shortname, depth, timepoint, startime), by="shortname")
+vegan::adonis2(
+  areas_mat~timepoint+startime+depth, data = adon_expl, by = "margin", 
+  permutations = 999, method = "manhattan")
 ```

### 2023-10-03 mzMLs/msconvert.R
Removing nit rerun skip code

@@ -8,9 +8,6 @@ folder_paths <- c(
 
 sapply(folder_paths, function(path_i){
   sapply(c("positive", "negative"), function(polarity){
-    if(path_i=="Z:/1_QEdata/2021/2021_Will_PARAGON/210813_Will_NIT-Expt_PARAGON_HILIC" & polarity=="positive"){
-      return(0)
-    }
     message(paste("Converting", basename(path_i), polarity))
     short_pol <- gsub("[ia]tive", "", polarity)
     match_pattern <- paste0("wIS.raw|noIS-", short_pol, ".*.raw|_Std_.*raw")

### 2023-10-04 metadata/parametadata_control.Rmd
Removing mzML ending from shortname for other_filedata

@@
   mutate(shortname=str_remove(shortname, ".*_Poo_|.*_Std_(4uMStds)?")) %>%
+  mutate(shortname=str_remove(shortname, ".mzML")) %>%
   mutate(samp_type=str_extract(filename, "Poo|Std")) %>%

### 2023-10-04 mzMLs/msconvert.R
Now converting over wIS pooled samps

@@ -10,7 +10,7 @@ sapply(folder_paths, function(path_i){
   sapply(c("positive", "negative"), function(polarity){
     message(paste("Converting", basename(path_i), polarity))
     short_pol <- gsub("[ia]tive", "", polarity)
-    match_pattern <- paste0("wIS.raw|noIS-", short_pol, ".*.raw|_Std_.*raw")
+    match_pattern <- paste0("wIS.raw|noIS-", short_pol, ".*.raw|_Std_.*raw|wIS-(Full|Half)")
     files_to_convert <- list.files(path_i, pattern = match_pattern, full.names = TRUE)
     mscommand <- paste(
       "msconvert",

### 2023-10-04 mzMLs/msconvert.R
Now should be run from PARAGON project dir instead of mzML folder

@@ -16,7 +16,7 @@ sapply(folder_paths, function(path_i){
       "msconvert",
       paste(files_to_convert, collapse = " "),
       "--mzML",
-      paste0('-o ', short_pol),
+      paste0('-o mzMLs/', short_pol),
       '--filter "peakPicking true 1-"',
       paste('--filter "polarity', polarity, '"')
     )
@@ -28,7 +28,7 @@ sapply(folder_paths, function(path_i){
       "msconvert",
       paste(msms_to_convert, collapse = " "),
       "--mzML",
-      paste0('-o ', short_pol, "/MSMS"),
+      paste0('-o mzMLs/', short_pol, "/MSMS"),
       '--filter "peakPicking true 1-"',
       paste('--filter "polarity', polarity, '"')
     )

### 2023-10-04 metadata/parametadata_control.Rmd
Adding pooled samp wIS to metadata

@@
   distinct()
-parametadata <- all_filedata %>%
+parametadata_all <- all_filedata %>%
   left_join(mzml_timestamps, by="filename")
@@
 ```{r write out}
-write_csv(parametadata, file = "metadata/parametadata.csv")
+write_csv(parametadata_all, file = "metadata/parametadata.csv")
 ```

### 2023-10-05 untargeted/paruntargeted_control.Rmd
Finished rescuing Arginine targets

@@
   if(polarity=="pos"){
-
+    rescue_feats <- c()
   } else if(polarity=="neg"){
-
+    rescue_feats <- c()
   } else {
@@
   if(polarity=="pos"){
-
+    rescue_feats <- c("FT0945", "FT1138", "FT1346", "FT1413", "FT1476",
+                      "FT1505", "FT1533", "FT1935", "FT1945", "FT2029",
+                      "FT2425", "FT2774", "FT2999", "FT3668")
   } else if(polarity=="neg"){
-
+    rescue_feats <- c("FT0790", "FT1022", "FT1031", "FT1387")
   } else {
@@
 anno_manu <- readxl::read_excel("untargeted/manual_annotations.xlsx") %>%
-  filter(polarity==!!polarity) %>%
   filter(amendment==!!amendment)
@@
 all_feat_data %>%
-  filter(mzmed%between%pmppm(132.1025, 10))
+  filter(mzmed%between%pmppm(138.9701 , 10))
 ```

### 2023-10-05 untargeted/paruntargeted_control.Rmd
Adding a bunch of isotope visualizations

@@
 iso_mzs %>%
-  distinct(compound_name, feature, n_count, labeled_mzmed) %>%
-  filter(compound_name=="Taurine") %>%
+  filter(compound_name=="Guanine") %>%
   pmap(function(...){
@@
   bind_rows() %>%
-  # filter(rt%between%c(15, 17)) %>%
+  filter(rt%between%c(7, 9)) %>%
   left_join(parametadata) %>%
@@
   geom_line(aes(x=rt, y=int, group=filename, color=timepoint)) +
-  facet_grid(timepoint~n_count)
+  facet_grid(timepoint~n_count) +
+  ggtitle("Surface morning spiked samples")
 ```
@@
 
-cmpd_name_i <- "Taurine"
+cmpd_name_i <- "Guanine"
 single_cmpd_iso_data <- iso_areas %>%
@@
   geom_boxplot(aes(x=labeled_n_count, y=area, color=timepoint)) +
-  facet_grid(startime~depth)
+  facet_grid(depth~startime, scales="free_y") +
+  ggtitle("Raw peak areas across depth and time")
 
@@
   geom_boxplot(aes(x=labeled_n_count, y=area, color=timepoint)) +
-  facet_grid(startime~depth)
+  facet_grid(depth~startime) +
+  ggtitle("Peak areas normalized to the max within each isotope")
 
@@
                   fill=labeled_n_count, group=labeled_n_count)) +
-  facet_grid(startime~depth) +
+  facet_grid(depth~startime) +
   scale_fill_brewer(type = "qual", palette = 3) +
-  theme_bw()
+  theme_bw() +
+  labs(title = "Median triplicate peak area as a fraction of total for each timepoint",
+       x="Timepoint", fill="15N count", 
+       y="Fraction of peak area in each isotopologue")
+
+single_cmpd_iso_data %>%
+  # group_by(depth, startime, labeled_n_count, timepoint) %>%
+  # summarize(area=median(area)) %>%
+  ggplot() +
+  geom_col(aes(x=timepoint, y=area, fill=labeled_n_count), color="grey50") +
+  facet_grid(depth~startime, scales="free_y") +
+  scale_fill_brewer(type = "qual", palette = 3) +
+  theme_bw() +
+  labs(x="Timepoint (hours after spike)", y="Peak area with 15N label",
+       fill="Number of\n15N atoms in\nisotopologue",
+       title = "Summed peak area by timepoint")
 ```
@@
 
+```{r special classes of isotope trends}
+iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv"))
+nucleos <- c("Adenine", "Guanine", "Cytosine", "Hypoxanthine", 
+             "Adenosine", "Guanosine", "Cytidine", "Inosine",
+             "Deoxyadenosine", "Deoxyguanosine?", "Deoxycytidine?")
+iso_areas %>%
+  filter(compound_name%in%nucleos) %>%
+  mutate(compound_name=factor(compound_name, levels=nucleos)) %>%
+  left_join(parametadata) %>%
+  filter(!is.na(depth)) %>%
+  mutate(labeled_n_count=factor(labeled_n_count)) %>%
+  filter(depth=="Surf") %>%
+  # filter(compound_name=="Guanine") %>%
+  group_by(compound_name, labeled_n_count, timepoint) %>%
+  summarize(area=median(area)) %>%
+  group_by(compound_name, timepoint) %>%
+  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
+  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
+  ggplot() +
+  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
+                  fill=labeled_n_count, group=labeled_n_count)) +
+  facet_wrap(~compound_name, nrow=3) +
+  scale_fill_brewer(type = "qual", palette = 3) +
+  theme_bw() +
+  labs(x="Timepoint (hours after spike)", 
+       y="Fraction of peak area with 15N label",
+       fill="Number of\n15N atoms in\nisotopologue")
+
+nbases <- c("Adenine", "Guanine", "Cytosine", "Hypoxanthine")
+iso_areas %>%
+  filter(compound_name%in%nbases) %>%
+  mutate(compound_name=factor(compound_name, levels=nbases)) %>%
+  left_join(parametadata) %>%
+  filter(!is.na(depth)) %>%
+  filter(labeled_n_count>0) %>%
+  mutate(labeled_n_count=factor(labeled_n_count)) %>%
+  filter(timepoint=="T26") %>%
+  ggplot() +
+  geom_boxplot(aes(x=labeled_n_count, y=area)) +
+  ggh4x::facet_grid2(depth~compound_name, scales="free_y", independent = "y")
+
+aminos <- c("Glycine", "L-Alanine", "L-Serine", "L-Proline", 
+            "L-Threonine", "L-Leucine", "L-Isoleucine", "L-Tyrosine", 
+            "L-Aspartic acid", "L-Asparagine", "beta-Glutamic acid", 
+            "L-Glutamine", "L-Methionine", "L-Histidine", "L-Arginine")
+iso_areas %>%
+  filter(compound_name%in%aminos) %>%
+  mutate(compound_name=factor(compound_name, levels=aminos)) %>%
+  left_join(parametadata) %>%
+  filter(!is.na(depth)) %>%
+  mutate(labeled_n_count=factor(labeled_n_count)) %>%
+  filter(depth=="Surf") %>%
+  group_by(compound_name, labeled_n_count, timepoint) %>%
+  summarize(area=median(area)) %>%
+  group_by(compound_name, timepoint) %>%
+  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
+  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
+  ggplot() +
+  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
+                  fill=labeled_n_count, group=labeled_n_count)) +
+  facet_wrap(~compound_name, ncol=5) +
+  scale_fill_brewer(type = "qual", palette = 3) +
+  theme_bw() +
+  labs(x="Timepoint (hours after spike)", 
+       y="Fraction of peak area with 15N label",
+       fill="Number of\n15N atoms in\nisotopologue")
+
+MAAs <- c("Palythine?", "Mycosporine-glycine?", "Asterina-330?", 
+          "Shinorine?", "Porphyra-334?", "Mycosporine-2-glycine?")
+iso_areas %>%
+  filter(compound_name%in%MAAs) %>%
+  mutate(compound_name=factor(compound_name, levels=MAAs)) %>%
+  left_join(parametadata) %>%
+  filter(!is.na(depth)) %>%
+  mutate(labeled_n_count=factor(labeled_n_count)) %>%
+  filter(depth=="Surf") %>%
+  group_by(compound_name, startime, labeled_n_count, timepoint) %>%
+  summarize(area=median(area)) %>%
+  group_by(compound_name, startime, timepoint) %>%
+  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
+  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
+  ggplot() +
+  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
+                  fill=labeled_n_count, group=labeled_n_count)) +
+  facet_grid(startime~compound_name) +
+  scale_fill_brewer(type = "qual", palette = 3) +
+  coord_cartesian(ylim = c(0.5, 1)) +
+  theme_bw() +
+  labs(x="Timepoint (hours after spike)", 
+       y="Fraction of peak area with 15N label",
+       fill="Number of\n15N atoms in\nisotopologue")
+
+others <- c("Theanine?", "Citrulline", "N-Methyl-L-glutamic acid?", 
+            "Ectoine", "5-Hydroxyectoine")
+iso_areas %>%
+  filter(compound_name%in%others) %>%
+  mutate(compound_name=factor(compound_name, levels=others)) %>%
+  left_join(parametadata) %>%
+  filter(!is.na(depth)) %>%
+  mutate(labeled_n_count=factor(labeled_n_count)) %>%
+  group_by(compound_name, depth, labeled_n_count, timepoint) %>%
+  summarize(area=median(area)) %>%
+  group_by(compound_name, depth, timepoint) %>%
+  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
+  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
+  ggplot() +
+  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
+                  fill=labeled_n_count, group=labeled_n_count)) +
+  facet_grid(depth~compound_name) +
+  scale_fill_brewer(type = "qual", palette = 3) +
+  theme_bw() +
+  labs(x="Timepoint (hours after spike)", 
+       y="Fraction of peak area with 15N label",
+       fill="Number of\n15N atoms in\nisotopologue")
+```
+
+```{r special classes of isotope trends alt visualization total area}
+chosen_cmpds <- MAAs
+iso_areas %>%
+  mutate(compound_name=factor(compound_name, levels=chosen_cmpds)) %>%
+  drop_na() %>%
+  left_join(parametadata) %>%
+  filter(!is.na(depth)) %>%
+  filter(depth=="Surf") %>%
+  mutate(labeled_n_count=factor(labeled_n_count, levels = 5:0)) %>%
+  group_by(depth, compound_name, labeled_n_count, timepoint, startime) %>%
+  summarize(area=median(area)) %>%
+  ggplot() +
+  geom_col(aes(x=timepoint, y=area, fill=labeled_n_count)) +
+  ggh4x::facet_grid2(startime~compound_name, scales="free_y", independent="y") +
+  # facet_grid(startime~compound_name) +
+  scale_fill_brewer(type = "qual", palette = 3, direction = -1) +
+  theme_bw() +
+  theme(legend.position = "none") +
+  labs(x="Timepoint (hours after spike)", y="Peak area with 15N label",
+       fill="Number of\n15N atoms in\nisotopologue")
+```
+
 # Multivariate approaches
@@
   # filter(labeled_n_count>0) %>%
+  # filter(labeled_n_count==0) %>%
   left_join(parametadata) %>%
@@
   geom_hline(yintercept = 0) +
-  geom_point(aes(x=MDS1, y=MDS2, fill=timepoint), color="black", size=4, pch=21) +
+  geom_point(aes(x=MDS1, y=MDS2, fill=timepoint, shape=startime), color="black", size=4) +
+  scale_shape_manual(values = c(22, 23), breaks = c("Morn", "Eve")) +
   facet_wrap(~depth, ncol=1) +
@@
 ```
+
+# Univariate followup

### 2023-10-05 untargeted/paruntargeted_control.Rmd
Added some basic BMIS visuals

@@
   select(compound_name, filename, area, labeled_n_count)
+output_csv_write(iso_areas)
+```
 
-output_csv_write(iso_areas)
+# BMIS and quant
+
+```{r get IS areas from Skyline}
+n_only_is <- c("L-Isoleucine, 15N", "AMP, 15N5", "GMP, 15N5", 
+               "Nitrate, 15N", "DL-Histidine, 15N")
+if(system("untargeted/IS_integrations/SkylineRunner.exe", show.output.on.console = FALSE)){
+  # Website = https://skyline.ms/wiki/home/software/Skyline/page.view?name=SkylineInstall_64_21-1&submit=false
+  stop("Please grab a copy of SkylineRunner from the website")
+}
+if(!file.exists("untargeted/IS_integrations/GMP_pos.csv")){
+  success_vec <- "untargeted/IS_integrations" %>%
+  list.files(pattern = "sky$", full.names = TRUE) %>%
+  sapply(skyline_files, function(skyfile){
+    polarity <- str_extract(skyfile, "POSITIVE|NEGATIVE")
+    polarity <- switch(polarity, "POSITIVE"="pos", "NEGATIVE"="neg")
+    amendment <- str_extract(skyfile, "Amm|Nit|GMP|Arg")
+
+    sky_cmd <- paste0(
+      "untargeted/IS_integrations/SkylineRunner.exe",
+      " --in=", skyfile,
+      ' --report-name="Molecule Peak Boundaries Modified"', 
+      ' --report-file=untargeted/IS_integrations/', amendment, "_", polarity, ".csv"
+    )
+    failed <- system(sky_cmd)
+    if(failed){
+      warning(paste("Extraction failed for file", skyfile))
+    }
+
+    new_csv_name <- paste0("untargeted/IS_integrations/", amendment, "_", polarity, ".csv")
+    new_csv <- read.csv(new_csv_name)
+    new_csv$polarity <- polarity
+    new_csv$amendment <- amendment
+    write.csv(new_csv, new_csv_name, row.names = FALSE)
+  })
+}
+IS_areas_all <- "untargeted/IS_integrations/" %>%
+  list.files(pattern = "(Amm|Nit|Arg|GMP)_(pos|neg).csv", full.names = TRUE) %>%
+  lapply(read.csv) %>%
+  do.call(what = "rbind") %>%
+  mutate(Area=as.numeric(Area)) %>%
+  mutate(Area=ifelse(is.na(Area), 0, Area)) %>%
+  select(filename=File.Name, IS_name=Molecule, rtmin=Min.Start.Time,
+         rtmax=Max.End.Time, IS_area=Area, polarity, amendment) %>%
+  mutate(filename=str_replace(filename, ".raw", ".mzML")) %>%
+  filter(!IS_name%in%n_only_is) %>%
+  filter(!IS_name=="Succinic acid, 2H4") %>%
+  mutate(lead_IS_area=lead(IS_area), lag_IS_area=lag(IS_area)) %>%
+  left_join(parametadata_all) %>%
+  mutate(IS_area=case_when(
+    shortname=="Amm-Surf-Eve-T3-A"~(lead_IS_area+lag_IS_area)/2,
+    shortname=="Nit-Deep-Eve-T0-A"~(lead_IS_area+lag_IS_area)/2,
+    TRUE~IS_area
+  )) %>%
+  select(IS_name, filename, IS_area)
+
+# write.csv(IS_areas_all, "untargeted/IS_integrations/IS_areas.csv", row.names = FALSE)
+
+IS_areas_all %>%
+  slice(1) %>%
+  left_join(parametadata_all)
+  filter(amendment==!!amendment) %>%
+  arrange(timestamp) %>%
+  mutate(shortname=fct_inorder(shortname)) %>%
+  ggplot() +
+  geom_col(aes(x=shortname, y=IS_area, fill=samp_type), width=1) +
+  facet_wrap(~IS_name, scales="free_y", ncol=3)
+```
+
+```{r get PheTau areas from raw files}
+
+```
+
+# Isotope visualization
+
+```{r load isotope mzs and areas}
+all_stans <- read_csv("metadata/all_stans.csv") %>%
+  filter(polarity==!!polarity) %>%
+  filter(compound_type!="Internal Standard") %>%
+  select(compound_name, formula)
+manual_anno <- readxl::read_excel("untargeted/manual_annotations.xlsx") %>%
+  filter(polarity==!!polarity) %>%
+  filter(amendment==!!amendment) %>%
+  select(feature, compound_name)
+best_feat_data <- output_folder %>%
+  paste0("best_feat_data.csv") %>%
+  read_csv() %>%
+  left_join(manual_anno) %>%
+  left_join(all_stans)
+iso_mzs <- read_csv(paste0(output_folder, "iso_mzs.csv"))
+iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv"))
 ```
@@
 single_cmpd_iso_req <- iso_mzs %>%
-  filter(compound_name=="Taurine") %>%
-  distinct(compound_name, n_count, labeled_mzmed) %>%
+  filter(compound_name=="Guanine") %>%
   mutate(labeled_mzmed=round(labeled_mzmed, 5)) %>%

### 2023-10-05 untargeted/paruntargeted_control.Rmd
Cleaning and tidying metadata

@@
 
-polarity <- "pos"
-# polarity <- "neg"
-amendment <- "Amm"
+# polarity <- "pos"
+polarity <- "neg"
+# amendment <- "Amm"
 # amendment <- "Nit"
 # amendment <- "GMP"
-# amendment <- "Arg"
+amendment <- "Arg"
 IS_type <- "noIS"
@@
 
-parametadata <- read_csv("metadata/parametadata.csv") %>% 
-  filter(polarity==!!polarity) %>%
-  filter(amendment==!!amendment) %>%
-  filter(IS_type==!!IS_type) %>%
+parametadata_all <- read_csv("metadata/parametadata.csv") %>% 
   mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
@@
   mutate(timepoint=factor(timepoint, levels=paste0("T", c(0,1,3,10,26,73))))
+parametadata <- parametadata_all %>%
+  filter(polarity==!!polarity) %>%
+  filter(amendment==!!amendment) %>%
+  filter(IS_type==!!IS_type)
 
@@
 xcms_file_data <- parametadata %>%
-  filter(amendment==!!amendment) %>%
-  filter(polarity==!!polarity) %>%
-  filter(IS_type==!!IS_type) %>%
   mutate(samp_group=paste(amendment, depth, startime, timepoint, sep="-"))
@@
 
-# Isotope exploration
+# Isotope detection
 
@@
   mutate(labeled_mzmed=mzmed+0.997025*(n_count-1)) %>%
-  select(compound_name, feature, labeled_mzmed, n_count) %>%
+  select(compound_name, feature, labeled_mzmed, n_count)
+output_csv_write(iso_mzs)
+
+iso_mz_bounds <- iso_mzs %>%
   left_join(peak_rt_bounds, relationship = "many-to-many") %>%
@@
   select(compound_name, feature, n_count, labeled_mzmed, filename, rtmin, rtmax)
-
-iso_areas <- iso_mzs %>%
+iso_areas <- iso_mz_bounds %>%
   split(.$filename) %>% 

### 2023-10-06 untargeted/paruntargeted_control.Rmd
Saved GMP samples, done annotating

@@
 # amendment <- "Nit"
-# amendment <- "GMP"
-amendment <- "Arg"
+amendment <- "GMP"
+# amendment <- "Arg"
 IS_type <- "noIS"
@@
   if(polarity=="pos"){
-    rescue_feats <- c()
+    rescue_feats <- c("FT1422", "FT1580", "FT1585", "FT1603", "FT3030")
   } else if(polarity=="neg"){
-    rescue_feats <- c()
+    rescue_feats <- c("FT0865")
   } else {
@@
 all_feat_data %>%
-  filter(mzmed%between%pmppm(138.9701 , 10))
+  filter(mzmed%between%pmppm(111.0195 , 10))
 ```

### 2023-10-09 arrowscript.R
Moving multiple small scripts to global run_analysis.R

@@ -1,38 +1,0 @@
-library(data.table)
-library(RaMS)
-library(arrow)
-
-parametadata <- fread("metadata/parametadata.csv")[
-  IS_type=="noIS",c("filepath", "filename", "samp_type", "amendment", "polarity")
-]
-parametadata[, .N, by=c("samp_type", "amendment", "polarity")]
-
-for(samp_type_i in unique(parametadata$samp_type)){
-  for(amendment_i in unique(parametadata$amendment)){
-    for(polarity_i in unique(parametadata$polarity)){
-      ms_files <- parametadata[samp_type==samp_type_i][amendment==amendment_i][polarity==polarity_i]$filepath
-      # ms_files <- head(ms_files, 3)
-      msdata <- grabMSdata(ms_files, grab_what = "MS1")
-      parq_dir <- paste0("tmzMLs/pqds",
-                         "/samp_type=", samp_type_i, 
-                         "/amendment=", amendment_i, 
-                         "/polarity=", polarity_i)
-      if(!dir.exists(parq_dir))dir.create(parq_dir, recursive = TRUE)
-      parq_name <- paste0(parq_dir, "/part-0.parquet")
-      print(parq_name)
-      write_parquet(msdata$MS1, parq_name)
-    }
-  }
-}
-
-# library(tidyverse)
-# open_dataset("tmzMLs/pqds") %>%
-#   filter(amendment=="Amm") %>%
-#   filter(polarity=="neg") %>%
-#   filter(samp_type=="Smp") %>%
-#   filter(mz%between%pmppm(150.041585, 5)) %>%
-#   filter(rt%between%c(7.5, 9)) %>%
-#   collect() %>%
-#   left_join(parametadata) %>%
-#   ggplot() +
-#   geom_line(aes(x=rt, y=int, group=filename, color=samp_type))

### 2023-10-09 eicscript.R

@@ -1,66 +1,0 @@
-library(tidyverse)
-library(arrow)
-library(RaMS)
-
-IS_type <- "noIS"
-
-all_stans <- read_csv("metadata/all_stans.csv") %>%
-  mutate(polarity=factor(polarity, levels=c("pos", "neg"))) %>%
-  arrange(polarity) %>%
-  filter(compound_type!="Internal Standard")
-
-for(polarity in c("pos", "neg")){
-  for(amendment in c("Amm", "Nit", "GMP", "Arg")){
-    parametadata <- read_csv("metadata/parametadata.csv") %>% 
-      filter(polarity==!!polarity) %>%
-      filter(amendment==!!amendment) %>%
-      filter(IS_type==!!IS_type)
-    output_folder <- paste(amendment, polarity, IS_type, "output", sep = "_") %>%
-      paste0("untargeted/", ., "/")
-    best_feat_data <- read_csv(paste0(output_folder, "peak_data.csv")) %>%
-      left_join(parametadata) %>%
-      filter(samp_type=="Smp") %>%
-      group_by(feature) %>%
-      summarize(mzmed=unique(feat_mzmed), rtmed=unique(feat_rtmed)/60, n=n(),
-                med_cor=median(beta_cor, na.rm=TRUE), 
-                med_snr=median(beta_snr, na.rm=TRUE)) %>%
-      filter(med_cor>0.75) %>%
-      select(-med_cor, -med_snr)
-    
-    pdf(paste0(output_folder, "anno_eics.pdf"), width = 8, height = 4)
-    all_stans %>%
-      filter(polarity==!!polarity) %>%
-      mutate(mz=ifelse(compound_name=="Carnitine", mz[which(compound_name=="Threonine betaine?")], mz)) %>%
-      group_by(mz) %>%
-      summarise(cmpd_lab=paste0(compound_name, " (", mix, " ,", rt, ")", collapse = ", ")) %>%
-      pwalk(function(...){
-        row_data <- list(...)
-        mass_feats <- best_feat_data %>% filter(mzmed%between%pmppm(row_data$mz, 5))
-        chosen_eic <- open_dataset("tmzMLs/pqds") %>%
-          filter(polarity==!!polarity) %>%
-          filter(amendment==!!amendment) %>%
-          filter(samp_type=="Poo" | samp_type=="Std") %>%
-          filter(mz%between%pmppm(row_data$mz, 5)) %>%
-          filter(rt%between%c(2, 20)) %>%
-          dplyr::collect()
[…]

### 2023-10-09 run_analysis.R

@@ -1,0 +1,121 @@
+
+library(tidyverse)
+
+# Generate metadata ----
+if(!file.exists("metadata/parametadata.csv")){
+  knitr::knit(input = "metadata/parametadata_control.Rmd", 
+              output = "metadata/parametadata_control.R", 
+              tangle = TRUE)
+}
+
+
+# Generate XCMS outputs ----
+if(!file.exists("untargeted/Arg_neg_noIS_output/best_feat_data.csv")){
+  for(amendment in c("Amm", "Nit", "GMP", "Arg")){
+    for(polarity in c("pos", "neg")){
+      source("untargeted/paruntargeted_control.R")
+    }
+  }
+}
+
+
+# Generate parquet files for fast EIC extraction ----
+if(!dir.exists("tmzMLs/pqds")){
+  library(data.table)
+  library(RaMS)
+  library(arrow)
+  parametadata <- fread("metadata/parametadata.csv")[
+    IS_type=="noIS",c("filepath", "filename", "samp_type", "amendment", "polarity")
+  ]
+  parametadata[, .N, by=c("samp_type", "amendment", "polarity")]
+  
+  for(samp_type_i in unique(parametadata$samp_type)){
+    for(amendment_i in unique(parametadata$amendment)){
+      for(polarity_i in unique(parametadata$polarity)){
+        ms_files <- parametadata[samp_type==samp_type_i][amendment==amendment_i][polarity==polarity_i]$filepath
+        # ms_files <- head(ms_files, 3)
+        msdata <- grabMSdata(ms_files, grab_what = "MS1")
+        parq_dir <- paste0("tmzMLs/pqds",
+                           "/samp_type=", samp_type_i, 
+                           "/amendment=", amendment_i, 
+                           "/polarity=", polarity_i)
+        if(!dir.exists(parq_dir))dir.create(parq_dir, recursive = TRUE)
+        parq_name <- paste0(parq_dir, "/part-0.parquet")
+        print(parq_name)
+        write_parquet(msdata$MS1, parq_name)
+      }
+    }
+  }
+}
+
+
+# Generate annotation chromatogram pdfs ----
+if(!file.exists("untargeted/Arg_neg_noIS_output/anno_eics.pdf")){
+  library(tidyverse)
+  library(arrow)
+  library(RaMS)
+  
+  IS_type <- "noIS"
+  
+  all_stans <- read_csv("metadata/all_stans.csv") %>%
[…]

### 2023-10-09 untargeted/paruntargeted_control.Rmd
Cleaning and tidying

@@
 
-# polarity <- "pos"
-polarity <- "neg"
-# amendment <- "Amm"
+polarity <- "pos"
+# polarity <- "neg"
+amendment <- "Amm"
 # amendment <- "Nit"
-amendment <- "GMP"
+# amendment <- "GMP"
 # amendment <- "Arg"
@@
 all_feat_data %>%
-  filter(mzmed%between%pmppm(111.0195 , 10))
+  filter(mzmed%between%pmppm(135.030686 , 10))
 ```
@@
   slice(1) %>%
-  left_join(parametadata_all)
+  left_join(parametadata_all) %>%
   filter(amendment==!!amendment) %>%

### 2023-10-09 mzMLs/msconvert.R
Moving msconvert to run_analysis as well

@@ -1,37 +1,0 @@
-
-folder_paths <- c(
-  "Z:/1_QEdata/2021/2021_Will_PARAGON/210813_Will_NIT-Expt_PARAGON_HILIC",
-  "Z:/1_QEdata/2021/2021_Will_PARAGON/210916_Will_Amm-Expt_PARAGON_HILIC",
-  "Z:/1_QEdata/2022/220715_Will_GMP-Expt_PARAGON_HILIC",
-  "Z:/1_QEdata/2022/220729_Will_Arg-Expt_PARAGON_HILIC"
-)
-
-sapply(folder_paths, function(path_i){
-  sapply(c("positive", "negative"), function(polarity){
-    message(paste("Converting", basename(path_i), polarity))
-    short_pol <- gsub("[ia]tive", "", polarity)
-    match_pattern <- paste0("wIS.raw|noIS-", short_pol, ".*.raw|_Std_.*raw|wIS-(Full|Half)")
-    files_to_convert <- list.files(path_i, pattern = match_pattern, full.names = TRUE)
-    mscommand <- paste(
-      "msconvert",
-      paste(files_to_convert, collapse = " "),
-      "--mzML",
-      paste0('-o mzMLs/', short_pol),
-      '--filter "peakPicking true 1-"',
-      paste('--filter "polarity', polarity, '"')
-    )
-    system(mscommand)
-    
-    msmsmatch_pattern <- paste0("_DDA", short_pol, ".*.raw")
-    msms_to_convert <- list.files(path_i, pattern = msmsmatch_pattern, full.names = TRUE)
-    msmscommand <- paste(
-      "msconvert",
-      paste(msms_to_convert, collapse = " "),
-      "--mzML",
-      paste0('-o mzMLs/', short_pol, "/MSMS"),
-      '--filter "peakPicking true 1-"',
-      paste('--filter "polarity', polarity, '"')
-    )
-    system(msmscommand)
-  })
-})

### 2023-10-09 run_analysis.R

@@ -1,6 +1,47 @@
 
 library(tidyverse)
 
+# Generate mzMLs from .raw via msconvert ----
+if(!file.exists("mzMLs/pos/220729_Smp_Arg-Surf-Morn-T73-C-wIS.mzML")){
+  folder_paths <- c(
+    "Z:/1_QEdata/2021/2021_Will_PARAGON/210813_Will_NIT-Expt_PARAGON_HILIC",
+    "Z:/1_QEdata/2021/2021_Will_PARAGON/210916_Will_Amm-Expt_PARAGON_HILIC",
+    "Z:/1_QEdata/2022/220715_Will_GMP-Expt_PARAGON_HILIC",
+    "Z:/1_QEdata/2022/220729_Will_Arg-Expt_PARAGON_HILIC"
+  )
+  sapply(folder_paths, function(path_i){
+    sapply(c("positive", "negative"), function(polarity){
+      message(paste("Converting", basename(path_i), polarity))
+      short_pol <- gsub("[ia]tive", "", polarity)
+      match_pattern <- paste0("wIS.raw|noIS-", short_pol, ".*.raw|_Std_.*raw|wIS-(Full|Half)")
+      files_to_convert <- list.files(path_i, pattern = match_pattern, full.names = TRUE)
+      mscommand <- paste(
+        "msconvert",
+        paste(files_to_convert, collapse = " "),
+        "--mzML",
+        paste0('-o mzMLs/', short_pol),
+        '--filter "peakPicking true 1-"',
+        paste('--filter "polarity', polarity, '"')
+      )
+      system(mscommand)
+      
+      msmsmatch_pattern <- paste0("_DDA", short_pol, ".*.raw")
+      msms_to_convert <- list.files(path_i, pattern = msmsmatch_pattern, full.names = TRUE)
+      msmscommand <- paste(
+        "msconvert",
+        paste(msms_to_convert, collapse = " "),
+        "--mzML",
+        paste0('-o mzMLs/', short_pol, "/MSMS"),
+        '--filter "peakPicking true 1-"',
+        paste('--filter "polarity', polarity, '"')
+      )
+      system(msmscommand)
+    })
+  })
+}
+
+
+
 # Generate metadata ----
 if(!file.exists("metadata/parametadata.csv")){
   knitr::knit(input = "metadata/parametadata_control.Rmd", 

### 2023-10-09 untargeted/paruntargeted_control.Rmd
Finished annotation of PheTau IS

@@
   filter(amendment==!!amendment) %>%
-  filter(IS_type==!!IS_type)
+  filter(IS_type==!!IS_type | samp_type=="Std")
 
@@
 all_feat_data <- read_csv(paste0(output_folder, "all_feat_data.csv"))
+best_feat_data <- read_csv(paste0(output_folder, "best_feat_data.csv"))
 anno_manu %>%
@@
 all_feat_data %>%
-  filter(mzmed%between%pmppm(135.030686 , 10))
+  filter(mzmed%between%pmppm(128.0325, 10))
+# 128.0325, 130.047052, 174.1365
+mass_check(128.0325)
 ```
@@
 mass_check(c("Cytosine"=244.093347))
-mass_check(c("Proline"=116.0712))
+mass_check(c("Proline"=174.1365))
 ```

### 2023-10-09 untargeted/paruntargeted_control.Rmd
Fixing some erroneous annotations

@@
 best_feat_data <- read_csv(paste0(output_folder, "best_feat_data.csv"))
-anno_manu %>%
-  filter(str_detect(notes, "[Rr]escue"))
-all_feat_data %>%
-  filter(mzmed%between%pmppm(128.0325, 10))
-# 128.0325, 130.047052, 174.1365
-mass_check(128.0325)
+anno_manu %>% filter(str_detect(notes, "[Rr]escue"))
+all_feat_data %>% filter(mzmed%between%pmppm(128.0325, 10))
 ```
@@
   scale_color_viridis_c()
+
+anno_manu %>%
+  select(compound_name, feature, mz) %>%
+  filter(!feature=="NA") %>%
+  left_join(all_feat_data) %>%
+  ggplot(aes(label=feature)) +
+  geom_point(aes(x=mz, y=mzmed-mz)) +
+  geom_abline(slope = 1, intercept = 0)
 

### 2023-10-12 untargeted/paruntargeted_control.Rmd
Now searching for isos in standards and pooled as well
Allows quantification

@@
 
-# polarity <- "pos"
-polarity <- "neg"
-# amendment <- "Amm"
+polarity <- "pos"
+# polarity <- "neg"
+amendment <- "Amm"
 # amendment <- "Nit"
 # amendment <- "GMP"
-amendment <- "Arg"
+# amendment <- "Arg"
 IS_type <- "noIS"
@@
   group_vec
+}
+trapz <- function(x, y) {
+  y <- y[order(x)]
+  x <- x[order(x)]
+  m <- length(x)
+  xp <- c(x, x[m:1])
+  yp <- c(numeric(m), y[m:1])
+  n <- 2*m
+  p1 <- sum(xp[1:(n-1)]*yp[2:n]) + xp[n]*yp[1]
+  p2 <- sum(xp[2:n]*yp[1:(n-1)]) + xp[1]*yp[n]
+
+  return(0.5*(p1-p2))
 }
@@
   left_join(parametadata) %>%
-  filter(samp_type=="Smp") %>%
+  # filter(samp_type=="Smp") %>%
   select(compound_name, feature, n_count, labeled_mzmed, filename, rtmin, rtmax)
@@
   mutate(labeled_n_count=n_count-1) %>%
-  select(compound_name, filename, area, labeled_n_count)
+  left_join(parametadata) %>%
+  select(filename, samp_type, compound_name, labeled_n_count, area)
 output_csv_write(iso_areas)
@@
 iso_mzs <- read_csv(paste0(output_folder, "iso_mzs.csv"))
-iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv"))
+iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv")) %>%
+  filter(samp_type=="Smp")
 ```

### 2023-10-23 untargeted/paruntargeted_control.Rmd
Splitting control doc into analysis and sandbox

@@
 ```{r get PheTau areas from raw files}
+pdf("untargeted/IS_integrations/IS_area_vis.pdf", width = 7, height = 4)
+IS_eics <- open_dataset("tmzMLs/pqds") %>%
+  filter(samp_type%in%c("Smp", "Poo")) %>%
+  filter(mz%between%pmppm(174.1365, 10) & rt%between%c(3, 7.5) | 
+           mz%between%pmppm(128.0325, 10) & rt%between%c(8, 12) | 
+           mz%between%pmppm(130.0471, 10) & rt%between%c(8, 12)) %>%
+  dplyr::collect()
+IS_eics %>%
+  filter(amendment!="Nit") %>%
+  mutate(compound_name=case_when(
+    mz<129~"Taurine, 2H4, neg",
+    mz<140~"Taurine, 2H4, pos",
+    TRUE~"Phenylalanine, 2H8, pos"
+  )) %>%
+  left_join(parametadata_all) %>%
+  group_by(amendment) %>%
+  mutate(timestamp=as.numeric(timestamp-min(timestamp))/as.numeric(max(timestamp)-min(timestamp))) %>%
+  ggplot() +
+  geom_line(aes(x=rt, y=int, group=filename, color=timestamp)) +
+  facet_grid(compound_name~amendment, scales="free")
+IS_eics %>%
+  mutate(compound_name=case_when(
+    mz<129~"Taurine, 2H4, neg",
+    mz<140~"Taurine, 2H4, pos",
+    TRUE~"Phenylalanine, 2H8, pos"
+  )) %>%
+  filter(amendment!="Nit") %>%
+  group_by(filename, compound_name) %>%
+  summarise(into=trapz(rt, int)*60, .groups = "drop") %>%
+  left_join(parametadata_all) %>%
+  arrange(timestamp) %>%
+  mutate(shortname=fct_inorder(paste0(amendment, shortname))) %>%
+  ggplot() +
+  geom_col(aes(x=shortname, y=into, fill=samp_type), width=1) +
+  facet_grid(compound_name~amendment, scales="free") +
+  ggtitle("Reintegrated IS from raw EICs")
 
-```
-
-# Isotope visualization
-
-```{r load isotope mzs and areas}
-all_stans <- read_csv("metadata/all_stans.csv") %>%
-  filter(polarity==!!polarity) %>%
-  filter(compound_type!="Internal Standard") %>%
-  select(compound_name, formula)
-manual_anno <- readxl::read_excel("untargeted/manual_annotations.xlsx") %>%
-  filter(polarity==!!polarity) %>%
-  filter(amendment==!!amendment) %>%
-  select(feature, compound_name)
-best_feat_data <- output_folder %>%
-  paste0("best_feat_data.csv") %>%
-  read_csv() %>%
-  left_join(manual_anno) %>%
-  left_join(all_stans)
-iso_mzs <- read_csv(paste0(output_folder, "iso_mzs.csv"))
-iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv")) %>%
-  filter(samp_type=="Smp")
-```
-
-```{r individual isotope EIC}
-single_cmpd_iso_req <- iso_mzs %>%
-  filter(compound_name=="Guanine") %>%
-  mutate(labeled_mzmed=round(labeled_mzmed, 5)) %>%
-  summarise(req_cmd=paste0("mz%between%pmppm(", labeled_mzmed, ", 10)", collapse="|")) %>%
-  pull(req_cmd)
-full_arrow_req <- paste(
-  'open_dataset("tmzMLs/pqds") %>%',
-  'filter(amendment==!!amendment) %>%',
-  'filter(samp_type=="Smp") %>%',
-  'filter(polarity==!!polarity) %>% ',
-  'filter(', single_cmpd_iso_req, ') %>% ',
-  'dplyr::collect()'
-)
-single_cmpd_iso_eic <- eval(parse(text=full_arrow_req))
-
-iso_mzs %>%
-  filter(compound_name=="Guanine") %>%
-  pmap(function(...){
-    row_data <- list(...)
-    single_cmpd_iso_eic %>%
-      filter(mz%between%pmppm(row_data$labeled_mzmed, 10)) %>%
-      mutate(n_count=row_data$n_count)
-  }) %>%
-  bind_rows() %>%
-  filter(rt%between%c(7, 9)) %>%
-  left_join(parametadata) %>%
-  filter(depth=="Surf" & startime=="Morn") %>%
+PheTau_all_amendments <- readxl::read_excel("untargeted/manual_annotations.xlsx") %>%
+  filter(compound_name%in%c("Taurine, 2H4", "Phenylalanine, 2H8")) %>%
+  mutate(output_folder=paste(amendment, polarity, "noIS_output/", sep="_")) %>%
+  nest(IS_info=-output_folder) %>%
+  mutate(IS_data=map(paste0("untargeted/", output_folder, "peak_data.csv"), read_csv)) %>%
+  unnest(IS_info) %>%
+  mutate(IS_data=map2(feature, IS_data, ~filter(.y, feature==.x))) %>%
+  select(-feature, -mz, -rt) %>%
+  unnest(cols = c(IS_data)) %>%
+  select(amendment, polarity, compound_name, filename, into)
+PheTau_all_amendments %>%
+  mutate(compound_name=paste(compound_name, polarity, sep=", ")) %>%
+  arrange(desc(into)) %>%
+  group_by(compound_name, filename) %>%
+  mutate(n=n()) %>%
+  slice(1) %>%
+  ungroup() %>%
+  left_join(parametadata_all) %>%
+  filter(samp_type!="Std") %>%
+  arrange(timestamp) %>%
+  mutate(shortname=fct_inorder(paste0(amendment, shortname))) %>%
   ggplot() +
-  geom_line(aes(x=rt, y=int, group=filename, color=timepoint)) +
-  facet_grid(timepoint~n_count) +
-  ggtitle("Surface morning spiked samples")
-```
-
-```{r individual isotope area plots}
-iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv"))
-
-cmpd_name_i <- "Guanine"
-single_cmpd_iso_data <- iso_areas %>%
-  filter(compound_name==cmpd_name_i) %>%
-  left_join(parametadata) %>%
-  filter(!is.na(depth)) %>%
-  mutate(labeled_n_count=factor(labeled_n_count))
-
-single_cmpd_iso_data %>%
-  ggplot() +
-  geom_boxplot(aes(x=labeled_n_count, y=area, color=timepoint)) +
-  facet_grid(depth~startime, scales="free_y") +
-  ggtitle("Raw peak areas across depth and time")
-
-single_cmpd_iso_data %>%
-  group_by(depth, startime, labeled_n_count) %>%
-  mutate(area=area/sum(area)) %>%
-  ggplot() +
-  geom_boxplot(aes(x=labeled_n_count, y=area, color=timepoint)) +
-  facet_grid(depth~startime) +
-  ggtitle("Peak areas normalized to the max within each isotope")
-
-single_cmpd_iso_data %>%
-  group_by(depth, startime, labeled_n_count, timepoint) %>%
-  summarize(area=median(area)) %>%
-  group_by(depth, startime, timepoint) %>%
-  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
-  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
-  ggplot() +
-  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
-                  fill=labeled_n_count, group=labeled_n_count)) +
-  facet_grid(depth~startime) +
-  scale_fill_brewer(type = "qual", palette = 3) +
-  theme_bw() +
-  labs(title = "Median triplicate peak area as a fraction of total for each timepoint",
-       x="Timepoint", fill="15N count", 
-       y="Fraction of peak area in each isotopologue")
-
-single_cmpd_iso_data %>%
-  # group_by(depth, startime, labeled_n_count, timepoint) %>%
-  # summarize(area=median(area)) %>%
-  ggplot() +
-  geom_col(aes(x=timepoint, y=area, fill=labeled_n_count), color="grey50") +
-  facet_grid(depth~startime, scales="free_y") +
-  scale_fill_brewer(type = "qual", palette = 3) +
-  theme_bw() +
-  labs(x="Timepoint (hours after spike)", y="Peak area with 15N label",
-       fill="Number of\n15N atoms in\nisotopologue",
-       title = "Summed peak area by timepoint")
-```
-
-```{r isotopes for all targ cmpds}
-pdf(paste0(output_folder, "targ_cmpd_mids.pdf"), width = 6, height = 3)
-walk(unique(iso_areas$compound_name), function(cmpd_name_i){
-  gp <- iso_areas %>%
-    filter(compound_name==cmpd_name_i) %>%
-    left_join(parametadata, by="filename") %>%
-    filter(!is.na(depth)) %>%
-    mutate(labeled_n_count=factor(labeled_n_count)) %>%
-    group_by(depth, startime, labeled_n_count, timepoint) %>%
-    summarize(area=median(area), .groups = "drop") %>%
-    group_by(depth, startime, timepoint) %>%
-    mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
-    mutate(upper_bound=lead(lower_bound, default = 1)) %>%
-    ggplot() +
-    geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
-                    fill=labeled_n_count, group=labeled_n_count)) +
-    facet_grid(startime~depth) +
-    scale_fill_brewer(type = "qual", palette = 3) +
-    labs(x="Timepoint", fill="15N count", y="Fraction of peak area in\neach isotopologue") +
-    theme_bw() +
-    ggtitle(cmpd_name_i)
-  plot(gp)
-}, .progress = TRUE)
+  geom_col(aes(x=shortname, y=into, fill=samp_type), width=1) +
+  facet_grid(compound_name~amendment, scales="free") +
+  ggtitle("IS areas from XCMS")
 dev.off()
 ```
-
-```{r special classes of isotope trends}
-iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv"))
-nucleos <- c("Adenine", "Guanine", "Cytosine", "Hypoxanthine", 
-             "Adenosine", "Guanosine", "Cytidine", "Inosine",
-             "Deoxyadenosine", "Deoxyguanosine?", "Deoxycytidine?")
-iso_areas %>%
-  filter(compound_name%in%nucleos) %>%
-  mutate(compound_name=factor(compound_name, levels=nucleos)) %>%
-  left_join(parametadata) %>%
-  filter(!is.na(depth)) %>%
-  mutate(labeled_n_count=factor(labeled_n_count)) %>%
-  filter(depth=="Surf") %>%
-  # filter(compound_name=="Guanine") %>%
-  group_by(compound_name, labeled_n_count, timepoint) %>%
-  summarize(area=median(area)) %>%
-  group_by(compound_name, timepoint) %>%
-  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
-  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
-  ggplot() +
-  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
-                  fill=labeled_n_count, group=labeled_n_count)) +
-  facet_wrap(~compound_name, nrow=3) +
-  scale_fill_brewer(type = "qual", palette = 3) +
-  theme_bw() +
-  labs(x="Timepoint (hours after spike)", 
-       y="Fraction of peak area with 15N label",
-       fill="Number of\n15N atoms in\nisotopologue")
-
-nbases <- c("Adenine", "Guanine", "Cytosine", "Hypoxanthine")
-iso_areas %>%
-  filter(compound_name%in%nbases) %>%
-  mutate(compound_name=factor(compound_name, levels=nbases)) %>%
-  left_join(parametadata) %>%
-  filter(!is.na(depth)) %>%
-  filter(labeled_n_count>0) %>%
-  mutate(labeled_n_count=factor(labeled_n_count)) %>%
-  filter(timepoint=="T26") %>%
-  ggplot() +
-  geom_boxplot(aes(x=labeled_n_count, y=area)) +
-  ggh4x::facet_grid2(depth~compound_name, scales="free_y", independent = "y")
-
-aminos <- c("Glycine", "L-Alanine", "L-Serine", "L-Proline", 
-            "L-Threonine", "L-Leucine", "L-Isoleucine", "L-Tyrosine", 
-            "L-Aspartic acid", "L-Asparagine", "beta-Glutamic acid", 
-            "L-Glutamine", "L-Methionine", "L-Histidine", "L-Arginine")
-iso_areas %>%
-  filter(compound_name%in%aminos) %>%
-  mutate(compound_name=factor(compound_name, levels=aminos)) %>%
-  left_join(parametadata) %>%
-  filter(!is.na(depth)) %>%
-  mutate(labeled_n_count=factor(labeled_n_count)) %>%
-  filter(depth=="Surf") %>%
-  group_by(compound_name, labeled_n_count, timepoint) %>%
-  summarize(area=median(area)) %>%
-  group_by(compound_name, timepoint) %>%
-  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
-  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
-  ggplot() +
-  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
-                  fill=labeled_n_count, group=labeled_n_count)) +
-  facet_wrap(~compound_name, ncol=5) +
-  scale_fill_brewer(type = "qual", palette = 3) +
-  theme_bw() +
-  labs(x="Timepoint (hours after spike)", 
-       y="Fraction of peak area with 15N label",
-       fill="Number of\n15N atoms in\nisotopologue")
-
-MAAs <- c("Palythine?", "Mycosporine-glycine?", "Asterina-330?", 
-          "Shinorine?", "Porphyra-334?", "Mycosporine-2-glycine?")
-iso_areas %>%
-  filter(compound_name%in%MAAs) %>%
-  mutate(compound_name=factor(compound_name, levels=MAAs)) %>%
-  left_join(parametadata) %>%
-  filter(!is.na(depth)) %>%
-  mutate(labeled_n_count=factor(labeled_n_count)) %>%
-  filter(depth=="Surf") %>%
-  group_by(compound_name, startime, labeled_n_count, timepoint) %>%
-  summarize(area=median(area)) %>%
-  group_by(compound_name, startime, timepoint) %>%
-  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
-  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
-  ggplot() +
-  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
-                  fill=labeled_n_count, group=labeled_n_count)) +
-  facet_grid(startime~compound_name) +
-  scale_fill_brewer(type = "qual", palette = 3) +
-  coord_cartesian(ylim = c(0.5, 1)) +
-  theme_bw() +
-  labs(x="Timepoint (hours after spike)", 
-       y="Fraction of peak area with 15N label",
-       fill="Number of\n15N atoms in\nisotopologue")
-
-others <- c("Theanine?", "Citrulline", "N-Methyl-L-glutamic acid?", 
-            "Ectoine", "5-Hydroxyectoine")
-iso_areas %>%
-  filter(compound_name%in%others) %>%
-  mutate(compound_name=factor(compound_name, levels=others)) %>%
-  left_join(parametadata) %>%
-  filter(!is.na(depth)) %>%
-  mutate(labeled_n_count=factor(labeled_n_count)) %>%
-  group_by(compound_name, depth, labeled_n_count, timepoint) %>%
-  summarize(area=median(area)) %>%
-  group_by(compound_name, depth, timepoint) %>%
-  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
-  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
-  ggplot() +
-  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
-                  fill=labeled_n_count, group=labeled_n_count)) +
-  facet_grid(depth~compound_name) +
-  scale_fill_brewer(type = "qual", palette = 3) +
-  theme_bw() +
-  labs(x="Timepoint (hours after spike)", 
-       y="Fraction of peak area with 15N label",
-       fill="Number of\n15N atoms in\nisotopologue")
-```
-
-```{r special classes of isotope trends alt visualization total area}
-chosen_cmpds <- MAAs
-iso_areas %>%
-  mutate(compound_name=factor(compound_name, levels=chosen_cmpds)) %>%
-  drop_na() %>%
-  left_join(parametadata) %>%
-  filter(!is.na(depth)) %>%
-  filter(depth=="Surf") %>%
-  mutate(labeled_n_count=factor(labeled_n_count, levels = 5:0)) %>%
-  group_by(depth, compound_name, labeled_n_count, timepoint, startime) %>%
-  summarize(area=median(area)) %>%
-  ggplot() +
-  geom_col(aes(x=timepoint, y=area, fill=labeled_n_count)) +
-  ggh4x::facet_grid2(startime~compound_name, scales="free_y", independent="y") +
-  # facet_grid(startime~compound_name) +
-  scale_fill_brewer(type = "qual", palette = 3, direction = -1) +
-  theme_bw() +
-  theme(legend.position = "none") +
-  labs(x="Timepoint (hours after spike)", y="Peak area with 15N label",
-       fill="Number of\n15N atoms in\nisotopologue")
-```
-
-# Multivariate approaches
-
-```{r overall analysis}
-iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv"))
-areas_scaled <- iso_areas %>%
-  # filter(labeled_n_count>0) %>%
-  # filter(labeled_n_count==0) %>%
-  left_join(parametadata) %>%
-  filter(!is.na(depth)) %>%
-  mutate(compound_iso=paste0(compound_name, " 15N", labeled_n_count)) %>%
-  arrange(desc(area)) %>%
-  select(shortname, compound_iso, area) %>%
-  group_by(shortname, compound_iso) %>%
-  slice(1) %>%
-  ungroup() %>%
-  complete(shortname, compound_iso, fill = list(area=0)) %>%
-  group_by(compound_iso) %>%
-  # mutate(norm_area=(area-min(area))/(max(area)-min(area))) %>%
-  # mutate(norm_area=rank(area)) %>%
-  mutate(norm_area=scale(area)[,1]) %>%
-  ungroup() %>%
-  select(shortname, compound_iso, norm_area)
-
-areas_scaled %>%
-  pivot_wider(names_from = shortname, values_from = norm_area) %>%
-  column_to_rownames("compound_iso") %>%
-  data.matrix() %>%
-  t() %>%
-  prcomp() %>%
-  pluck("x") %>%
-  as.data.frame() %>%
-  rownames_to_column("shortname") %>%
-  select(shortname, PC1:PC3) %>%
-  left_join(parametadata) %>%
-  arrange(timepoint) %>%
-  mutate(shortname=fct_inorder(shortname)) %>%
-  pivot_longer(starts_with("PC")) %>%
-  filter(name%in%paste0("PC", 1:3)) %>%
-  ggplot() +
-  geom_col(aes(x=shortname, y=value, fill=timepoint)) +
-  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5)) +
-  facet_wrap(~name, ncol=1)
-
-areas_mat <- areas_scaled %>%
-  left_join(parametadata, by="shortname") %>%
-  select(compound_iso, shortname, norm_area) %>%
-  group_by(compound_iso) %>%
-  filter(sum(norm_area==0)<n()*0.9) %>%
-  pivot_wider(names_from = "compound_iso", values_from = norm_area) %>%
-  column_to_rownames("shortname") %>%
-  data.matrix()
-
-adon_expl <- data.frame(shortname=rownames(areas_mat)) %>%
-  left_join(parametadata %>% select(shortname, depth, timepoint, startime), by="shortname")
-vegan::adonis2(
-  areas_mat~timepoint+startime+depth, data = adon_expl, by = "margin", 
-  permutations = 999, method = "manhattan")
-```
-
-```{r multivar analysis by depth}
-multivar_analysis <- areas_scaled %>%
-  left_join(parametadata, by="shortname") %>%
-  nest(data=-depth) %>%
-  mutate(wide_peaks=map(data, function(metab_df){
-    metab_df %>%
-      select(compound_iso, shortname, norm_area) %>%
-      group_by(compound_iso) %>%
-      filter(sum(norm_area==0)<n()*0.9) %>%
-      pivot_wider(names_from = "compound_iso", values_from = norm_area) %>%
-      column_to_rownames("shortname") %>%
-      data.matrix()
-  })) %>%
-  mutate(nmds_output=map(wide_peaks, function(metab_mat){
-    # env_vec <- pull(parametadata, timepoint, shortname)[rownames(metab_mat)]
-    vegan::metaMDS(metab_mat, k = 2, distance = "manhattan", trace = 0, 
-                   autotransform = FALSE, noshare = FALSE, wascores = FALSE)
-    # vegan::MDSrotate(env_vec)
-  })) %>%
-  mutate(nmds_stress=map_dbl(nmds_output, pluck, "stress")) %>%
-  mutate(perm_output=map(wide_peaks, function(metab_mat){
-    adon_expl <- data.frame(shortname=rownames(metab_mat)) %>%
-      left_join(parametadata %>% select(shortname, timepoint, startime), by="shortname")
-    vegan::adonis2(
-      metab_mat~timepoint+startime, data = adon_expl, by = "margin", 
-      permutations = 999, method = "manhattan")
-  })) %>%
-  mutate(pca_output=map(wide_peaks, prcomp)) %>%
-  mutate(hclust_obj=map(wide_peaks, ~hclust(dist(.x))))
-
-multivar_analysis %>%
-  mutate(nmds_points=map(nmds_output, function(nmds_output_i){
-    nmds_output_i %>% pluck("points") %>% as.data.frame()
-  })) %>%
-  select(depth, nmds_points) %>%
-  unnest_longer(nmds_points, indices_to = "shortname") %>%
-  unnest_wider(nmds_points) %>%
-  left_join(parametadata %>% distinct(shortname, shortname, timepoint, startime)) %>%
-  # mutate(MDS2=ifelse(str_detect(depth,"Deep"), MDS2*-1, MDS2)) %>%
-  ggplot(aes(label=shortname)) +
-  geom_vline(xintercept = 0) +
-  geom_hline(yintercept = 0) +
-  geom_point(aes(x=MDS1, y=MDS2, fill=timepoint, shape=startime), color="black", size=4) +
-  scale_shape_manual(values = c(22, 23), breaks = c("Morn", "Eve")) +
-  facet_wrap(~depth, ncol=1) +
-  theme_bw() +
-  coord_fixed()
-
-multivar_analysis$perm_output
-```
-
-# Univariate followup

### 2023-10-23 untargeted/paruntargeted_sandbox.Rmd
Splitting control doc into analysis and sandbox

---
title: "figure sandbox"
author: "William Kumler"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Isotope visualization

```{r load isotope mzs and areas}
all_stans <- read_csv("metadata/all_stans.csv") %>%
  filter(polarity==!!polarity) %>%
  filter(compound_type!="Internal Standard") %>%
  select(compound_name, formula)
manual_anno <- readxl::read_excel("untargeted/manual_annotations.xlsx") %>%
  filter(polarity==!!polarity) %>%
  filter(amendment==!!amendment) %>%
  select(feature, compound_name)
best_feat_data <- output_folder %>%
  paste0("best_feat_data.csv") %>%
  read_csv() %>%
  left_join(manual_anno) %>%
  left_join(all_stans)
iso_mzs <- read_csv(paste0(output_folder, "iso_mzs.csv"))
iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv")) %>%
  filter(samp_type=="Smp")
```

```{r individual isotope EIC}
single_cmpd_iso_req <- iso_mzs %>%
  filter(compound_name=="Guanine") %>%
  mutate(labeled_mzmed=round(labeled_mzmed, 5)) %>%
  summarise(req_cmd=paste0("mz%between%pmppm(", labeled_mzmed, ", 10)", collapse="|")) %>%
  pull(req_cmd)
full_arrow_req <- paste(
  'open_dataset("tmzMLs/pqds") %>%',
  'filter(amendment==!!amendment) %>%',
  'filter(samp_type=="Smp") %>%',
  'filter(polarity==!!polarity) %>% ',
  'filter(', single_cmpd_iso_req, ') %>% ',
  'dplyr::collect()'
)
single_cmpd_iso_eic <- eval(parse(text=full_arrow_req))

iso_mzs %>%
  filter(compound_name=="Guanine") %>%
  pmap(function(...){
    row_data <- list(...)
    single_cmpd_iso_eic %>%
      filter(mz%between%pmppm(row_data$labeled_mzmed, 10)) %>%
      mutate(n_count=row_data$n_count)
  }) %>%
  bind_rows() %>%
  filter(rt%between%c(7, 9)) %>%
  left_join(parametadata) %>%
  filter(depth=="Surf" & startime=="Morn") %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename, color=timepoint)) +
  facet_grid(timepoint~n_count) +
  ggtitle("Surface morning spiked samples")
```

```{r individual isotope area plots}
iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv"))

cmpd_name_i <- "Guanine"
single_cmpd_iso_data <- iso_areas %>%
  filter(compound_name==cmpd_name_i) %>%
  left_join(parametadata) %>%
  filter(!is.na(depth)) %>%
  mutate(labeled_n_count=factor(labeled_n_count))

single_cmpd_iso_data %>%
  ggplot() +
  geom_boxplot(aes(x=labeled_n_count, y=area, color=timepoint)) +
  facet_grid(depth~startime, scales="free_y") +
  ggtitle("Raw peak areas across depth and time")

single_cmpd_iso_data %>%
  group_by(depth, startime, labeled_n_count) %>%
  mutate(area=area/sum(area)) %>%
  ggplot() +
  geom_boxplot(aes(x=labeled_n_count, y=area, color=timepoint)) +
  facet_grid(depth~startime) +
  ggtitle("Peak areas normalized to the max within each isotope")

single_cmpd_iso_data %>%
  group_by(depth, startime, labeled_n_count, timepoint) %>%
  summarize(area=median(area)) %>%
  group_by(depth, startime, timepoint) %>%
  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
  ggplot() +
  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
                  fill=labeled_n_count, group=labeled_n_count)) +
  facet_grid(depth~startime) +
  scale_fill_brewer(type = "qual", palette = 3) +
  theme_bw() +
  labs(title = "Median triplicate peak area as a fraction of total for each timepoint",
       x="Timepoint", fill="15N count", 
       y="Fraction of peak area in each isotopologue")

single_cmpd_iso_data %>%
  # group_by(depth, startime, labeled_n_count, timepoint) %>%
  # summarize(area=median(area)) %>%
  ggplot() +
  geom_col(aes(x=timepoint, y=area, fill=labeled_n_count), color="grey50") +
  facet_grid(depth~startime, scales="free_y") +
  scale_fill_brewer(type = "qual", palette = 3) +
  theme_bw() +
  labs(x="Timepoint (hours after spike)", y="Peak area with 15N label",
       fill="Number of\n15N atoms in\nisotopologue",
       title = "Summed peak area by timepoint")
```

```{r isotopes for all targ cmpds}
pdf(paste0(output_folder, "targ_cmpd_mids.pdf"), width = 6, height = 3)
walk(unique(iso_areas$compound_name), function(cmpd_name_i){
  gp <- iso_areas %>%
    filter(compound_name==cmpd_name_i) %>%
    left_join(parametadata, by="filename") %>%
    filter(!is.na(depth)) %>%
    mutate(labeled_n_count=factor(labeled_n_count)) %>%
    group_by(depth, startime, labeled_n_count, timepoint) %>%
    summarize(area=median(area), .groups = "drop") %>%
    group_by(depth, startime, timepoint) %>%
    mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
    mutate(upper_bound=lead(lower_bound, default = 1)) %>%
    ggplot() +
    geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
                    fill=labeled_n_count, group=labeled_n_count)) +
    facet_grid(startime~depth) +
    scale_fill_brewer(type = "qual", palette = 3) +
    labs(x="Timepoint", fill="15N count", y="Fraction of peak area in\neach isotopologue") +
    theme_bw() +
    ggtitle(cmpd_name_i)
  plot(gp)
}, .progress = TRUE)
dev.off()
```

```{r special classes of isotope trends}
iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv"))
nucleos <- c("Adenine", "Guanine", "Cytosine", "Hypoxanthine", 
             "Adenosine", "Guanosine", "Cytidine", "Inosine",
             "Deoxyadenosine", "Deoxyguanosine?", "Deoxycytidine?")
iso_areas %>%
  filter(compound_name%in%nucleos) %>%
  mutate(compound_name=factor(compound_name, levels=nucleos)) %>%
  left_join(parametadata) %>%
  filter(!is.na(depth)) %>%
  mutate(labeled_n_count=factor(labeled_n_count)) %>%
  filter(depth=="Surf") %>%
  # filter(compound_name=="Guanine") %>%
  group_by(compound_name, labeled_n_count, timepoint) %>%
  summarize(area=median(area)) %>%
  group_by(compound_name, timepoint) %>%
  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
  ggplot() +
  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
                  fill=labeled_n_count, group=labeled_n_count)) +
  facet_wrap(~compound_name, nrow=3) +
  scale_fill_brewer(type = "qual", palette = 3) +
  theme_bw() +
  labs(x="Timepoint (hours after spike)", 
       y="Fraction of peak area with 15N label",
       fill="Number of\n15N atoms in\nisotopologue")

nbases <- c("Adenine", "Guanine", "Cytosine", "Hypoxanthine")
iso_areas %>%
  filter(compound_name%in%nbases) %>%
  mutate(compound_name=factor(compound_name, levels=nbases)) %>%
  left_join(parametadata) %>%
  filter(!is.na(depth)) %>%
  filter(labeled_n_count>0) %>%
  mutate(labeled_n_count=factor(labeled_n_count)) %>%
  filter(timepoint=="T26") %>%
  ggplot() +
  geom_boxplot(aes(x=labeled_n_count, y=area)) +
  ggh4x::facet_grid2(depth~compound_name, scales="free_y", independent = "y")

aminos <- c("Glycine", "L-Alanine", "L-Serine", "L-Proline", 
            "L-Threonine", "L-Leucine", "L-Isoleucine", "L-Tyrosine", 
            "L-Aspartic acid", "L-Asparagine", "beta-Glutamic acid", 
            "L-Glutamine", "L-Methionine", "L-Histidine", "L-Arginine")
iso_areas %>%
  filter(compound_name%in%aminos) %>%
  mutate(compound_name=factor(compound_name, levels=aminos)) %>%
  left_join(parametadata) %>%
  filter(!is.na(depth)) %>%
  mutate(labeled_n_count=factor(labeled_n_count)) %>%
  filter(depth=="Surf") %>%
  group_by(compound_name, labeled_n_count, timepoint) %>%
  summarize(area=median(area)) %>%
  group_by(compound_name, timepoint) %>%
  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
  ggplot() +
  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
                  fill=labeled_n_count, group=labeled_n_count)) +
  facet_wrap(~compound_name, ncol=5) +
  scale_fill_brewer(type = "qual", palette = 3) +
  theme_bw() +
  labs(x="Timepoint (hours after spike)", 
       y="Fraction of peak area with 15N label",
       fill="Number of\n15N atoms in\nisotopologue")

MAAs <- c("Palythine?", "Mycosporine-glycine?", "Asterina-330?", 
          "Shinorine?", "Porphyra-334?", "Mycosporine-2-glycine?")
iso_areas %>%
  filter(compound_name%in%MAAs) %>%
  mutate(compound_name=factor(compound_name, levels=MAAs)) %>%
  left_join(parametadata) %>%
  filter(!is.na(depth)) %>%
  mutate(labeled_n_count=factor(labeled_n_count)) %>%
  filter(depth=="Surf") %>%
  group_by(compound_name, startime, labeled_n_count, timepoint) %>%
  summarize(area=median(area)) %>%
  group_by(compound_name, startime, timepoint) %>%
  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
  ggplot() +
  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
                  fill=labeled_n_count, group=labeled_n_count)) +
  facet_grid(startime~compound_name) +
  scale_fill_brewer(type = "qual", palette = 3) +
  coord_cartesian(ylim = c(0.5, 1)) +
  theme_bw() +
  labs(x="Timepoint (hours after spike)", 
       y="Fraction of peak area with 15N label",
       fill="Number of\n15N atoms in\nisotopologue")

others <- c("Theanine?", "Citrulline", "N-Methyl-L-glutamic acid?", 
            "Ectoine", "5-Hydroxyectoine")
iso_areas %>%
  filter(compound_name%in%others) %>%
  mutate(compound_name=factor(compound_name, levels=others)) %>%
  left_join(parametadata) %>%
  filter(!is.na(depth)) %>%
  mutate(labeled_n_count=factor(labeled_n_count)) %>%
  group_by(compound_name, depth, labeled_n_count, timepoint) %>%
  summarize(area=median(area)) %>%
  group_by(compound_name, depth, timepoint) %>%
  mutate(lower_bound=cumsum(lag(area, default = 0))/sum(area)) %>%
  mutate(upper_bound=lead(lower_bound, default = 1)) %>%
  ggplot() +
  geom_ribbon(aes(x=timepoint, ymin=lower_bound, ymax=upper_bound, 
                  fill=labeled_n_count, group=labeled_n_count)) +
  facet_grid(depth~compound_name) +
  scale_fill_brewer(type = "qual", palette = 3) +
  theme_bw() +
  labs(x="Timepoint (hours after spike)", 
       y="Fraction of peak area with 15N label",
       fill="Number of\n15N atoms in\nisotopologue")
```

```{r special classes of isotope trends alt visualization total area}
chosen_cmpds <- MAAs
iso_areas %>%
  mutate(compound_name=factor(compound_name, levels=chosen_cmpds)) %>%
  drop_na() %>%
  left_join(parametadata) %>%
  filter(!is.na(depth)) %>%
  filter(depth=="Surf") %>%
  mutate(labeled_n_count=factor(labeled_n_count, levels = 5:0)) %>%
  group_by(depth, compound_name, labeled_n_count, timepoint, startime) %>%
  summarize(area=median(area)) %>%
  ggplot() +
  geom_col(aes(x=timepoint, y=area, fill=labeled_n_count)) +
  ggh4x::facet_grid2(startime~compound_name, scales="free_y", independent="y") +
  # facet_grid(startime~compound_name) +
  scale_fill_brewer(type = "qual", palette = 3, direction = -1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(x="Timepoint (hours after spike)", y="Peak area with 15N label",
       fill="Number of\n15N atoms in\nisotopologue")
```

# Multivariate approaches

```{r overall analysis}
iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv"))
areas_scaled <- iso_areas %>%
  # filter(labeled_n_count>0) %>%
  # filter(labeled_n_count==0) %>%
  left_join(parametadata) %>%
  filter(!is.na(depth)) %>%
  mutate(compound_iso=paste0(compound_name, " 15N", labeled_n_count)) %>%
  arrange(desc(area)) %>%
  select(shortname, compound_iso, area) %>%
  group_by(shortname, compound_iso) %>%
  slice(1) %>%
  ungroup() %>%
  complete(shortname, compound_iso, fill = list(area=0)) %>%
  group_by(compound_iso) %>%
  # mutate(norm_area=(area-min(area))/(max(area)-min(area))) %>%
  # mutate(norm_area=rank(area)) %>%
  mutate(norm_area=scale(area)[,1]) %>%
  ungroup() %>%
  select(shortname, compound_iso, norm_area)

areas_scaled %>%
  pivot_wider(names_from = shortname, values_from = norm_area) %>%
  column_to_rownames("compound_iso") %>%
  data.matrix() %>%
  t() %>%
  prcomp() %>%
  pluck("x") %>%
  as.data.frame() %>%
  rownames_to_column("shortname") %>%
  select(shortname, PC1:PC3) %>%
  left_join(parametadata) %>%
  arrange(timepoint) %>%
  mutate(shortname=fct_inorder(shortname)) %>%
  pivot_longer(starts_with("PC")) %>%
  filter(name%in%paste0("PC", 1:3)) %>%
  ggplot() +
  geom_col(aes(x=shortname, y=value, fill=timepoint)) +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5)) +
  facet_wrap(~name, ncol=1)

areas_mat <- areas_scaled %>%
  left_join(parametadata, by="shortname") %>%
  select(compound_iso, shortname, norm_area) %>%
  group_by(compound_iso) %>%
  filter(sum(norm_area==0)<n()*0.9) %>%
  pivot_wider(names_from = "compound_iso", values_from = norm_area) %>%
  column_to_rownames("shortname") %>%
  data.matrix()

adon_expl <- data.frame(shortname=rownames(areas_mat)) %>%
  left_join(parametadata %>% select(shortname, depth, timepoint, startime), by="shortname")
vegan::adonis2(
  areas_mat~timepoint+startime+depth, data = adon_expl, by = "margin", 
  permutations = 999, method = "manhattan")
```

```{r multivar analysis by depth}
multivar_analysis <- areas_scaled %>%
  left_join(parametadata, by="shortname") %>%
  nest(data=-depth) %>%
  mutate(wide_peaks=map(data, function(metab_df){
    metab_df %>%
      select(compound_iso, shortname, norm_area) %>%
      group_by(compound_iso) %>%
      filter(sum(norm_area==0)<n()*0.9) %>%
      pivot_wider(names_from = "compound_iso", values_from = norm_area) %>%
      column_to_rownames("shortname") %>%
      data.matrix()
  })) %>%
  mutate(nmds_output=map(wide_peaks, function(metab_mat){
    # env_vec <- pull(parametadata, timepoint, shortname)[rownames(metab_mat)]
    vegan::metaMDS(metab_mat, k = 2, distance = "manhattan", trace = 0, 
                   autotransform = FALSE, noshare = FALSE, wascores = FALSE)
    # vegan::MDSrotate(env_vec)
  })) %>%
  mutate(nmds_stress=map_dbl(nmds_output, pluck, "stress")) %>%
  mutate(perm_output=map(wide_peaks, function(metab_mat){
    adon_expl <- data.frame(shortname=rownames(metab_mat)) %>%
      left_join(parametadata %>% select(shortname, timepoint, startime), by="shortname")
    vegan::adonis2(
      metab_mat~timepoint+startime, data = adon_expl, by = "margin", 
      permutations = 999, method = "manhattan")
  })) %>%
  mutate(pca_output=map(wide_peaks, prcomp)) %>%
  mutate(hclust_obj=map(wide_peaks, ~hclust(dist(.x))))

multivar_analysis %>%
  mutate(nmds_points=map(nmds_output, function(nmds_output_i){
    nmds_output_i %>% pluck("points") %>% as.data.frame()
  })) %>%
  select(depth, nmds_points) %>%
  unnest_longer(nmds_points, indices_to = "shortname") %>%
  unnest_wider(nmds_points) %>%
  left_join(parametadata %>% distinct(shortname, shortname, timepoint, startime)) %>%
  # mutate(MDS2=ifelse(str_detect(depth,"Deep"), MDS2*-1, MDS2)) %>%
  ggplot(aes(label=shortname)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  geom_point(aes(x=MDS1, y=MDS2, fill=timepoint, shape=startime), color="black", size=4) +
  scale_shape_manual(values = c(22, 23), breaks = c("Morn", "Eve")) +
  facet_wrap(~depth, ncol=1) +
  theme_bw() +
  coord_fixed()

multivar_analysis$perm_output
```

# Univariate followup

### 2023-11-07 untargeted/paruntargeted_sandbox.Rmd
Adding isotope viz and some metadata clarifications

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+knitr::opts_chunk$set(echo = TRUE, purl=FALSE)
+options(pillar.sigfig=7)
+options(readr.num_columns = 0)
+options(readr.show_col_types = FALSE)
+library(tidyverse)
+library(data.table)
+library(arrow)
+library(xcms)
+library(RaMS)
 
-# Isotope visualization
+# polarity <- "pos"
+# polarity <- "neg"
+# amendment <- "Amm"
+# amendment <- "Nit"
+# amendment <- "GMP"
+# amendment <- "Arg"
+# IS_type <- "noIS"
+# IS_type <- "wIS"
 
-```{r load isotope mzs and areas}
+parametadata_all <- read_csv("metadata/parametadata.csv") %>% 
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
+  mutate(timepoint=factor(timepoint, levels=paste0("T", c(0,1,3,10,26,73))))
+parametadata <- parametadata_all %>%
+  filter(polarity==!!polarity) %>%
+  filter(amendment==!!amendment) %>%
+  filter(IS_type==!!IS_type | samp_type=="Std")
+all_stans <- read_csv("metadata/all_stans.csv")
+
+output_folder <- paste(amendment, polarity, IS_type, "output", sep = "_") %>%
+  paste0("untargeted/", ., "/")
+
 all_stans <- read_csv("metadata/all_stans.csv") %>%
@@
   left_join(all_stans)
+
+all_areas <- map(c("pos", "neg"), function(pol_i){
+  pol_i_areas <- map(c("Amm", "Nit", "GMP", "Arg"), function(amendment_i){
+    file_path <- paste0("G:/My Drive/PARAGON/untargeted/", amendment_i, "_", 
+                        pol_i, "_noIS_output/iso_areas.csv")
+      read_csv(file_path) %>%
+      mutate(amendment=amendment_i)
+  }) %>% bind_rows() %>%
+    mutate(polarity=pol_i)
+}) %>% bind_rows()
+all_rfs <- all_areas %>% 
+  group_by(amendment, polarity) %>%
+  filter(samp_type=="Std") %>%
+  filter(labeled_n_count==0) %>%
+  left_join(all_stans %>% select(compound_name, mix, conc_um, polarity)) %>%
+  mutate(samp_type=str_extract(filename, paste0(mix, "InMatrix|H2OInMatrix"))) %>%
+  drop_na() %>%
+  mutate(samp_type=str_remove(samp_type, "(?<=Mix)\\d")) %>%
+  select(compound_name, samp_type, area, conc_um) %>%
+  pivot_wider(names_from=samp_type, values_from = area, values_fn = mean) %>%
+  mutate(rf=(MixInMatrix-H2OInMatrix)/conc_um) %>%
+  select(compound_name, rf) %>%
+  filter(!is.na(rf)) %>%
+  filter(rf>0)
+all_concs <- all_areas %>%
+  right_join(all_rfs) %>%
+  mutate(uM_in_vial=area/rf) %>%
+  mutate(uM_in_env=(uM_in_vial*(400/1e6))/2) %>%
+  mutate(nM_in_env=uM_in_env*1e3) %>%
+  select(compound_name, filename, amendment, polarity, labeled_n_count, nM_in_env)
+```
+
+# Isotope visualization
+
+## Targeted things
+
+```{r load isotope mzs and areas}
 iso_mzs <- read_csv(paste0(output_folder, "iso_mzs.csv"))
@@
   ggtitle("Surface morning spiked samples")
+```
+
+```{r individual isotope unknown eic}
+unknown_eic <- open_dataset("tmzMLs/pqds") %>%
+  filter(amendment=="Amm") %>%
+  filter(samp_type=="Smp") %>%
+  filter(polarity=="pos") %>%
+  filter(mz%between%pmppm(84.04496, 10) | mz%between%pmppm(85.04196, 10)) %>%
+  dplyr::collect()
+unknown_eic %>%
+  mutate(labeled=ifelse(mz<85, "Unlabeled", "Labeled")) %>%
+  filter(rt%between%c(10, 13)) %>%
+  left_join(parametadata) %>%
+  filter(timepoint=="T26") %>%
+  ggplot() +
+  geom_line(aes(x=rt, y=int, group=filename)) +
+  facet_wrap(~labeled, ncol=1, scales="free_y")
+unknown_eic %>%
+  mutate(labeled=ifelse(mz<85, "Unlabeled", "Labeled")) %>%
+  filter(rt%between%c(10, 13)) %>%
+  left_join(parametadata) %>%
+  arrange(timestamp) %>%
+  arrange(timepoint) %>%
+  mutate(shortname=fct_inorder(shortname)) %>%
+  ggplot() +
+  geom_point(aes(x=rt, y=shortname, group=filename, color=log10(int))) +
+  facet_wrap(~labeled, ncol=1, scales="free_y") +
+  scale_color_viridis_c()
 ```
@@
 
+## Among good peak isotope pairs
+
+```{r make iso_pairs}
+peak_data <- read_csv(paste0(output_folder, "peak_data.csv"))
+iso_pairs <- read_csv(paste0(output_folder, "iso_pairs.csv"))
+
+clean_iso_data <- iso_pairs %>% 
+  select(feature, feature_iso) %>%
+  left_join(peak_data %>% select(feature, filename, into), by="feature") %>%
+  left_join(parametadata, by="filename") %>%
+  filter(samp_type=="Smp") %>%
+  filter(!str_detect(shortname, "Blk")) %>%
+  left_join(peak_data %>% select(feature, filename, into), 
+            by=c(feature_iso="feature", "filename")) %>%
+  mutate(label_ratio=into.y/into.x) %>%
+  group_by(feature, feature_iso, untripl) %>%
+  mutate(label_ratio=ifelse(is.na(label_ratio), median(label_ratio, na.rm=TRUE), label_ratio)) %>%
+  group_by(feature, feature_iso, depth) %>%
+  # mutate(label_ratio=rank(label_ratio)) %>%
+  ungroup() %>%
+  select(feature, feature_iso, shortname, label_ratio, timepoint, timepoint_num, depth, startime)
+```
+
+```{r render all iso_pairs ratios}
+pdf(paste0(output_folder, "iso_untarg_ratios.pdf"), width = 6, height = 6)
+pb <- txtProgressBar(max = nrow(iso_pairs), style = 3)
+for(ind_i in seq_len(nrow(iso_pairs))){
+  setTxtProgressBar(pb, ind_i)
+  iso_gp <- iso_pairs %>% 
+    slice(ind_i) %>%
+    select(feature, feature_iso) %>%
+    left_join(peak_data %>% select(feature, filename, into), by="feature") %>%
+    left_join(parametadata, by="filename") %>%
+    filter(samp_type=="Smp") %>%
+    filter(!str_detect(shortname, "Blk")) %>%
+    left_join(peak_data %>% select(feature, filename, into), by=c(feature_iso="feature", "filename")) %>%
+    mutate(label_ratio=into.y/into.x) %>%
+    group_by(untripl) %>%
+    mutate(label_ratio=ifelse(is.na(label_ratio), median(label_ratio, na.rm=TRUE), label_ratio)) %>%
+    ungroup() %>%
+    arrange(depth, startime, timepoint) %>%
+    mutate(shortname=fct_inorder(shortname)) %>%
+    ggplot() +
+    geom_col(aes(x=label_ratio, y=shortname, fill=timepoint)) +
+    scale_x_log10() +
+    scale_fill_discrete(drop=FALSE) +
+    facet_wrap(~depth, ncol=1, scales="free") +
+    ggtitle(paste0(iso_pairs$feature[ind_i], ": ", 
+                  ifelse(is.na(iso_pairs$compound_name[ind_i]), 
+                         round(iso_pairs$mzmed[ind_i], 5), 
+                         iso_pairs$compound_name[ind_i])))
+  print(iso_gp)
+}
+close(pb)
+dev.off()
+```
+
+```{r make clean_iso_data}
+clean_iso_data %>%
+  filter(feature%in%c("FT0085", "FT0239", "FT0240")) %>%
+  ggplot(aes(x=timepoint_num, y=label_ratio)) +
+  geom_smooth(method=minpack.lm::nlsLM, color="black", formula="y~a*x^b+d", se = FALSE, 
+              method.args=list(start=list(a=1, b=2, d=1))) +
+  geom_point(aes(color=timepoint, shape=startime)) +
+  # geom_boxplot(aes(color=timepoint, group=timepoint_num)) +
+  ggh4x::facet_grid2(depth~feature, scales="free_y", independent = "y")
+```
+
+```{r manual curve log/lin/exp classification}
+shape_classes <- clean_iso_data %>%
+  # filter(depth=="Surf") %>%
+  # filter(feature%in%c("FT0085", "FT0239", "FT0240")) %>%
+  nest(data=-c(feature, depth)) %>%
+  mutate(lin_test=map(data, ~broom::tidy(lm(.x$label_ratio~.x$timepoint_num)))) %>%
+  mutate(exp_test=map(data, ~broom::tidy(lm(log(.x$label_ratio)~.x$timepoint_num)))) %>%
+  mutate(log_test=map(data, ~broom::tidy(lm(exp(.x$label_ratio)~.x$timepoint_num)))) %>%
+  unnest(ends_with("test"), names_sep = "_") %>%
+  filter(lin_test_term==".x$timepoint_num") %>%
+  select(feature, depth, data, ends_with("p.value")) %>%
+  pivot_longer(ends_with("p.value")) %>%
+  group_by(feature, depth) %>%
+  arrange(value) %>%
+  slice(1) %>%
+  ungroup() %>%
+  mutate(name=str_remove(name, "_test_p.value")) %>%
+  select(feature, depth, cluster=name)
+shape_classes %>%
+  left_join(clean_iso_data %>% select(feature, label_ratio, shortname, timepoint, depth)) %>%
+  group_by(feature) %>%
+  mutate(label_ratio=scale(label_ratio)[,1]) %>%
+  ungroup() %>%
+  group_by(cluster, shortname, timepoint, depth) %>%
+  summarise(label_ratio=mean(label_ratio, na.rm=TRUE)) %>%
+  ggplot(aes(x=timepoint, y=label_ratio, color=timepoint)) +
+  geom_hline(yintercept = 0) +
+  geom_boxplot() +
+  facet_grid(depth~cluster)
+shape_classes %>%
+  left_join(iso_anno) %>%
+  filter(!is.na(compound_name)) %>%
+  print(n=Inf)
+```
+
+```{r kmeans isotope curve classification}
+kmeans_obj <- clean_iso_data %>%
+  select(feature, shortname, label_ratio) %>%
+  complete(feature, shortname) %>%
+  group_by(feature) %>%
+  filter(sum(is.na(label_ratio))==0) %>%
+  mutate(label_ratio=scale(label_ratio)[,1]) %>%
+  ungroup() %>%
+  pivot_wider(names_from = shortname, values_from = label_ratio) %>%
+  column_to_rownames("feature") %>%
+  kmeans(centers = 3)
+data.frame(cluster=kmeans_obj$cluster) %>%
+  mutate(cluster=paste0("Cluster ", cluster, ", n=", kmeans_obj$size[cluster])) %>%
+  rownames_to_column("feature") %>%
+  left_join(clean_iso_data %>% select(feature, label_ratio, shortname, timepoint, depth)) %>%
+  group_by(feature) %>%
+  mutate(label_ratio=scale(label_ratio)[,1]) %>%
+  ungroup() %>%
+  # group_by(cluster, shortname, timepoint, depth) %>%
+  # summarise(label_ratio=mean(label_ratio)) %>%
+  ggplot(aes(x=timepoint, y=label_ratio, color=timepoint)) +
+  geom_hline(yintercept = 0) +
+  geom_boxplot() +
+  facet_grid(depth~cluster)
+```
+
+```{r nls curve classification w minpack.lm package}
+library(minpack.lm)
+clean_iso_data %>%
+  complete(feature, shortname) %>%
+  mutate(untripl=str_remove(shortname, "-[A-C]")) %>%
+  group_by(feature) %>%
+  filter(sum(is.na(label_ratio))==0) %>%
+  nest(data=-c(feature, depth)) %>%
+  mutate(nls_test=map(data, ~broom::tidy(nlsLM(label_ratio~a*timepoint_num^b+d, 
+                                             start = list(a=1, b=2, d=1), data=.x,
+                                             control=nls.lm.control(maxiter = 500))))) %>%
+  unnest(nls_test) %>%
+  select(feature, depth, data, term, estimate) %>%
+  pivot_wider(names_from = term, values_from = estimate) %>%
+  # left_join(shape_classes) %>%
+  mutate(a_sign=factor(ifelse(a<0, "neg", "pos"), levels=c("pos", "neg"))) %>%
+  mutate(a=abs(a)) %>%
+  left_join(iso_pairs %>% select(feature, compound_name) %>% drop_na()) %>%
+  mutate(feature=ifelse(is.na(compound_name), feature, compound_name)) %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_vline(xintercept = 1) +
+  geom_text(aes(x=b, y=a, label=feature, color=depth), hjust=0.2) +
+  annotate("label", x=1, y=Inf, hjust=0, vjust=1, label="Exponential", label.r=unit(0, "in")) +
+  annotate("label", x=1, y=Inf, hjust=1, vjust=1, label="Logarithmic", label.r=unit(0, "in")) +
+  scale_y_log10() +
+  facet_wrap(~a_sign, ncol=1, scales = "free_y")
+ggsave(filename = paste0(output_folder, "nls_coefs.pdf"), device = "pdf", 
+       width = 10, height = 10, units = "in")
+
+clean_iso_data %>%
+  left_join(iso_pairs %>% select(feature, compound_name) %>% drop_na()) %>%
+  mutate(feature=ifelse(is.na(compound_name), feature, compound_name)) %>%
+  ggplot(aes(x=timepoint_num, y=label_ratio)) +
+  geom_hline(yintercept = 0) +
+  geom_smooth(method="nlsLM", color="black", formula="y~a*x^b+d", se = FALSE, 
+              method.args=list(start=list(a=1, b=2, d=1))) +
+  geom_point(aes(color=timepoint, shape=startime)) +
+  ggh4x::facet_grid2(depth~feature, scales="free_y", independent = "y")
+ggsave(filename = paste0(output_folder, "nls_fits.pdf"), device = "pdf",
+       width = 80, height = 4, units = "in", limitsize = FALSE)
+```
+
+## For all untargeted compounds from isotope envelopes
+
+```{r}
+raw_iso_envelopes <- read_csv(paste0(output_folder, "raw_iso_envelopes.csv"))
+
+label_frac <- raw_iso_envelopes %>%
+  group_by(feature, iso_count) %>%
+  mutate(med_corr=median(corr, na.rm=TRUE)) %>%
+  filter(med_corr>0.8) %>%
+  select(feature, filename, iso_count, area) %>%
+  # ggplot() + geom_boxplot(aes(x=iso_count, y=area, group=iso_count))
+  mutate(label_type=ifelse(iso_count==0, "unlabeled", "labeled")) %>%
+  select(-iso_count) %>%
+  group_by(feature, filename, label_type) %>%
+  summarise(area=sum(area, na.rm=TRUE)) %>%
+  pivot_wider(names_from=label_type, values_from = area) %>%
+  mutate(labeled_ratio=labeled/(unlabeled+labeled)) %>%
+  left_join(parametadata) %>%
+  drop_na()
+label_frac %>%
+  # filter(feature=="FT0915") %>%
+  # filter(feature=="FT1705") %>%
+  filter(feature=="FT0239") %>%
+  ggplot() + 
+  geom_hline(yintercept = 0) +
+  geom_vline(xintercept = 0) +
+  geom_abline(slope = 1, intercept = 0) +
+  geom_point(aes(x=labeled, y=unlabeled, color=timepoint, shape=startime)) +
+  facet_wrap(~depth, ncol=1, scales="free")
+label_frac %>%
+  filter(feature=="FT0239") %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(x=timepoint, y=labeled_ratio, color=timepoint, shape=startime)) +
+  facet_wrap(~depth, ncol=1, scales="free")
+```
+
 # Multivariate approaches
@@
 
+```{r kmeans clustering of isotope signals}
+
+```
+
 # Univariate followup
+
+```{r}
+amend_plot <- function(cmpd_name_i){
+  amend_comp_concs %>%
+    filter(compound_name==cmpd_name_i) %>%
+    ggplot() +
+    geom_col(aes(x=timepoint, y=nM_in_env, fill=labeled_n_count), color="black") +
+    facet_grid2(depth~amendment, scales="free_y", space = "free", 
+                strip = strip_themed(
+                  background_x=elem_list_rect(fill = scales::hue_pal()(4)[c(1,4,3,2)])
+                )) +
+    scale_fill_gradient(low = "grey20", high = "grey80") +
+    scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
+    scale_x_discrete(breaks=c("T0", "T1", "T3", "T10", "T26", "T73"), labels=c("T0", 1,3,10,26, 73)) +
+    theme_bw() +
+    theme(text = element_text(size=14), legend.text = element_text(size=18), 
+          legend.position = "none", strip.text = element_text(size=18)) +
+    labs(x="Timepoint (hours after spike)", 
+         y="Concentration (nM)\nin each 15N label",
+         title = cmpd_name_i)
+}
+```

### 2023-11-07 metadata/parametadata_control.Rmd
Adding additional metadata columns

@@
   mutate(timepoint=str_extract(shortname, "T\\d+")) %>%
-  mutate(tripl=str_extract(shortname, "[A-C]$"))
+  mutate(timepoint_num=as.numeric(str_remove(timepoint, "T"))) %>%
+  mutate(tripl=str_extract(shortname, "[A-C]$")) %>%
+  mutate(untripl=str_remove(shortname, "-[A-C]$"))
 

### 2023-11-07 untargeted/paruntargeted_control.Rmd
Restructured isotope detection

@@
 
-# polarity <- "pos"
-polarity <- "neg"
+polarity <- "pos"
+# polarity <- "neg"
 # amendment <- "Amm"
 # amendment <- "Nit"
-# amendment <- "GMP"
-amendment <- "Arg"
+amendment <- "GMP"
+# amendment <- "Arg"
 IS_type <- "noIS"
@@
 }
-trapz <- function(x, y) {
+trapz <- function(x, y, na.rm=TRUE) {
   y <- y[order(x)]
@@
   n <- 2*m
-  p1 <- sum(xp[1:(n-1)]*yp[2:n]) + xp[n]*yp[1]
-  p2 <- sum(xp[2:n]*yp[1:(n-1)]) + xp[1]*yp[n]
+  p1 <- sum(xp[1:(n-1)]*yp[2:n], na.rm = na.rm) + xp[n]*yp[1]
+  p2 <- sum(xp[2:n]*yp[1:(n-1)], na.rm = na.rm) + xp[1]*yp[n]
 
@@
   select(feature, feat_mzmed, feat_rtmed, feat_npeaks, filename,
-         mz, mzmin, mzmax, rt, rtmin, rtmax, into, sn, beta_cor, beta_snr)
+         mz, mzmin, mzmax, rt, rtmin, rtmax, into, sn, beta_cor, beta_snr) %>%
+  arrange(desc(into)) %>%
+  group_by(feature, filename) %>%
+  slice(1) %>%
+  ungroup()
 output_csv_write(peak_data)
@@
                       Dexpanthenol="FT2426", Cytidine="FT2792",
-                      Mycogly='FT2814', Inosine="FT2960", Guanosine="FT3015")
+                      Mycogly='FT2814', Inosine="FT2960", Guanosine="FT3015",
+                      Hydroxyproline="FT1237")
   } else if(polarity=="neg"){
@@
   if(polarity=="pos"){
-    rescue_feats <- c("FT1422", "FT1580", "FT1585", "FT1603", "FT3030")
+    rescue_feats <- c("FT1422", "FT1580", "FT1585", "FT1603", "FT3030", "FT1569")
   } else if(polarity=="neg"){
@@
 
+```{r find isotope pairs among good peaks}
+best_feat_data <- read_csv(paste0(output_folder, "best_feat_data.csv"))
+peak_data <- read_csv(paste0(output_folder, "peak_data.csv"))
+manual_anno <- readxl::read_excel("untargeted/manual_annotations.xlsx") %>%
+  filter(polarity==!!polarity) %>%
+  filter(amendment==!!amendment) %>%
+  select(feature, compound_name)
+iso_mzs <- read_csv(paste0(output_folder, "iso_mzs.csv"))
+
+single_lab_matching <- manual_anno %>%
+  left_join(iso_pairs) %>%
+  mutate(compound_name=paste(compound_name, "x 15N1")) %>%
+  drop_na() %>%
+  select(feature=feature_iso, compound_name)
+double_lab_matching <- single_lab_matching %>%
+  left_join(iso_pairs %>% select(-compound_name)) %>%
+  mutate(compound_name=str_replace(compound_name, "x 15N1", "x 15N2")) %>%
+  drop_na() %>%
+  select(feature=feature_iso, compound_name)
+treble_lab_matching <- double_lab_matching %>%
+  left_join(iso_pairs %>% select(-compound_name)) %>%
+  mutate(compound_name=str_replace(compound_name, "x 15N2", "x 15N3")) %>%
+  drop_na() %>%
+  select(feature=feature_iso, compound_name)
+quad_lab_matching <- treble_lab_matching %>%
+  left_join(iso_pairs %>% select(-compound_name)) %>%
+  mutate(compound_name=str_replace(compound_name, "x 15N3", "x 15N4")) %>%
+  drop_na() %>%
+  select(feature=feature_iso, compound_name)
+quint_lab_matching <- quad_lab_matching %>%
+  left_join(iso_pairs %>% select(-compound_name)) %>%
+  mutate(compound_name=str_replace(compound_name, "x 15N4", "x 15N5")) %>%
+  drop_na() %>%
+  select(feature=feature_iso, compound_name)
+iso_anno <- manual_anno %>%
+  bind_rows(single_lab_matching, double_lab_matching, treble_lab_matching, 
+            quad_lab_matching, quint_lab_matching)
+
+carbon_pairs <- best_feat_data %>%
+  select(feature, mzmed) %>%
+  pull(mzmed, feature) %>%
+  dist() %>%
+  as.matrix() %>%
+  {which(abs(.-1.003355)<0.001, arr.ind = TRUE)} %>%
+  as.data.frame() %>%
+  filter(row<col) %>%
+  left_join(best_feat_data %>% select(feature, mzmed, rtmed) %>% mutate(row=row_number()),
+            by="row") %>%
+  left_join(best_feat_data %>% select(feature, rtmed, mzmed) %>% mutate(col=row_number()),
+            by="col", suffix=c("", "_iso")) %>%
+  filter(abs(rtmed_iso-rtmed)<0.1) %>%
+  distinct(feature=feature_iso)
+
+iso_pairs <- best_feat_data %>%
+  select(feature, mzmed) %>%
+  pull(mzmed, feature) %>%
+  dist() %>%
+  as.matrix() %>%
+  {which(abs(.-0.997035)<0.001, arr.ind = TRUE)} %>%
+  as.data.frame() %>%
+  filter(row<col) %>%
+  left_join(best_feat_data %>% select(feature, mzmed, rtmed) %>% mutate(row=row_number()),
+            by="row") %>%
+  left_join(best_feat_data %>% select(feature, rtmed, mzmed) %>% mutate(col=row_number()),
+            by="col", suffix=c("", "_iso")) %>%
+  filter(abs(rtmed_iso-rtmed)<0.1) %>%
+  select(-row, -col) %>%
+  left_join(iso_anno) %>%
+  anti_join(carbon_pairs, by="feature")
+
+output_csv_write(iso_pairs)
+
+# (select_iso <- iso_pairs %>% filter(feature=="FT2466"))
+# FT0085_eics <- open_dataset("tmzMLs/pqds") %>%
+#   filter(samp_type=="Smp") %>%
+#   filter(amendment==!!amendment) %>%
+#   filter(polarity==!!polarity) %>%
+#   filter(mz%between%pmppm(select_iso$mzmed) | mz%between%pmppm(select_iso$mzmed_iso)) %>%
+#   dplyr::collect()
+# FT0085_eics %>%
+#   filter(rt%between%(select_iso$rtmed+c(-1, 1))) %>%
+#   mutate(iso=ifelse(mz<mean(mz), "Unlabeled", "Labeled")) %>%
+#   left_join(parametadata) %>%
+#   ggplot() +
+#   geom_line(aes(x=rt, y=int, group=filename, color=timepoint)) +
+#   facet_wrap(~iso, ncol=1, scales="free_y")
+```
+
+```{r find isotope envelopes for all untarg}
+best_feat_data <- read_csv(paste0(output_folder, "best_feat_data.csv"))
+peak_data <- read_csv(paste0(output_folder, "peak_data.csv"))
+
+library(furrr)
+plan(multisession, workers = 6)
+n_psb_isos <- 6
+n_ok_nas <- 10
+
+raw_iso_envelopes <- parametadata %>%
+  filter(samp_type=="Smp") %>%
+  filter(!str_detect(filename, "Blk")) %>%
+  arrange(timepoint, depth, startime) %>%
+  pull(filename) %>%
+  future_map(function(filename_i){
+    file_msdata <- open_dataset("tmzMLs/pqds") %>%
+      filter(filename==filename_i) %>%
+      dplyr::collect()
+    pmap(best_feat_data, function(...){
+      row_data <- list(...)
+      # row_data <- filter(best_feat_data, feature=="FT0769")
+      # row_data <- slice(best_feat_data, 326)
+      rt_bounds <- peak_data %>% 
+        filter(feature==row_data$feature) %>%
+        filter(filename==filename_i) %>%
+        reframe(rtmin=rtmin/60, rtmax=rtmax/60)
+      if(nrow(rt_bounds)==0)return(NULL)
+      file_rt_eic <- file_msdata %>%
+        .[rt%between%c(rt_bounds$rtmin, rt_bounds$rtmax)] %>%
+        .[mz%between%c(row_data$mzmed-(n_psb_isos+2), row_data$mzmed+(n_psb_isos+2))]
+
+      eic_df <- c(rev(seq(row_data$mzmed, row_data$mzmed-n_psb_isos, -0.997035)[-1]), 
+                   seq(row_data$mzmed, row_data$mzmed+n_psb_isos, 0.997035)) %>%
+        sapply(pmppm) %>%
+        t() %>%
+        as.data.frame() %>%
+        set_names(c("mzmin", "mzmax")) %>%
+        mutate(iso_count=seq(-n_psb_isos, n_psb_isos)) %>%
+        left_join(file_rt_eic, by=join_by(mzmin<mz, mzmax>mz)) %>%
+        drop_na(rt) %>%
+        complete(iso_count, rt) %>%
+        group_by(iso_count, rt) %>%
+        summarise(int=max(int), .groups = "drop_last") %>%
+        filter(sum(!is.na(int))>=n_ok_nas)
+      eic_mat <- eic_df %>%
+        pivot_wider(names_from = iso_count, values_from = int) %>%
+        arrange(rt) %>%
+        column_to_rownames("rt")
+      if(nrow(eic_mat)==0)return(NULL)
+      # scale(eic_mat) %>% matplot(type="l")
+      # pairs(eic_mat)
+      iso_area_df <- eic_df %>%
+        drop_na() %>%
+        group_by(iso_count) %>%
+        summarise(area=trapz(rt, int))
+      eic_mat %>%
+        cor(use = "pairwise.complete") %>%
+        as.data.frame() %>%
+        rownames_to_column("init_count") %>%
+        pivot_longer(-init_count, names_to = "iso_count", values_to = "corr") %>%
+        mutate(init_count=as.numeric(init_count), iso_count=as.numeric(iso_count)) %>%
+        filter(init_count==0) %>%
+        select(iso_count, corr) %>%
+        mutate(feature=row_data$feature) %>%
+        left_join(iso_area_df, by="iso_count")
+    }) %>%
+      bind_rows() %>% 
+      mutate(filename=filename_i) %>%
+      select(feature, filename, iso_count, corr, area)
+  }, .progress = TRUE) %>%
+  bind_rows()
+output_csv_write(raw_iso_envelopes)
+```
+
 # BMIS and quant
@@
 ```
+
+# Export final peak areas

### 2023-11-07 untargeted/paruntargeted_sandbox.Rmd
Fixed names in manual_typing

@@
   filter(IS_type==!!IS_type | samp_type=="Std")
-all_stans <- read_csv("metadata/all_stans.csv")
+manual_typing <- readxl::read_excel("metadata/manual_typing.xlsx") %>%
+  mutate(manual_type=ifelse(is.na(manual_type), "Others", manual_type)) %>%
+  add_row(compound_name="Thymidine", manual_type="Nucleobases +")
+all_stans <- read_csv("metadata/all_stans.csv") %>%
+  left_join(manual_typing)
 

### 2023-11-10 untargeted/paruntargeted_sandbox.Rmd
Adding contour plot fitting and kmeans/RF things

--- 2023-11-07 untargeted/paruntargeted_sandbox.Rmd
@@
 
-# polarity <- "pos"
+polarity <- "pos"
 # polarity <- "neg"
-# amendment <- "Amm"
+amendment <- "Amm"
 # amendment <- "Nit"
@@
 # amendment <- "Arg"
-# IS_type <- "noIS"
+IS_type <- "noIS"
 # IS_type <- "wIS"
@@
   filter(IS_type==!!IS_type | samp_type=="Std")
-all_stans <- read_csv("metadata/all_stans.csv")
+manual_typing <- readxl::read_excel("metadata/manual_typing.xlsx") %>%
+  mutate(manual_type=ifelse(is.na(manual_type), "Other", manual_type)) %>%
+  add_row(compound_name="Thymidine", manual_type="Nucleobases +")
+all_stans <- read_csv("metadata/all_stans.csv") %>%
+  left_join(manual_typing)
 
@@
 
+```{r logistic curve fitting contour plot and curve fitting}
+glm_coef_data <- all_concs %>%
+  left_join(parametadata_all) %>%
+  filter(amendment=="Amm") %>%
+  filter(samp_type=="Smp") %>%
+  filter(!str_detect(filename, "Blk")) %>%
+  mutate(label_type=ifelse(labeled_n_count==0, "Unlabeled", "Labeled")) %>%
+  select(compound_name, filename, label_type, nM_in_env, timepoint_num, depth, startime) %>%
+  pivot_wider(names_from = label_type, values_from=nM_in_env, values_fn = sum) %>%
+  mutate(frac_labeled=Labeled/(Labeled+Unlabeled)) %>%
+  nest(data=-c(compound_name, depth)) %>%
+  mutate(nls_test=map(data, ~broom::tidy(glm(.x$frac_labeled~.x$timepoint_num, family=quasibinomial)))) %>%
+  unnest(nls_test) %>%
+  select(compound_name, depth, data, term, estimate) %>%
+  pivot_wider(names_from = term, values_from = estimate) %>%
+  dplyr::rename(intercept=`(Intercept)`, slope=`.x$timepoint_num`) %>%
+  mutate(intercept=1-(exp(-intercept)/(1+exp(-intercept)))) %>%
+  left_join(manual_typing)
+```
+
 ```{r kmeans clustering of isotope signals}
+frac_labeled_df <- all_areas %>%
+  left_join(parametadata_all) %>%
+  filter(amendment=="Amm") %>%
+  filter(samp_type=="Smp") %>%
+  filter(!str_detect(filename, "Blk")) %>%
+  filter(polarity=="pos") %>%
+  mutate(label_type=ifelse(labeled_n_count==0, "Unlabeled", "Labeled")) %>%
+  select(compound_name, shortname, label_type, area, timepoint_num, depth, startime) %>%
+  pivot_wider(names_from = label_type, values_from=area, values_fn = sum) %>%
+  mutate(frac_labeled=Labeled/(Labeled+Unlabeled)) %>%
+  ungroup() %>%
+  arrange(depth, startime, timepoint_num) %>%
+  mutate(shortname=fct_inorder(shortname))
 
+frac_labeled_df %>%
+  filter(compound_name=="L-Glutamine") %>%
+  ggplot() +
+  geom_point(aes(x=timepoint_num, y=frac_labeled)) +
+  facet_grid(depth~startime)
+
+frac_mat <- frac_labeled_df %>%
+  select(compound_name, shortname, frac_labeled) %>%
+  pivot_wider(names_from=shortname, values_from=frac_labeled) %>%
+  mutate(across(everything(), function(x)ifelse(is.na(x), 0, x))) %>%
+  # filter(compound_name=="L-Glutamine") %>%
+  # filter(is.na(`210916_Smp_Amm-Deep-Eve-T0-A-noIS-pos.mzML`)) %>%
+  column_to_rownames("compound_name")
+
+# heatmaply::heatmaply(frac_mat, Colv=FALSE)
+kmeans_obj <- kmeans(frac_mat, centers = 4)
+mapping <- `names<-`(1:6, order(kmeans_obj$size, decreasing = TRUE))
+old_names <- names(kmeans_obj$cluster)
+kmeans_obj$cluster <- mapping[as.character(kmeans_obj$cluster)]
+names(kmeans_obj$cluster) <- old_names
+kmeans_obj$size <- sort(kmeans_obj$size, decreasing = TRUE)
+table(kmeans_obj$cluster)
+
+data.frame(cluster=kmeans_obj$cluster) %>%
+  # split(.$cluster)
+  mutate(cluster=paste0("Cluster ", cluster, " n=", kmeans_obj$size[cluster])) %>%
+  rownames_to_column("compound_name") %>%
+  left_join(frac_labeled_df) %>%
+  # ggplot() +
+  # geom_boxplot(aes(x=timepoint_num, y=frac_labeled, color=startime, group=interaction(timepoint_num, startime))) +
+  # facet_grid(depth~cluster)
+  group_by(cluster, shortname) %>%
+  summarise(mean_labeled=mean(frac_labeled, na.rm=TRUE)) %>%
+  left_join(parametadata) %>%
+  ggplot() +
+  geom_point(aes(x=timepoint_num, y=mean_labeled, color=startime)) +
+  facet_grid(depth~cluster)
+
+library(randomForest)
+library(rpart)
+# library(rpart.plot)
+
+(rfmodel <- randomForest(frac_mat, y=factor(kmeans_obj$cluster), ntree = 5000, 
+                        importance=TRUE, proximity = TRUE, keep.forest = TRUE))
+# varImpPlot(rfmodel, n.var = 60)
 ```

### 2024-05-13 metadata/parametadata_control.Rmd
Organizing to factor by depth and startime

@@
 
-magick::image_read("http://scope.soest.hawaii.edu/data/scope2021/Paragon1CTDchlp.pdf", density = 300) %>%
-  magick::image_trim() %>%
-  magick::image_write("metadata/raw_data/chl_pdf_url.png", format = "png")
+# magick::image_read("http://scope.soest.hawaii.edu/data/scope2021/Paragon1CTDchlp.pdf", density = 300) %>%
+#   magick::image_trim() %>%
+#   magick::image_write("metadata/raw_data/chl_pdf_url.png", format = "png")
 ```
@@
 
-```{r visualize}
+```{r visualize, eval=FALSE}
 my_casts <- c(4, 5, 9, 11, 13, 14, 16, 18, 25, 27, 31, 32, 34, 36, 39, 41)
@@
 ```{r write out}
-write_csv(parametadata_all, file = "metadata/parametadata.csv")
+parametadata_all %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
+  arrange(polarity, amendment, depth, startime, timepoint_num, tripl) %>%
+  write_csv(file = "metadata/parametadata.csv")
 ```

### 2024-05-13 metadata/parametadata_control.Rmd
Adding Blk as separate samp_type

@@
   mutate(shortname=str_remove(shortname, ".mzML$")) %>%
-  mutate(samp_type="Smp") %>%
+  mutate(samp_type=ifelse(str_detect(shortname, "Blk"), "Blk", "Smp")) %>%
   mutate(IS_type=str_extract(filepath, "wIS|noIS")) %>%

### 2024-07-03 tmzMLs/arrowscript.R
Recompiling pqds to be in m/z order and use blanks

@@ -3,7 +3,7 @@ library(RaMS)
 library(arrow)
 
 parametadata <- fread("metadata/parametadata.csv")[
-  IS_type=="noIS",c("filepath", "filename", "samp_type", "amendment", "polarity")
+  IS_type=="noIS"|is.na(IS_type),c("filepath", "filename", "samp_type", "amendment", "polarity")
 ]
 parametadata[, .N, by=c("samp_type", "amendment", "polarity")]
 
@@ -12,6 +12,7 @@ for(samp_type_i in unique(parametadata$samp_type)){
     for(polarity_i in unique(parametadata$polarity)){
       ms_files <- parametadata[samp_type==samp_type_i][amendment==amendment_i][polarity==polarity_i]$filepath
       # ms_files <- head(ms_files, 3)
+      if(length(ms_files)==0)next
       msdata <- grabMSdata(ms_files, grab_what = "MS1")
       parq_dir <- paste0("tmzMLs/pqds",
                          "/samp_type=", samp_type_i, 
@@ -20,7 +21,7 @@ for(samp_type_i in unique(parametadata$samp_type)){
       if(!dir.exists(parq_dir))dir.create(parq_dir, recursive = TRUE)
       parq_name <- paste0(parq_dir, "/part-0.parquet")
       print(parq_name)
-      write_parquet(msdata$MS1, parq_name)
+      write_parquet(msdata$MS1[order(mz)], parq_name)
     }
   }
 }
@@ -36,3 +37,27 @@ for(samp_type_i in unique(parametadata$samp_type)){
 #   left_join(parametadata) %>%
 #   ggplot() +
 #   geom_line(aes(x=rt, y=int, group=filename, color=samp_type))
+
+
+
+
+# library(DBI)
+# ms_files <- parametadata[polarity=="pos"]$filepath
+# ms_file_list <- split(ms_files, 1:10)
+# msduck <- dbConnect(duckdb::duckdb(), "tmzMLs/duckdb")
+# dbWriteTable(msduck, "MS1_pos", grabMSdata(ms_file_list[[1]], grab_what = "MS1")$MS1[order(mz)])
+# pbapply::pblapply(ms_file_list[-1], function(files_i){
+#   ms1_data <- grabMSdata(files_i, grab_what = "MS1", verbosity = 0)$MS1[order(mz)]
[…]

### 2024-07-03 metadata/parametadata_control.Rmd
Now only filling amendment bc IS_type should be NA

@@
   mutate(amendment=str_extract(shortname, "Nit|Amm|Arg|GMP")) %>%
-  fill(IS_type, amendment, .direction = "down")
+  fill(amendment)
 

### 2024-11-04 metadata/parametadata_control.Rmd
Adding n_C and n_N columns to all_stans

@@
   mutate(polarity=ifelse(z<0, "neg", "pos")) %>%
-  arrange(mz, rt)
+  arrange(mz, rt) %>%
+  mutate(n_C=as.numeric(str_extract(formula, "(?<=C)\\d+"))) %>%
+  mutate(n_C=case_when(
+    !str_detect(formula, "C")~0,
+    str_detect(formula, "C") & is.na(n_C)~1,
+    TRUE~n_C
+  )) %>%
+  mutate(n_N=as.numeric(str_extract(formula, "(?<=N)\\d+"))) %>%
+  mutate(n_N=case_when(
+    !str_detect(formula, "C")~0,
+    str_detect(formula, "C") & is.na(n_N)~1,
+    TRUE~n_N
+  ))
 write_csv(all_stans, "metadata/all_stans.csv")

### 2024-11-04 tmzMLs/arrowscript.R
Adding neg mode pqds and ignoring

@@ -3,7 +3,7 @@ library(RaMS)
 library(arrow)
 
 parametadata <- fread("metadata/parametadata.csv")[
-  IS_type=="noIS"|is.na(IS_type),c("filepath", "filename", "samp_type", "amendment", "polarity")
+  IS_type=="wIS"|is.na(IS_type),c("filepath", "filename", "samp_type", "amendment", "polarity")
 ]
 parametadata[, .N, by=c("samp_type", "amendment", "polarity")]
 
@@ -14,7 +14,7 @@ for(samp_type_i in unique(parametadata$samp_type)){
       # ms_files <- head(ms_files, 3)
       if(length(ms_files)==0)next
       msdata <- grabMSdata(ms_files, grab_what = "MS1")
-      parq_dir <- paste0("tmzMLs/pqds",
+      parq_dir <- paste0("tmzMLs/pqds_wIS",
                          "/samp_type=", samp_type_i, 
                          "/amendment=", amendment_i, 
                          "/polarity=", polarity_i)

### 2024-11-04 metadata/parametadata_control.Rmd
Adding cast_datetime to metadata (in UTC)

@@
 ) %>%
-  mutate(station=2)
+  mutate(station=2) %>%
+  left_join(para_core %>% distinct(station, cast, cast_datetime=time))
 

### 2025-05-05 run_analysis.R
Cleaning repo of outdated scripts

@@ -1,162 +1,0 @@
-
-library(tidyverse)
-
-# Generate mzMLs from .raw via msconvert ----
-if(!file.exists("mzMLs/pos/220729_Smp_Arg-Surf-Morn-T73-C-wIS.mzML")){
-  folder_paths <- c(
-    "Z:/1_QEdata/2021/2021_Will_PARAGON/210813_Will_NIT-Expt_PARAGON_HILIC",
-    "Z:/1_QEdata/2021/2021_Will_PARAGON/210916_Will_Amm-Expt_PARAGON_HILIC",
-    "Z:/1_QEdata/2022/220715_Will_GMP-Expt_PARAGON_HILIC",
-    "Z:/1_QEdata/2022/220729_Will_Arg-Expt_PARAGON_HILIC"
-  )
-  sapply(folder_paths, function(path_i){
-    sapply(c("positive", "negative"), function(polarity){
-      message(paste("Converting", basename(path_i), polarity))
-      short_pol <- gsub("[ia]tive", "", polarity)
-      match_pattern <- paste0("wIS.raw|noIS-", short_pol, ".*.raw|_Std_.*raw|wIS-(Full|Half)")
-      files_to_convert <- list.files(path_i, pattern = match_pattern, full.names = TRUE)
-      mscommand <- paste(
-        "msconvert",
-        paste(files_to_convert, collapse = " "),
-        "--mzML",
-        paste0('-o mzMLs/', short_pol),
-        '--filter "peakPicking true 1-"',
-        paste('--filter "polarity', polarity, '"')
-      )
-      system(mscommand)
-      
-      msmsmatch_pattern <- paste0("_DDA", short_pol, ".*.raw")
-      msms_to_convert <- list.files(path_i, pattern = msmsmatch_pattern, full.names = TRUE)
-      msmscommand <- paste(
-        "msconvert",
-        paste(msms_to_convert, collapse = " "),
-        "--mzML",
-        paste0('-o mzMLs/', short_pol, "/MSMS"),
-        '--filter "peakPicking true 1-"',
-        paste('--filter "polarity', polarity, '"')
-      )
-      system(msmscommand)
-    })
-  })
-}
-
-
-
-# Generate metadata ----
-if(!file.exists("metadata/parametadata.csv")){
-  knitr::knit(input = "metadata/parametadata_control.Rmd", 
-              output = "metadata/parametadata_control.R", 
-              tangle = TRUE)
-}
-
-
[…]

### 2025-05-05 untargeted/paruntargeted_sandbox.Rmd
Updating paruntargeted sandbox, needs review

@@
        fill="Number of\n15N atoms in\nisotopologue")
+```
+
+```{r nested geom_tile plot overview multicompound comparison}
+library(ggh4x)
+iso_areas %>%
+  filter(!str_detect(filename, "Blk")) %>%
+  filter(compound_name%in%c("Guanine", "Guanosine", "Adenine", "Adenosine",
+                            "Cytosine", "Cytidine", "Thymine", "Thymidine",
+                            "Uracil", "Uridine", "Hypoxanthine", "Inosine")) %>%
+  left_join(parametadata_all) %>%
+  group_by(compound_name, timepoint, depth, startime, tripl) %>%
+  mutate(area=area/sum(area)) %>%
+  ggplot() +
+  geom_tile(aes(x=tripl, y=labeled_n_count, fill=area)) +
+  ggh4x::facet_nested(compound_name~depth+startime+timepoint, space = "free_y", scales = "free_y") +
+  scale_fill_viridis_c(labels=scales::label_percent()) +
+  labs(y="Number of labeled nitrogens", fill="Fraction\nof total\npeak area") +
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
+        axis.ticks.x=element_blank())
+
+iso_areas %>%
+  filter(!str_detect(filename, "Blk")) %>%
+  filter(compound_name%in%c("Guanine", "Guanosine", "Adenine", "Adenosine")) %>%
+  left_join(parametadata_all) %>%
+  ggplot() +
+  geom_boxplot(aes(x=factor(labeled_n_count), y=area, color=timepoint)) +
+  ggh4x::facet_nested(compound_name~depth+startime, scales="free_y", independent = "y") +
+  labs(x="Number of labeled nitrogens", y="Peak area", color="Timepoint") +
+  theme_bw()
+
+iso_areas %>%
+  filter(!str_detect(filename, "Blk")) %>%
+  filter(compound_name%in%c("Guanine", "Guanosine", "Adenine", "Adenosine")) %>%
+  left_join(parametadata_all) %>%
+  group_by(compound_name, depth, startime, timepoint, labeled_n_count) %>%
+  summarise(med_area=median(area), .groups = "drop_last") %>%
+  mutate(med_area=med_area/sum(med_area)) %>%
+  mutate(labeled_n_count=factor(labeled_n_count, levels=5:0)) %>%
+  ggplot() +
+  geom_col(aes(x=timepoint, y=med_area, fill=labeled_n_count), position = "stack") +
+  facet_nested(compound_name~depth+startime) +
+  scale_y_continuous(labels=scales::percent_format()) +
+  scale_fill_viridis_d(direction = -1) +
+  labs(x="Timepoint", y="Isotope distribution", fill="Number of\nlabeled\nnitrogens")
 ```
@@
 ```
+
+```{r}
+parametadata <- read_csv("metadata/parametadata.csv") %>% 
+  mutate(timepoint=fct_inorder(timepoint)) %>%
+  mutate(depth=fct_inorder(depth)) %>%
+  mutate(startime=fct_inorder(startime))
+amm_iso_areas <- read_csv("untargeted/Amm_pos_noIS_output/iso_areas.csv") %>%
+  select(-samp_type)
+
+amm_iso_areas %>%
+  filter(compound_name=="L-Glutamine") %>% # Or l-glutamine, or carnitine,
+  left_join(parametadata) %>%
+  filter(samp_type=="Smp") %>%
+  ggplot() +
+  # geom_col(aes(x=tripl, y=area, fill=labeled_n_count), position = "fill") +
+  geom_col(aes(x=tripl, y=area, fill=labeled_n_count)) +
+  facet_grid(depth~startime+timepoint, scales = "free_y") +
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), 
+        axis.ticks.x=element_blank())
+
+```
+
+Demo written up for PARAGON group meeting:
+
+```{r}
+library(tidyverse)
+amm_iso_areas <- read_csv("untargeted/Amm_pos_noIS_output/iso_areas.csv") %>%
+  mutate(amend="Amm")
+nit_iso_areas <- read_csv("untargeted/Nit_pos_noIS_output/iso_areas.csv") %>%
+  mutate(amend="Nit")
+parametadata <- read_csv("metadata/parametadata.csv")
+
+library(ggh4x)
+bind_rows(amm_iso_areas, nit_iso_areas) %>%
+  filter(compound_name%in%c("L-Glutamine", "Guanine", "Shinorine?")) %>%
+  mutate(compound_name=factor(compound_name, levels=c("L-Glutamine", "Guanine", "Shinorine?"))) %>%
+  filter(samp_type=="Smp") %>%
+  mutate(label_boolean=ifelse(labeled_n_count>0, "Labeled", "Unlabeled")) %>%
+  left_join(parametadata) %>%
+  # Shinorine isn't found in deep samples
+  filter(!(compound_name=="Shinorine?" & depth=="Deep")) %>%
+  # Add factors to order things correctly
+  arrange(timepoint_num) %>%
+  mutate(timepoint=fct_inorder(timepoint)) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"), labels=c("25m", "175m"))) %>%
+  # Calculate label fractions
+  group_by(amend, compound_name, timepoint, timepoint_num, label_boolean, labeled_n_count, tripl, depth, startime) %>%
+  summarize(area=median(area)) %>%
+  group_by(amend, compound_name, timepoint, timepoint_num, tripl, depth, startime) %>%
+  mutate(area=area/sum(area)) %>%
+  group_by(amend, compound_name, timepoint, timepoint_num, label_boolean, tripl, depth, startime) %>%
+  summarise(area=sum(area)) %>%
+  drop_na() %>%
+  ggplot() +
+  geom_col(aes(x=tripl, y=area, fill=label_boolean), position = "fill", width = 1, color="black") +
+  facet_nested(compound_name+depth~amend+startime+timepoint) +
+  scale_y_continuous(labels = scales::label_percent(), expand = expansion()) +
+  scale_x_discrete(expand = expansion()) +
+  scale_fill_manual(breaks=c("Unlabeled", "Labeled"), values = c("#a6cde3", "#1d7cb4")) +
+  theme_bw() +
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), 
+        axis.ticks.x = element_blank(), legend.position = "top") +
+  labs(y="Fraction labeled", fill=NULL)
+
+bind_rows(amm_iso_areas, nit_iso_areas) %>%
+  filter(compound_name%in%c("L-Glutamine", "Guanine")) %>%
+  mutate(compound_name=factor(compound_name, levels=c("L-Glutamine", "Guanine"))) %>%
+  filter(samp_type=="Smp") %>%
+  mutate(label_boolean=ifelse(labeled_n_count>0, "Labeled", "Unlabeled")) %>%
+  left_join(parametadata) %>%
+  filter(startime=="Morn") %>%
+  filter(depth=="Surf") %>%
+  arrange(timepoint_num) %>%
+  mutate(timepoint=fct_inorder(timepoint)) %>%
+  group_by(amend, compound_name, timepoint, timepoint_num, label_boolean, labeled_n_count, tripl) %>%
+  summarize(area=median(area)) %>%
+  group_by(amend, compound_name, timepoint, timepoint_num, tripl) %>%
+  mutate(area=area/sum(area)) %>%
+  group_by(amend, compound_name, timepoint, timepoint_num, label_boolean, tripl) %>%
+  summarise(area=sum(area)) %>%
+  ggplot() +
+  geom_col(aes(x=tripl, y=area, fill=label_boolean), position = "fill", width = 1, color="black") +
+  facet_nested(compound_name~amend+timepoint) +
+  scale_y_continuous(labels = scales::label_percent(), expand = expansion()) +
+  scale_x_discrete(expand = expansion()) +
+  scale_fill_manual(breaks=c("Unlabeled", "Labeled"), values = c("#a6cde3", "#1d7cb4")) +
+  theme_bw() +
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), 
+        axis.ticks.x = element_blank(), legend.position = "top") +
+  labs(y="Fraction labeled", fill=NULL)
+
+bind_rows(amm_iso_areas, nit_iso_areas) %>%
+  filter(compound_name%in%c("L-Glutamine", "Guanine")) %>%
+  mutate(compound_name=factor(compound_name, levels=c("L-Glutamine", "Guanine"))) %>%
+  filter(samp_type=="Smp") %>%
+  mutate(label_boolean=ifelse(labeled_n_count>0, "Labeled", "Unlabeled")) %>%
+  left_join(parametadata) %>%
+  filter(startime=="Morn") %>%
+  filter(depth=="Surf") %>%
+  arrange(timepoint_num) %>%
+  mutate(timepoint=fct_inorder(timepoint)) %>%
+  group_by(amend, compound_name, timepoint, timepoint_num, label_boolean, labeled_n_count) %>%
+  summarize(area=median(area)) %>%
+  group_by(amend, compound_name, timepoint, timepoint_num) %>%
+  mutate(area=area/sum(area)) %>%
+  group_by(amend, compound_name, timepoint, timepoint_num, label_boolean) %>%
+  summarise(area=sum(area)) %>%
+  ggplot() +
+  geom_col(aes(x=timepoint, y=area, fill=label_boolean), position = "fill", width = 1, color="black") +
+  facet_nested(compound_name~amend) +
+  scale_y_continuous(labels = scales::label_percent(), expand = expansion()) +
+  scale_x_discrete(expand = expansion()) +
+  scale_fill_manual(breaks=c("Labeled", "Unlabeled"), values = c("#1d7cb4", "#a6cde3")) +
+  theme_bw() +
+  theme(legend.position = "top") +
+  labs(y="Fraction labeled", fill=NULL, x="Timepoint")
+```

### 2025-05-05 targeted/paratargeted_control.Rmd
Adding targeted analysis

---
title: "PARAGON targeted analysis"
author: "William Kumler"
date: "2024-05-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RaMS)
library(arrow)
library(ggh4x)
options(pillar.sigfig=7)

parametadata <- read_csv("metadata/parametadata.csv") %>%
  mutate(depth=fct_inorder(depth)) %>%
  mutate(startime=fct_inorder(startime)) %>%
  mutate(timepoint=fct_inorder(timepoint))
all_stans <- read_csv("metadata/all_stans.csv") %>%
  mutate(n_C=as.numeric(str_extract(formula, "(?<=C)\\d+"))) %>%
  mutate(n_C=case_when(
    !str_detect(formula, "C")~0,
    str_detect(formula, "C") & is.na(n_C)~1,
    TRUE~n_C
  )) %>%
  mutate(n_N=as.numeric(str_extract(formula, "(?<=N)\\d+"))) %>%
  mutate(n_N=case_when(
    !str_detect(formula, "N")~0,
    str_detect(formula, "N") & is.na(n_N)~1,
    TRUE~n_N
  ))
iso_stans <- all_stans %>%
  select(compound_name, mz, n_C, n_N, polarity) %>%
  pmap(function(...){
    row_data <- list(...)
    expand_grid(n_C=0:row_data$n_C, n_N=0:row_data$n_N, 
                amendment=c("Amm", "Nit", "Arg", "GMP")) %>%
      mutate(compound_name=row_data$compound_name, mz=row_data$mz,
             polarity=row_data$polarity)
  }) %>%
  bind_rows() %>%
  filter(!(amendment%in%c("Amm", "Nit") & n_C>0)) %>%
  mutate(mz=mz+1.003355*n_C+0.997035*n_N) %>%
  mutate(iso_name = paste0(compound_name, ", 13C", n_C, ", 15N", n_N))

splividis <- function(n, ...){
  vir <- viridis::viridis(n, ...)
  if(n%%2==0){
    vir[as.numeric(matrix(1:n, ncol=2, byrow = TRUE))]
  } else {
    index_vec <- suppressWarnings(as.numeric(matrix(1:n, ncol=2, byrow = TRUE)))
    vir[head(index_vec, -1)]
  }
}
```

## Integrations for isotopes
### Make graphs for peak boundaries

```{r extract EICs and convert to PDF using 0-20 RT}
all_stans <- read_csv("metadata/all_stans.csv") %>%
  mutate(polarity=factor(polarity, levels=c("pos", "neg"))) %>%
  arrange(polarity) %>%
  filter(compound_type!="Internal Standard")

all_stans %>%
  select(polarity, compound_name, mz, rt, date_added, mix) %>%
  expand_grid(amendment=c("Amm", "Nit", "GMP", "Arg"), .) %>%
  mutate(rtstart=0, rtend=20, notes=NA) %>%
  write_csv("targeted/peak_bounds.csv")

pdf("targeted/feat_bounds.pdf", onefile = TRUE, width = 7, height = 4)
all_stans %>%
  expand_grid(amendment=c("Amm", "Nit", "GMP", "Arg"), .) %>%
  # head() %>%
  pwalk(function(...){
    row_data <- list(...)
    eic_i <- open_dataset("tmzMLs/pqds") %>%
      filter(polarity==row_data$polarity) %>%
      filter(samp_type%in%c("Poo", "Std")) %>%
      filter(amendment==row_data$amendment) %>%
      filter(mz%between%pmppm(row_data$mz, 10)) %>%
      filter(rt%between%c(0.5, 20)) %>%
      collect()
    gp <- eic_i %>%
      mutate(mixtype=str_extract(filename, "Poo|Mix1|Mix2")) %>%
      filter(!str_detect(filename, "H2OInMatrix")) %>%
      left_join(parametadata, by = join_by(filename, samp_type, amendment, polarity)) %>%
      mutate(mixtype=factor(mixtype, levels=c("Poo", "Mix1", "Mix2"))) %>%
      ggplot() +
      geom_vline(xintercept = 1:19, color="grey70") +
      geom_vline(xintercept = row_data$rt, color="red") +
      geom_line(aes(x=rt, y=int, group=filename, color=mixtype), linewidth=0.8) +
      facet_wrap(~mixtype, ncol=1, scales="free") +
      ggtitle(paste0(row_data$amendment, " ", row_data$compound_name, 
                     "  (", row_data$mix, ")")) +
      theme_bw() +
      theme(legend.position = "none") +
      scale_x_continuous(breaks = 0:20, limits = c(0, 20)) +
      scale_color_manual(breaks=c("Mix1", "Mix2", "Poo"), 
                         values = c("purple", "mediumorchid1", "blue"))
    print(gp)
  }, .progress = TRUE)
dev.off()
```

```{r check integration consistency between runs}
bound_stans <- readxl::read_excel("targeted/peak_bounds.xlsx")

bound_stans %>%
  filter(rtstart>0) %>%
  ggplot() +
  geom_segment(aes(x=rtstart, xend=rtend, y=mz, yend=mz, color=polarity)) +
  facet_wrap(~amendment)

library(GGally)
bound_stans %>%
  filter(polarity=="pos") %>%
  select(compound_name, amendment, rtstart) %>%
  pivot_wider(names_from = amendment, values_from = rtstart) %>%
  column_to_rownames("compound_name") %>%
  ggpairs(aes(alpha=0.5), diag = list(continuous = "blankDiag"))

bound_stans %>%
  filter(polarity=="pos") %>%
  select(compound_name, amendment, rtstart) %>%
  pivot_wider(names_from = amendment, values_from = rtstart) %>%
  filter(Amm>0|GMP>0) %>%
  ggplot(aes(x=Amm, y=GMP, label=compound_name)) +
  geom_abline(slope=1) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point() +
  geom_text() +
  scale_x_continuous(expand = expansion(0.2))
```

```{r check integration bounds by plotting chroms}
bound_stans <- readxl::read_excel("targeted/peak_bounds.xlsx")

eic_data <- bound_stans %>%
  filter(rtstart>0) %>%
  # filter(compound_name%in%head(unique(compound_name))) %>%
  nest(row_data=c(amendment, rtstart, rtend, in_pooled, notes)) %>%
  mutate(eic=pbapply::pbmapply(function(mz_i, polarity_i){
    open_dataset("tmzMLs/pqds") %>%
      filter(polarity==polarity_i) %>%
      filter(mz%between%pmppm(mz_i, 10)) %>%
      filter(rt%between%c(0.5, 20)) %>%
      collect()
    }, .$mz, .$polarity, SIMPLIFY = FALSE))
saveRDS(eic_data, "targeted/eic_data.rds")

eic_data <- readRDS("targeted/eic_data.rds")
pdf("targeted/samp_int_check.pdf", onefile = TRUE, width = 7, height = 4)
for(i in 1:nrow(eic_data)){
  gp <- left_join(eic_data$row_data[[i]], eic_data$eic[[i]], by=join_by(
    amendment, dplyr::between(y$rt, x$rtstart, x$rtend))
    ) %>%
    filter(samp_type%in%c("Smp","Poo")) %>%
    qplotMS1data() +
    ggtitle(eic_data$compound_name[[i]]) +
    facet_wrap(~amendment, scales="free_y")
  print(gp)
}
dev.off()
```

### Integrate areas

```{r get feature eics for integration}
cmpd_bounds <- readxl::read_xlsx("targeted/peak_bounds.xlsx") %>%
  filter(rtstart>0) %>%
  filter(is.na(notes) | notes!="Remove") %>%
  mutate(mz=ifelse(compound_name=="Homarine", 138.0541, mz)) %>%
  select(amendment, polarity, compound_name, rtstart, rtend)

# Update iso_stans to reflect the changes made during the Excel annotation
updated_iso_stans <- iso_stans %>%
  filter(!compound_name=="beta-Glutamic acid") %>%
  mutate(compound_name=str_replace(compound_name, "L-Glutamic acid", "L-Glutamic acid/beta-Glutamic acid")) %>%
  filter(!compound_name=="L-Isoleucine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Leucine", "L-Leucine/L-Isoleucine")) %>%
  filter(!compound_name=="L-Homoserine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Threonine", "L-Threonine/L-Homoserine")) %>%
  filter(!compound_name=="Allopurinol") %>%
  mutate(compound_name=str_replace(compound_name, "Hypoxanthine", "Hypoxanthine/Allopurinol")) %>%
  distinct(amendment, compound_name, iso_name, polarity, mz, n_C, n_N) %>%
  filter(compound_name%in%unique(cmpd_bounds$compound_name))
new_pory_entry <- updated_iso_stans %>% 
  filter(compound_name=="Porphyra-334?") %>% 
  mutate(compound_name="Also Porphyra-334?") %>%
  mutate(iso_name=str_replace(iso_name, "Porphyra-334\\?", "Also Porphyra-334\\?"))
new_paly_entry <- updated_iso_stans %>% 
  filter(compound_name=="Palythene/Usujirene?") %>% 
  mutate(compound_name="Another Palythene/Usujirene?") %>%
  mutate(iso_name=str_replace(iso_name, "Palythene/Usujirene\\?", "Another Palythene/Usujirene\\?"))
updated_iso_stans <- updated_iso_stans %>% 
  bind_rows(new_pory_entry) %>% 
  bind_rows(new_paly_entry)

final_bounds <- cmpd_bounds %>%
  left_join(updated_iso_stans)

cmpd_eics <- final_bounds %>%
  nest(amend_specs=-c(compound_name, polarity)) %>%
  # head() %>%
  # slice(1)  -> row_data
  # filter(compound_name=="Theanine?") -> row_data
  pmap(function(...){
    row_data <- list(...)
    # print(row_data$compound_name)
    amend_specs <- row_data$amend_specs %>%
      rowwise() %>%
      mutate(mzmin=pmppm(mz, 10)[1]) %>%
      mutate(mzmax=pmppm(mz, 10)[2]) %>%
      ungroup()
    mzmin <- min(amend_specs$mz)-0.5
    mzmax <- max(amend_specs$mz)+0.5
    rtmin <- min(amend_specs$rtstart)
    rtmax <- max(amend_specs$rtend)
    all_eic <- open_dataset("tmzMLs/pqds") %>%
      filter(polarity==row_data$polarity) %>%
      filter(mz%between%c(mzmin, mzmax)) %>%
      filter(rt%between%c(rtmin, rtmax)) %>%
      collect()
    cmpd_eic <- left_join(amend_specs, all_eic, by=join_by(
      amendment, dplyr::between(y$rt, x$rtstart, x$rtend), 
      dplyr::between(y$mz, x$mzmin, x$mzmax)
    )) %>%
      distinct() %>%
      group_by(amendment, filename, iso_name, rt) %>%
      # nest() %>% ungroup() %>% slice(1) %>% unnest(data)
      slice_max(int) %>%
      ungroup() %>%
      mutate(compound_name=row_data$compound_name) %>%
      select(amendment, compound_name, iso_name, rt, mz=mz.y, int, filename, polarity)
    cmpd_eic
  }, .progress = TRUE) %>%
  bind_rows() %>%
  ungroup()
saveRDS(cmpd_eics, "targeted/cmpd_eics.rds")
```

```{r single compound demo check}
cmpd_eics <- readRDS("targeted/cmpd_eics.rds")
cmpd_eics %>%
  ungroup() %>%
  filter(compound_name=="Theanine?") %>%
  left_join(iso_stans %>% select(-mz)) %>%
  # filter(filename=="210916_Poo_TruePooAmm-noIS-pos-Full_5.mzML") %>%
  filter(n_C==0 & n_N<=1) %>%
  left_join(parametadata) %>%
  mutate(iso_name=str_remove(iso_name, ".*?, ")) %>%
  qplotMS1data(color_col = "timepoint") + 
  facet_grid(iso_name~amendment, scales = "free")
```

```{r integrate areas}
cmpd_eics <- readRDS("targeted/cmpd_eics.rds")
raw_areas <- cmpd_eics %>%
  left_join(updated_iso_stans %>% select(-mz)) %>%
  split(.$compound_name) %>%
  pbapply::pbsapply(function(cmpd_eic_i){
    unlabeled_eic <- cmpd_eic_i %>%
      group_by(filename) %>%
      filter(n_C==0 & n_N==0) %>%
      mutate(unlab_n=n()) %>%
      select(filename, rt, unlab_int=int, unlab_n) %>%
      ungroup()
    cmpd_eic_i %>%
      inner_join(unlabeled_eic, by = join_by(rt, filename)) %>%
      group_by(filename, iso_name) %>%
      filter(n()>5) %>%
      arrange(filename, iso_name, rt) %>%
      summarise(cor_model=list(cor.test(int, unlab_int)), 
                cor_coef=cor_model[[1]]$estimate,
                cor_pval=cor_model[[1]]$p.value,
                overlap_frac=n()/unique(unlab_n), 
                area=trapz(rt, int, baseline="square"),
                .groups = "drop")
}, simplify = FALSE) %>% 
  bind_rows(.id = "compound_name") %>%
  group_by(compound_name) %>%
  mutate(fdr_pval=p.adjust(cor_pval, method="fdr")) %>%
  mutate(holm_pval=p.adjust(cor_pval, method="holm")) %>%
  mutate(compound_name=factor(compound_name, levels=unique(cmpd_eics$compound_name))) %>%
  arrange(compound_name) %>%
  ungroup()
```

```{r inspect raw peaks to decide on thresholds}
raw_areas %>%
  # filter(compound_name=="L-Arginine") %>%
  filter(compound_name=="L-Glutamic acid/beta-Glutamic acid") %>%
  left_join(parametadata %>% distinct(filename, amendment, startime, depth, timepoint)) %>%
  arrange(timepoint) %>%
  filter(amendment=="Arg") %>%
  filter(startime=="Morn" & depth=="Surf") %>%
  distinct(filename, compound_name, amendment) %>%
  drop_na() %>% 
  # select(filename, overlap_frac, cor_coef) %>% print(n=Inf)
  pwalk(function(...){
    row_data <- data.frame(...)
    gp <- row_data %>% 
      left_join(cmpd_eics, by = join_by(compound_name, amendment, filename)) %>%
      group_by(iso_name) %>%
      ggplot() +
      geom_line(aes(x=rt, y=int, group = interaction(iso_name, filename), color=iso_name)) +
      facet_wrap(~iso_name, ncol=1, scales="free_y") +
      ggtitle(row_data$filename) +
      theme(legend.position="none")
    print(gp)
  }, .progress = TRUE)
raw_areas %>%
  # filter(compound_name=="L-Arginine") %>%
  filter(compound_name=="L-Glutamic acid/beta-Glutamic acid") %>%
  left_join(parametadata %>% distinct(filename, amendment, startime, depth, shortname)) %>%
  filter(amendment=="Arg") %>%
  filter(startime=="Morn" & depth=="Surf") %>%
  drop_na() %>% 
  select(shortname, iso_name, overlap_frac, cor_coef, holm_pval, area) %>%
  print(n=Inf)
```

```{r define peak areas by filtering out intrusive ions}
deionized_areas <- raw_areas %>%
  filter(cor_pval<0.01) %>%
  select(filename, compound_name, iso_name, area)

raw_areas %>%
  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
  filter(compound_name=="L-Glutamine") %>%
  filter(amendment=="Arg") %>%
  filter(depth=="Deep") %>%
  filter(startime=="Morn") %>%
  filter(iso_name=="L-Glutamine, 13C5, 15N1") %>%
  select(filename, iso_name, cor_pval, area) %>%
  print(n=Inf)
```

```{r visualize final peak areas}
peak_areas %>%
  filter(compound_name=="L-Glutamic acid/beta-Glutamic acid") %>%
  left_join(parametadata) %>%
  filter(samp_type=="Smp") %>%
  # filter(amendment%in%c("Arg", "GMP")) %>%
  ggplot() +
  geom_col(aes(x=tripl, y=area, fill=iso_name)) +
  facet_nested(depth+amendment~startime+timepoint, scales="free_y")
# cmpd_eics %>%
#   filter(compound_name=="L-Glutamic acid/beta-Glutamic acid") %>%
#   left_join(parametadata) %>%
#   filter(samp_type=="Smp") %>%
#   # filter(amendment%in%c("Arg", "GMP")) %>%
#   filter(timepoint=="T3") %>%
#   ggplot() +
#   geom_line(aes(x=rt, y=int, group=filename, color=amendment)) +
#   facet_wrap(~iso_name, ncol=2, scales="free_y")
```

### Fix problematic areas from IS contamination

```{r}
is_contam_shortnames <- c("Amm-Surf-Eve-T3-A", "Amm-Surf-Eve-T3-B", 
                          "Amm-Surf-Morn-T26-A", "Arg-Deep-Eve-T10-A", 
                          "Nit-Deep-Eve-T0-A")
problem_stans <- all_stans %>%
  filter(compound_type=="Internal Standard") %>%
  select(compound_name, polarity) %>%
  mutate(unlab_name=str_remove(compound_name, ", .*")) %>%
  filter(str_detect(compound_name, "15N|13C")) %>%
  mutate(n_C=str_extract(compound_name, "(?<=13C)\\d+")) %>%
  mutate(n_C=case_when(
    !str_detect(compound_name, "13C")~"0",
    str_detect(compound_name, "13C") & is.na(n_C)~"1",
    TRUE~n_C
  )) %>%
  mutate(n_N=str_extract(compound_name, "(?<=15N)\\d+")) %>%
  mutate(n_N=case_when(
    !str_detect(compound_name, "15N")~"0",
    str_detect(compound_name, "15N") & is.na(n_N)~"1",
    TRUE~n_N
  )) %>%
  mutate(contam_name=paste0(unlab_name, ", 13C", n_C, ", 15N", n_N)) %>%
  expand_grid(shortname=is_contam_shortnames) %>%
  left_join(parametadata %>% distinct(shortname, filename)) %>%
  select(iso_name=contam_name, filename)

decontam_areas <- deionized_areas %>%
  anti_join(problem_stans)
```

```{r write out}
write_csv(decontam_areas, "targeted/all_peak_areas_w_isos.csv")
```

### Some visualizations

```{r}
peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv")
```

```{r Stacked barplots}
peak_areas %>%
  filter(compound_name%in%c("Theanine?")) %>%
  inner_join(parametadata %>% filter(samp_type=="Smp")) %>%
  ggplot() +
  geom_col(aes(x=tripl, y=area, fill=iso_name), position = position_fill(), 
           color="black", width=1) +
  facet_nested(amendment+depth~compound_name+startime+timepoint, scales="free_y")
```

```{r Logistic regression curves}
peak_areas %>%
  filter(compound_name=="L-Glutamine") %>%
  mutate(labeled_frac=1-area/sum(area), .by = c("filename", "compound_name")) %>%
  select(compound_name, iso_name, filename, labeled_frac) %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name) %>%
  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
  filter(!is.na(startime)) %>%
  ggplot(aes(x=timepoint_num, y=labeled_frac)) +
  geom_point() +
  geom_smooth(method = "glm", formula = y~x, method.args = list(family = "quasibinomial")) +
  facet_nested(amendment+depth~startime)
```

```{r Logistic regression coefficients}
peak_areas %>%
  filter(compound_name=="L-Glutamine") %>%
  mutate(labeled_frac=1-area/sum(area), .by = c("filename", "compound_name")) %>%
  select(compound_name, iso_name, filename, labeled_frac) %>%
  left_join(parametadata %>% distinct(filename, samp_type, amendment, depth, startime, timepoint_num)) %>%
  filter(samp_type=="Smp") %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name, -samp_type) %>%
  nest(data=c(filename, timepoint_num, labeled_frac)) %>%
  mutate(log_fit=map(data, ~glm(.x$labeled_frac~.x$timepoint_num, family="quasibinomial")),
         coefs=map(log_fit, function(x)broom::tidy(x))) %>%
  unnest(coefs) %>%
  # filter(p.adjust(p.value, "fdr")<0.05) %>%
  select(-statistic, -p.value, -std.error) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  rename(intercept=`(Intercept)`, slope=`.x$timepoint_num`) %>%
  mutate(half_max=-intercept/slope, scale=1/slope) %>%
  mutate(frac_at_T24 = 1/(1+exp(-(intercept+slope*24)))) %>%
  ggplot() +
  geom_point(aes(x=half_max, y=frac_at_T24, color=amendment, shape=depth, size=startime), alpha=0.5) +
  scale_size_manual(values=c(5, 3))
```

```{r Logreg attempting to propagate errors via delta correction}
library(msm)
peak_areas %>%
  filter(compound_name=="L-Glutamine") %>%
  mutate(labeled_frac=1-area/sum(area), .by = c("filename", "compound_name")) %>%
  select(compound_name, iso_name, filename, labeled_frac) %>%
  left_join(parametadata %>% distinct(filename, samp_type, amendment, depth, startime, timepoint_num)) %>%
  filter(samp_type=="Smp") %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name, -samp_type) %>%
  nest(data=c(filename, timepoint_num, labeled_frac)) %>%
  mutate(log_fit=map(data, ~glm(.x$labeled_frac~.x$timepoint_num, family="quasibinomial")),
         coefs=map(log_fit, function(x)broom::tidy(x))) %>%
  unnest(coefs) %>%
  select(-statistic, -p.value, -std.error) %>%
  mutate(term=ifelse(term=="(Intercept)", "intercept", "slope")) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  mutate(half_max=-intercept/slope) %>%
  mutate(frac_at_T24 = 1/(1+exp(-(intercept+slope*24)))) %>%
  rowwise() %>%
  mutate(half_max_se = msm::deltamethod(~ -x1 / x2, coef(log_fit), vcov(log_fit))) %>%
  mutate(frac24_se = msm::deltamethod(~ 1 / (1 + exp(-(x1 + x2 * 24))), coef(log_fit), vcov(log_fit))) %>%
  ungroup() %>%
  ggplot(aes(x=half_max, y=frac_at_T24, color=amendment)) +
  geom_linerange(aes(ymin=frac_at_T24-2*frac24_se, ymax=frac_at_T24+2*frac24_se)) +
  geom_linerange(aes(xmin=half_max-2*half_max_se, xmax=half_max+2*half_max_se)) +
  geom_point(aes(shape=depth, size=startime), alpha=0.5) +
  scale_size_manual(values=c(5, 3))
```

```{r Logreg attempting error distributions via bootstrapping}
set.seed(123)
all_bootstraps <- peak_areas %>%
  filter(compound_name == "Glycine betaine") %>%
  mutate(labeled_frac = 1 - area / sum(area), .by = c("filename", "compound_name")) %>%
  select(compound_name, iso_name, filename, labeled_frac) %>%
  left_join(parametadata %>% distinct(filename, samp_type, amendment, depth, startime, timepoint_num)) %>%
  filter(samp_type == "Smp") %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name, -samp_type) %>%
  nest(data = c(filename, timepoint_num, labeled_frac)) %>%
  mutate(log_fit = map(data, ~glm(.x$labeled_frac ~ .x$timepoint_num, family = "quasibinomial"))) %>%
  mutate(bootstrap_results = map(data, function(data_i) {
    boot_samples <- map(1:100, function(i) {
      boot_data <- data_i[sample(nrow(data_i), replace = TRUE), ]
      boot_model <- glm(labeled_frac ~ timepoint_num, data = boot_data, family = "quasibinomial")
      half_max_boot <- -coef(boot_model)[1] / coef(boot_model)[2]
      frac24_boot <- 1 / (1 + exp(-(coef(boot_model)[1] + coef(boot_model)[2] * 24)))
      data.frame(half_max_boot=half_max_boot, frac24_boot=frac24_boot)
    }) %>%
      bind_rows()
  }))
# Plot raw data colored by amendment
all_bootstraps %>%
  unnest(bootstrap_results) %>%
  ggplot(aes(x = half_max_boot, y = frac24_boot, color = amendment)) +
  geom_point(alpha = 0.5) +
  scale_size_manual(values = c(5, 3))
# Plot summarized data as mean +/- SD
all_bootstraps %>%
  unnest(bootstrap_results) %>%
  summarise(half_max_mean=mean(half_max_boot), half_max_sd=sd(half_max_boot),
            frac24_mean=mean(frac24_boot), frac24_sd=sd(frac24_boot),
            .by = c(amendment, depth, startime)) %>%
  ggplot(aes(x=half_max_mean, y=frac24_mean, color=amendment, label=paste(depth, startime))) +
  geom_point() +
  geom_text(nudge_y = 0.02) +
  geom_linerange(aes(ymin=frac24_mean-2*frac24_sd, ymax=frac24_mean+2*frac24_sd)) +
  geom_linerange(aes(xmin=half_max_mean-2*half_max_sd, xmax=half_max_mean+2*half_max_sd))
# Plot raw data faceted by amendment and colored by exp factor
all_bootstraps %>%
  unnest(bootstrap_results) %>%
  ggplot(aes(x = half_max_boot, y = frac24_boot, color = interaction(depth, startime))) +
  geom_point(alpha = 0.5) +
  scale_size_manual(values = c(5, 3)) +
  facet_wrap(~amendment, scales="free")
```

```{r Plotting peak areas of each isotope as a heatmap instead}
peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv")
peak_areas %>% 
  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
  filter(samp_type=="Smp") %>%
  filter(compound_name=="Adenine") %>%
  left_join(iso_stans) %>%
  arrange(depth, startime, timepoint) %>%
  mutate(filename=fct_inorder(filename)) %>%
  mutate(area=area/sum(area), .by=filename) %>%
  mutate(n_N=factor(n_N, levels=5:0)) %>%
  mutate(n_C=factor(n_C, levels=10:0)) %>%
  filter(amendment%in%c("Amm", "Nit")) %>%
  # filter(samp_type=="Smp") %>%
  # filter(startime=="Morn") %>%
  # filter(n_N==1) %>%
  ggplot() +
  geom_tile(aes(x=timepoint, y=tripl, fill=area)) +
  scale_fill_viridis_c() +
  ggh4x::facet_nested(amendment+n_C+n_N~depth+startime, 
                      scales = "free", space = "free")
```

```{r Raw chromatogram to double check}
v <- open_dataset("tmzMLs/pqds") %>%
  filter(polarity=="pos") %>%
  filter(mz%between%pmppm(174.1365, 10)) %>%
  collect()

v %>%
  # filter(amendment=="Amm") %>% filter(int>1e6) %>% distinct(filename)
  filter(!filename%in%c(
    "210916_Smp_Amm-Surf-Eve-T3-A-noIS-pos.mzML",
    "210916_Smp_Amm-Surf-Eve-T3-B-noIS-pos.mzML",
    "210916_Smp_Amm-Surf-Morn-T26-A-noIS-pos.mzML"
  )) %>%
  filter(str_detect(filename, "Smp")) %>%
  left_join(parametadata) %>%
  qplotMS1data(color_col = "timepoint") +
  facet_wrap(~amendment, scales = "free_y")
```

```{r}
peak_areas %>%
  group_by(filename, compound_name) %>%
  mutate(area=area/sum(area)) %>%
  mutate(label_type=ifelse(str_detect(iso_name, ", 13C0, 15N0"), "unlabeled", "labeled")) %>%
  group_by(filename, compound_name, label_type) %>%
  summarise(area_unlabeled=sum(area)) %>%
  filter(label_type=="unlabeled") %>%
  left_join(parametadata) %>%
  filter(startime=="Morn") %>%
  filter(depth=="Deep") %>%
  # filter(compound_name=="L-Glutamine") %>%
  group_by(compound_name, amendment, timepoint) %>%
  summarise(labeled_above_thresh=median(area_unlabeled)<0.5) %>% #Less than X peak area is in unlabeled fraction
  group_by(amendment, timepoint) %>%
  summarise(n_labeled=sum(labeled_above_thresh), n_total=n()) %>%
  select(amendment, timepoint, n_labeled, total=n_total) %>%
  ungroup() %>%
  mutate(total=median(total), .by = amendment) %>%
  pivot_wider(names_from = timepoint, values_from = n_labeled)
```

## PheTau things (not currently used for anything)

Just confirming that all the injections went off successfully

```{r IS areas}
IS_eics <- open_dataset("tmzMLs/pqds") %>%
      filter(mz%between%c(128.0308598, 128.0321402) |
               mz%between%c(130.0464498, 130.0477502) |
               mz%between%c(174.1356293, 174.1373707)) %>%
      collect()

IS_bounds <- tribble(
  ~amendment, ~polarity, ~compound_name, ~mz, ~rtstart, ~rtend, 
  "Amm", "neg", "Taurine, 2H4", 128.0315, 9.8, 11.5,
  "Arg", "neg", "Taurine, 2H4", 128.0315, 8.5, 11,
  "GMP", "neg", "Taurine, 2H4", 128.0315, 9.5, 11.5,
  "Nit", "neg", "Taurine, 2H4", 128.0315, 10, 10.5,
  "Amm", "pos", "Taurine, 2H4", 130.0471, 9.8, 11.5,
  "Arg", "pos", "Taurine, 2H4", 130.0471, 8.5, 11,
  "GMP", "pos", "Taurine, 2H4", 130.0471, 9.5, 11.5,
  "Nit", "pos", "Taurine, 2H4", 130.0471, 10, 10.5,
  "Amm", "pos", "L-Phenylalanine, 2H8", 174.1365, 4, 6.5,
  "Arg", "pos", "L-Phenylalanine, 2H8", 174.1365, 3.6, 6,
  "GMP", "pos", "L-Phenylalanine, 2H8", 174.1365, 4, 7.2,
  "Nit", "pos", "L-Phenylalanine, 2H8", 174.1365, 3.5, 5
)

IS_areas <- IS_bounds %>%
  rowwise() %>%
  mutate(mzmin = pmppm(mz, 5)[1], mzmax = pmppm(mz, 5)[2]) %>%
  ungroup() %>%
  select(-mz) %>%
  left_join(IS_eics, by=join_by(
      amendment, polarity, dplyr::between(y$rt, x$rtstart, x$rtend), 
      dplyr::between(y$mz, x$mzmin, x$mzmax)
    )) %>%
  distinct() %>%
  # qplotMS1data(color_col = "amendment") + facet_wrap(~compound_name+polarity, ncol = 1, scales = "free")
  group_by(compound_name, amendment, filename, rt) %>%
  slice_max(int) %>%
  group_by(amendment, filename, compound_name, polarity) %>%
  arrange(rt) %>%
  summarise(area=trapz(rt, int), .groups="drop")

bad_IS_files <- c(
  "210916_Smp_Amm-Deep-Eve-T1-C-noIS",
  "220729_Smp_Arg-Deep-Eve-T10-A-noIS",
  "220715_Smp_GMP-Deep-Eve-T26-B-noIS",
  "220715_Smp_GMP-Surf-Eve-T3-C-noIS",
  "220729_Smp_Arg-Surf-Morn-T73-C-noIS"
) %>%
  rep(each=2) %>%
  paste0(c("-pos.mzML", "-neg.mzML"))

fixed_IS_areas <- IS_areas %>%
  left_join(parametadata) %>%
  filter(samp_type=="Smp") %>%
  arrange(compound_name, polarity, amendment, timestamp) %>%
  group_by(compound_name, polarity, amendment) %>%
  mutate(lag_area=lag(area), lead_area=lead(area)) %>%
  mutate(area=ifelse(filename%in%bad_IS_files, (lag_area+lead_area)/2, area)) %>%
  select(amendment, filename, compound_name, polarity, area)
IS_areas <- rows_update(IS_areas, fixed_IS_areas, by = c("amendment", "filename", "compound_name", "polarity"))

IS_areas %>%
  select(amendment, filename, compound_name, polarity, area) %>%
  filter(amendment!="Nit") %>%
  write_csv("targeted/PheTau_peak_areas.csv")
```

```{r vis IS areas}
IS_areas <- read_csv("targeted/PheTau_peak_areas.csv")
IS_areas %>%
  left_join(parametadata) %>%
  arrange(timestamp) %>%
  mutate(filename=str_remove(filename, "-pos|-neg")) %>%
  mutate(filename=fct_inorder(filename)) %>%
  filter(samp_type%in%c("Poo", "Smp")) %>%
  ggplot() +
  geom_col(aes(x=filename, y=area, fill=amendment), width=1) +
  facet_grid(compound_name+polarity~samp_type, scales = "free", space = "free_x")

# Check for missing samples
# parametadata %>%
#   filter(IS_type=="noIS") %>%
#   anti_join(distinct(IS_areas, filename)) %>%
#   distinct(filename) %>%
#   print(n=Inf)

# Test correlation between IS
# IS_areas %>%
#   filter(amendment!="Nit") %>%
#   filter(str_detect(filename, "Poo")) %>%
#   mutate(filename=str_remove(filename, "-pos|-neg")) %>%
#   pivot_wider(names_from=c(compound_name, polarity), values_from = area) %>%
#   # select(3:5) %>% cor() %>% round(3) %>% `diag<-`("")
#   ggplot() +
#   geom_point(aes(x=`Taurine, 2H4_neg`, y=`L-Phenylalanine, 2H8_pos`, color=amendment))
```

```{r}
IS_areas <- read_csv("targeted/PheTau_peak_areas.csv")
peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv")

pooled_metadata <- parametadata %>%
  filter(samp_type=="Poo") %>%
  filter(IS_type=="noIS") %>%
  filter(amendment!="Nit") %>%
  select(filename)
pooled_metadata %>%
  left_join(IS_areas)
pooled_metadata %>%
  left_join(peak_areas)
```

## Quant (requires starting over for the wIS samples)

### Get initial peak areas (for both IS and all compounds)

```{r plot chroms to estimate peak bounds, eval=FALSE}
is_chroms <- read_csv("metadata/all_stans.csv") %>%
  filter(compound_type=="Internal Standard") %>%
  filter(date_added<200101) %>%
  select(compound_name, polarity, mz) %>%
  mutate(is_name=compound_name) %>%
  mutate(compound_name=str_remove(compound_name, ",.*")) %>%
  rename(is_mz=mz) %>%
  left_join(readxl::read_xlsx("targeted/peak_bounds.xlsx")) %>%
  filter(compound_name=="L-Cysteic acid, 2H3")

  drop_na("amendment") %>%
  select(compound_name=is_name, mz=is_mz, amendment, polarity, rtstart, rtend) %>%
  rowwise() %>%
  mutate(mzmin=pmppm(mz, 10)[1]) %>%
  mutate(mzmax=pmppm(mz, 10)[2]) %>%
  ungroup() %>%
  # slice(1) -> row_data
  pmap(function(...){
    row_data <- list(...)
    open_dataset("tmzMLs/pqds_wIS") %>%
      filter(amendment==row_data$amendment) %>%
      filter(polarity==row_data$polarity) %>%
      filter(mz%between%c(row_data$mzmin, row_data$mzmax)) %>%
      filter(rt%between%c(row_data$rtstart, row_data$rtend)) %>%
      collect() %>%
      filter(str_detect(filename, "T0")) %>%
      slice_max(int, by=c(filename, rt)) %>%
      mutate(compound_name=row_data$compound_name,
             polarity=row_data$polarity) %>%
      select(compound_name, amendment, filename, rt, int, polarity)
  }, .progress = TRUE) %>%
  bind_rows()
is_chroms %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  # filter(compound_name==unique(compound_name)[3]) %>%
  qplotMS1data(color_col="amendment") +
  facet_wrap(~compound_name, scales = "free")
plotly::ggplotly()

read_csv("metadata/all_stans.csv") %>%
  filter(compound_type=="Internal Standard") %>%
  filter(date_added<200101) %>%
  select(compound_name, polarity, mz) %>%
  mutate(is_name=compound_name) %>%
  mutate(compound_name=str_remove(compound_name, ",.*")) %>%
  rename(is_mz=mz) %>%
  left_join(readxl::read_xlsx("targeted/peak_bounds.xlsx")) %>%
  drop_na("amendment") %>%
  select(compound_name=is_name, mz=is_mz, amendment, polarity, rtstart, rtend) %>%
  write_csv("targeted/IS_bounds.csv")
```

```{r integrate IS areas, eval=FALSE}
is_chroms <- readxl::read_xlsx("targeted/IS_bounds.xlsx") %>%
  rowwise() %>%
  mutate(mzmin=pmppm(mz, 10)[1]) %>%
  mutate(mzmax=pmppm(mz, 10)[2]) %>%
  ungroup() %>%
  # slice(1) -> row_data
  pmap(function(...){
    row_data <- list(...)
    open_dataset("tmzMLs/pqds_wIS") %>%
      filter(amendment==row_data$amendment) %>%
      filter(polarity==row_data$polarity) %>%
      filter(mz%between%c(row_data$mzmin, row_data$mzmax)) %>%
      filter(rt%between%c(row_data$rtstart, row_data$rtend)) %>%
      collect() %>%
      # distinct(filename)
      slice_max(int, by=c(filename, rt)) %>%
      mutate(compound_name=row_data$compound_name, 
             polarity=row_data$polarity) %>%
      select(compound_name, amendment, filename, rt, int, polarity)
  }, .progress = TRUE) %>%
  bind_rows()
is_chroms %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  # filter(compound_name==unique(compound_name)[3]) %>%
  qplotMS1data(color_col="amendment") +
  facet_wrap(~compound_name, scales = "free")

is_areas <- is_chroms %>%
  arrange(rt) %>%
  summarise(IS_area=trapz(rt, int), .by=c(amendment, filename, compound_name, polarity)) %>%
  left_join(parametadata %>% distinct(filename, timestamp, samp_type, polarity)) %>%
  complete(filename, compound_name) %>%
  arrange(filename) %>%
  group_by(filename) %>%
  fill(timestamp, amendment, samp_type, .direction="downup") %>%
  ungroup() %>%
  arrange(compound_name, timestamp) %>%
  # filter(filename=="210813_Smp_Nit-Deep-Eve-T0-A-wIS.mzML")
  mutate(IS_area=case_when(
    filename=="210813_Smp_Nit-Deep-Eve-T0-A-wIS.mzML"~(lag(IS_area)+lead(IS_area))/2,
    TRUE~IS_area
  )) %>%
  select(amendment, filename, polarity, IS_name=compound_name, IS_area)
write_csv(is_areas, "targeted/IS_areas.csv")
```

```{r integrate nonIS unlabeled areas from wIS files}
wIS_bounds <- readxl::read_xlsx("targeted/peak_bounds.xlsx") %>%
  filter(rtstart>0) %>%
  filter(is.na(notes) | notes!="Remove") %>%
  filter(compound_name!="Homarine") %>% # Can't separate from Adenine IS sufficiently
  select(amendment, polarity, compound_name, mz, rtstart, rtend)

wIS_eics <- wIS_bounds %>%
  nest(amend_specs=-c(compound_name, polarity)) %>%
  # head() %>%
  # slice(1)  -> row_data
  # filter(compound_name=="Theanine?") -> row_data
  pmap(function(...){
    row_data <- list(...)
    # print(row_data$compound_name)
    amend_specs <- row_data$amend_specs %>%
      rowwise() %>%
      mutate(mzmin=pmppm(mz, 10)[1]) %>%
      mutate(mzmax=pmppm(mz, 10)[2]) %>%
      ungroup()
    mzmin <- min(amend_specs$mz)-0.5
    mzmax <- max(amend_specs$mz)+0.5
    rtmin <- min(amend_specs$rtstart)
    rtmax <- max(amend_specs$rtend)
    all_eic <- open_dataset("tmzMLs/pqds_wIS") %>%
      filter(polarity==row_data$polarity) %>%
      filter(mz%between%c(mzmin, mzmax)) %>%
      filter(rt%between%c(rtmin, rtmax)) %>%
      collect()
    cmpd_eic <- left_join(amend_specs, all_eic, by=join_by(
      amendment, dplyr::between(y$rt, x$rtstart, x$rtend), 
      dplyr::between(y$mz, x$mzmin, x$mzmax)
    )) %>%
      distinct() %>%
      group_by(amendment, filename, rt) %>%
      # nest() %>% ungroup() %>% slice(1) %>% unnest(data)
      slice_max(int) %>%
      ungroup() %>%
      mutate(compound_name=row_data$compound_name) %>%
      select(amendment, compound_name, rt, mz=mz.y, int, filename, polarity)
    cmpd_eic
  }, .progress = TRUE) %>%
  bind_rows() %>%
  ungroup()
saveRDS(wIS_eics, "targeted/wIS_eics.rds")

wIS_eics <- readRDS("targeted/wIS_eics.rds")
wIS_areas <- wIS_eics %>%
  slice_max(int, by=c(filename, compound_name, rt)) %>%
  arrange(rt) %>%
  summarise(area=trapz(rt, int, baseline = "square"), 
            .by = c(amendment, filename, polarity, compound_name)) %>%
  mutate(compound_name=factor(compound_name, levels=unique(wIS_eics$compound_name))) %>%
  arrange(compound_name, filename)
wIS_areas %>%
  write_csv("targeted/wIS_areas.csv")
```

### Refine IS peak areas due to contamination

```{r Replace wonky values with average of previous and next}
is_areas <- read_csv("targeted/IS_areas.csv")

IS_mix_one <- c("Thymine, 2H4", "Uracil, 15N2, 2H2", "Adenine, 15N2",
                "Cytosine, 13C2, 15N3", "Glycine betaine, 13C5, 15N",
                "Guanine, 13C, 15N2", "Homarine, 2H3")

fixed_IS <- parametadata %>%
  filter(samp_type%in%c("Smp", "Poo", "Blk")) %>%
  filter(IS_type=="wIS") %>%
  distinct(filename, shortname, timestamp) %>%
  left_join(is_areas %>% select(filename, amendment, IS_name, IS_area)) %>%
  arrange(IS_name, timestamp) %>%
  mutate(filename=fct_inorder(filename)) %>%
  # filter(shortname=="Nit-Deep-Eve-T1-A") %>%
  mutate(IS_area=case_when(
    shortname=="Nit-Deep-Eve-T1-A" & IS_name%in%IS_mix_one ~ 
      (lag(IS_area) + lead(IS_area))/2,
    shortname=="Amm-Surf-Eve-T3-A" ~ (lag(IS_area) + lead(IS_area))/2,
    shortname=="Arg-Deep-Eve-T10-A" ~ (lag(IS_area) + lead(IS_area))/2,
    shortname=="GMP-Surf-Eve-T3-C" ~ (lag(IS_area) + lead(IS_area, 2))/2,
    shortname=="GMP-Surf-Eve-T3-C" & IS_name=="L-Histidine, 15N" ~ 
      (lag(IS_area) + lead(IS_area, 2))/2,
    shortname=="GMP-Surf-Morn-T26-B" & IS_name=="L-Histidine, 15N" ~ 
      (lag(IS_area, 2) + lead(IS_area))/2,
    shortname=="Arg-Surf-Eve-T10-B" & IS_name=="Sulfoacetic acid, 13C2" ~ 
      (lag(IS_area) + lead(IS_area, 2))/2,
    shortname=="Arg-Deep-Eve-T3-A" & IS_name=="Sulfoacetic acid, 13C2" ~ 
      (lag(IS_area) + lead(IS_area, 2))/2,
    TRUE ~ IS_area
  ))
fixed_IS %>%
  select(filename, amendment, IS_name, IS_area) %>%
  write_csv("targeted/IS_areas.csv")
```

```{r double-checking that IS fixes worked via plots and PCA}
fixed_IS %>%
  # filter(str_detect(IS_name, "Homarine")) %>%
  left_join(parametadata %>% distinct(filename, timestamp, amendment)) %>%
  # filter(amendment=="Amm") %>%
  arrange(timestamp) %>%
  mutate(filename=fct_inorder(filename)) %>%
  ggplot() +
  geom_col(aes(x=filename, y=IS_area, fill=amendment), width=1) +
  facet_wrap(~IS_name, ncol=3, scales="free_y")

wide_IS <- fixed_IS %>%
  select(shortname, IS_name, IS_area) %>%
  # filter(!str_detect(shortname, "PooGMP")) %>%
  pivot_wider(names_from=IS_name, values_from=IS_area) %>%
  column_to_rownames("shortname")

# wide_IS %>% write.csv("~/../Desktop/fixed_IS.csv")

wide_IS %>%
  data.matrix() %>%
  prcomp() %>%
  pluck("x") %>%
  as.data.frame() %>%
  rownames_to_column("shortname") %>%
  left_join(parametadata %>% filter(IS_type=="wIS")) %>%
  ggplot() +
  geom_text(aes(x=PC1, y=PC2, label=filename, color=amendment))
wide_IS %>%
  data.matrix() %>%
  t() %>%
  prcomp() %>%
  pluck("x") %>%
  as.data.frame() %>%
  rownames_to_column("IS_name") %>%
  ggplot() +
  geom_text(aes(x=PC1, y=PC2, label=IS_name))

wide_IS %>%
  data.matrix() %>%
  t() %>%
  prcomp() %>%
  pluck("rotation") %>%
  as.data.frame() %>%
  rownames_to_column("shortname") %>%
  select(shortname, PC1:PC3) %>%
  pivot_longer(PC1:PC3) %>%
  left_join(parametadata %>% filter(IS_type=="wIS")) %>%
  ggplot() +
  geom_col(aes(x=shortname, y=value, fill=amendment), width=1) +
  facet_wrap(~name,ncol=1)
```

### Choose BMIS

```{r identify the BMIS for each compound}
IS_areas <- read_csv("targeted/IS_areas.csv")

chosen_IS <- IS_areas %>%
  inner_join(parametadata %>% distinct(filename, amendment, samp_type)) %>%
  filter(samp_type=="Poo") %>%
  select(filename, IS_name, amendment, IS_area) %>%
  left_join(wIS_areas, relationship = "many-to-many") %>%
  arrange(compound_name, filename) %>%
  group_by(amendment, compound_name, IS_name) %>%
  mutate(norm_area=area/IS_area*mean(IS_area)) %>%
  summarise(cv=sd(norm_area)/mean(norm_area), .groups = "drop") %>%
  slice_min(cv, by = c(compound_name, amendment)) %>%
  # print(n=Inf)
  # with(table(IS_name, compound_name))
  select(amendment, compound_name, IS_name)
count(chosen_IS, amendment, IS_name) %>%
  pivot_wider(names_from=amendment, values_from=n) %>%
  print(n=Inf)

IS_with_unlabeled <- IS_areas %>%
  distinct(IS_name, amendment) %>%
  mutate(compound_name=str_remove(IS_name, ", .*"))
chosen_IS %>%
  filter(compound_name%in%IS_with_unlabeled$compound_name) %>%
  count(compound_name, IS_name) %>%
  print(n=Inf)

chosen_IS <- chosen_IS %>%
  rows_update(IS_with_unlabeled, by = c("compound_name", "amendment"), unmatched = "ignore")
write_csv(chosen_IS, "targeted/chosen_IS.csv")
```

### Calculate RFs

```{r create response factor dataframe from standards}
wIS_areas <- read_csv("targeted/wIS_areas.csv")
stan_patched <- all_stans %>%
  filter(!compound_name=="beta-Glutamic acid") %>%
  mutate(compound_name=str_replace(compound_name, "L-Glutamic acid", "L-Glutamic acid/beta-Glutamic acid")) %>%
  filter(!compound_name=="L-Isoleucine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Leucine", "L-Leucine/L-Isoleucine")) %>%
  filter(!compound_name=="L-Homoserine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Threonine", "L-Threonine/L-Homoserine")) %>%
  filter(!compound_name=="Allopurinol") %>%
  mutate(compound_name=str_replace(compound_name, "Hypoxanthine", "Hypoxanthine/Allopurinol"))

rf_df <- wIS_areas %>%
  left_join(parametadata %>% select(filename, samp_type, polarity)) %>%
  filter(samp_type=="Std") %>%
  left_join(stan_patched %>% distinct(compound_name, mix, conc_um)) %>%
  filter(str_detect(filename, paste0(mix, "InMatrix")) | 
           str_detect(filename, "H2OInMatrix")) %>%
  select(compound_name, conc_um, amendment, filename, area) %>%
  mutate(filename=str_extract(filename, "Mix|H2O")) %>%
  pivot_wider(names_from = filename, values_from=area, values_fn = mean) %>%
  mutate(area_diff=Mix-H2O) %>%
  # filter(rf<0)
  filter(area_diff>0) %>%
  mutate(rf=area_diff/conc_um) %>%
  select(compound_name, amendment, rf)

write_csv(rf_df, "targeted/rf_df.csv")
rf_df %>%
  arrange(rf) %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  ggplot() +
  geom_boxplot(aes(x=rf, y=compound_name)) +
  scale_x_log10()
```

### Finalize concentrations and drop low-quality compounds

```{r quantify}
wIS_areas <- read_csv("targeted/wIS_areas.csv")
chosen_IS <- read_csv("targeted/chosen_IS.csv")
IS_areas <- read_csv("targeted/IS_areas.csv")
rf_df <- read_csv("targeted/rf_df.csv")

env_concs <- wIS_areas %>%
  filter(!str_detect(compound_name, "\\?$")) %>%
  filter(!compound_name%in%c("Urea", "L-Valine", "Oxalic acid", "Malic acid", 
                             "Acetylcholine", "Homarine")) %>%
  inner_join(parametadata %>% filter(samp_type=="Smp") %>% distinct(filename)) %>%
  left_join(chosen_IS) %>%
  left_join(IS_areas) %>%
  mutate(norm_area=area/IS_area*mean(IS_area), .by = c(amendment, compound_name)) %>%
  left_join(rf_df) %>%
  mutate(conc_um_in_vial=norm_area/rf) %>%
  select(compound_name, filename, conc_um_in_vial) %>%
  mutate(conc_nm_in_env=conc_um_in_vial*400/2e6*1000)
write_csv(env_concs, "targeted/final_env_concs.csv")
```

```{r quant plots for confirmation}
env_concs %>%
  filter(compound_name%in%c("Gonyol", "L-Glutamine", "Guanosine",
                            "L-Glutamic acid/beta-Glutamic acid")) %>%
  left_join(parametadata, relationship = "many-to-many") %>%
  ggplot() +
  geom_hline(yintercept = 0) +
  geom_boxplot(aes(x=amendment, y=conc_nm_in_env, color=depth)) +
  facet_wrap(~compound_name, scales="free_y")

env_concs %>%
  filter(conc_nm_in_env==0) %>%
  left_join(parametadata, relationship = "many-to-many") %>%
  count(compound_name, amendment) %>%
  complete(compound_name, amendment, fill = list(n=0)) %>%
  pivot_wider(names_from = amendment, values_from=n)

env_concs %>%
  mutate(avg_conc=mean(conc_nm_in_env, na.rm=TRUE), .by=compound_name) %>%
  arrange(avg_conc) %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  left_join(parametadata, relationship = "many-to-many") %>%
  filter(conc_nm_in_env>0) %>%
  ggplot() +
  geom_boxplot(aes(x=conc_nm_in_env, y=compound_name, color=depth)) +
  scale_x_log10(breaks=c(0.000001, 0.001, 1), labels=c("1 fM", "1 pM", "1 nM"))
```

### 2025-05-05 manuscript/figscript.R
Adding manuscript skeleton

@@ -1,0 +1,64 @@
+
+# Setup ----
+
+library(tidyverse)
+library(ggh4x)
+library(patchwork)
+
+parametadata <- read_csv("metadata/parametadata.csv") %>%
+  # filter(IS_type=="noIS") %>%
+  filter(polarity=="pos") %>%
+  filter(samp_type=="Smp") %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
+  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"))) %>%
+  arrange(amendment, depth, timepoint_num, startime) %>%
+  mutate(timepoint=fct_inorder(timepoint))
+peak_areas <- read_csv("targeted/all_peak_areas_w_isos_incl_13C_for_inorg.csv") %>%
+  left_join(parametadata %>% distinct(filename, shortname))
+all_stans <- read_csv("metadata/all_stans.csv")
+samp_areas <- parametadata %>%
+  filter(samp_type=="Smp") %>%
+  distinct(shortname, depth) %>%
+  left_join(peak_areas) %>%
+  mutate(shortname=fct_inorder(shortname)) %>%
+  filter(!compound_name%in%c("Melamine", "Uridine", "Asterina-330?", "Palythene/Usujirene?",
+                             "Mycosporine-2-glycine?", "Porphyra-334?", "Itaconic acid?",
+                             "Inosine", "Dexpanthenol", "Choline sulfate", "5-Hydroxyectoine")) %>%
+  filter(!(depth=="Deep" & compound_name%in%c("Also Palythene/Usujirene?", "Shinorine?",
+                                              "Mycosporine-glycine?", "Palythine?"))) %>%
+  # filter(!str_detect(shortname, "Nit-Deep-Eve-T10")) %>%
+  select(-depth)
+labeled_fracs <- samp_areas %>%
+  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
+  select(compound_name, iso_name, shortname, labeled_frac) %>%
+  filter(str_detect(iso_name, "13C0, 15N0")) %>%
+  select(-iso_name)
+n_labeled_fracs <- samp_areas %>%
+  filter(str_detect(iso_name, "13C0")) %>%
[…]

### 2025-05-05 manuscript/manuscript.Rmd
Adding rmarkdown skeleton, using FMarS template for now

---
title: ""
author:
  - William Kumler
  - Laura T. Carlson
  - Anitra E. Ingalls
output: 
  word_document:
    reference_docx: Frontiers_Template.docx
bibliography:
- Exported Items.bib
- r-packages.bib
csl: frontiers-in-marine-science.csl
---

```{r Run analysis, include=FALSE}
# TBD
```

```{r setup}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(tidyverse)
library(ggh4x)
library(patchwork)
library(ggtext)
library(minpack.lm)
```

## Abstract

## Introduction

## Results

## Discussion

## Conclusion

## Methods

## Acknowledgements

## Data availability

## References

```{r packages, eval=FALSE}
packages_used <- list.files("..", pattern = "*\\.R$|*\\.Rmd$", recursive = TRUE) %>%
  str_subset("old|scripts", negate=TRUE) %>%
  map(function(file_name){
    v <- readLines(paste0("../", file_name))
    libs <- unique(str_extract(v, "(?<=library\\().*?(?=\\)|,)"))
    refs <- unique(str_extract(v, "[[:alnum:]]*(?=::)"))
    pkgs <- unique(c(na.omit(libs), na.omit(refs)))
  }) %>% unlist() %>% unique()
knitr::write_bib(packages_used[nchar(packages_used)>0], file = "r-packages.bib")
```

<div id="refs"></div>

## Supplement

### 2025-05-05 manuscript/manuscript.Rmd
Adding data import code from figure_gallery

@@
 library(minpack.lm)
+
+parametadata <- read_csv("metadata/parametadata.csv") %>%
+  # filter(IS_type=="noIS") %>%
+  filter(polarity=="pos") %>%
+  filter(samp_type=="Smp") %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
+  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"))) %>%
+  arrange(amendment, depth, timepoint_num, startime) %>%
+  mutate(timepoint=fct_inorder(timepoint))
+peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv") %>%
+  left_join(parametadata %>% distinct(filename, shortname))
+all_stans <- read_csv("metadata/all_stans.csv")
+maa_names <- c("Also Palythene/Usujirene?", "Shinorine?", "Mycosporine-glycine?", 
+               "Palythine?", "Another Palythene/Usujirene?", "Porphyra-334?", 
+               "Also Porphyra-334?", "Asterina-330?", "Palythene/Usujirene?",
+               "Mycosporine-2-glycine?")
+nucleo_names <- c("5-Methylcytosine", "Adenosine", "Cytidine", "Adenine", "Guanosine",
+                  "Cytosine", "Hypoxanthine/Allopurinol", "Guanine", "Deoxycytidine?",
+                  "Deoxyadenosine", "Deoxyguanosine?")
+betaine_names <- c("beta-Alaninebetaine", "Glycine betaine", "Homarine", "Trigonelline",
+                   "Choline", "Glycerophosphocholine", "Homoserine betaine?", 
+                   "Threonine betaine?", "Proline betaine",
+                   "(3-Carboxypropyl)trimethylammonium", "O-Acetylcarnitine", 
+                   "Betonicine", "Acetylcholine", "Phosphocholine")
+non_n_names <- c("Dimethylsulfonioacetate", "Dimethylsulfoniopropionate",
+                 "Arsenobetaine", "Dimethylsulfoxonium propionate?", "Gonyol",
+                 "2-O-alpha-D-Glucosylglycerol")
+aa_names <- c("Glutamine", "Aspartic acid", "Glutamic acid/beta-Glutamic acid",
+              "Asparagine", "Alanine", "Methionine", "Arginine", "Proline",
+              "Lysine", "Threonine/Homoserine", "Serine", "Glycine", "Histidine",
+              "Leucine/Isoleucine", "Valine", "Arginine")
+bad_cmpds <- c("Melamine", "Uridine", "Itaconic acid?", "Inosine", "Dexpanthenol", 
+               "Choline sulfate", "5-Hydroxyectoine", "L-Tyrosine")
+samp_areas <- parametadata %>%
+  filter(samp_type=="Smp") %>%
+  distinct(shortname, depth) %>%
+  left_join(peak_areas) %>%
+  mutate(shortname=fct_inorder(shortname)) %>%
+  filter(!compound_name%in%bad_cmpds) %>%
+  filter(!(depth=="Deep" & compound_name%in%maa_names)) %>%
+  select(-depth)
+labeled_fracs <- samp_areas %>%
+  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
+  select(compound_name, iso_name, shortname, labeled_frac) %>%
+  filter(str_detect(iso_name, "13C0, 15N0")) %>%
+  select(-iso_name)
+n_labeled_fracs <- samp_areas %>%
+  filter(str_detect(iso_name, "13C0")) %>%
+  filter(!str_detect(iso_name, "15N0") | str_detect(iso_name, "13C0, 15N0")) %>%
+  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
+  select(compound_name, iso_name, shortname, labeled_frac) %>%
+  filter(str_detect(iso_name, "13C0, 15N0")) %>%
+  select(-iso_name)
+log_coef_df <- labeled_fracs %>%
+  left_join(parametadata %>% distinct(shortname, amendment, depth, timepoint_num)) %>%
+  nest(data=c(shortname, timepoint_num, labeled_frac)) %>%
+  mutate(log_fit=map(data, ~glm(.x$labeled_frac~.x$timepoint_num, family="quasibinomial")),
+         coefs=map(log_fit, function(x)broom::tidy(x))) %>%
+  unnest(coefs) %>%
+  select(compound_name, amendment, depth, term, estimate) %>%
+  pivot_wider(names_from = term, values_from = estimate) %>%
+  rename(intercept=`(Intercept)`, slope=`.x$timepoint_num`) %>%
+  mutate(half_max=-intercept/slope, scale=1/slope) %>%
+  mutate(frac_at_T24 = 1/(1+exp(-(intercept+slope*24))))
+splividis <- function(n, ...){
+  vir <- viridis::viridis(n, ...)
+  if(n%%2==0){
+    vir[as.numeric(matrix(1:n, ncol=2, byrow = TRUE))]
+  } else {
+    index_vec <- suppressWarnings(as.numeric(matrix(1:n, ncol=2, byrow = TRUE)))
+    vir[head(index_vec, -1)]
+  }
+}
 ```

### 2025-05-05 manuscript/manuscript.Rmd
Adding figure 1 heatmap

--- 2025-05-05 manuscript/manuscript.Rmd
@@
 library(minpack.lm)
+
+parametadata <- read_csv("metadata/parametadata.csv") %>%
+  # filter(IS_type=="noIS") %>%
+  filter(polarity=="pos") %>%
+  filter(samp_type=="Smp") %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
+  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"))) %>%
+  arrange(amendment, depth, timepoint_num, startime) %>%
+  mutate(timepoint=fct_inorder(timepoint))
+peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv") %>%
+  left_join(parametadata %>% distinct(filename, shortname))
+all_stans <- read_csv("metadata/all_stans.csv")
+maa_names <- c("Also Palythene/Usujirene?", "Shinorine?", "Mycosporine-glycine?", 
+               "Palythine?", "Another Palythene/Usujirene?", "Porphyra-334?", 
+               "Also Porphyra-334?", "Asterina-330?", "Palythene/Usujirene?",
+               "Mycosporine-2-glycine?")
+nucleo_names <- c("5-Methylcytosine", "Adenosine", "Cytidine", "Adenine", "Guanosine",
+                  "Cytosine", "Hypoxanthine/Allopurinol", "Guanine", "Deoxycytidine?",
+                  "Deoxyadenosine", "Deoxyguanosine?")
+betaine_names <- c("beta-Alaninebetaine", "Glycine betaine", "Homarine", "Trigonelline",
+                   "Choline", "Glycerophosphocholine", "Homoserine betaine?", 
+                   "Threonine betaine?", "Proline betaine",
+                   "(3-Carboxypropyl)trimethylammonium", "O-Acetylcarnitine", 
+                   "Betonicine", "Acetylcholine", "Phosphocholine")
+non_n_names <- c("Dimethylsulfonioacetate", "Dimethylsulfoniopropionate",
+                 "Arsenobetaine", "Dimethylsulfoxonium propionate?", "Gonyol",
+                 "2-O-alpha-D-Glucosylglycerol")
+aa_names <- c("Glutamine", "Aspartic acid", "Glutamic acid/beta-Glutamic acid",
+              "Asparagine", "Alanine", "Methionine", "Arginine", "Proline",
+              "Lysine", "Threonine/Homoserine", "Serine", "Glycine", "Histidine",
+              "Leucine/Isoleucine", "Valine", "Arginine")
+bad_cmpds <- c("Melamine", "Uridine", "Itaconic acid?", "Inosine", "Dexpanthenol", 
+               "Choline sulfate", "5-Hydroxyectoine", "L-Tyrosine")
+samp_areas <- parametadata %>%
+  filter(samp_type=="Smp") %>%
+  distinct(shortname, depth) %>%
+  left_join(peak_areas) %>%
+  mutate(shortname=fct_inorder(shortname)) %>%
+  filter(!compound_name%in%bad_cmpds) %>%
+  filter(!(depth=="Deep" & compound_name%in%maa_names)) %>%
+  select(-depth)
+labeled_fracs <- samp_areas %>%
+  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
+  select(compound_name, iso_name, shortname, labeled_frac) %>%
+  filter(str_detect(iso_name, "13C0, 15N0")) %>%
+  select(-iso_name)
+n_labeled_fracs <- samp_areas %>%
+  filter(str_detect(iso_name, "13C0")) %>%
+  filter(!str_detect(iso_name, "15N0") | str_detect(iso_name, "13C0, 15N0")) %>%
+  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
+  select(compound_name, iso_name, shortname, labeled_frac) %>%
+  filter(str_detect(iso_name, "13C0, 15N0")) %>%
+  select(-iso_name)
+log_coef_df <- labeled_fracs %>%
+  left_join(parametadata %>% distinct(shortname, amendment, depth, timepoint_num)) %>%
+  nest(data=c(shortname, timepoint_num, labeled_frac)) %>%
+  mutate(log_fit=map(data, ~glm(.x$labeled_frac~.x$timepoint_num, family="quasibinomial")),
+         coefs=map(log_fit, function(x)broom::tidy(x))) %>%
+  unnest(coefs) %>%
+  select(compound_name, amendment, depth, term, estimate) %>%
+  pivot_wider(names_from = term, values_from = estimate) %>%
+  rename(intercept=`(Intercept)`, slope=`.x$timepoint_num`) %>%
+  mutate(half_max=-intercept/slope, scale=1/slope) %>%
+  mutate(frac_at_T24 = 1/(1+exp(-(intercept+slope*24))))
 ```
@@
 ## Results
+
+```{r figure 1 }
+class_levels <- c("Glutamine", "Glutamic acid", "Aspartic acid", "Asparagine", "Other amino acids", "Theanine?", "Citrulline", "Ectoine", "Guanine", "Other nucleic acids", "MAAs", "Betaines", "All others")
+
+labeled_fracs %>%
+  left_join(parametadata %>% distinct(shortname, timepoint, timepoint_num, depth, startime, amendment)) %>%
+  filter(amendment=="Amm") %>%
+  mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
+  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
+  mutate(compound_class=case_when(
+    compound_name%in%class_levels ~ compound_name,
+    compound_name%in%maa_names~"MAAs",
+    compound_name%in%aa_names~"Other amino acids",
+    compound_name%in%nucleo_names~"Other nucleic acids",
+    compound_name%in%betaine_names~"Betaines",
+    TRUE~"All others"
+  )) %>%
+  mutate(compound_class=factor(compound_class, levels=class_levels)) %>%
+  summarise(avg_frac=mean(labeled_frac), .by = c(compound_class, timepoint_num, startime, depth)) %>%
+  mutate(color_group=paste(depth, startime)) %>%
+  ggplot(aes(x=timepoint_num, y=avg_frac, color=color_group, fill=color_group)) +
+  geom_smooth(method = "glm", formula = y~x, method.args = list(family = "quasibinomial"),
+              fill="grey80") +
+  geom_point(pch=21, color="black") +
+  facet_wrap(~compound_class, ncol=5) +
+  scale_y_continuous(labels = scales::label_percent(), breaks = c(0, 0.5, 1)) +
+  scale_x_continuous(breaks = c(0.3, 1, 3, 10, 26), labels = c(0.3, 1, 3, 10, 26)) +
+  scale_color_manual(breaks = c("Surf Morn", "Surf Eve", "Deep Morn", "Deep Eve"),
+                     values = c("#FFD53D", "#DBA20A", "#256A76", "#003743")) +
+  scale_fill_manual(breaks = c("Surf Morn", "Surf Eve", "Deep Morn", "Deep Eve"),
+                     values = c("#FFD53D", "#DBA20A", "#256A76", "#003743")) +
+  labs(x="Hours after 15N added", y="Fraction of metabolite labeled", color=NULL) +
+  theme_bw() +
+  guides(color = guide_legend(nrow = 2, override.aes = list(linewidth=2)),
+         fill = "none") +
+  theme(legend.position = "inside",
+        legend.position.inside = c(0.8, 1/6),
+        legend.location = "plot")
+
+labeled_fracs %>%
+  left_join(parametadata %>% distinct(shortname, timepoint, timepoint_num, depth, startime, amendment, tripl)) %>%
+  filter(amendment=="Amm") %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"), labels=c("Surface seawater (25 meters)", "Seawater from below DCM (175m)"))) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start"))) %>%
+  mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
+  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
+  mutate(compound_class=case_when(
+    compound_name%in%c("Glutamine", "Aspartic acid", "Asparagine", "Glutamic acid", "Ectoine", "Theanine?", "Citrulline", "Guanine") ~ compound_name,
+    compound_name%in%maa_names~"MAAs",
+    compound_name%in%aa_names~"Other amino acids",
+    compound_name%in%nucleo_names~"Other nucleic acids",
+    compound_name%in%betaine_names~"Betaines",
+    TRUE~"All others"
+  )) %>%
+  mutate(compound_class=factor(compound_class, levels=class_levels)) %>%
+  summarise(avg_frac=mean(labeled_frac), .by = c(compound_class, timepoint, startime, depth, tripl)) %>%
+  ggplot() +
+  geom_raster(aes(x=timepoint, y=tripl, fill=avg_frac)) +
+  geom_vline(xintercept = c(1.5:4.5), color="grey80") +
+  facet_nested(compound_class~depth+startime, scales = "free_y", switch="y") +
+  scale_fill_gradientn(colours = viridis::viridis(10, alpha = seq(0, 1, length.out=10), direction = -1),
+                       limits=c(0, 1), breaks=c(0, 0.25, 0.5, 0.75, 1),
+                       labels=c("Fully unlabeled", "", "50% labeled", "", "Fully labeled")) +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_discrete(expand = expansion()) +
+  labs(x=NULL, y=NULL, fill=NULL) +
+  theme_bw() +
+  theme(panel.spacing.x=unit(0.5, "lines"), 
+        panel.spacing.y = unit(0, "lines"),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        strip.text.y.left = element_text(angle=0, hjust = 1),
+        strip.background.y = element_rect(fill=NA, color=NA),
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        legend.position = "top", 
+        legend.key.width = unit(1, "null")) +
+  guides(
+    fill=guide_colorbar(theme=theme(
+      legend.frame = element_rect(color = "black"),
+      legend.ticks = element_line(color="black"),
+      legend.text = ggtext::element_markdown(hjust=c(0, 0, 0.5, 1, 1)))
+    )
+  )
+```
 

### 2025-05-05 manuscript/manuscript.Rmd
Adding full heatmap as supplemental fig 1

@@
 ## Supplement
+
+## Supp fig 1: full heatmap of all compounds and timepoints
+
+```{r}
+cmpd_order <- labeled_fracs %>%
+  left_join(parametadata, relationship = "many-to-many") %>%
+  filter(amendment=="Amm") %>%
+  filter(depth=="Surf") %>%
+  filter(startime=="Morn") %>%
+  summarise(mean_label=mean(labeled_frac, na.rm=TRUE), .by=compound_name) %>%
+  arrange(desc(mean_label)) %>%
+  pull(compound_name) %>%
+  str_remove_all("L-")
+
+heatmap_gp <- labeled_fracs %>%
+  left_join(parametadata, relationship = "many-to-many") %>%
+  arrange(amendment, depth, timepoint_num, startime) %>%
+  mutate(shortname=fct_inorder(shortname)) %>%
+  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"),
+                          labels = c("+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"), labels=c("25m", "175m"))) %>%
+  mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
+  mutate(compound_name=factor(compound_name, cmpd_order)) %>%
+  mutate(compound_class=case_when(
+    compound_name%in%maa_names~"MAA",
+    compound_name%in%aa_names~"Amino\nacid",
+    compound_name%in%nucleo_names~"Nucleic\nacid",
+    compound_name%in%non_n_names~"N-free",
+    compound_name%in%betaine_names~"Betaine",
+    TRUE~"Other"
+  )) %>%
+  mutate(compound_class=factor(compound_class, levels=c("Amino\nacid", "Nucleic\nacid", "MAA", "Betaine", "N-free", "Other"))) %>%
+  ggplot() +
+  geom_raster(aes(x=timepoint, y=tripl, fill=labeled_frac)) +
+  scale_fill_gradientn(colours = viridis::viridis(10, alpha = seq(0, 1, length.out=10), direction = -1), limits=c(0, 1)) +
+  scale_y_discrete(expand=expansion()) +
+  scale_x_discrete(guide = guide_axis(n.dodge = 3), expand=expansion()) +
+  facet_nested(compound_class+compound_name~amendment+depth+startime, scales = "free_x", switch="y", strip=strip_nested(size = "variable")) +
+  labs(x="Timepoint", y="Metabolite", fill="Fraction\nlabeled") +
+  theme_bw() +
+  theme(panel.spacing.x=unit(0.5, "lines"), 
+        panel.spacing.y = unit(0, "lines"),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        strip.text.y.left = element_text(angle=0, hjust = 1),
+        strip.background.y = element_rect(fill=NA, color=NA),
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        legend.position = "top", 
+        legend.key.width = unit(1, "null")) +
+  guides(
+    fill=guide_colorbar(theme=theme(
+      legend.frame = element_rect(color = "black"),
+      legend.ticks = element_line(color="black"),
+      legend.text = ggtext::element_markdown(hjust=c(0, 0, 0.5, 1, 1)))
+    )
+  )
+
+# ggsave(plot = heatmap_gp, "~/../Desktop/para_map.pdf", device = "pdf", 
+#        width = 10, height = 12)
+```

### 2025-05-05 manuscript/manuscript.Rmd
Adding fig 2 MAA many options

@@
 
+```{r fig 2 MAA labeling patterns}
+chosen_maas <- c("Mycosporine-glycine?", "Shinorine?", "Palythine?",
+                 "Palythene/Usujirene?", "Also Palythene/Usujirene?", 
+                 "Also Porphyra-334?")
+labeled_fracs %>%
+  filter(compound_name%in%chosen_maas) %>%
+  mutate(compound_name=factor(compound_name, levels=chosen_maas)) %>%
+  arrange(compound_name) %>%
+  mutate(compound_name=str_remove(compound_name, "\\?$")) %>%
+  mutate(compound_name=str_replace(compound_name, "Also Palythene/Usujirene", "Palythene or\n usujirene")) %>%
+  mutate(compound_name=str_replace(compound_name, "Palythene/Usujirene", "Usujirene or\n palythene")) %>%
+  mutate(compound_name=str_remove(compound_name, "Also ")) %>%
+  mutate(compound_name=str_replace(compound_name, "-glycine", "-\nglycine")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  left_join(parametadata, relationship = "many-to-many") %>%
+  filter(samp_type=="Smp") %>%
+  filter(IS_type=="noIS") %>%
+  filter(amendment%in%c("Amm", "Nit")) %>%
+  mutate(amendment=factor(amendment, levels=c("Amm", "Nit"), labels=c("+Ammonia", "+Nitrate"))) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start"))) %>%
+  ggplot(aes(x=timepoint, y=labeled_frac, color=startime)) +
+  geom_vline(xintercept = 1:5, color="grey80") +
+  geom_boxplot(fill="grey95") +
+  geom_point(aes(fill=startime), position = position_dodge(width=0.7), color="black", pch=21) +
+  scale_color_manual(values=c("#e0670b", "#580770"), breaks=c("6AM start", "6PM start"),
+                     aesthetics = c("color", "fill")) +
+  ggpubr::stat_compare_means(label.y = 0.9, method = "t.test", 
+                             label = "p.signif", hide.ns = TRUE) +
+  facet_grid(compound_name~amendment) +
+  scale_y_continuous(limits=c(0, 1), labels=scales::label_percent()) +
+  labs(y="Fraction labeled with at least one 15N atom", x=NULL, color=NULL, fill=NULL) +
+  theme_bw()
+
+labeled_fracs %>%
+  filter(compound_name%in%chosen_maas) %>%
+  filter(compound_name!="Glycine") %>%
+  mutate(compound_name=factor(compound_name, levels=chosen_maas)) %>%
+  left_join(parametadata, relationship = "many-to-many") %>%
+  filter(samp_type=="Smp") %>%
+  filter(amendment=="Amm") %>%
+  ggplot(aes(x=timepoint_num, y=labeled_frac, color=startime, fill=startime)) +
+  geom_smooth(method="nlsLM", formula=y~a*x^b, method.args=list(start = c(a=1, b=1)), se=FALSE) +
+  # geom_smooth(method="nlsLM", formula=y~(a*x)/(b+x), method.args=list(start = c(a=1, b=1)), se=FALSE) +
+  # geom_smooth(method="nlsLM", formula=y~(a*x)/(b+x), method.args=list(start = c(a=1, b=1)), se=FALSE) +
+  # geom_smooth(method="glm", formula=y~x, method.args=list(family="quasibinomial")) +
+  geom_point(color="black", pch=21, size=3, alpha=0.5, stroke=1) +
+  facet_wrap(~compound_name, ncol=2) +
+  scale_y_continuous(limits=c(0, 1), labels=scales::label_percent()) +
+  scale_x_continuous(breaks = c(0, 3, 10, 26)) +
+  scale_color_manual(values=c("#e0670b", "#580770"), breaks=c("Morn", "Eve"), aesthetics = c("fill", "color")) +
+  labs(x="Hours after 15NH4 addition", y="Fraction labeled with at least one 15N atom") +
+  theme_bw()
+
+peak_areas %>%
+  filter(compound_name%in%chosen_maas) %>%
+  mutate(compound_name=factor(compound_name, levels=chosen_maas)) %>%
+  left_join(parametadata, relationship = "many-to-many") %>%
+  filter(samp_type=="Smp") %>%
+  filter(depth=="Surf") %>%
+  filter(amendment%in%c("Nit", "Amm")) %>%
+  mutate(n_N=paste0(str_extract(iso_name, "(?<=15N)\\d"), "x15N")) %>%
+  mutate(area=area/1e4) %>%
+  ggplot() +
+  geom_col(aes(x=tripl, y=area, fill=n_N), color="black", width=1) +
+  facet_nested(compound_name~amendment+startime+timepoint, scales="free_y") +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_continuous(expand = expansion(mult=c(0, 0.1))) +
+  labs(x=NULL, y="Peak area (x10^4)", fill="Number of\n15N labels") +
+  theme_bw() +
+  theme(panel.spacing.y = unit(0.1, "lines"), 
+        panel.spacing.x = unit(0, "lines"),
+        strip.text.y = element_text(angle=0, hjust = 0),
+        panel.grid.major.x = element_blank(), 
+        panel.grid.minor = element_blank())
+
+labeled_fracs %>%
+  filter(compound_name%in%chosen_maas) %>%
+  mutate(compound_name=factor(compound_name, levels=chosen_maas)) %>%
+  arrange(compound_name) %>%
+  mutate(compound_name=str_remove(compound_name, "\\?$")) %>%
+  mutate(compound_name=str_replace(compound_name, "Also Palythene/Usujirene", "Palythene or\n usujirene")) %>%
+  mutate(compound_name=str_replace(compound_name, "Palythene/Usujirene", "Usujirene or\n palythene")) %>%
+  mutate(compound_name=str_remove(compound_name, "Also ")) %>%
+  mutate(compound_name=str_replace(compound_name, "-glycine", "-\nglycine")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  left_join(parametadata, relationship = "many-to-many") %>%
+  filter(samp_type=="Smp") %>%
+  filter(IS_type=="noIS") %>%
+  filter(amendment%in%c("Amm", "Nit")) %>%
+  mutate(amendment=factor(amendment, levels=c("Amm", "Nit"), labels=c("+Ammonia", "+Nitrate"))) %>%
+  mutate(clock_time=ifelse(startime=="Morn", 6+timepoint_num, 18+timepoint_num)) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start"))) %>%
+  ggplot() +
+  annotate("rect", xmin=-Inf, xmax = 6, ymin=-Inf, ymax=Inf, fill="grey80") +
+  # annotate("rect", xmin=6, xmax = 18, ymin=-Inf, ymax=Inf, fill="#d0bd65") +
+  annotate("rect", xmin=18, xmax = 30, ymin=-Inf, ymax=Inf, fill="grey80") +
+  # annotate("rect", xmin=30, xmax = 42, ymin=-Inf, ymax=Inf, fill="#d0bd65") +
+  annotate("rect", xmin=42, xmax = Inf, ymin=-Inf, ymax=Inf, fill="grey80") +
+  geom_hline(yintercept = c(0.25, 0.5, 0.75, 1), color="grey90") +
+  # annotate("rect", xmin=-Inf, xmax=Inf, ymin=-Inf, ymax = Inf, fill="white", alpha=0.3) +
+  geom_smooth(aes(x=clock_time, y=labeled_frac, color=startime),
+              method=nlsLM, formula="y~a*x^b+d", se = FALSE,
+              method.args=list(start=list(a=1, b=2, d=1),
+                               control=nls.lm.control(maxiter = 500))) +
+  # geom_smooth(aes(x=clock_time, y=labeled_frac, group=startime),
+  #             method=glm, formula = "y~x", method.args=list(family="quasibinomial"), 
+  #             color="grey10", se=FALSE) +
+  # geom_point(aes(x=clock_time, y=labeled_frac, shape=startime), color="black", 
+  #            fill="white", size=3) +
+  geom_point(aes(x=clock_time, y=labeled_frac, shape=startime, fill=startime),
+             color="black", size=2, alpha=0.7) +
+  facet_nested(compound_name~amendment) +
+  scale_x_continuous(breaks = c(6, 18, 30, 42), labels = c("6AM", "6PM", "6AM", "6PM")) +
+  scale_y_continuous(labels = scales::label_percent(),
+                     limits = c(0, 1), breaks = c(0, 0.5, 1)) +
+  scale_shape_manual(values=c(24, 25), breaks = c("6AM start", "6PM start")) +
+  scale_fill_manual(values=c("#f4bb23", "#0b505c"), breaks = c("6AM start", "6PM start"),
+                    aesthetics = c("color", "fill")) +
+  labs(x=NULL, color=NULL, shape=NULL, fill=NULL, y="Fraction of MAA containing at least one 15N atom") +
+  theme_bw() +
+  theme(axis.text.y = element_markdown(vjust=c(0, 0.5, 1)),
+        axis.title.x = element_blank(),
+        legend.position = "inside",
+        legend.margin = margin(0.1,0.1,0.1,0.1, unit="lines"),
+        legend.position.inside = c(0, 1),
+        legend.justification = c(0, 1), 
+        legend.background = element_rect(color="black"), 
+        legend.direction = "horizontal") +
+  guides(shape=guide_legend(override.aes = list(size=3, stroke=1.2, alpha=1, linetype=0)))
+```
+
 ## Discussion

### 2025-05-05 manuscript/manuscript.Rmd
Tidying figure code and removing old MAA plots

--- 2025-05-05 manuscript/manuscript.Rmd
@@
 
-```{r figure 1 }
+```{r figure 1 small heatmap compound groups}
 class_levels <- c("Glutamine", "Glutamic acid", "Aspartic acid", "Asparagine", "Other amino acids", "Theanine?", "Citrulline", "Ectoine", "Guanine", "Other nucleic acids", "MAAs", "Betaines", "All others")
@@
 
+```{r fig 2 MAA labeling patterns}
+chosen_maas <- c("Mycosporine-glycine?", "Shinorine?", "Palythine?",
+                 "Palythene/Usujirene?", "Also Palythene/Usujirene?", 
+                 "Also Porphyra-334?")
+labeled_fracs %>%
+  filter(compound_name%in%chosen_maas) %>%
+  mutate(compound_name=factor(compound_name, levels=chosen_maas)) %>%
+  arrange(compound_name) %>%
+  mutate(compound_name=str_remove(compound_name, "\\?$")) %>%
+  mutate(compound_name=str_replace(compound_name, "Also Palythene/Usujirene", "Palythene or\n usujirene")) %>%
+  mutate(compound_name=str_replace(compound_name, "Palythene/Usujirene", "Usujirene or\n palythene")) %>%
+  mutate(compound_name=str_remove(compound_name, "Also ")) %>%
+  mutate(compound_name=str_replace(compound_name, "-glycine", "-\nglycine")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  left_join(parametadata, relationship = "many-to-many") %>%
+  filter(samp_type=="Smp") %>%
+  filter(IS_type=="noIS") %>%
+  filter(amendment%in%c("Amm", "Nit")) %>%
+  mutate(amendment=factor(amendment, levels=c("Amm", "Nit"), labels=c("+Ammonia", "+Nitrate"))) %>%
+  mutate(clock_time=ifelse(startime=="Morn", 6+timepoint_num, 18+timepoint_num)) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start"))) %>%
+  ggplot() +
+  annotate("rect", xmin=-Inf, xmax = 6, ymin=-Inf, ymax=Inf, fill="grey80") +
+  # annotate("rect", xmin=6, xmax = 18, ymin=-Inf, ymax=Inf, fill="#d0bd65") +
+  annotate("rect", xmin=18, xmax = 30, ymin=-Inf, ymax=Inf, fill="grey80") +
+  # annotate("rect", xmin=30, xmax = 42, ymin=-Inf, ymax=Inf, fill="#d0bd65") +
+  annotate("rect", xmin=42, xmax = Inf, ymin=-Inf, ymax=Inf, fill="grey80") +
+  geom_hline(yintercept = c(0.25, 0.5, 0.75, 1), color="grey90") +
+  # annotate("rect", xmin=-Inf, xmax=Inf, ymin=-Inf, ymax = Inf, fill="white", alpha=0.3) +
+  geom_smooth(aes(x=clock_time, y=labeled_frac, color=startime),
+              method=nlsLM, formula="y~a*x^b+d", se = FALSE,
+              method.args=list(start=list(a=1, b=2, d=1),
+                               control=nls.lm.control(maxiter = 500))) +
+  # geom_smooth(aes(x=clock_time, y=labeled_frac, group=startime),
+  #             method=glm, formula = "y~x", method.args=list(family="quasibinomial"), 
+  #             color="grey10", se=FALSE) +
+  # geom_point(aes(x=clock_time, y=labeled_frac, shape=startime), color="black", 
+  #            fill="white", size=3) +
+  geom_point(aes(x=clock_time, y=labeled_frac, shape=startime, fill=startime),
+             color="black", size=2, alpha=0.7) +
+  facet_nested(compound_name~amendment) +
+  scale_x_continuous(breaks = c(6, 18, 30, 42), labels = c("6AM", "6PM", "6AM", "6PM")) +
+  scale_y_continuous(labels = scales::label_percent(),
+                     limits = c(0, 1), breaks = c(0, 0.5, 1)) +
+  scale_shape_manual(values=c(24, 25), breaks = c("6AM start", "6PM start")) +
+  scale_fill_manual(values=c("#f4bb23", "#0b505c"), breaks = c("6AM start", "6PM start"),
+                    aesthetics = c("color", "fill")) +
+  labs(x=NULL, color=NULL, shape=NULL, fill=NULL, y="Fraction of MAA containing at least one 15N atom") +
+  theme_bw() +
+  theme(axis.text.y = element_markdown(vjust=c(0, 0.5, 1)),
+        axis.title.x = element_blank(),
+        legend.position = "inside",
+        legend.margin = margin(0.1,0.1,0.1,0.1, unit="lines"),
+        legend.position.inside = c(0, 1),
+        legend.justification = c(0, 1), 
+        legend.background = element_rect(color="black"), 
+        legend.direction = "horizontal") +
+  guides(shape=guide_legend(override.aes = list(size=3, stroke=1.2, alpha=1, linetype=0)))
+```
+
 ## Discussion
@@
 
-## Supp fig 1: full heatmap of all compounds and timepoints
+### Supp fig 1: full heatmap of all compounds and timepoints
 
-```{r}
+```{r suppfig full heatmap}
 cmpd_order <- labeled_fracs %>%

### 2025-05-05 manuscript/manuscript.Rmd
Adding an absolute mess of an intro

@@
 ## Introduction
+
+Marine carbon is fixed into biomass at a rate equivalent to ten fully loaded Boeing 747 airplanes crashing into the ocean every second of every day [ref 50Pg IPCC(?), ref Boeing weights].
+In most of the surface ocean this flux is controlled by the availability of nitrogen, an element with a complex and incompletely characterized biogeochemical cycle [@Capone2008; @Moore2013; @Hutchins2022].
+Organic nitrogen in particular is often treated as a black box despite substantial variability in its bioavailability and chemical nature.
+While extensive work has been done to characterize the forms and fluxes of nitrogen within the inorganic pool, the equivalent for the organic pool is woefully underdeveloped [@ref, prob NME2008 or Hansell2024, maybe @Hutchins2022 for visual?].
+This is largely due to the difficulty of comprehensively measuring the many organic molecules that contain nitrogen in the marine environment [@Boysen2018; @Moran2022].
+The bioavailability of the nitrogen atom is in the environment is largely determined by the form of the organic material it composes.
+
+The mechanisms by which dissolved nitrogen becomes biomass and vice versa are major factors in our ability to predict ecosystem productivity and therefore carbon fixation and export.
+Traditionally, the conversion of inorganic material into organic substrate was thought to be limited to autotrophs while remineralization was performed by the heterotrophic community.
+Today, it’s known that many photosynthetic organisms are able to take up and use organic nitrogen [@Anita1991; @Morando2018; @Hugo2021] and that only a select subset of phototrophs are capable of nitrate reduction or nitrogen fixation [@ref], while inorganic nitrogen is directly accessible to heterotrophs as well.
+This means that the transformations possible for a given N-containing compound are a function of the microbial community and its environment while the microbial community is in turn also expected to be a function of the N-containing molecules available.
+This creates a recursive network that’s difficult to untangle and cannot necessarily be extrapolated from axenic culture studies.
+
+One way to reveal the pathways and transformations that marine N undergoes in a natural community is via 15N labeling paired with metabolomics.
+Metabolites that incorporate the labeled nitrogen can then be separated on the mass spectrometer and quantified separately from the unlabeled pool.
+Here, we trace the uptake and use of labeled ammonia, nitrate, arginine, and guanosine monophosphate in a natural community from the NPSG at multiple depths and diel conditions.
+Our application of metabolomics to these samples allows us to map out the pathways and restructuring that organic nitrogen experiences in the largest biomes on the planet, quantifying the pool sizes and turnover rates for various low molecular weight compounds that serve as building blocks and intermediates of cell biology.
+
+### Other things to maybe discuss:
+
+  -	mention the expected differences with depth and diel effects
+  -	detail which compounds we chose and why?
+  -	expand on DON-PON interactions and importance?
+  - Extracellular enzymes, organic use as ammonia vs as-is?
+  -	expand on the importance of biogeochemical models and why we need pathways mapped out?
+  -	mention basic biosynthetic pathways (maybe better suited to discussion?)
+  -	new (nitrate) vs regenerated (ammonia and DON) production (maybe also discussion?)
+
+### Other soundbites:
+  - It is assumed that most fixed nitrogen becomes protein, presumably based on the large fraction of cell nitrogen in protein and previous research showing how quickly it enters the (free?) amino acid pool.
+However, 1) important intermediates molecules 2) doesn’t explain where biounavailable N comes from 3) makes gene-based molecular modeling really difficult.
+  - In each case, assimilation into organic matter is typically done via the glutamine synthetase (GS)/glutamate synthase (GOGAT) or glutamate dehydrogenase (GDH) pathways to produce glutamate from α-ketoglutarate [@ref, see bronk chapter].
+Glutamate then serves as a nitrogen supply in an enormous number of biochemical pathways in both primary and secondary metabolism [@Walker2016].
+  - nitrate amendments typically resulting in diatom blooms while ammonia incubations favor cyanobacteria [@Glibert2016].
+  - Communities to distinguish are 1) surface (largely regenerated production) and 2) 175m (largely new production(??))
+  - use of organics vs synthesis of organics
+  - Compounds such as non-proteinogenic amino acids (e.g., ornithine, citrulline, creatine, MAAs), osmolytes like betaines, nucleobases, and sulfur-containing molecules like taurine represent significant but poorly quantified pools.
+Understanding the rates and pathways by which these compounds are utilized is crucial, as they form the fundamental unit linking genetic potential to elemental cycling in biogeochemical models.
+  - Nitrogen use varies significantly between the sunlit surface where productivity is largely regenerated through recycling of ammonia and organic nitrogen, and deeper waters near the base of the euphotic zone, where nitrate supports “new” production.
+Quantifying rates of nitrogen assimilation into ammonia and downstream products across these gradients is therefore essential for predicting marine productivity and nutrient cycling.
+  - Nonetheless, how the community uses a substrate is arguably more important than whether it can.
 

### 2025-05-05 manuscript/manuscript.Rmd
Rendering figures as pngs

@@
 
-```{r figure 1 small heatmap compound groups}
+```{r fig small heatmap compound groups}
 class_levels <- c("Glutamine", "Glutamic acid", "Aspartic acid", "Asparagine", "Other amino acids", "Theanine?", "Citrulline", "Ectoine", "Guanine", "Other nucleic acids", "MAAs", "Betaines", "All others")
@@
   )
+
+ggsave(filename = "grouped_heatmap.png", device = "png", path = "manuscript/figures",
+       width = 6.5, height = 4, units = "in", dpi = 300)
 ```
 
-```{r fig 2 MAA labeling patterns}
+![](figures/grouped_heatmap.png)
+
+*Figure XX: Fraction of each metabolite or metabolite group containing at least one 15N label during the treatments amended with 15N ammonia after the number of hours shown on the x-axis.
+Darker colors indicate a larger percentage of the compound signal was isotopically labeled.
+Separate incubations were started at 6AM and 6PM with water from both 25 meters and 175 meters near Station ALOHA.
+Three biological replicates are stacked vertically at each timepoint.*
+
+```{r fig MAA labeling patterns}
 chosen_maas <- c("Mycosporine-glycine?", "Shinorine?", "Palythine?",
@@
   facet_nested(compound_name~amendment) +
-  scale_x_continuous(breaks = c(6, 18, 30, 42), labels = c("6AM", "6PM", "6AM", "6PM")) +
+  scale_x_continuous(breaks = c(6, 18, 30, 42), labels = c("6AM", "6PM", "6AM", "")) +
   scale_y_continuous(labels = scales::label_percent(),
@@
         legend.position = "inside",
-        legend.margin = margin(0.1,0.1,0.1,0.1, unit="lines"),
-        legend.position.inside = c(0, 1),
-        legend.justification = c(0, 1), 
-        legend.background = element_rect(color="black"), 
-        legend.direction = "horizontal") +
+        legend.margin = margin(0.1,0.25,0.1,0.1, unit="lines"),
+        legend.position.inside = c(1, 1),
+        legend.justification = c(1, 1), 
+        legend.background = element_rect(fill="#FFFFFFAA", color="black", linewidth = 0)) +
   guides(shape=guide_legend(override.aes = list(size=3, stroke=1.2, alpha=1, linetype=0)))
+
+ggsave(filename = "MAA_curves.png", device = "png", path = "manuscript/figures",
+       width = 3, height = 6, units = "in", dpi = 300)
 ```
+
+![](figures/MAA_curves.png)
+
+*Figure XX: Fraction of UV-reactive mycosporine-like amino acids (MAAs) containing at least one 15N label over one day during the treatments amended with 15N labeled ammonia or nitrate.
+Incubations started at 6AM are shown in yellow while incubations started at 6PM are shown in blue, with grey background regions highlighting nighttime.
+Three replicates are included at each of the 5 timepoints and best-fit power law curves have been fit behind the data.*
 

### 2025-05-06 manuscript/manuscript.Rmd
Adding potential deep vs surface labeling comparison figure

@@
   select(-iso_name)
-n_labeled_fracs <- samp_areas %>%
-  filter(str_detect(iso_name, "13C0")) %>%
-  filter(!str_detect(iso_name, "15N0") | str_detect(iso_name, "13C0, 15N0")) %>%
-  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
-  select(compound_name, iso_name, shortname, labeled_frac) %>%
-  filter(str_detect(iso_name, "13C0, 15N0")) %>%
-  select(-iso_name)
 log_coef_df <- labeled_fracs %>%
+  filter(!compound_name%in%maa_names) %>%
   left_join(parametadata %>% distinct(shortname, amendment, depth, timepoint_num)) %>%
@@
 Three biological replicates are stacked vertically at each timepoint.*
+
+```{r fig surface vs deep labeling rates}
+labeled_fracs %>%
+  filter(!compound_name%in%maa_names) %>%
+  left_join(distinct(parametadata, shortname, amendment, depth, timepoint_num)) %>%
+  filter(amendment!="Nit") %>%
+  summarise(avg_label=mean(labeled_frac), .by=c(compound_name, amendment, depth, timepoint_num)) %>%
+  pivot_wider(names_from = depth, values_from=avg_label) %>%
+  mutate(timepoint_num=ifelse(timepoint_num==0, 0.3, timepoint_num)) %>%
+  ggplot(aes(x=timepoint_num, y=Surf-Deep, color=compound_name, group=compound_name)) +
+  geom_hline(yintercept = 0) +
+  geom_line() +
+  geom_point() +
+  facet_wrap(~amendment, nrow=1) +
+  scale_x_log10() +
+  theme_bw() +
+  theme(legend.position = "none")
+ggsave(filename = "deep_vs_surf_labeling.png", device = "png", path = "manuscript/figures",
+       width = 6.5, height = 4, units = "in", dpi = 300)
+```
+
+![](figures/deep_vs_surf_labeling.png)
+
+*Figure XX: Line plots of the metabolite fraction labeled at the surface (25 meters) minus the fraction labeled at 175 meters.
+Incubations with added ammonia, arginine, and guanosine monophosphate (GMP) are shown.
+Points are the average of six biological replicates (incubations started in the morning and the evening are combined for additional power) and colors have been added solely for disambiguating purposes.
+Positive values indicate increased synthesis in the surface samples relative to those at depth.*
+
+```{r regression attempts currently failing, eval=FALSE}
+labeled_fracs %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(compound_name=="Creatinine") %>%
+  filter(amendment=="GMP") %>%
+  mutate(timepoint_num=ifelse(timepoint_num==0, 0.3, timepoint_num)) %>%
+  ggplot(aes(x=timepoint_num, y=labeled_frac, color=startime)) +
+  geom_point() +
+  geom_smooth(method=glm, formula=y~x, method.args=list(family=quasibinomial)) +
+  facet_wrap(~depth, ncol=1)
+
+reg_data <- labeled_fracs %>%
+  filter(!compound_name%in%maa_names) %>%
+  left_join(distinct(parametadata, shortname, amendment, depth, timepoint_num)) %>%
+  mutate(max_frac=max(labeled_frac[timepoint_num==max(timepoint_num)]), .by=c(amendment, compound_name, depth)) %>%
+  mutate(t10_labfrac = mean(labeled_frac[timepoint_num==sort(unique(timepoint_num))[4]]), .by=c(amendment, compound_name, depth))
+
+reg_data %>%
+  distinct(compound_name, amendment, depth, max_frac, t10_labfrac) %>%
+  mutate(category=ifelse(max_frac < 0.1, "Unlabeled", "Good")) %>%
+  mutate(category=ifelse(t10_labfrac <= 0, "Only T26 labeled", category)) %>%
+  mutate(category=ifelse(is.na(category), "DMSOP (unlabeled)", category)) %>%
+  count(amendment, category, depth) %>%
+  print(n=Inf)
+
+reg_fits <- reg_data %>%
+  filter(max_frac>0.1) %>%
+  filter(t10_labfrac>0) %>%
+  filter(compound_name!="Betonicine") %>%
+  filter(compound_name!="Arsenobetaine") %>%
+  filter(compound_name!="Gonyol") %>%
+  mutate(timepoint_num=ifelse(timepoint_num==0, 0.3, timepoint_num)) %>%
+  nest(data=c(shortname, timepoint_num, labeled_frac)) %>%
+  mutate(log_fit=map(data, ~glm(.x$labeled_frac~.x$timepoint_num, family="quasibinomial")),
+         coefs=map(log_fit, function(x)broom::tidy(x))) %>%
+  unnest(coefs) %>%
+  mutate(p_adj=p.adjust(p.value, method = "fdr"), .by = term) %>%
+  filter(p_adj<0.001) %>%
+  select(compound_name, amendment, depth, term, estimate) %>%
+  pivot_wider(names_from = term, values_from = estimate) %>%
+  rename(intercept=`(Intercept)`, slope=`.x$timepoint_num`) %>%
+  mutate(half_max=-intercept/slope, scale=1/slope) %>%
+  mutate(frac_at_T24 = 1/(1+exp(-(intercept+slope*24))))
+# reg_fits %>%
+#   select(compound_name, amendment, depth, intercept, slope) %>%
+#   pivot_longer(intercept:slope) %>%
+#   ggplot(aes(x=depth, y=value)) +
+#   geom_boxplot() +
+#   facet_grid(name~amendment, scales="free_y")
+# reg_fits %>%
+#   select(compound_name, amendment, depth, intercept, slope) %>%
+#   pivot_longer(intercept:slope) %>%
+#   pivot_wider(names_from=depth) %>%
+#   filter(!is.na(Deep)) %>%
+#   nest(data=c(compound_name, Surf, Deep)) %>%
+#   mutate(test=map(data, ~broom::tidy(t.test(.x$Surf, .x$Deep, paired=TRUE)))) %>%
+#   unnest(test)
+# labeled_fracs %>%
+#   left_join(parametadata %>% distinct(shortname, amendment, depth, timepoint_num)) %>%
+#   filter(timepoint_num==max(timepoint_num), .by=amendment) %>%
+#   summarise(avg_Tf_label=mean(labeled_frac), .by=c(compound_name, amendment, depth)) %>%
+#   ggplot() +
+#   geom_boxplot(aes(x=amendment, y=avg_Tf_label, color=depth))
+# 
+# labeled_fracs %>%
+#   left_join(parametadata %>% distinct(shortname, amendment, depth, timepoint_num)) %>%
+#   filter(timepoint_num==max(timepoint_num), .by=amendment) %>%
+#   summarise(avg_Tf_label=mean(labeled_frac), .by=c(compound_name, amendment, depth)) %>%
+#   pivot_wider(names_from=depth, values_from=avg_Tf_label) %>%
+#   ggplot() +
+#   geom_segment(aes(y=Surf, yend=Deep), x=1, xend=2) +
+#   facet_wrap(~amendment, nrow=1)
+# reg_fits %>%
+#   ggplot() +
+#   geom_segment(aes(y=frac_at_T24_Surf, yend=frac_at_T24_Deep,
+#                    x=half_max_Surf, xend=half_max_Deep, color=compound_name),
+#                data = reg_fits %>%
+#                  select(compound_name, depth, half_max, frac_at_T24) %>% 
+#                  pivot_wider(names_from = depth, values_from=c(half_max, frac_at_T24)) %>%
+#                  filter(!is.na(half_max_Deep))) +
+#   geom_point(aes(x=half_max, y=frac_at_T24, fill=depth), pch=21) +
+#   scale_fill_manual(values=c("black", "white"), breaks=c("Deep", "Surf"))
+
+reg_fits %>%
+  filter(amendment!="Nit") %>%
+  ggplot() +
+  geom_segment(aes(y=frac_at_T24_Surf, yend=frac_at_T24_Deep,
+                   x=half_max_Surf, xend=half_max_Deep, color=compound_name),
+               data = reg_fits %>%
+                 select(compound_name, amendment, depth, half_max, frac_at_T24) %>% 
+                 pivot_wider(names_from = depth, values_from=c(half_max, frac_at_T24)) %>%
+                 filter(!is.na(half_max_Deep))) +
+  geom_point(aes(x=half_max, y=frac_at_T24, fill=depth, label=compound_name), pch=21) +
+  scale_fill_manual(values=c("black", "white"), breaks=c("Deep", "Surf")) +
+  facet_wrap(~amendment, scales="free_x") +
+  theme(legend.position = "none")
+```
 

### 2025-05-08 thaas/thaa_analysis.R
Adding THAA data to analysis

@@ -1,0 +1,283 @@
+
+
+thaa_mzs <- read_csv("thaas/AminoAcid_masses_paragon1.csv") %>%
+  set_names(c("compound_name", "cmpd_mz", "cmpd_rt")) %>%
+  mutate(compound_name=str_replace(compound_name, "Phenylalaine", "Phenylalanine"))
+
+ms_files <- list.files("thaas/mzMLs", pattern = "mzML", full.names = TRUE)
+msdata <- grabMSdata(ms_files, grab_what=c("EIC", "metadata"), mz = thaa_mzs$cmpd_mz, ppm = 20)
+msdata$BPC %>% qplotMS1data(color_col="filename")
+
+metathaa <- data.frame(filename=basename(ms_files)) %>%
+  mutate(samp_type=str_extract(filename, "Blk|Smp|Std|Poo")) %>%
+  mutate(amendment=str_extract(filename, "Amm|Nit|Arg|GMP")) %>%
+  mutate(timepoint=str_extract(filename, "T\\d+")) %>%
+  mutate(timepoint=factor(timepoint, levels=c("T0", "T1", "T3", "T10", "T26", "T73"))) %>%
+  mutate(depth=str_extract(filename, "Surf|Deep")) %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
+  mutate(tripl=str_extract(filename, "[A-F](?=\\.mzML)")) %>%
+  mutate(startime=str_extract(filename, "Morn|Eve")) %>%
+  mutate(startime=case_when(
+    samp_type=="Smp" & tripl%in%c("D", "E", "F")~"Morn",
+    TRUE~startime
+  )) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
+  mutate(tripl=case_when(
+    tripl=="D"~"A",
+    tripl=="E"~"B",
+    tripl=="F"~"C",
+    TRUE~tripl
+  )) %>%
+  left_join(msdata$metadata %>% distinct(filename, timestamp))
+write_csv(metathaa, "thaas/metathaa.csv")
+
+
+
+thaa_mzs %>%
+  pwalk(function(compound_name, cmpd_mz, cmpd_rt){
+    gp <- msdata$MS1[mz%between%pmppm(cmpd_mz, ppm = 10)] %>%
+      filter(rt%between%c(cmpd_rt-0.5, cmpd_rt+0.5)) %>%
+      qplotMS1data() +
[…]

### 2025-05-08 thaas/thaa_analysis.R
Adding more THAA plots/analysis

@@ -248,7 +248,10 @@ quant_data <- all_iso_areas %>%
   left_join(cal_curves) %>%
   mutate(conc_um_in_vial=norm_area/slope) %>%
   mutate(env_conc_nm=conc_um_in_vial/100*1000) %>%
-  filter(iso_name!="Proline, 13C2, 15N1")
+  filter(iso_name!="Proline, 13C2, 15N1") %>%
+  filter(iso_name!="Proline, 13C3, 15N1") %>%
+  filter(iso_name!="Proline, 13C4, 15N1") %>%
+  filter(compound_name!="Lysine")
 write_csv(quant_data, "thaas/quant_data.csv")
 
 
@@ -281,3 +284,75 @@ quant_data %>%
            position="fill",
            color="black", width=1) +
   facet_nested(startime+depth~amendment+timepoint)
+
+pdf("thaas/all_cmpd_labels.pdf", width=9, height = 5)
+for(cmpd_i in setdiff(unique(quant_data$compound_name), c("Citrulline", "Ornithine", "Taurine"))){
+  print(cmpd_i)
+  gp <- quant_data %>%
+    filter(compound_name==cmpd_i) %>%
+    left_join(metathaa) %>%
+    drop_na() %>%
+    ggplot() +
+    geom_col(aes(x=tripl, y=env_conc_nm, fill=iso_name), 
+             position="fill",
+             color="black", width=1) +
+    facet_nested(depth+startime~amendment+timepoint) +
+    ggtitle(cmpd_i)
+  print(gp)
+}
+dev.off()
+
+
+
+
+quant_data %>%
+  # filter(compound_name=="Histidine") %>%
+  filter(compound_name=="Tyrosine") %>%
+  left_join(metathaa) %>%
+  drop_na() %>%
+  ggplot(aes(x=tripl, y=env_conc_nm, fill=iso_name, label=iso_name)) +
+  geom_col(position="fill", color="black", width=1) +
+  facet_nested(depth+startime~amendment+timepoint)
+plotly::ggplotly()
+# Tyrosine, 13C4, 15N1 from GMP?
+# Phenylalanine, 13C4, 15N1 from GMP?
+# Histidine, 13C5, 15N3 from GMP?
+
+
+
+quant_data %>%
+  filter(compound_name%in%c("Phenylalanine", "Tyrosine")) %>%
[…]

### 2025-05-08 manuscript/manuscript.Rmd
Adding big heatmap to supplement

@@
   pull(compound_name) %>%
-  str_remove_all("L-")
+  str_remove_all("^L-|(?<=)/L-") %>%
+  c("5-Methylcytosine")
 
@@
   mutate(depth=factor(depth, levels=c("Surf", "Deep"), labels=c("25m", "175m"))) %>%
-  mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
+  mutate(compound_name=str_remove_all(compound_name, "^L-|(?<=/)L-")) %>%
   mutate(compound_name=factor(compound_name, cmpd_order)) %>%
@@
   mutate(compound_class=factor(compound_class, levels=c("Amino\nacid", "Nucleic\nacid", "MAA", "Betaine", "N-free", "Other"))) %>%
+  # filter(compound_class=="Other") %>%
   ggplot() +
   geom_raster(aes(x=timepoint, y=tripl, fill=labeled_frac)) +
-  scale_fill_gradientn(colours = viridis::viridis(10, alpha = seq(0, 1, length.out=10), direction = -1), limits=c(0, 1)) +
+  scale_fill_gradientn(colours = viridis::viridis(10, alpha = seq(0, 1, length.out=10), direction = -1), limits=c(0, 1), breaks=c(0, 0.25, 0.5, 0.75, 1), labels=c("0%<br>(Unlabeled)", "", "50%", "", "100%<br>(Fully labeled)")) +
   scale_y_discrete(expand=expansion()) +
-  scale_x_discrete(guide = guide_axis(n.dodge = 3), expand=expansion()) +
+  scale_x_discrete(guide = guide_axis(n.dodge = 5), expand=expansion()) +
   facet_nested(compound_class+compound_name~amendment+depth+startime, scales = "free_x", switch="y", strip=strip_nested(size = "variable")) +
@@
   theme_bw() +
-  theme(panel.spacing.x=unit(0.5, "lines"), 
+  theme(panel.spacing.x=unit(0.2, "lines"), 
         panel.spacing.y = unit(0, "lines"),
@@
         axis.ticks.y = element_blank(),
+        # axis.ticks.length.x=unit(1:5, "lines"),
+        # axis.text.x = element_markdown(vjust = -5:-1),
         strip.text.y.left = element_text(angle=0, hjust = 1),
-        strip.background.y = element_rect(fill=NA, color=NA),
+        # strip.background.y = element_rect(fill=NA, color=NA),
         panel.grid.major = element_blank(), 
@@
       legend.ticks = element_line(color="black"),
-      legend.text = ggtext::element_markdown(hjust=c(0, 0, 0.5, 1, 1)))
+      legend.text = ggtext::element_markdown(hjust=c(0, 0.5, 0.5, 0.5, 1)))
     )
@@
 
-# ggsave(plot = heatmap_gp, "~/../Desktop/para_map.pdf", device = "pdf", 
-#        width = 10, height = 12)
+ggsave(plot = heatmap_gp, "manuscript/figures/supp_para_map.pdf", device = "pdf",
+       width = 10, height = 12)
 ```

### 2025-05-08 manuscript/manuscript.Rmd
Adding MAA results

@@
 
+Same basic pattern in the evening except MAAs, guanosine, proline?
+
+```{r fig MAA labeling patterns}
+chosen_maas <- c("Mycosporine-glycine?", "Shinorine?", "Palythine?",
+                 "Palythene/Usujirene?", "Also Palythene/Usujirene?", 
+                 "Also Porphyra-334?")
+labeled_fracs %>%
+  filter(compound_name%in%chosen_maas) %>%
+  mutate(compound_name=factor(compound_name, levels=chosen_maas)) %>%
+  arrange(compound_name) %>%
+  mutate(compound_name=str_remove(compound_name, "\\?$")) %>%
+  mutate(compound_name=str_replace(compound_name, "Also Palythene/Usujirene", "Palythene or\n usujirene")) %>%
+  mutate(compound_name=str_replace(compound_name, "Palythene/Usujirene", "Usujirene or\n palythene")) %>%
+  mutate(compound_name=str_remove(compound_name, "Also ")) %>%
+  mutate(compound_name=str_replace(compound_name, "-glycine", "-\nglycine")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  left_join(parametadata, relationship = "many-to-many") %>%
+  filter(samp_type=="Smp") %>%
+  filter(IS_type=="noIS") %>%
+  filter(amendment%in%c("Amm", "Nit")) %>%
+  mutate(amendment=factor(amendment, levels=c("Amm", "Nit"), labels=c("+Ammonia", "+Nitrate"))) %>%
+  mutate(clock_time=ifelse(startime=="Morn", 6+timepoint_num, 18+timepoint_num)) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start"))) %>%
+  ggplot() +
+  annotate("rect", xmin=-Inf, xmax = 6, ymin=-Inf, ymax=Inf, fill="grey80") +
+  # annotate("rect", xmin=6, xmax = 18, ymin=-Inf, ymax=Inf, fill="#d0bd65") +
+  annotate("rect", xmin=18, xmax = 30, ymin=-Inf, ymax=Inf, fill="grey80") +
+  # annotate("rect", xmin=30, xmax = 42, ymin=-Inf, ymax=Inf, fill="#d0bd65") +
+  annotate("rect", xmin=42, xmax = Inf, ymin=-Inf, ymax=Inf, fill="grey80") +
+  geom_hline(yintercept = c(0.25, 0.5, 0.75, 1), color="grey90") +
+  # annotate("rect", xmin=-Inf, xmax=Inf, ymin=-Inf, ymax = Inf, fill="white", alpha=0.3) +
+  geom_smooth(aes(x=clock_time, y=labeled_frac, color=startime),
+              method=nlsLM, formula="y~a*x^b+d", se = FALSE,
+              method.args=list(start=list(a=1, b=2, d=1),
+                               control=nls.lm.control(maxiter = 500))) +
+  # geom_smooth(aes(x=clock_time, y=labeled_frac, group=startime),
+  #             method=glm, formula = "y~x", method.args=list(family="quasibinomial"), 
+  #             color="grey10", se=FALSE) +
+  # geom_point(aes(x=clock_time, y=labeled_frac, shape=startime), color="black", 
+  #            fill="white", size=3) +
+  geom_point(aes(x=clock_time, y=labeled_frac, shape=startime, fill=startime),
+             color="black", size=2, alpha=0.7) +
+  facet_nested(compound_name~amendment) +
+  scale_x_continuous(breaks = c(6, 18, 30, 42), labels = c("6AM", "6PM", "6AM", "")) +
+  scale_y_continuous(labels = scales::label_percent(),
+                     limits = c(0, 1), breaks = c(0, 0.5, 1)) +
+  scale_shape_manual(values=c(24, 25), breaks = c("6AM start", "6PM start")) +
+  scale_fill_manual(values=c("#f4bb23", "#0b505c"), breaks = c("6AM start", "6PM start"),
+                    aesthetics = c("color", "fill")) +
+  labs(x=NULL, color=NULL, shape=NULL, fill=NULL, y="Fraction of MAA containing at least one 15N atom") +
+  theme_bw() +
+  theme(axis.text.y = element_markdown(vjust=c(0, 0.5, 1)),
+        axis.title.x = element_blank(),
+        legend.position = "inside",
+        legend.margin = margin(0.1,0.25,0.1,0.1, unit="lines"),
+        legend.position.inside = c(1, 1),
+        legend.justification = c(1, 1), 
+        legend.background = element_rect(fill="#FFFFFFAA", color="black", linewidth = 0)) +
+  guides(shape=guide_legend(override.aes = list(size=3, stroke=1.2, alpha=1, linetype=0)))
+
+ggsave(filename = "MAA_curves.png", device = "png", path = "manuscript/figures",
+       width = 3, height = 6, units = "in", dpi = 300)
+```
+
+![](figures/MAA_curves.png)
+
+*Figure XX: Fraction of UV-reactive mycosporine-like amino acids (MAAs) containing at least one 15N label over one day during the treatments amended with 15N labeled ammonia or nitrate.
+Incubations started at 6AM are shown in yellow while incubations started at 6PM are shown in blue, with grey background regions highlighting nighttime.
+Three replicates are included at each of the 5 timepoints and best-fit power law curves have been fit behind the data.*
+
+Deep labeling is slower and much more limited for ammonia
+
+  - Asp/Asn especially
+  - Ectoine major exception (ref mcparland for bacterial signal)
+
+Nitrate surface morning surprisingly limited
+
+  - "out of order" labeling in MAAs and guanine, maybe also arginine, taurine, NAc-lysine, maybe maybe theanine being unlabeled because it would mean diatoms don’t make it
+
+Organic compounds equally bioavailable at surface and deep and used very similarly to each other
+
+  - Exceptions: guanosine (not guanine!) for GMP amendments
+  - Exceptions: arginine, citrulline, maybe ornithine for Arg (signal goes away after a while and unclear why??)
+
+Organics seem to be largely used for ammonia rather than as-is
+
+  - Carbon backbone traced into glutamate from arginine
+  - Carbon backbone traced into other nucleobases from GMP
+  - Possibly due to way too much of a given compound, useless to have that much Arg/GMP
+  - Not very much fully-labeled guanine despite guanosine being easily cleavable, instead it’s mostly made from scratch (CHECK THIS)
+  - Creatine for being mostly 2x15N 1x13C labeled from arginine
+
 ```{r fig surface vs deep labeling rates}
@@
 
-```{r fig MAA labeling patterns}
-chosen_maas <- c("Mycosporine-glycine?", "Shinorine?", "Palythine?",
-                 "Palythene/Usujirene?", "Also Palythene/Usujirene?", 
-                 "Also Porphyra-334?")
-labeled_fracs %>%
-  filter(compound_name%in%chosen_maas) %>%
-  mutate(compound_name=factor(compound_name, levels=chosen_maas)) %>%
-  arrange(compound_name) %>%
-  mutate(compound_name=str_remove(compound_name, "\\?$")) %>%
-  mutate(compound_name=str_replace(compound_name, "Also Palythene/Usujirene", "Palythene or\n usujirene")) %>%
-  mutate(compound_name=str_replace(compound_name, "Palythene/Usujirene", "Usujirene or\n palythene")) %>%
-  mutate(compound_name=str_remove(compound_name, "Also ")) %>%
-  mutate(compound_name=str_replace(compound_name, "-glycine", "-\nglycine")) %>%
-  mutate(compound_name=fct_inorder(compound_name)) %>%
-  left_join(parametadata, relationship = "many-to-many") %>%
-  filter(samp_type=="Smp") %>%
-  filter(IS_type=="noIS") %>%
-  filter(amendment%in%c("Amm", "Nit")) %>%
-  mutate(amendment=factor(amendment, levels=c("Amm", "Nit"), labels=c("+Ammonia", "+Nitrate"))) %>%
-  mutate(clock_time=ifelse(startime=="Morn", 6+timepoint_num, 18+timepoint_num)) %>%
-  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start"))) %>%
-  ggplot() +
-  annotate("rect", xmin=-Inf, xmax = 6, ymin=-Inf, ymax=Inf, fill="grey80") +
-  # annotate("rect", xmin=6, xmax = 18, ymin=-Inf, ymax=Inf, fill="#d0bd65") +
-  annotate("rect", xmin=18, xmax = 30, ymin=-Inf, ymax=Inf, fill="grey80") +
-  # annotate("rect", xmin=30, xmax = 42, ymin=-Inf, ymax=Inf, fill="#d0bd65") +
-  annotate("rect", xmin=42, xmax = Inf, ymin=-Inf, ymax=Inf, fill="grey80") +
-  geom_hline(yintercept = c(0.25, 0.5, 0.75, 1), color="grey90") +
-  # annotate("rect", xmin=-Inf, xmax=Inf, ymin=-Inf, ymax = Inf, fill="white", alpha=0.3) +
-  geom_smooth(aes(x=clock_time, y=labeled_frac, color=startime),
-              method=nlsLM, formula="y~a*x^b+d", se = FALSE,
-              method.args=list(start=list(a=1, b=2, d=1),
-                               control=nls.lm.control(maxiter = 500))) +
-  # geom_smooth(aes(x=clock_time, y=labeled_frac, group=startime),
-  #             method=glm, formula = "y~x", method.args=list(family="quasibinomial"), 
-  #             color="grey10", se=FALSE) +
-  # geom_point(aes(x=clock_time, y=labeled_frac, shape=startime), color="black", 
-  #            fill="white", size=3) +
-  geom_point(aes(x=clock_time, y=labeled_frac, shape=startime, fill=startime),
-             color="black", size=2, alpha=0.7) +
-  facet_nested(compound_name~amendment) +
-  scale_x_continuous(breaks = c(6, 18, 30, 42), labels = c("6AM", "6PM", "6AM", "")) +
-  scale_y_continuous(labels = scales::label_percent(),
-                     limits = c(0, 1), breaks = c(0, 0.5, 1)) +
-  scale_shape_manual(values=c(24, 25), breaks = c("6AM start", "6PM start")) +
-  scale_fill_manual(values=c("#f4bb23", "#0b505c"), breaks = c("6AM start", "6PM start"),
-                    aesthetics = c("color", "fill")) +
-  labs(x=NULL, color=NULL, shape=NULL, fill=NULL, y="Fraction of MAA containing at least one 15N atom") +
-  theme_bw() +
-  theme(axis.text.y = element_markdown(vjust=c(0, 0.5, 1)),
-        axis.title.x = element_blank(),
-        legend.position = "inside",
-        legend.margin = margin(0.1,0.25,0.1,0.1, unit="lines"),
-        legend.position.inside = c(1, 1),
-        legend.justification = c(1, 1), 
-        legend.background = element_rect(fill="#FFFFFFAA", color="black", linewidth = 0)) +
-  guides(shape=guide_legend(override.aes = list(size=3, stroke=1.2, alpha=1, linetype=0)))
-
-ggsave(filename = "MAA_curves.png", device = "png", path = "manuscript/figures",
-       width = 3, height = 6, units = "in", dpi = 300)
-```
-
-![](figures/MAA_curves.png)
-
-*Figure XX: Fraction of UV-reactive mycosporine-like amino acids (MAAs) containing at least one 15N label over one day during the treatments amended with 15N labeled ammonia or nitrate.
-Incubations started at 6AM are shown in yellow while incubations started at 6PM are shown in blue, with grey background regions highlighting nighttime.
-Three replicates are included at each of the 5 timepoints and best-fit power law curves have been fit behind the data.*
+Untargeted metabolites? Clustering their % labeled patterns with the targeted metabolites to see what groups they might belong to
 

### 2025-05-08 manuscript/manuscript.Rmd
Adding preliminary results skeleton

@@
 ```{r setup}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE)
 library(tidyverse)
@@
 library(minpack.lm)
+options(pillar.sigfig = 7, pillar.print_max = 100)
 
@@
   mutate(frac_at_T24 = 1/(1+exp(-(intercept+slope*24))))
+
+splividis <- function(n, ...){
+  vir <- viridis::viridis(n, ...)
+  if(n%%2==0){
+    vir[as.numeric(matrix(1:n, ncol=2, byrow = TRUE))]
+  } else {
+    index_vec <- suppressWarnings(as.numeric(matrix(1:n, ncol=2, byrow = TRUE)))
+    vir[head(index_vec, -1)]
+  }
+}
 ```
@@
 ## Results
+
+### Environmental context of PARAGON cruise
+
+Make plots from Rhea's data? Would love to add her as collaborator
+
+Maybe just ref to PARAGON synthesis paper, should include Angel's IFCB data showing mostly Hemiaulus and its disappearance over the course of the cruise
+
+Ref to Fuyan's published paper or remake the graph here?
+
+### Rapid uptake and use of all N substrates
+
+```{r}
+env_concs <- read_csv("targeted/final_env_concs.csv")
+env_concs %>%
+  mutate(mean_nm=mean(conc_nm_in_env), .by=compound_name) %>%
+  arrange(desc(mean_nm)) %>%
+  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", "Glutamate", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:19], compound_name, "Other")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  summarise(conc_nm=sum(conc_nm_in_env), .by=c(compound_name, filename)) %>%
+  left_join(parametadata) %>%
+  ggplot(aes(x=tripl, y=conc_nm, fill=compound_name)) +
+  geom_col(color="black", width=1) +
+  facet_nested(depth+startime~amendment+timepoint, switch="x") +
+  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
+  scale_fill_manual(values = c(splividis(19), "grey")) +
+  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
+  theme_bw() +
+  theme(
+    panel.spacing.x = unit(0, "lines"),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
+```
+
+```{r}
+peak_areas %>%
+  filter(!str_detect(iso_name, "13C0, 15N0")) %>%
+  filter(!compound_name%in%c("Urea", "L-Valine", "Oxalic acid", "Malic acid", 
+                             "Acetylcholine", "Homarine")) %>%
+  left_join(parametadata %>% distinct(filename, samp_type, amendment)) %>%
+  filter(samp_type=="Smp") %>%
+  left_join(read_csv("targeted/rf_df.csv")) %>%
+  mutate(conc_nm=area/rf) %>%
+  filter(!is.na(conc_nm)) %>%
+  mutate(mean_nm=mean(conc_nm), .by=compound_name) %>%
+  arrange(desc(mean_nm)) %>%
+  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", 
+                              "Glutamate", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:19], 
+                              compound_name, "Other")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  summarise(conc_nm=sum(conc_nm), .by=c(compound_name, filename)) %>%
+  left_join(parametadata) %>%
+  ggplot(aes(x=tripl, y=conc_nm, fill=compound_name)) +
+  geom_col(color="black", width=1) +
+  facet_nested(depth+startime~amendment+timepoint, switch="x") +
+  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
+  scale_fill_manual(values = c(splividis(19), "grey")) +
+  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
+  theme_bw() +
+  theme(
+    panel.spacing.x = unit(0, "lines"),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
+```
+
+```{r}
+cbind(
+  peak_areas %>%
+    # filter(iso_name=="L-Glutamine, 13C0, 15N0") %>%
+    filter(iso_name=="Glycine betaine, 13C0, 15N0") %>%
+    filter(str_detect(filename, "T0")) %>%
+    select(filename, area),
+  read_csv("targeted/wIS_areas.csv") %>%
+    # filter(compound_name=="L-Glutamine") %>%
+    filter(compound_name=="Glycine betaine") %>%
+    filter(str_detect(filename, "T0")) %>%
+    select(wIS_filename=filename, wIS_area=area)
+) %>%
+  # with(lm(wIS_area~area)) %>% summary()
+  ggplot() +
+  geom_point(aes(x=area, y=wIS_area)) +
+  geom_abline(slope = 1.5, intercept = 0)
+```
+
+Total fraction of PN that was visible in metabolites, maybe big stacked barplot like usual?
+
+Total number of compounds labeled over time (line plot?)
+
+```{r thaa data}
+thaa_quant <- read_csv("thaas/quant_data.csv")
+metathaa <- read_csv("thaas/metathaa.csv") %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve")))
+aa_ncs <- all_stans %>%
+  filter(compound_type=="Amino Acid") %>%
+  mutate(compound_name=str_remove(compound_name, "^L-")) %>%
+  distinct(compound_name, n_C, n_N) %>%
+  add_row(compound_name="Phenylalanine", n_C=9, n_N=1)
+
+all_thaa_plot <- thaa_quant %>%
+  filter(!str_detect(iso_name, "2H")) %>%
+  left_join(metathaa) %>%
+  filter(samp_type=="Smp") %>%
+  left_join(aa_ncs) %>%
+  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
+  mutate(n_N_lab=as.numeric(str_extract(iso_name, "(?<=15N)\\d+"))) %>%
+  mutate(n_N_unlab=n_N-n_N_lab) %>%
+  mutate(nM_N_lab=env_conc_nm*n_N_lab, nM_N_unlab=env_conc_nm*n_N_unlab) %>%
+  select(filename, compound_name, iso_name, nM_N_lab, nM_N_unlab) %>%
+  pivot_longer(c(nM_N_lab, nM_N_unlab)) %>%
+  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
+  left_join(metathaa) %>%
+  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
+  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
+  summarise(total_uM=mean(total_uM), .by=c(amendment, depth, startime, timepoint, name)) %>%
+  arrange(depth, startime) %>%
+  mutate(x_axis=paste(depth, startime)) %>%
+  mutate(x_axis=fct_inorder(x_axis)) %>%
+  ggplot(aes(x=x_axis, y=total_uM, fill=name)) +
+  geom_col(color="black", width=1) +
+  facet_nested(.~amendment, scales="free_y") +
+  theme_bw() +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
+  scale_fill_manual(breaks = c("nM_N_unlab", "nM_N_lab"), values=c("grey80", "grey20"), 
+                    labels=c("Unlabeled", "Labeled")) +
+  labs(x=NULL, y="Total uM N in THAAs", fill=NULL) +
+  theme(
+    axis.text.x=element_blank()
+  )
+
+thaa_plot_data <- thaa_quant %>%
+  left_join(metathaa) %>%
+  filter(samp_type=="Smp") %>%
+  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
+  summarise(avg_env_conc_nm=mean(env_conc_nm, na.rm=TRUE), .by=c(compound_name, iso_name, amendment, timepoint, startime, depth)) %>%
+  filter(compound_name%in%c("Glutamic acid", "Arginine")) %>%
+  mutate(iso_name=case_when(
+    str_detect(iso_name, "13C0, 15N0")~"Unlabeled",
+    str_detect(iso_name, "13C0, 15N1")~"Singly 15N labeled",
+    str_detect(iso_name, "13C1, 15N0")~"Singly 13C labeled",
+    str_detect(iso_name, "Glutamic acid, 13C5, 15N1")~"Fully C and N\nlabeled glutamate",
+    str_detect(iso_name, "Arginine, 13C0, 15N3")~"3x 15N\nlabeled arginine",
+    str_detect(iso_name, "Arginine, 13C0, 15N4")~"Fully N\nlabeled arginine",
+    str_detect(iso_name, "Arginine, 13C6, 15N4")~"Fully C and N\nlabeled arginine",
+    TRUE~"Other labeled"
+  )) %>%
+  mutate(iso_name=factor(iso_name, levels=c(
+    "Unlabeled", "Singly 15N labeled", "Singly 13C labeled", 
+    "3x 15N\nlabeled arginine", "Fully N\nlabeled arginine", 
+    "Fully C and N\nlabeled arginine", "Fully C and N\nlabeled glutamate", 
+    "Other labeled"))) %>%
+  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
+  arrange(depth, startime) %>%
+  mutate(x_axis=paste(depth, startime)) %>%
+  mutate(x_axis=fct_inorder(x_axis))
+cpal_breaks <- levels(thaa_plot_data$iso_name)
+cpal_cols <- c("grey80", viridis::viridis(length(cpal_breaks)-2), "grey20")
+arg_glut_thaa_plot <- ggplot(thaa_plot_data) +
+  geom_col(aes(x=x_axis, y=avg_env_conc_nm, fill=iso_name),
+           position="fill",
+           color="black", width=1) +
+  facet_wrap2(compound_name~amendment, ncol=5, strip = strip_nested(bleed = TRUE)) +
+  theme_bw() +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_continuous(labels=scales::label_percent(), breaks=c(0, 0.5, 1),
+                     expand = expansion()) +
+  scale_fill_manual(breaks = cpal_breaks, values = cpal_cols) +
+  labs(x=NULL, y="Fraction labeled", fill=NULL) +
+  theme(
+    axis.text.x=element_text(angle=90, hjust=1, vjust = 0.5)
+  )
+total_thaa_gp <- all_thaa_plot / arg_glut_thaa_plot + plot_layout(heights = c(0.4, 0.6))
+ggsave("total_thaa_gp.png", total_thaa_gp, device = "png", path = "manuscript/figures",
+       width = 6.5, height = 5.5, units = "in", dpi = 300)
+```
+
+![](figures/total_thaa_gp.png)
+  - THAAs were 30-50% of total PC (~2 uM, need actual data) and 70-100% PN (~0.35uM)
+  - Maybe more labeling happening from inorganics in evening?
+  - 50% turnover of arginine in proteins in 3 days per +Arg data, less at surface than deep (??)
+  - 30-50% Glu/Gln turnover in proteins in one day per Amm data
+  - Glutamate, proline fully labeled labeled from arginine
+  - Phenylalanine, tyrosine, histidine(?) labeled from GMP
+  - How is there so much nitrate labeling when I don't see that in the metabolites?
+
+```{r thaa supplement figures}
+thaa_quant %>%
+  left_join(metathaa) %>%
+  filter(samp_type=="Smp") %>%
+  ggplot() +
+  geom_col(aes(x=tripl, y=env_conc_nm, fill=compound_name), 
+           color="black", width=1, position="fill") +
+  facet_nested(depth+startime~amendment+timepoint) +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_continuous(labels=scales::label_percent(), breaks=c(0, 0.5, 1),
+                     expand = expansion()) +
+  labs(x="Triplicate", y="Fraction labeled", fill=NULL) +
+  theme_bw() +
+  theme(
+    axis.text.y=element_markdown(vjust=c(0, 0.5, 1))
+  )
+
+THAA_mol_C_gp <- thaa_quant %>%
+  filter(!str_detect(iso_name, "2H")) %>%
+  left_join(metathaa) %>%
+  filter(samp_type=="Smp") %>%
+  left_join(aa_ncs) %>%
+  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
+  mutate(n_C_lab=as.numeric(str_extract(iso_name, "(?<=13C)\\d+"))/n_C) %>%
+  mutate(n_C_unlab=n_C-n_C_lab) %>%
+  mutate(nM_C_lab=env_conc_nm*n_C_lab, nM_C_unlab=env_conc_nm*n_C_unlab) %>%
+  select(filename, compound_name, iso_name, nM_C_lab, nM_C_unlab) %>%
+  pivot_longer(c(nM_C_lab, nM_C_unlab)) %>%
+  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
+  left_join(metathaa) %>%
+  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
+  ggplot(aes(x=tripl, y=total_uM, fill=name)) +
+  geom_col(color="black", width=1) +
+  facet_nested(depth+startime~amendment+timepoint) +
+  theme_bw() +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
+  scale_fill_manual(breaks = c("nM_C_unlab", "nM_C_lab"), values=c("grey80", "grey20"), 
+                    labels=c("Unlabeled", "Labeled")) +
+  labs(x="Triplicate", y="Total C in THAAs (uM)", fill=NULL) +
+  theme(
+    legend.position = "top"
+  )
+THAA_mol_N_gp <- thaa_quant %>%
+  filter(!str_detect(iso_name, "2H")) %>%
+  left_join(metathaa) %>%
+  filter(samp_type=="Smp") %>%
+  left_join(aa_ncs) %>%
+  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
+  mutate(n_N_lab=as.numeric(str_extract(iso_name, "(?<=15N)\\d+"))) %>%
+  mutate(n_N_unlab=n_N-n_N_lab) %>%
+  mutate(nM_N_lab=env_conc_nm*n_N_lab, nM_N_unlab=env_conc_nm*n_N_unlab) %>%
+  select(filename, compound_name, iso_name, nM_N_lab, nM_N_unlab) %>%
+  pivot_longer(c(nM_N_lab, nM_N_unlab)) %>%
+  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
+  left_join(metathaa) %>%
+  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
+  ggplot(aes(x=tripl, y=total_uM, fill=name)) +
+  geom_col(color="black", width=1) +
+  facet_nested(depth+startime~amendment+timepoint) +
+  theme_bw() +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
+  scale_fill_manual(breaks = c("nM_N_unlab", "nM_N_lab"), values=c("grey80", "grey20"), 
+                    labels=c("Unlabeled", "Labeled")) +
+  labs(x="Triplicate", y="Total N in THAAs (uM)", fill=NULL) +
+  theme(
+    legend.position = "top"
+  )
+all_thaa_NC_gp <- guide_area() + THAA_mol_C_gp + THAA_mol_N_gp + plot_layout(guides = "collect", heights = c(0.1, 0.45, 0.45), design = c("1\n2\n3"))
+ggsave("thaas/all_thaa_NC_gp.png", width=6.5, height=8)
+
+pdf("thaas/all_cmpd_labels.pdf", width=9, height = 5)
+for(cmpd_i in setdiff(unique(thaa_quant$compound_name), c("Citrulline", "Ornithine", "Taurine"))){
+  print(cmpd_i)
+  gp <- thaa_quant %>%
+    filter(compound_name==cmpd_i) %>%
+    left_join(metathaa) %>%
+    drop_na() %>%
+    ggplot() +
+    geom_col(aes(x=tripl, y=env_conc_nm, fill=iso_name), 
+             position="fill",
+             color="black", width=1) +
+    facet_nested(depth+startime~amendment+timepoint) +
+    ggtitle(cmpd_i)
+  print(gp)
+}
+dev.off()
+```
+
+### Universal bioavailability of ammonia at surface
+
+The most rapid and extensive labeling was obtained during the ammonia-amended incubations of surface seawater.
+We observed significant labeling of glutamine, [and others] even at the T0 timepoint, indicating turnover times shorter than the time required to filter and flash-freeze the samples in liquid nitrogen (~20 minutes).
+
+DISCUSSION: the speed at which ammonia labeling happens hints at a high flux, low concentration ammonia cycle in the surface ocean since uptake was too fast and too widespread to be a transcriptional response, indicating that they have ammonia transporters plus GS/GOGAT/GDH always active.
+
+Synthesis rates?
+
+Aspartate labeled faster than glutamate - discuss pool sizes
+
+Slow labeling of osmolytes - discuss pool sizes
 

### 2025-05-09 metadata/parametadata_control.Rmd
Fixing bug in formula extractor

@@
   mutate(n_N=case_when(
-    !str_detect(formula, "C")~0,
-    str_detect(formula, "C") & is.na(n_N)~1,
+    !str_detect(formula, "N")~0,
+    str_detect(formula, "N") & is.na(n_N)~1,
     TRUE~n_N

### 2025-05-09 manuscript/manuscript.Rmd
Very rough first draft of the results

@@
     reference_docx: Frontiers_Template.docx
+csl: frontiers-in-marine-science.csl
 bibliography:
-- Exported Items.bib
   - r-packages.bib
-csl: frontiers-in-marine-science.csl
 ---
@@
 
-```{r setup}
+```{r setup, echo=FALSE}
 knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE)
@@
 library(minpack.lm)
-options(pillar.sigfig = 7, pillar.print_max = 100)
+options(pillar.sigfig = 7)
+options(pillar.print_min = 100)
 
@@
 all_stans <- read_csv("metadata/all_stans.csv")
+stan_patched <- all_stans %>%
+  filter(!compound_name=="beta-Glutamic acid") %>%
+  mutate(compound_name=str_replace(compound_name, "L-Glutamic acid", "L-Glutamic acid/beta-Glutamic acid")) %>%
+  filter(!compound_name=="L-Isoleucine") %>%
+  mutate(compound_name=str_replace(compound_name, "L-Leucine", "L-Leucine/L-Isoleucine")) %>%
+  filter(!compound_name=="L-Homoserine") %>%
+  mutate(compound_name=str_replace(compound_name, "L-Threonine", "L-Threonine/L-Homoserine")) %>%
+  filter(!compound_name=="Allopurinol") %>%
+  mutate(compound_name=str_replace(compound_name, "Hypoxanthine", "Hypoxanthine/Allopurinol"))
+
 maa_names <- c("Also Palythene/Usujirene?", "Shinorine?", "Mycosporine-glycine?", 
@@
                "Choline sulfate", "5-Hydroxyectoine", "L-Tyrosine")
+aa_ncs <- all_stans %>%
+  filter(compound_type=="Amino Acid") %>%
+  mutate(compound_name=str_remove(compound_name, "^L-")) %>%
+  distinct(compound_name, n_C, n_N) %>%
+  add_row(compound_name="Phenylalanine", n_C=9, n_N=1)
+
 samp_areas <- parametadata %>%
@@
 
-## Abstract
+# Abstract
 
-## Introduction
+Nitrogen availability controls productivity in much of the surface ocean.
+The form of this nitrogen controls who it is available to and how it is used.
+Here, we explored how four different forms of nitrogen (ammonia, nitrate, arginine, and guanosine monophosphate) were taken up and transformed in the North Pacific Subtropical Gyre using stable isotope labeling metabolomics.
+We repeated these experiments both at the surface and at the base of the euphotic zone as well as in the morning and evening.
+We found that ammonia is converted into glutamate in less than 20 minutes and results in labeling throughout the metabolome, though osmolytes like betaines had slow turnover rates.
+This process was slower at depth but independent of time of day.
+Nitrate amendments resulted in the labeling of only a small subset of molecules such as mycosporine-like amino acids, whose incorporation patterns suggest that this form of nitrogen was available to a subset of the community, likely eukaryotes.
+The two organic nitrogen amendments labeled the metabolome more slowly than the inorganic forms of nitrogen and were primarily converted into glutamate rather than used as intact substrates.
+They also showed the same usage at depth as they did at the surface, hinting at the major role that heterotrophic bacteria likely play in their uptake and use.
+These measurements reveal the forms and fluxes of organic nitrogen in the natural environment to provide constraints on biogeochemical models and better characterize the transformations between organic and inorganic nitrogen.
 
-Marine carbon is fixed into biomass at a rate equivalent to ten fully loaded Boeing 747 airplanes crashing into the ocean every second of every day [ref 50Pg IPCC(?), ref Boeing weights].
+# Introduction
+
+Marine carbon is fixed into biomass very quickly.
 In most of the surface ocean this flux is controlled by the availability of nitrogen, an element with a complex and incompletely characterized biogeochemical cycle [@Capone2008; @Moore2013; @Hutchins2022].
@@
 
-### Other things to maybe discuss:
+## Other things to maybe discuss:
 
@@
 
-### Other soundbites:
+## Other soundbites:
   - It is assumed that most fixed nitrogen becomes protein, presumably based on the large fraction of cell nitrogen in protein and previous research showing how quickly it enters the (free?) amino acid pool.
@@
 
-## Results
+# Results
 
-### Environmental context of PARAGON cruise
+## Environmental context of PARAGON cruise
 
-Make plots from Rhea's data? Would love to add her as collaborator
+PARAGON synthesis paper isn't out yet and I'm not sure it will be anytime soon
 
-Maybe just ref to PARAGON synthesis paper, should include Angel's IFCB data showing mostly Hemiaulus and its disappearance over the course of the cruise
+Particulate carbon and nitrogen numbers for PARAGON I (from Matt(?))
 
-Ref to Fuyan's published paper or remake the graph here?
+![](figures/borrowed/pcpn_PIandPII.png)
 
-### Rapid uptake and use of all N substrates
+Data from Rhea showing ammonia/nitrite concentrations - I sampled at 25m and 175m (nitrite max, below ammonia max). Does this mean I was getting a lot of AOA users?
 
-```{r}
-env_concs <- read_csv("targeted/final_env_concs.csv")
-env_concs %>%
-  mutate(mean_nm=mean(conc_nm_in_env), .by=compound_name) %>%
-  arrange(desc(mean_nm)) %>%
-  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", "Glutamate", compound_name)) %>%
-  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:19], compound_name, "Other")) %>%
-  mutate(compound_name=fct_inorder(compound_name)) %>%
-  summarise(conc_nm=sum(conc_nm_in_env), .by=c(compound_name, filename)) %>%
-  left_join(parametadata) %>%
-  ggplot(aes(x=tripl, y=conc_nm, fill=compound_name)) +
-  geom_col(color="black", width=1) +
-  facet_nested(depth+startime~amendment+timepoint, switch="x") +
-  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
-  scale_fill_manual(values = c(splividis(19), "grey")) +
-  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
-  theme_bw() +
-  theme(
-    panel.spacing.x = unit(0, "lines"),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()
-  )
-```
+![](figures/borrowed/rhea_ammonia_concs.png)
 
-```{r}
-peak_areas %>%
-  filter(!str_detect(iso_name, "13C0, 15N0")) %>%
-  filter(!compound_name%in%c("Urea", "L-Valine", "Oxalic acid", "Malic acid", 
-                             "Acetylcholine", "Homarine")) %>%
-  left_join(parametadata %>% distinct(filename, samp_type, amendment)) %>%
-  filter(samp_type=="Smp") %>%
-  left_join(read_csv("targeted/rf_df.csv")) %>%
-  mutate(conc_nm=area/rf) %>%
-  filter(!is.na(conc_nm)) %>%
-  mutate(mean_nm=mean(conc_nm), .by=compound_name) %>%
-  arrange(desc(mean_nm)) %>%
-  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", 
-                              "Glutamate", compound_name)) %>%
-  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:19], 
-                              compound_name, "Other")) %>%
-  mutate(compound_name=fct_inorder(compound_name)) %>%
-  summarise(conc_nm=sum(conc_nm), .by=c(compound_name, filename)) %>%
-  left_join(parametadata) %>%
-  ggplot(aes(x=tripl, y=conc_nm, fill=compound_name)) +
-  geom_col(color="black", width=1) +
-  facet_nested(depth+startime~amendment+timepoint, switch="x") +
-  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
-  scale_fill_manual(values = c(splividis(19), "grey")) +
-  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
-  theme_bw() +
-  theme(
-    panel.spacing.x = unit(0, "lines"),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()
-  )
-```
+Angel's IFCB data showing Hemiaulus disappearing over the course of the cruise (most of the Bacillariophyceae are Hemiaulus and she shows that in another plot)
 
-```{r}
-cbind(
-  peak_areas %>%
-    # filter(iso_name=="L-Glutamine, 13C0, 15N0") %>%
-    filter(iso_name=="Glycine betaine, 13C0, 15N0") %>%
-    filter(str_detect(filename, "T0")) %>%
-    select(filename, area),
-  read_csv("targeted/wIS_areas.csv") %>%
-    # filter(compound_name=="L-Glutamine") %>%
-    filter(compound_name=="Glycine betaine") %>%
-    filter(str_detect(filename, "T0")) %>%
-    select(wIS_filename=filename, wIS_area=area)
-) %>%
-  # with(lm(wIS_area~area)) %>% summary()
-  ggplot() +
-  geom_point(aes(x=area, y=wIS_area)) +
-  geom_abline(slope = 1.5, intercept = 0)
-```
+![](figures/borrowed/angel_hemiaulus.png)
 
-Total fraction of PN that was visible in metabolites, maybe big stacked barplot like usual?
+Fuyan reports the abundances of various bacterial groups (with the caveat that the SASVs have to have been found in the 4000m depth traps?)
 
-Total number of compounds labeled over time (line plot?)
+![](figures/borrowed/fuyan_composition_bact.png)
 
-```{r thaa data}
-thaa_quant <- read_csv("thaas/quant_data.csv")
-metathaa <- read_csv("thaas/metathaa.csv") %>%
-  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
-  mutate(startime=factor(startime, levels=c("Morn", "Eve")))
-aa_ncs <- all_stans %>%
-  filter(compound_type=="Amino Acid") %>%
-  mutate(compound_name=str_remove(compound_name, "^L-")) %>%
-  distinct(compound_name, n_C, n_N) %>%
-  add_row(compound_name="Phenylalanine", n_C=9, n_N=1)
-
-all_thaa_plot <- thaa_quant %>%
-  filter(!str_detect(iso_name, "2H")) %>%
-  left_join(metathaa) %>%
-  filter(samp_type=="Smp") %>%
-  left_join(aa_ncs) %>%
-  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
-  mutate(n_N_lab=as.numeric(str_extract(iso_name, "(?<=15N)\\d+"))) %>%
-  mutate(n_N_unlab=n_N-n_N_lab) %>%
-  mutate(nM_N_lab=env_conc_nm*n_N_lab, nM_N_unlab=env_conc_nm*n_N_unlab) %>%
-  select(filename, compound_name, iso_name, nM_N_lab, nM_N_unlab) %>%
-  pivot_longer(c(nM_N_lab, nM_N_unlab)) %>%
-  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
-  left_join(metathaa) %>%
-  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
-  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
-  summarise(total_uM=mean(total_uM), .by=c(amendment, depth, startime, timepoint, name)) %>%
-  arrange(depth, startime) %>%
-  mutate(x_axis=paste(depth, startime)) %>%
-  mutate(x_axis=fct_inorder(x_axis)) %>%
-  ggplot(aes(x=x_axis, y=total_uM, fill=name)) +
-  geom_col(color="black", width=1) +
-  facet_nested(.~amendment, scales="free_y") +
-  theme_bw() +
-  scale_x_discrete(expand = expansion()) +
-  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
-  scale_fill_manual(breaks = c("nM_N_unlab", "nM_N_lab"), values=c("grey80", "grey20"), 
-                    labels=c("Unlabeled", "Labeled")) +
-  labs(x=NULL, y="Total uM N in THAAs", fill=NULL) +
-  theme(
-    axis.text.x=element_blank()
-  )
-
-thaa_plot_data <- thaa_quant %>%
-  left_join(metathaa) %>%
-  filter(samp_type=="Smp") %>%
-  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
-  summarise(avg_env_conc_nm=mean(env_conc_nm, na.rm=TRUE), .by=c(compound_name, iso_name, amendment, timepoint, startime, depth)) %>%
-  filter(compound_name%in%c("Glutamic acid", "Arginine")) %>%
-  mutate(iso_name=case_when(
-    str_detect(iso_name, "13C0, 15N0")~"Unlabeled",
-    str_detect(iso_name, "13C0, 15N1")~"Singly 15N labeled",
-    str_detect(iso_name, "13C1, 15N0")~"Singly 13C labeled",
-    str_detect(iso_name, "Glutamic acid, 13C5, 15N1")~"Fully C and N\nlabeled glutamate",
-    str_detect(iso_name, "Arginine, 13C0, 15N3")~"3x 15N\nlabeled arginine",
-    str_detect(iso_name, "Arginine, 13C0, 15N4")~"Fully N\nlabeled arginine",
-    str_detect(iso_name, "Arginine, 13C6, 15N4")~"Fully C and N\nlabeled arginine",
-    TRUE~"Other labeled"
-  )) %>%
-  mutate(iso_name=factor(iso_name, levels=c(
-    "Unlabeled", "Singly 15N labeled", "Singly 13C labeled", 
-    "3x 15N\nlabeled arginine", "Fully N\nlabeled arginine", 
-    "Fully C and N\nlabeled arginine", "Fully C and N\nlabeled glutamate", 
-    "Other labeled"))) %>%
-  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
-  arrange(depth, startime) %>%
-  mutate(x_axis=paste(depth, startime)) %>%
-  mutate(x_axis=fct_inorder(x_axis))
-cpal_breaks <- levels(thaa_plot_data$iso_name)
-cpal_cols <- c("grey80", viridis::viridis(length(cpal_breaks)-2), "grey20")
-arg_glut_thaa_plot <- ggplot(thaa_plot_data) +
-  geom_col(aes(x=x_axis, y=avg_env_conc_nm, fill=iso_name),
-           position="fill",
-           color="black", width=1) +
-  facet_wrap2(compound_name~amendment, ncol=5, strip = strip_nested(bleed = TRUE)) +
-  theme_bw() +
-  scale_x_discrete(expand = expansion()) +
-  scale_y_continuous(labels=scales::label_percent(), breaks=c(0, 0.5, 1),
-                     expand = expansion()) +
-  scale_fill_manual(breaks = cpal_breaks, values = cpal_cols) +
-  labs(x=NULL, y="Fraction labeled", fill=NULL) +
-  theme(
-    axis.text.x=element_text(angle=90, hjust=1, vjust = 0.5)
-  )
-total_thaa_gp <- all_thaa_plot / arg_glut_thaa_plot + plot_layout(heights = c(0.4, 0.6))
-ggsave("total_thaa_gp.png", total_thaa_gp, device = "png", path = "manuscript/figures",
-       width = 6.5, height = 5.5, units = "in", dpi = 300)
-```
-
-![](figures/total_thaa_gp.png)
-  - THAAs were 30-50% of total PC (~2 uM, need actual data) and 70-100% PN (~0.35uM)
-  - Maybe more labeling happening from inorganics in evening?
-  - 50% turnover of arginine in proteins in 3 days per +Arg data, less at surface than deep (??)
-  - 30-50% Glu/Gln turnover in proteins in one day per Amm data
-  - Glutamate, proline fully labeled labeled from arginine
-  - Phenylalanine, tyrosine, histidine(?) labeled from GMP
-  - How is there so much nitrate labeling when I don't see that in the metabolites?
-
-```{r thaa supplement figures}
-thaa_quant %>%
-  left_join(metathaa) %>%
-  filter(samp_type=="Smp") %>%
-  ggplot() +
-  geom_col(aes(x=tripl, y=env_conc_nm, fill=compound_name), 
-           color="black", width=1, position="fill") +
-  facet_nested(depth+startime~amendment+timepoint) +
-  scale_x_discrete(expand = expansion()) +
-  scale_y_continuous(labels=scales::label_percent(), breaks=c(0, 0.5, 1),
-                     expand = expansion()) +
-  labs(x="Triplicate", y="Fraction labeled", fill=NULL) +
-  theme_bw() +
-  theme(
-    axis.text.y=element_markdown(vjust=c(0, 0.5, 1))
-  )
-
-THAA_mol_C_gp <- thaa_quant %>%
-  filter(!str_detect(iso_name, "2H")) %>%
-  left_join(metathaa) %>%
-  filter(samp_type=="Smp") %>%
-  left_join(aa_ncs) %>%
-  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
-  mutate(n_C_lab=as.numeric(str_extract(iso_name, "(?<=13C)\\d+"))/n_C) %>%
-  mutate(n_C_unlab=n_C-n_C_lab) %>%
-  mutate(nM_C_lab=env_conc_nm*n_C_lab, nM_C_unlab=env_conc_nm*n_C_unlab) %>%
-  select(filename, compound_name, iso_name, nM_C_lab, nM_C_unlab) %>%
-  pivot_longer(c(nM_C_lab, nM_C_unlab)) %>%
-  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
-  left_join(metathaa) %>%
-  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
-  ggplot(aes(x=tripl, y=total_uM, fill=name)) +
-  geom_col(color="black", width=1) +
-  facet_nested(depth+startime~amendment+timepoint) +
-  theme_bw() +
-  scale_x_discrete(expand = expansion()) +
-  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
-  scale_fill_manual(breaks = c("nM_C_unlab", "nM_C_lab"), values=c("grey80", "grey20"), 
-                    labels=c("Unlabeled", "Labeled")) +
-  labs(x="Triplicate", y="Total C in THAAs (uM)", fill=NULL) +
-  theme(
-    legend.position = "top"
-  )
-THAA_mol_N_gp <- thaa_quant %>%
-  filter(!str_detect(iso_name, "2H")) %>%
-  left_join(metathaa) %>%
-  filter(samp_type=="Smp") %>%
-  left_join(aa_ncs) %>%
-  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
-  mutate(n_N_lab=as.numeric(str_extract(iso_name, "(?<=15N)\\d+"))) %>%
-  mutate(n_N_unlab=n_N-n_N_lab) %>%
-  mutate(nM_N_lab=env_conc_nm*n_N_lab, nM_N_unlab=env_conc_nm*n_N_unlab) %>%
-  select(filename, compound_name, iso_name, nM_N_lab, nM_N_unlab) %>%
-  pivot_longer(c(nM_N_lab, nM_N_unlab)) %>%
-  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
-  left_join(metathaa) %>%
-  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
-  ggplot(aes(x=tripl, y=total_uM, fill=name)) +
-  geom_col(color="black", width=1) +
-  facet_nested(depth+startime~amendment+timepoint) +
-  theme_bw() +
-  scale_x_discrete(expand = expansion()) +
-  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
-  scale_fill_manual(breaks = c("nM_N_unlab", "nM_N_lab"), values=c("grey80", "grey20"), 
-                    labels=c("Unlabeled", "Labeled")) +
-  labs(x="Triplicate", y="Total N in THAAs (uM)", fill=NULL) +
-  theme(
-    legend.position = "top"
-  )
-all_thaa_NC_gp <- guide_area() + THAA_mol_C_gp + THAA_mol_N_gp + plot_layout(guides = "collect", heights = c(0.1, 0.45, 0.45), design = c("1\n2\n3"))
-ggsave("thaas/all_thaa_NC_gp.png", width=6.5, height=8)
-
-pdf("thaas/all_cmpd_labels.pdf", width=9, height = 5)
-for(cmpd_i in setdiff(unique(thaa_quant$compound_name), c("Citrulline", "Ornithine", "Taurine"))){
-  print(cmpd_i)
-  gp <- thaa_quant %>%
-    filter(compound_name==cmpd_i) %>%
-    left_join(metathaa) %>%
-    drop_na() %>%
-    ggplot() +
-    geom_col(aes(x=tripl, y=env_conc_nm, fill=iso_name), 
-             position="fill",
-             color="black", width=1) +
-    facet_nested(depth+startime~amendment+timepoint) +
-    ggtitle(cmpd_i)
-  print(gp)
-}
-dev.off()
-```
-
-### Universal bioavailability of ammonia at surface
+## Universal bioavailability of ammonia at surface
 
@@
 
-Synthesis rates?
+Synthesis rates, especially for betaines?
 
@@
   - Not very much fully-labeled guanine despite guanosine being easily cleavable, instead it’s mostly made from scratch (CHECK THIS)
-  - Creatine for being mostly 2x15N 1x13C labeled from arginine
+  - Creatine for being mostly 2x15N 1x13C labeled from arginine (CHECK THIS)
 
@@
 
-Untargeted metabolites? Clustering their % labeled patterns with the targeted metabolites to see what groups they might belong to
+## Rapid uptake and use of all N substrates
 
-## Discussion
+Total fraction of PN that was visible in metabolites
 
-## Conclusion
+```{r}
+env_concs <- read_csv("targeted/final_env_concs.csv")
 
-## Methods
+# More compounds, triplicates shown
+env_concs %>%
+  mutate(mean_nm=mean(conc_nm_in_env), .by=compound_name) %>%
+  filter(!compound_name%in%c("Thymine", "Thymidine")) %>%
+  arrange(desc(mean_nm)) %>%
+  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", "Glutamate", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:19], compound_name, "Other")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  summarise(conc_nm=sum(conc_nm_in_env), .by=c(compound_name, filename)) %>%
+  left_join(parametadata) %>%
+  ggplot(aes(x=tripl, y=conc_nm, fill=compound_name)) +
+  geom_col(color="black", width=1) +
+  facet_nested(depth+startime~amendment+timepoint, switch="x") +
+  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
+  scale_fill_manual(values = c(splividis(19), "grey")) +
+  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
+  theme_bw() +
+  theme(
+    panel.spacing.x = unit(0, "lines"),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
 
-## Acknowledgements
+# Summarized version
+env_concs %>%
+  mutate(mean_nm=mean(conc_nm_in_env), .by=compound_name) %>%
+  filter(!compound_name%in%c("Thymine", "Thymidine")) %>%
+  arrange(desc(mean_nm)) %>%
+  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", 
+                              "Glutamate", compound_name)) %>%
+  mutate(compound_name=str_remove_all(compound_name, "^L-|(?<=/)L-")) %>%
+  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:9], 
+                              compound_name, "Other")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  summarise(conc_nm=sum(conc_nm_in_env), .by=c(compound_name, filename)) %>%
+  left_join(parametadata) %>%
+  summarise(conc_nm=mean(conc_nm), .by=c(compound_name, amendment, depth, timepoint)) %>%
+  ggplot(aes(x=timepoint, y=conc_nm, fill=compound_name)) +
+  geom_col(color="black", width=1) +
+  facet_grid(depth~amendment, scales="free") +
+  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
+  scale_fill_manual(values = c(splividis(9), "grey")) +
+  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
+  theme_bw()
 
-## Data availability
+# In N moles
+env_concs %>%
+  left_join(stan_patched %>% distinct(compound_name, n_N)) %>%
+  mutate(nM_N=conc_nm_in_env*n_N) %>%
+  mutate(mean_nm=mean(nM_N), .by=compound_name) %>%
+  filter(!compound_name%in%c("Thymine", "Thymidine")) %>%
+  arrange(desc(mean_nm)) %>%
+  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", 
+                              "Glutamate", compound_name)) %>%
+  mutate(compound_name=str_remove_all(compound_name, "^L-|(?<=/)L-")) %>%
+  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:9], 
+                              compound_name, "Other")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  summarise(conc_nm=sum(nM_N), .by=c(compound_name, filename)) %>%
+  left_join(parametadata) %>%
+  summarise(conc_nm=mean(conc_nm), .by=c(compound_name, amendment, depth, timepoint)) %>%
+  ggplot(aes(x=timepoint, y=conc_nm, fill=compound_name)) +
+  geom_col(color="black", width=1) +
+  facet_grid(depth~amendment, scales="free") +
+  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
+  scale_fill_manual(values = c(splividis(9), "grey")) +
+  labs(x=NULL, y="Nitrogen in unlabeled metabolites (nM)", fill=NULL) +
+  theme_bw()
+```
 
-## References
+```{r}
+peak_areas %>%
+  filter(str_detect(iso_name, "13C0, 15N0")) %>%
+  filter(!compound_name%in%c("Urea", "L-Valine", "Oxalic acid", "Malic acid", 
+                             "Acetylcholine", "Homarine")) %>%
+  left_join(parametadata %>% distinct(filename, samp_type, amendment)) %>%
+  filter(samp_type=="Smp") %>%
+  left_join(read_csv("targeted/rf_df.csv")) %>%
+  mutate(conc_nm=area/rf) %>%
+  filter(!is.na(conc_nm)) %>%
+  mutate(mean_nm=mean(conc_nm), .by=compound_name) %>%
+  arrange(desc(mean_nm)) %>%
+  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", 
+                              "Glutamate", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:19], 
+                              compound_name, "Other")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  summarise(conc_nm=sum(conc_nm), .by=c(compound_name, filename)) %>%
+  left_join(parametadata) %>%
+  ggplot(aes(x=tripl, y=conc_nm, fill=compound_name)) +
+  geom_col(color="black", width=1) +
+  facet_nested(depth+startime~amendment+timepoint, switch="x") +
+  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
+  scale_fill_manual(values = c(splividis(19), "grey")) +
+  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
+  theme_bw() +
+  theme(
+    panel.spacing.x = unit(0, "lines"),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
+```
+
+## Labeling traced into protein via total hydrolyzable amino acids
+
+```{r thaa data}
+thaa_quant <- read_csv("thaas/quant_data.csv")
+metathaa <- read_csv("thaas/metathaa.csv") %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve")))
+
+all_thaa_plot <- thaa_quant %>%
+  filter(!str_detect(iso_name, "2H")) %>%
+  left_join(metathaa) %>%
+  filter(samp_type=="Smp") %>%
+  left_join(aa_ncs) %>%
+  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
+  mutate(n_N_lab=as.numeric(str_extract(iso_name, "(?<=15N)\\d+"))) %>%
+  mutate(n_N_unlab=n_N-n_N_lab) %>%
+  mutate(nM_N_lab=env_conc_nm*n_N_lab, nM_N_unlab=env_conc_nm*n_N_unlab) %>%
+  select(filename, compound_name, iso_name, nM_N_lab, nM_N_unlab) %>%
+  pivot_longer(c(nM_N_lab, nM_N_unlab)) %>%
+  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
+  left_join(metathaa) %>%
+  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
+  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
+  summarise(total_uM=mean(total_uM), .by=c(amendment, depth, startime, timepoint, name)) %>%
+  arrange(depth, startime) %>%
+  mutate(x_axis=paste(depth, startime)) %>%
+  mutate(x_axis=fct_inorder(x_axis)) %>%
+  ggplot(aes(x=x_axis, y=total_uM, fill=name)) +
+  geom_col(color="black", width=1) +
+  facet_nested(.~amendment, scales="free_y") +
+  theme_bw() +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
+  scale_fill_manual(breaks = c("nM_N_unlab", "nM_N_lab"), values=c("grey80", "grey20"), 
+                    labels=c("Unlabeled", "Labeled")) +
+  labs(x=NULL, y="Total uM N in THAAs", fill=NULL) +
+  theme(
+    axis.text.x=element_blank()
+  )
+
+thaa_plot_data <- thaa_quant %>%
+  left_join(metathaa) %>%
+  filter(samp_type=="Smp") %>%
+  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
+  summarise(avg_env_conc_nm=mean(env_conc_nm, na.rm=TRUE), .by=c(compound_name, iso_name, amendment, timepoint, startime, depth)) %>%
+  filter(compound_name%in%c("Glutamic acid", "Arginine")) %>%
+  mutate(iso_name=case_when(
+    str_detect(iso_name, "13C0, 15N0")~"Unlabeled",
+    str_detect(iso_name, "13C0, 15N1")~"Singly 15N labeled",
+    str_detect(iso_name, "13C1, 15N0")~"Singly 13C labeled",
+    str_detect(iso_name, "Glutamic acid, 13C5, 15N1")~"Fully C and N\nlabeled glutamate",
+    str_detect(iso_name, "Arginine, 13C0, 15N3")~"3x 15N\nlabeled arginine",
+    str_detect(iso_name, "Arginine, 13C0, 15N4")~"Fully N\nlabeled arginine",
+    str_detect(iso_name, "Arginine, 13C6, 15N4")~"Fully C and N\nlabeled arginine",
+    TRUE~"Other labeled"
+  )) %>%
+  mutate(iso_name=factor(iso_name, levels=c(
+    "Unlabeled", "Singly 15N labeled", "Singly 13C labeled", 
+    "3x 15N\nlabeled arginine", "Fully N\nlabeled arginine", 
+    "Fully C and N\nlabeled arginine", "Fully C and N\nlabeled glutamate", 
+    "Other labeled"))) %>%
+  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
+  arrange(depth, startime) %>%
+  mutate(x_axis=paste(depth, startime)) %>%
+  mutate(x_axis=fct_inorder(x_axis))
+cpal_breaks <- levels(thaa_plot_data$iso_name)
+cpal_cols <- c("grey80", viridis::viridis(length(cpal_breaks)-2), "grey20")
+arg_glut_thaa_plot <- ggplot(thaa_plot_data) +
+  geom_col(aes(x=x_axis, y=avg_env_conc_nm, fill=iso_name),
+           position="fill",
+           color="black", width=1) +
+  facet_wrap2(compound_name~amendment, ncol=5, strip = strip_nested(bleed = TRUE)) +
+  theme_bw() +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_continuous(labels=scales::label_percent(), breaks=c(0, 0.5, 1),
+                     expand = expansion()) +
+  scale_fill_manual(breaks = cpal_breaks, values = cpal_cols) +
+  labs(x=NULL, y="Fraction labeled", fill=NULL) +
+  theme(
+    axis.text.x=element_text(angle=90, hjust=1, vjust = 0.5)
+  )
+total_thaa_gp <- all_thaa_plot / arg_glut_thaa_plot + plot_layout(heights = c(0.4, 0.6))
+ggsave("total_thaa_gp.png", total_thaa_gp, device = "png", path = "manuscript/figures",
+       width = 6.5, height = 5.5, units = "in", dpi = 300)
+```
+
+![](figures/total_thaa_gp.png)
+
+  - THAAs were 30-50% of total PC (~2 uM, need actual data) and 70-100% PN (~0.35uM)
+  - Maybe more labeling happening from inorganics in evening?
+  - 50% turnover of arginine in proteins in 3 days per +Arg data, less at surface than deep (??)
+  - 30-50% Glu/Gln turnover in proteins in one day per Amm data
+  - Glutamate, proline fully labeled labeled from arginine
+  - Phenylalanine, tyrosine, histidine(?) labeled from GMP
+  - How is there so much nitrate labeling when I don't see that in the metabolites?
+
+# Discussion
+
+# Conclusion
+
+# Methods
+
+# Acknowledgements
+
+# Data availability
+
+# References
 
 ```{r packages, eval=FALSE}
-packages_used <- list.files("..", pattern = "*\\.R$|*\\.Rmd$", recursive = TRUE) %>%
+packages_used <- list.files(pattern = "*\\.R$|*\\.Rmd$", recursive = TRUE, full.names = TRUE) %>%
   str_subset("old|scripts", negate=TRUE) %>%
   map(function(file_name){
-    v <- readLines(paste0("../", file_name))
+    v <- readLines(file_name)
     libs <- unique(str_extract(v, "(?<=library\\().*?(?=\\)|,)"))
@@
     pkgs <- unique(c(na.omit(libs), na.omit(refs)))
-  }) %>% unlist() %>% unique()
-knitr::write_bib(packages_used[nchar(packages_used)>0], file = "r-packages.bib")
+  }, .progress = TRUE) %>% unlist() %>% unique()
+knitr::write_bib(packages_used[nchar(packages_used)>0], file = "manuscript/r-packages.bib")
 ```
@@
 
-## Supplement
+# Supplement
 
-### Supp fig 1: full heatmap of all compounds and timepoints
+## Supp fig: full heatmap of all compounds and timepoints
 
@@
 ```
+
+## Supp fig: THAA labeling broken down by triplicate
+
+```{r thaa supplement figures}
+thaa_quant %>%
+  left_join(metathaa) %>%
+  filter(samp_type=="Smp") %>%
+  ggplot() +
+  geom_col(aes(x=tripl, y=env_conc_nm, fill=compound_name), 
+           color="black", width=1, position="fill") +
+  facet_nested(depth+startime~amendment+timepoint) +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_continuous(labels=scales::label_percent(), breaks=c(0, 0.5, 1),
+                     expand = expansion()) +
+  labs(x="Triplicate", y="Fraction labeled", fill=NULL) +
+  theme_bw() +
+  theme(
+    axis.text.y=element_markdown(vjust=c(0, 0.5, 1))
+  )
+
+THAA_mol_C_gp <- thaa_quant %>%
+  filter(!str_detect(iso_name, "2H")) %>%
+  left_join(metathaa) %>%
+  filter(samp_type=="Smp") %>%
+  left_join(aa_ncs) %>%
+  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
+  mutate(n_C_lab=as.numeric(str_extract(iso_name, "(?<=13C)\\d+"))/n_C) %>%
+  mutate(n_C_unlab=n_C-n_C_lab) %>%
+  mutate(nM_C_lab=env_conc_nm*n_C_lab, nM_C_unlab=env_conc_nm*n_C_unlab) %>%
+  select(filename, compound_name, iso_name, nM_C_lab, nM_C_unlab) %>%
+  pivot_longer(c(nM_C_lab, nM_C_unlab)) %>%
+  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
+  left_join(metathaa) %>%
+  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
+  ggplot(aes(x=tripl, y=total_uM, fill=name)) +
+  geom_col(color="black", width=1) +
+  facet_nested(depth+startime~amendment+timepoint) +
+  theme_bw() +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
+  scale_fill_manual(breaks = c("nM_C_unlab", "nM_C_lab"), values=c("grey80", "grey20"), 
+                    labels=c("Unlabeled", "Labeled")) +
+  labs(x="Triplicate", y="Total C in THAAs (uM)", fill=NULL) +
+  theme(
+    legend.position = "top"
+  )
+THAA_mol_N_gp <- thaa_quant %>%
+  filter(!str_detect(iso_name, "2H")) %>%
+  left_join(metathaa) %>%
+  filter(samp_type=="Smp") %>%
+  left_join(aa_ncs) %>%
+  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
+  mutate(n_N_lab=as.numeric(str_extract(iso_name, "(?<=15N)\\d+"))) %>%
+  mutate(n_N_unlab=n_N-n_N_lab) %>%
+  mutate(nM_N_lab=env_conc_nm*n_N_lab, nM_N_unlab=env_conc_nm*n_N_unlab) %>%
+  select(filename, compound_name, iso_name, nM_N_lab, nM_N_unlab) %>%
+  pivot_longer(c(nM_N_lab, nM_N_unlab)) %>%
+  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
+  left_join(metathaa) %>%
+  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
+  ggplot(aes(x=tripl, y=total_uM, fill=name)) +
+  geom_col(color="black", width=1) +
+  facet_nested(depth+startime~amendment+timepoint) +
+  theme_bw() +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
+  scale_fill_manual(breaks = c("nM_N_unlab", "nM_N_lab"), values=c("grey80", "grey20"), 
+                    labels=c("Unlabeled", "Labeled")) +
+  labs(x="Triplicate", y="Total N in THAAs (uM)", fill=NULL) +
+  theme(
+    legend.position = "top"
+  )
+all_thaa_NC_gp <- guide_area() + THAA_mol_C_gp + THAA_mol_N_gp + plot_layout(guides = "collect", heights = c(0.1, 0.45, 0.45), design = c("1\n2\n3"))
+ggsave("thaas/all_thaa_NC_gp.png", width=6.5, height=8)
+
+pdf("thaas/all_cmpd_labels.pdf", width=9, height = 5)
+for(cmpd_i in setdiff(unique(thaa_quant$compound_name), c("Citrulline", "Ornithine", "Taurine"))){
+  print(cmpd_i)
+  gp <- thaa_quant %>%
+    filter(compound_name==cmpd_i) %>%
+    left_join(metathaa) %>%
+    drop_na() %>%
+    ggplot() +
+    geom_col(aes(x=tripl, y=env_conc_nm, fill=iso_name), 
+             position="fill",
+             color="black", width=1) +
+    facet_nested(depth+startime~amendment+timepoint) +
+    ggtitle(cmpd_i)
+  print(gp)
+}
+dev.off()
+```

### 2025-05-12 manuscript/manuscript.Rmd
Massively cleaning up manuscript directory

@@
 ---
-title: ""
+title: "Nitrogen's form determines its fate in the North Pacific Subtropical Gyre during PARAGON I"
 author:
@@
   - Anitra E. Ingalls
-output: 
-  word_document:
-    reference_docx: Frontiers_Template.docx
-csl: frontiers-in-marine-science.csl
-bibliography:
-  - r-packages.bib
+  - Other SCOPE collaborators
+output: word_document
 ---
@@
 
-pdf("thaas/all_cmpd_labels.pdf", width=9, height = 5)
+pdf("figures/all_cmpd_labels.pdf", width=9, height = 5)
 for(cmpd_i in setdiff(unique(thaa_quant$compound_name), c("Citrulline", "Ornithine", "Taurine"))){
@@
 ```
+
+## Supp fig: Fraction of AA as DFAA vs THAA
+
+```{r fraction of aa pool in df vs th}
+thaa_pool_sizes <- thaa_quant %>%
+  filter(!compound_name%in%c("Citrulline", "Ornithine", "Taurine")) %>%
+  filter(str_detect(iso_name, "13C0, 15N0")) %>%
+  filter(str_detect(filename, "Smp.*T0")) %>%
+  select(filename, compound_name, env_conc_nm) %>%
+  mutate(compound_name=ifelse(compound_name=="Isoleucine", "Leucine", compound_name)) %>%
+  summarise(env_conc_nm=sum(env_conc_nm), .by=c(filename, compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Leucine", "Leu/Ile", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Glutamic acid", "Glu/Gln", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Aspartic acid", "Asp/Asn", compound_name)) %>%
+  left_join(metathaa) %>%
+  summarise(mean_thaa_conc=mean(env_conc_nm), .by=c(amendment, depth, startime, compound_name))
+metab_pool_sizes <- env_concs %>%
+  filter(str_detect(filename, "Smp.*T0")) %>%
+  mutate(compound_name=str_remove_all(compound_name, "L-")) %>%
+  mutate(compound_name=str_remove(compound_name, "/Homoserine")) %>%
+  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
+  mutate(compound_name=ifelse(compound_name=="Glutamine", "Glutamic acid", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Asparagine", "Aspartic acid", compound_name)) %>%
+  summarise(conc_nm_in_env=sum(conc_nm_in_env), .by=c(filename, compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Glutamic acid", "Glu/Gln", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Aspartic acid", "Asp/Asn", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Leucine/Isoleucine", "Leu/Ile", compound_name)) %>%
+  filter(compound_name%in%thaa_pool_sizes$compound_name) %>%
+  left_join(parametadata) %>%
+  summarise(mean_metab_conc=mean(conc_nm_in_env), .by=c(amendment, depth, startime, compound_name))
+# Valine and phenylalanine failed QC
+thaa_pool_sizes %>%
+  left_join(metab_pool_sizes) %>%
+  ggplot() +
+  geom_point(aes(x=mean_thaa_conc, y=mean_metab_conc, color=compound_name)) +
+  facet_nested(depth+startime~amendment)
+thaa_pool_sizes %>%
+  left_join(metab_pool_sizes) %>%
+  ggplot() +
+  geom_point(aes(x=mean_thaa_conc, y=mean_metab_conc)) +
+  facet_wrap(~compound_name, scales="free")
+
+thaa_pool_sizes %>%
+  left_join(metab_pool_sizes) %>%
+  mutate(metab_frac=mean_metab_conc/mean_thaa_conc) %>%
+  ggplot() +
+  geom_density(aes(x=metab_frac, color=compound_name))
+thaa_pool_sizes %>%
+  left_join(metab_pool_sizes) %>%
+  mutate(metab_frac=mean_metab_conc/mean_thaa_conc) %>%
+  ggplot() +
+  geom_boxplot(aes(x=compound_name, y=metab_frac)) +
+  scale_y_continuous(labels = scales::label_percent()) +
+  labs(x=NULL, y='Fraction of AA in "free" pool: [metab] / [THAA]') +
+  coord_flip()
+```

### 2025-05-12 manuscript/figscript.R

@@ -1,64 +1,0 @@
-
-# Setup ----
-
-library(tidyverse)
-library(ggh4x)
-library(patchwork)
-
-parametadata <- read_csv("metadata/parametadata.csv") %>%
-  # filter(IS_type=="noIS") %>%
-  filter(polarity=="pos") %>%
-  filter(samp_type=="Smp") %>%
-  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
-  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
-  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"))) %>%
-  arrange(amendment, depth, timepoint_num, startime) %>%
-  mutate(timepoint=fct_inorder(timepoint))
-peak_areas <- read_csv("targeted/all_peak_areas_w_isos_incl_13C_for_inorg.csv") %>%
-  left_join(parametadata %>% distinct(filename, shortname))
-all_stans <- read_csv("metadata/all_stans.csv")
-samp_areas <- parametadata %>%
-  filter(samp_type=="Smp") %>%
-  distinct(shortname, depth) %>%
-  left_join(peak_areas) %>%
-  mutate(shortname=fct_inorder(shortname)) %>%
-  filter(!compound_name%in%c("Melamine", "Uridine", "Asterina-330?", "Palythene/Usujirene?",
-                             "Mycosporine-2-glycine?", "Porphyra-334?", "Itaconic acid?",
-                             "Inosine", "Dexpanthenol", "Choline sulfate", "5-Hydroxyectoine")) %>%
-  filter(!(depth=="Deep" & compound_name%in%c("Also Palythene/Usujirene?", "Shinorine?",
-                                              "Mycosporine-glycine?", "Palythine?"))) %>%
-  # filter(!str_detect(shortname, "Nit-Deep-Eve-T10")) %>%
-  select(-depth)
-labeled_fracs <- samp_areas %>%
-  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
-  select(compound_name, iso_name, shortname, labeled_frac) %>%
-  filter(str_detect(iso_name, "13C0, 15N0")) %>%
-  select(-iso_name)
-n_labeled_fracs <- samp_areas %>%
-  filter(str_detect(iso_name, "13C0")) %>%
[…]

### 2025-05-12 manuscript/manuscript.Rmd
Pivoting results section to be xamend first

@@
 
-Nitrogen availability controls productivity in much of the surface ocean.
-The form of this nitrogen controls who it is available to and how it is used.
+Nitrogen availability controls productivity in much of the surface ocean and the form of this nitrogen controls who it is available to and how it is used.
 Here, we explored how four different forms of nitrogen (ammonia, nitrate, arginine, and guanosine monophosphate) were taken up and transformed in the North Pacific Subtropical Gyre using stable isotope labeling metabolomics.
@@
 We found that ammonia is converted into glutamate in less than 20 minutes and results in labeling throughout the metabolome, though osmolytes like betaines had slow turnover rates.
-This process was slower at depth but independent of time of day.
+This process was slower at depth but largely independent of time of day.
 Nitrate amendments resulted in the labeling of only a small subset of molecules such as mycosporine-like amino acids, whose incorporation patterns suggest that this form of nitrogen was available to a subset of the community, likely eukaryotes.
-The two organic nitrogen amendments labeled the metabolome more slowly than the inorganic forms of nitrogen and were primarily converted into glutamate rather than used as intact substrates.
-They also showed the same usage at depth as they did at the surface, hinting at the major role that heterotrophic bacteria likely play in their uptake and use.
+The two organic nitrogen amendments labeled the metabolome more slowly than the inorganic forms of nitrogen and showed similar usage at depth and at the surface, hinting at the major role that heterotrophic bacteria likely play in their uptake and use.
+A large fraction of the organic nitrogen added was remineralized into ammonia and used as glutamate rather than as intact substrates, though labeled arginine was incorporated into proteins directly.
 These measurements reveal the forms and fluxes of organic nitrogen in the natural environment to provide constraints on biogeochemical models and better characterize the transformations between organic and inorganic nitrogen.

### 2025-05-12 manuscript/manuscript.Rmd
Pivoting results section to be xamend first

@@
                "Choline sulfate", "5-Hydroxyectoine", "L-Tyrosine")
+class_levels <- c("Glutamine", "Glutamic acid", "Aspartic acid", "Asparagine", "Arginine", 
+                  "Other amino acids", "Theanine?", "Citrulline", "Ectoine", "Guanine", 
+                  "Guanosine", "Other nucleic acids", "MAAs", "Betaines", "All others")
 aa_ncs <- all_stans %>%
@@
   select(-iso_name)
-log_coef_df <- labeled_fracs %>%
-  filter(!compound_name%in%maa_names) %>%
-  left_join(parametadata %>% distinct(shortname, amendment, depth, timepoint_num)) %>%
-  nest(data=c(shortname, timepoint_num, labeled_frac)) %>%
-  mutate(log_fit=map(data, ~glm(.x$labeled_frac~.x$timepoint_num, family="quasibinomial")),
-         coefs=map(log_fit, function(x)broom::tidy(x))) %>%
-  unnest(coefs) %>%
-  select(compound_name, amendment, depth, term, estimate) %>%
-  pivot_wider(names_from = term, values_from = estimate) %>%
-  rename(intercept=`(Intercept)`, slope=`.x$timepoint_num`) %>%
-  mutate(half_max=-intercept/slope, scale=1/slope) %>%
-  mutate(frac_at_T24 = 1/(1+exp(-(intercept+slope*24))))
-
-splividis <- function(n, ...){
-  vir <- viridis::viridis(n, ...)
-  if(n%%2==0){
-    vir[as.numeric(matrix(1:n, ncol=2, byrow = TRUE))]
-  } else {
-    index_vec <- suppressWarnings(as.numeric(matrix(1:n, ncol=2, byrow = TRUE)))
-    vir[head(index_vec, -1)]
-  }
-}
 ```
@@
 
-## Environmental context of PARAGON cruise
+## Differences in nitrogen labeling at the surface between amendments
 
-PARAGON synthesis paper isn't out yet and I'm not sure it will be anytime soon
-
-Particulate carbon and nitrogen numbers for PARAGON I (from Matt(?))
-
-![](figures/borrowed/pcpn_PIandPII.png)
-
-Data from Rhea showing ammonia/nitrite concentrations - I sampled at 25m and 175m (nitrite max, below ammonia max). Does this mean I was getting a lot of AOA users?
-
-![](figures/borrowed/rhea_ammonia_concs.png)
-
-Angel's IFCB data showing Hemiaulus disappearing over the course of the cruise (most of the Bacillariophyceae are Hemiaulus and she shows that in another plot)
-
-![](figures/borrowed/angel_hemiaulus.png)
-
-Fuyan reports the abundances of various bacterial groups (with the caveat that the SASVs have to have been found in the 4000m depth traps?)
-
-![](figures/borrowed/fuyan_composition_bact.png)
-
-## Universal bioavailability of ammonia at surface
-
-The most rapid and extensive labeling was obtained during the ammonia-amended incubations of surface seawater.
-We observed significant labeling of glutamine, [and others] even at the T0 timepoint, indicating turnover times shorter than the time required to filter and flash-freeze the samples in liquid nitrogen (~20 minutes).
-
-DISCUSSION: the speed at which ammonia labeling happens hints at a high flux, low concentration ammonia cycle in the surface ocean since uptake was too fast and too widespread to be a transcriptional response, indicating that they have ammonia transporters plus GS/GOGAT/GDH always active.
-
-Synthesis rates, especially for betaines?
-
-Aspartate labeled faster than glutamate - discuss pool sizes
-
-Slow labeling of osmolytes - discuss pool sizes
-
-```{r fig small heatmap compound groups}
-class_levels <- c("Glutamine", "Glutamic acid", "Aspartic acid", "Asparagine", "Other amino acids", "Theanine?", "Citrulline", "Ectoine", "Guanine", "Other nucleic acids", "MAAs", "Betaines", "All others")
-
-labeled_fracs %>%
-  left_join(parametadata %>% distinct(shortname, timepoint, timepoint_num, depth, startime, amendment)) %>%
-  filter(amendment=="Amm") %>%
+```{r}
+xamend_metab_gp <- labeled_fracs %>%
+  left_join(parametadata %>% distinct(shortname, timepoint, timepoint_num, depth, startime, amendment, tripl)) %>%
+  filter(depth=="Surf") %>%
   mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
@@
   mutate(compound_class=case_when(
-    compound_name%in%class_levels ~ compound_name,
+    compound_name%in%c("Glutamine", "Aspartic acid", "Asparagine", "Glutamic acid", "Ectoine", "Theanine?", "Citrulline", "Guanine", "Guanosine", "Arginine") ~ compound_name,
     compound_name%in%maa_names~"MAAs",
@@
   mutate(compound_class=factor(compound_class, levels=class_levels)) %>%
-  summarise(avg_frac=mean(labeled_frac), .by = c(compound_class, timepoint_num, startime, depth)) %>%
-  mutate(color_group=paste(depth, startime)) %>%
-  ggplot(aes(x=timepoint_num, y=avg_frac, color=color_group, fill=color_group)) +
-  geom_smooth(method = "glm", formula = y~x, method.args = list(family = "quasibinomial"),
-              fill="grey80") +
-  geom_point(pch=21, color="black") +
-  facet_wrap(~compound_class, ncol=5) +
-  scale_y_continuous(labels = scales::label_percent(), breaks = c(0, 0.5, 1)) +
-  scale_x_continuous(breaks = c(0.3, 1, 3, 10, 26), labels = c(0.3, 1, 3, 10, 26)) +
-  scale_color_manual(breaks = c("Surf Morn", "Surf Eve", "Deep Morn", "Deep Eve"),
-                     values = c("#FFD53D", "#DBA20A", "#256A76", "#003743")) +
-  scale_fill_manual(breaks = c("Surf Morn", "Surf Eve", "Deep Morn", "Deep Eve"),
-                     values = c("#FFD53D", "#DBA20A", "#256A76", "#003743")) +
-  labs(x="Hours after 15N added", y="Fraction of metabolite labeled", color=NULL) +
-  theme_bw() +
-  guides(color = guide_legend(nrow = 2, override.aes = list(linewidth=2)),
-         fill = "none") +
-  theme(legend.position = "inside",
-        legend.position.inside = c(0.8, 1/6),
-        legend.location = "plot")
-
-labeled_fracs %>%
-  left_join(parametadata %>% distinct(shortname, timepoint, timepoint_num, depth, startime, amendment, tripl)) %>%
-  filter(amendment=="Amm") %>%
-  mutate(depth=factor(depth, levels=c("Surf", "Deep"), labels=c("Surface seawater (25 meters)", "Seawater from below DCM (175m)"))) %>%
-  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start"))) %>%
-  mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
-  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
-  mutate(compound_class=case_when(
-    compound_name%in%c("Glutamine", "Aspartic acid", "Asparagine", "Glutamic acid", "Ectoine", "Theanine?", "Citrulline", "Guanine") ~ compound_name,
-    compound_name%in%maa_names~"MAAs",
-    compound_name%in%aa_names~"Other amino acids",
-    compound_name%in%nucleo_names~"Other nucleic acids",
-    compound_name%in%betaine_names~"Betaines",
-    TRUE~"All others"
-  )) %>%
-  mutate(compound_class=factor(compound_class, levels=class_levels)) %>%
-  summarise(avg_frac=mean(labeled_frac), .by = c(compound_class, timepoint, startime, depth, tripl)) %>%
+  summarise(avg_frac=mean(labeled_frac), .by = c(compound_class, amendment, timepoint, tripl)) %>%
+  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"),
+                          labels=c("+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
   ggplot() +
@@
   geom_vline(xintercept = c(1.5:4.5), color="grey80") +
-  facet_nested(compound_class~depth+startime, scales = "free_y", switch="y") +
+  facet_nested(compound_class~amendment, scales = "free_x", switch="y") +
   scale_fill_gradientn(colours = viridis::viridis(10, alpha = seq(0, 1, length.out=10), direction = -1),
@@
         panel.spacing.y = unit(0, "lines"),
+        axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
         axis.text.y = element_blank(),
@@
   )
-
-ggsave(filename = "grouped_heatmap.png", device = "png", path = "manuscript/figures",
-       width = 6.5, height = 4, units = "in", dpi = 300)
+ggsave("xamend_thaa_gp.png",plot = xamend_metab_gp, device = "png", path = "figures",
+       width = 6.5, height = 5.5, units = "in", dpi = 300)
 ```
 
-![](figures/grouped_heatmap.png)
+![](figures/xamend_thaa_gp.png)
 
-*Figure XX: Fraction of each metabolite or metabolite group containing at least one 15N label during the treatments amended with 15N ammonia after the number of hours shown on the x-axis.
+*Figure 1: Fraction of each metabolite or metabolite group containing at least one 15N label during the surface seawater amended various substrates after the number of hours shown on the x-axis.
 Darker colors indicate a larger percentage of the compound signal was isotopically labeled.
-Separate incubations were started at 6AM and 6PM with water from both 25 meters and 175 meters near Station ALOHA.
-Three biological replicates are stacked vertically at each timepoint.*
+Three biological replicates are stacked vertically at each timepoint and the categories with multiple entries have been averaged.
+MAA = mycosporine-like amino acids, GMP = guanosine monophosphate.*
 
-Same basic pattern in the evening except MAAs, guanosine, proline?
+  - Ammonia
+    - Really rapid labeling of glutamine/glutamate that aligns nicely with prior work
+      - Aspartate labeled faster than glutamate likely due to small aspartate pool size
+    - Surprisingly rapid labeling of citrulline that isn't reflected in arginine
+      - Pool size problem (arginine >> citrulline)?
+      - Indicates that urea cycle is only half-active?
+    - Putative annotation of theanine based on labeling patterns
+    - Rapid and heavy guanine labeling
+      - Potentially being used as a storage compound
+    - 10-40% of the N in the THAA (protein) pool ends up labeled by T26 (below)
+      - THAAs were 30-50% of total PC (~2 uM, need actual data) and 70-100% PN (~0.35uM)
+  - Nitrate
+    - Very little labeling, even in glutamate/glutamine
+      - Surface community not "expecting" nitrate?
+      - Redox problems with surface energy?
+    - Guanine and MAA labeling exceeds glutamate labeling
+      - Clear sign of a community *subset* using nitrate
+        - Specifically the MAA synthesizers (euks) and those storing N in guanine (diatoms/haptos/dinos)
+    - More extensive labeling (10-50% of the N) in the THAAs
+      - Must be a really high flux through a small AA pool subset, e.g. euks specifically
+  - Arginine
+    - Clear uptake at T0 (20 minutes), 80-90% of the pool taken up at T10
+      - Labeling signal decreases after 24 hours
+    - Label obviously making it into citrulline so urea cycle is active
+      - Don't really have good enough urea data to see labeling change there?
+    - Big jump in labeling of other metabolites at T26 not seen at T10
+      - No hypotheses why this is happening but it occurs in both morn/eve so not light signal
+    - Maybe less storage in guanine specifically, labeling is on par with other NAs unlike ammonia
+    - THAAs show clear usage of arginine in proteins
+      - 30-50% arg in THAAs labeled @73hrs
+      - Clear reuse of arginine backbone as glutamate (fully labeled glutamate) and maybe proline?
+  - Guanosine monophosphate
+    - Unable to measure GMP well enough directly?
+    - Guanosine labeling visible at T3, definitely T10, maybe some decrease at T73 like Arg?
+      - Guanosine pool never gets super heavily labeled
+    - Has a VERY similar pattern to arginine
+      - Also a big jump at T26 with basically no labeling at T10
+      - Implies mostly converted into ammonia as well and used that way
+    - Reuse of carbon backbone for THAAs though
+      - Histidine 13C5, 15N3
+      - Tyrosine 13C4, 15N1
+      - Phenylalanine 13C4, 15N1
 
-```{r fig MAA labeling patterns}
+```{r thaa data}
+thaa_quant <- read_csv("quant_data.csv")
+metathaa <- read_csv("metathaa.csv") %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve")))
+
+all_thaa_plot <- thaa_quant %>%
+  filter(!str_detect(iso_name, "2H")) %>%
+  left_join(metathaa) %>%
+  filter(samp_type=="Smp") %>%
+  left_join(aa_ncs) %>%
+  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
+  mutate(n_N_lab=as.numeric(str_extract(iso_name, "(?<=15N)\\d+"))) %>%
+  mutate(n_N_unlab=n_N-n_N_lab) %>%
+  mutate(nM_N_lab=env_conc_nm*n_N_lab, nM_N_unlab=env_conc_nm*n_N_unlab) %>%
+  select(filename, compound_name, iso_name, nM_N_lab, nM_N_unlab) %>%
+  pivot_longer(c(nM_N_lab, nM_N_unlab)) %>%
+  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
+  left_join(metathaa) %>%
+  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
+  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"),
+                          labels=c("Initial", "+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
+  summarise(total_uM=mean(total_uM), .by=c(amendment, timepoint, name)) %>%
+  mutate(name=factor(name, levels=c("nM_N_unlab", "nM_N_lab"), labels = c("Unlabeled", "Labeled"))) %>%
+  ggplot(aes(x=amendment, y=total_uM, fill=name)) +
+  geom_col(color="black", width=1) +
+  theme_bw() +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
+  scale_fill_manual(breaks = c("Unlabeled", "Labeled"), values=c("grey80", "grey20")) +
+  labs(x=NULL, y="Total uM N in THAAs", fill=NULL)
+
+plot_iso_levels <- c(
+  "Unlabeled", 
+  "<sup>15</sup>N<sub>1</sub>", 
+  "<sup>15</sup>N<sub>2</sub>", 
+  "<sup>15</sup>N<sub>3</sub>", 
+  "Entirely <sup>15</sup>N",
+  "<sup>13</sup>C<sub>1</sub>", 
+  "<sup>13</sup>C<sub>1</sub><sup>15</sup>N<sub>1</sub>",
+  "<sup>13</sup>C<sub>4</sub><sup>15</sup>N<sub>1</sub>", 
+  "<sup>13</sup>C<sub>5</sub><sup>15</sup>N<sub>3</sub>",
+  "Entirely <sup>13</sup>C<br>and <sup>15</sup>N", 
+  "Other labels")
+plot_iso_colors <- c(
+  "grey80",
+  "#e0ecf6",
+  "#97bfe0",
+  "#377EB8",
+  "#285b84",
+  "#ffb4b5",
+  "#d7b6dc",
+  "#FFFF33",
+  "#4DAF4A",
+  "#984EA3",
+  "grey20"
+)
+# "#E41A1C" "#377EB8" "#4DAF4A" "#984EA3" "#FF7F00" "#FFFF33" "#A65628" "#F781BF" "#999999"
+thaa_plot_data <- thaa_quant %>%
+  left_join(metathaa) %>%
+  filter(samp_type=="Smp") %>%
+  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
+  summarise(avg_env_conc_nm=mean(env_conc_nm, na.rm=TRUE), .by=c(compound_name, iso_name, amendment, timepoint)) %>%
+  filter(compound_name%in%c("Glutamic acid", "Arginine", "Histidine", "Phenylalanine")) %>%
+  mutate(compound_name=ifelse(compound_name=="Glutamic acid", "Glutamine + Glutamate", compound_name)) %>%
+  mutate(plot_iso_name=case_when(
+    str_detect(iso_name, "13C0, 15N0")~"Unlabeled",
+    str_detect(iso_name, "13C0, 15N1")~"<sup>15</sup>N<sub>1</sub>",
+    str_detect(iso_name, "13C0, 15N2")~"<sup>15</sup>N<sub>2</sub>",
+    str_detect(iso_name, "Histidine, 13C0, 15N3")~"Entirely <sup>15</sup>N",
+    str_detect(iso_name, "13C0, 15N3")~"<sup>15</sup>N<sub>3</sub>",
+    str_detect(iso_name, "Arginine, 13C0, 15N4")~"Entirely <sup>15</sup>N",
+    str_detect(iso_name, "13C1, 15N0")~"<sup>13</sup>C<sub>1</sub>",
+    str_detect(iso_name, "13C1, 15N1")~"<sup>13</sup>C<sub>1</sub><sup>15</sup>N<sub>1</sub>",
+    str_detect(iso_name, "Glutamic acid, 13C5, 15N1")~"Entirely <sup>13</sup>C and <sup>15</sup>N",
+    str_detect(iso_name, "Arginine, 13C6, 15N4")~"Entirely <sup>13</sup>C<br>and <sup>15</sup>N",
+    iso_name=="Phenylalanine, 13C4, 15N1"~"<sup>13</sup>C<sub>4</sub><sup>15</sup>N<sub>1</sub>",
+    iso_name=="Histidine, 13C5, 15N3"~"<sup>13</sup>C<sub>5</sub><sup>15</sup>N<sub>3</sub>",
+    TRUE~"Other labels"
+  )) %>%
+  mutate(plot_iso_name=factor(plot_iso_name, levels=plot_iso_levels)) %>%
+  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP")))
+arg_glut_thaa_plot <- ggplot(thaa_plot_data) +
+  geom_col(aes(x=amendment, y=avg_env_conc_nm, fill=plot_iso_name, label=iso_name),
+           position="fill",
+           color="black", width=1) +
+  facet_wrap(~compound_name) +
+  theme_bw() +
+  scale_x_discrete(expand = expansion()) +
+  scale_y_continuous(labels=scales::label_percent(), breaks=c(0, 0.5, 1),
+                     expand = expansion()) +
+  scale_fill_manual(breaks = plot_iso_levels, values = plot_iso_colors) +
+  labs(x=NULL, y="Fraction of THAA labeled", fill=NULL) +
+  theme(
+    # axis.text.x=element_text(angle=90, hjust=1, vjust = 0.5),
+    legend.text = element_markdown()
+  )
+total_thaa_gp <- all_thaa_plot / arg_glut_thaa_plot + plot_layout(heights = c(0.3, 0.7))
+total_thaa_gp
+ggsave("total_thaa_gp.png", total_thaa_gp, device = "png", path = "figures",
+       width = 6.5, height = 5, units = "in", dpi = 300)
+```
+
+![](figures/total_thaa_gp.png)
+
+*Figure 2: Stacked bar plots showing the fraction of various total hydrolyzable amino acids (THAAs) labeled at the end of the incubations across the four amendments.
+Total uM nitrogen (N) in the protein pool is shown in the top plot while the fraction of individual amino acids labeled is shown in the rows below.
+Colors denote specific labeling patterns detected in the THAAs.*
+
+## Minimal diel signal observed except in the UV-reactive mycosporine-like amino acids
+
+```{r fig MAA labeling new}
 chosen_maas <- c("Mycosporine-glycine?", "Shinorine?", "Palythine?",
@@
 
-ggsave(filename = "MAA_curves.png", device = "png", path = "manuscript/figures",
+ggsave(filename = "MAA_curves.png", device = "png", path = "figures",
        width = 3, height = 6, units = "in", dpi = 300)
@@
 
-*Figure XX: Fraction of UV-reactive mycosporine-like amino acids (MAAs) containing at least one 15N label over one day during the treatments amended with 15N labeled ammonia or nitrate.
+*Figure 3: Fraction of UV-reactive mycosporine-like amino acids (MAAs) containing at least one 15N label over one day during the treatments amended with 15N labeled ammonia or nitrate.
 Incubations started at 6AM are shown in yellow while incubations started at 6PM are shown in blue, with grey background regions highlighting nighttime.
@@
 
-Deep labeling is slower and much more limited for ammonia
+  - MAAs clearly show daytime synthesis during the day and not at night
+    - Very clear in +Ammonia, less so in +Nitrate
+    - Need to show this statistically somehow?
+  - High levels of labeling despite minimal glycine labeling and synthesis being 4DG+Gly for these
+  - Putative annotations based on labeling pattern (and only found at surface)
+    - Maybe also can do a search for fragments
 
-  - Asp/Asn especially
-  - Ectoine major exception (ref mcparland for bacterial signal)
+```{r}
+peak_areas %>%
+  filter(compound_name%in%chosen_maas) %>%
+  # filter(compound_name%in%maa_names) %>%
+  # filter(compound_name=="Shinorine?") %>%
+  left_join(parametadata, relationship = "many-to-many") %>%
+  filter(amendment%in%c("Nit", "Amm")) %>%
+  filter(depth=="Surf") %>%
+  mutate(n_N_lab=str_extract(iso_name, "(?<=15N)\\d+")) %>%
+  ggplot() +
+  geom_col(aes(x=tripl, y=area, fill=n_N_lab), color="black", width=1) +
+  facet_nested(compound_name~amendment+startime+timepoint, scales="free_y") +
+  theme_bw()
+```
 
-Nitrate surface morning surprisingly limited
-
-  - "out of order" labeling in MAAs and guanine, maybe also arginine, taurine, NAc-lysine, maybe maybe theanine being unlabeled because it would mean diatoms don’t make it
-
-Organic compounds equally bioavailable at surface and deep and used very similarly to each other
-
-  - Exceptions: guanosine (not guanine!) for GMP amendments
-  - Exceptions: arginine, citrulline, maybe ornithine for Arg (signal goes away after a while and unclear why??)
-
-Organics seem to be largely used for ammonia rather than as-is
-
-  - Carbon backbone traced into glutamate from arginine
-  - Carbon backbone traced into other nucleobases from GMP
-  - Possibly due to way too much of a given compound, useless to have that much Arg/GMP
-  - Not very much fully-labeled guanine despite guanosine being easily cleavable, instead it’s mostly made from scratch (CHECK THIS)
-  - Creatine for being mostly 2x15N 1x13C labeled from arginine (CHECK THIS)
+## Inorganics more bioavailable at surface, organics equally bioavailable at 25m and 175m
 
 ```{r fig surface vs deep labeling rates}
-labeled_fracs %>%
+surfdeep_demo_cmpd_gp <- labeled_fracs %>%
+  mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
+  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
+  mutate(compound_class=case_when(
+    compound_name%in%c("Glutamine", "Aspartic acid", "Asparagine", "Glutamic acid", "Ectoine", "Theanine?", "Citrulline", "Guanine", "Guanosine", "Arginine") ~ compound_name,
+    compound_name%in%aa_names~"Other amino acids",
+    compound_name%in%nucleo_names~"Other nucleic acids",
+    compound_name%in%betaine_names~"Betaines",
+    TRUE~"All others"
+  )) %>%
+  mutate(compound_class=factor(compound_class, levels=setdiff(class_levels, "MAAs"))) %>%
+  left_join(parametadata %>% distinct(shortname, timepoint_num, depth, startime, amendment, tripl)) %>%
+  filter(compound_name=="Glutamic acid" & amendment=="Amm" |
+           compound_name=="Arginine" & amendment=="Arg" |
+           compound_name=="Guanosine" & amendment=="GMP") %>%
+  mutate(timepoint_num=ifelse(timepoint_num==0, 0.3, timepoint_num)) %>%
+  mutate(amendment=factor(amendment, levels=c("Amm", "Arg", "GMP"),
+                          labels=c("+Ammonia", "+Arginine", "+GMP"))) %>%
+  select(-shortname) %>%
+  summarise(mean_label=mean(labeled_frac), .by=c(compound_class, amendment, timepoint_num, depth)) %>%
+  pivot_wider(names_from=depth, values_from=mean_label) %>%
+  ggplot(aes(x=timepoint_num, color=compound_class, shape=compound_class)) +
+  geom_segment(aes(xend=timepoint_num, y=Surf, yend=Deep), lwd=1) +
+  geom_point(aes(y=Surf), fill="grey80", size=3, stroke=1) +
+  geom_point(aes(y=Deep), fill="grey20", size=3, stroke=1) +
+  stat_summary(aes(y=Surf, label=paste0(round(Surf-Deep, 2)*100, "%"), vjust=-sign(Surf-Deep)*1.5+0.75), 
+               geom = "text", fun = mean) +
+  facet_wrap(~amendment, nrow=1) +
+  scale_color_discrete(drop=FALSE) +
+  scale_x_log10(breaks=c(0.3, 3, 30), labels=c("0.3", "3", "30"), expand=expansion(0.15)) +
+  scale_y_continuous(labels=scales::label_percent(), expand=expansion(mult=c(0, 0.1)),
+                     breaks=c(0, 0.5, 1), limits = c(0, 1)) +
+  scale_shape_manual(breaks=c("Glutamic acid", "Arginine", "Guanosine"), values=c(22, 25, 21)) +
+  labs(x="Hours after <sup>15</sup>N spike", color=NULL, fill=NULL, shape=NULL,
+       y="Percent metabolite\nlabeled") +
+  theme_bw() +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    legend.position = "none"
+  )
+
+surfdeep_all_cmpd_gp <- labeled_fracs %>%
+  left_join(parametadata %>% distinct(shortname, timepoint_num, depth, startime, amendment, tripl)) %>%
   filter(!compound_name%in%maa_names) %>%
-  left_join(distinct(parametadata, shortname, amendment, depth, timepoint_num)) %>%
+  mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
+  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
+  mutate(compound_class=case_when(
+    compound_name%in%c("Glutamine", "Aspartic acid", "Asparagine", "Glutamic acid", "Ectoine", "Theanine?", "Citrulline", "Guanine", "Guanosine", "Arginine") ~ compound_name,
+    compound_name%in%aa_names~"Other amino acids",
+    compound_name%in%nucleo_names~"Other nucleic acids",
+    compound_name%in%betaine_names~"Betaines",
+    TRUE~"All others"
+  )) %>%
+  mutate(compound_class=factor(compound_class, levels=class_levels)) %>%
+  summarise(avg_frac=mean(labeled_frac), .by = c(compound_class, amendment, timepoint_num, depth, tripl)) %>%
   filter(amendment!="Nit") %>%
-  summarise(avg_label=mean(labeled_frac), .by=c(compound_name, amendment, depth, timepoint_num)) %>%
+  summarise(avg_label=mean(avg_frac), .by=c(compound_class, amendment, depth, timepoint_num)) %>%
   pivot_wider(names_from = depth, values_from=avg_label) %>%
   mutate(timepoint_num=ifelse(timepoint_num==0, 0.3, timepoint_num)) %>%
-  ggplot(aes(x=timepoint_num, y=Surf-Deep, color=compound_name, group=compound_name)) +
+  mutate(amendment=factor(amendment, levels=c("Amm", "Arg", "GMP"),
+                          labels=c("+Ammonia", "+Arginine", "+GMP"))) %>%
+  ggplot(aes(x=timepoint_num, y=Surf-Deep, color=compound_class, 
+             fill=compound_class, shape=compound_class)) +
+  geom_line(lwd=0.8) +
+  geom_point(color="black", size=3) +
   geom_hline(yintercept = 0) +
-  geom_line() +
-  geom_point() +
   facet_wrap(~amendment, nrow=1) +
-  scale_x_log10() +
+  scale_x_log10(breaks=c(0.3, 3, 30), labels=c("0.3", "3", "30"), expand=expansion(0.15)) +
+  scale_y_continuous(labels=scales::label_percent()) +
+  scale_shape_manual(breaks=setdiff(class_levels, "MAAs"), values = rep(21:25, length.out=14)) +
+  labs(x="Hours after <sup>15</sup>N spike", color=NULL, fill=NULL, shape=NULL,
+       y="Percent metabolite labeled at 25m\nminus percent labeled at 175m") +
   theme_bw() +
-  theme(legend.position = "none")
-ggsave(filename = "deep_vs_surf_labeling.png", device = "png", path = "manuscript/figures",
-       width = 6.5, height = 4, units = "in", dpi = 300)
+  theme(
+    # axis.title.x = element_markdown(),
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    strip.text = element_blank(),
+    strip.background = element_blank()
+  )
+
+surfdeep_all_cmpd_boxplot_gp <- labeled_fracs %>%
+  left_join(parametadata %>% distinct(shortname, timepoint_num, depth, startime, amendment, tripl)) %>%
+  filter(!compound_name%in%maa_names) %>%
+  filter(amendment!="Nit") %>%
+  mutate(timepoint_num=ifelse(timepoint_num==0, 0.3, timepoint_num)) %>%
+  summarise(labeled_frac=mean(labeled_frac), .by=c(timepoint_num, compound_name, depth, amendment)) %>%
+  ggplot(aes(x=timepoint_num, y=labeled_frac, group=interaction(timepoint_num, depth), fill=depth)) +
+  geom_boxplot(color="black", outlier.shape=21) +
+  facet_wrap(~amendment, nrow=1) +
+  scale_x_log10(breaks=c(0.3, 3, 30), labels=c("0.3", "3", "30"), expand=expansion(0.15/2)) +
+  scale_y_continuous(labels = scales::label_percent(), breaks=c(0, 0.5, 1)) +
+  scale_fill_manual(breaks=c("Surf", "Deep"), values=c("grey90", "grey30"), labels=c("25 meters", "175 meters")) +
+  labs(x="Hours after <sup>15</sup>N spike", color=NULL, fill=NULL, shape=NULL,
+       y="Percent metabolite\nlabeled") +
+  theme_bw() +
+  theme(
+    axis.title.x = element_markdown(),
+    strip.text = element_blank(),
+    strip.background = element_blank()
+  )
+
+surfdeep_gp <- surfdeep_demo_cmpd_gp / surfdeep_all_cmpd_gp / surfdeep_all_cmpd_boxplot_gp + plot_layout(guides = "collect", heights = c(0.3, 0.5, 0.2))
+
+ggsave(filename = "deep_vs_surf_labeling.png", surfdeep_gp, device = "png", path = "figures",
+       width = 6.5, height = 6.5, units = "in", dpi = 300)
 ```
@@
 
-*Figure XX: Line plots of the metabolite fraction labeled at the surface (25 meters) minus the fraction labeled at 175 meters.
+*Figure 4: Plots of the metabolite fraction labeled at the surface (25 meters) and the fraction labeled at 175 meters.
 Incubations with added ammonia, arginine, and guanosine monophosphate (GMP) are shown.
-Points are the average of six biological replicates (incubations started in the morning and the evening are combined for additional power) and colors have been added solely for disambiguating purposes.
-Positive values indicate increased synthesis in the surface samples relative to those at depth.*
+Top row of plots show the calculation of % labeled at 25 meters minus % labeled at 175 meters for three example compounds.
+All compounds are shown in the middle row of plots, grouped by chemical similarity and labeling pattern where positive values indicate increased synthesis in the surface samples relative to those at depth.
+Finally, the bottom row of plots shows the relative difference between the surface and deep as boxplots for all compounds without aggregating.
+Points in the first two rows are the average of six biological replicates (incubations started in the morning and the evening are combined here for additional power) while points in the bottom row denote metabolites with labeled fractions falling outside 1.5 times the interquartile range.*
 
-```{r regression attempts currently failing, eval=FALSE}
-labeled_fracs %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(compound_name=="Creatinine") %>%
-  filter(amendment=="GMP") %>%
-  mutate(timepoint_num=ifelse(timepoint_num==0, 0.3, timepoint_num)) %>%
-  ggplot(aes(x=timepoint_num, y=labeled_frac, color=startime)) +
-  geom_point() +
-  geom_smooth(method=glm, formula=y~x, method.args=list(family=quasibinomial)) +
-  facet_wrap(~depth, ncol=1)
+  - Ammonia incubations result in much faster labeling at 25 meters than 175m
+    - Presumed to be a phototrophic signal
+    - Major exception is ectoine, a bacterial biosignal (McParland 2021)
+  - Organic N incubations result in equal amounts of labeling at surface and depth
+    - Ectoine again predominantly labeled at depth
+    - Labeling of guanosine at T0 in the +GMP treatment implies rapid uptake and P-cleavage
+  - Nitrate data unavailable due to T10 and T26 deep samples contaminated
+    - See supplemental figure 1 for the data anyway
 
-reg_data <- labeled_fracs %>%
-  filter(!compound_name%in%maa_names) %>%
-  left_join(distinct(parametadata, shortname, amendment, depth, timepoint_num)) %>%
-  mutate(max_frac=max(labeled_frac[timepoint_num==max(timepoint_num)]), .by=c(amendment, compound_name, depth)) %>%
-  mutate(t10_labfrac = mean(labeled_frac[timepoint_num==sort(unique(timepoint_num))[4]]), .by=c(amendment, compound_name, depth))
+## Other results from untargeted analysis
 
-reg_data %>%
-  distinct(compound_name, amendment, depth, max_frac, t10_labfrac) %>%
-  mutate(category=ifelse(max_frac < 0.1, "Unlabeled", "Good")) %>%
-  mutate(category=ifelse(t10_labfrac <= 0, "Only T26 labeled", category)) %>%
-  mutate(category=ifelse(is.na(category), "DMSOP (unlabeled)", category)) %>%
-  count(amendment, category, depth) %>%
-  print(n=Inf)
-
-reg_fits <- reg_data %>%
-  filter(max_frac>0.1) %>%
-  filter(t10_labfrac>0) %>%
-  filter(compound_name!="Betonicine") %>%
-  filter(compound_name!="Arsenobetaine") %>%
-  filter(compound_name!="Gonyol") %>%
-  mutate(timepoint_num=ifelse(timepoint_num==0, 0.3, timepoint_num)) %>%
-  nest(data=c(shortname, timepoint_num, labeled_frac)) %>%
-  mutate(log_fit=map(data, ~glm(.x$labeled_frac~.x$timepoint_num, family="quasibinomial")),
-         coefs=map(log_fit, function(x)broom::tidy(x))) %>%
-  unnest(coefs) %>%
-  mutate(p_adj=p.adjust(p.value, method = "fdr"), .by = term) %>%
-  filter(p_adj<0.001) %>%
-  select(compound_name, amendment, depth, term, estimate) %>%
-  pivot_wider(names_from = term, values_from = estimate) %>%
-  rename(intercept=`(Intercept)`, slope=`.x$timepoint_num`) %>%
-  mutate(half_max=-intercept/slope, scale=1/slope) %>%
-  mutate(frac_at_T24 = 1/(1+exp(-(intercept+slope*24))))
-# reg_fits %>%
-#   select(compound_name, amendment, depth, intercept, slope) %>%
-#   pivot_longer(intercept:slope) %>%
-#   ggplot(aes(x=depth, y=value)) +
-#   geom_boxplot() +
-#   facet_grid(name~amendment, scales="free_y")
-# reg_fits %>%
-#   select(compound_name, amendment, depth, intercept, slope) %>%
-#   pivot_longer(intercept:slope) %>%
-#   pivot_wider(names_from=depth) %>%
-#   filter(!is.na(Deep)) %>%
-#   nest(data=c(compound_name, Surf, Deep)) %>%
-#   mutate(test=map(data, ~broom::tidy(t.test(.x$Surf, .x$Deep, paired=TRUE)))) %>%
-#   unnest(test)
-# labeled_fracs %>%
-#   left_join(parametadata %>% distinct(shortname, amendment, depth, timepoint_num)) %>%
-#   filter(timepoint_num==max(timepoint_num), .by=amendment) %>%
-#   summarise(avg_Tf_label=mean(labeled_frac), .by=c(compound_name, amendment, depth)) %>%
-#   ggplot() +
-#   geom_boxplot(aes(x=amendment, y=avg_Tf_label, color=depth))
-# 
-# labeled_fracs %>%
-#   left_join(parametadata %>% distinct(shortname, amendment, depth, timepoint_num)) %>%
-#   filter(timepoint_num==max(timepoint_num), .by=amendment) %>%
-#   summarise(avg_Tf_label=mean(labeled_frac), .by=c(compound_name, amendment, depth)) %>%
-#   pivot_wider(names_from=depth, values_from=avg_Tf_label) %>%
-#   ggplot() +
-#   geom_segment(aes(y=Surf, yend=Deep), x=1, xend=2) +
-#   facet_wrap(~amendment, nrow=1)
-# reg_fits %>%
-#   ggplot() +
-#   geom_segment(aes(y=frac_at_T24_Surf, yend=frac_at_T24_Deep,
-#                    x=half_max_Surf, xend=half_max_Deep, color=compound_name),
-#                data = reg_fits %>%
-#                  select(compound_name, depth, half_max, frac_at_T24) %>% 
-#                  pivot_wider(names_from = depth, values_from=c(half_max, frac_at_T24)) %>%
-#                  filter(!is.na(half_max_Deep))) +
-#   geom_point(aes(x=half_max, y=frac_at_T24, fill=depth), pch=21) +
-#   scale_fill_manual(values=c("black", "white"), breaks=c("Deep", "Surf"))
-
-reg_fits %>%
-  filter(amendment!="Nit") %>%
-  ggplot() +
-  geom_segment(aes(y=frac_at_T24_Surf, yend=frac_at_T24_Deep,
-                   x=half_max_Surf, xend=half_max_Deep, color=compound_name),
-               data = reg_fits %>%
-                 select(compound_name, amendment, depth, half_max, frac_at_T24) %>% 
-                 pivot_wider(names_from = depth, values_from=c(half_max, frac_at_T24)) %>%
-                 filter(!is.na(half_max_Deep))) +
-  geom_point(aes(x=half_max, y=frac_at_T24, fill=depth, label=compound_name), pch=21) +
-  scale_fill_manual(values=c("black", "white"), breaks=c("Deep", "Surf")) +
-  facet_wrap(~amendment, scales="free_x") +
-  theme(legend.position = "none")
-```
-
-## Rapid uptake and use of all N substrates
-
-Total fraction of PN that was visible in metabolites
-
-```{r}
-env_concs <- read_csv("targeted/final_env_concs.csv")
-
-# More compounds, triplicates shown
-env_concs %>%
-  mutate(mean_nm=mean(conc_nm_in_env), .by=compound_name) %>%
-  filter(!compound_name%in%c("Thymine", "Thymidine")) %>%
-  arrange(desc(mean_nm)) %>%
-  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", "Glutamate", compound_name)) %>%
-  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:19], compound_name, "Other")) %>%
-  mutate(compound_name=fct_inorder(compound_name)) %>%
-  summarise(conc_nm=sum(conc_nm_in_env), .by=c(compound_name, filename)) %>%
-  left_join(parametadata) %>%
-  ggplot(aes(x=tripl, y=conc_nm, fill=compound_name)) +
-  geom_col(color="black", width=1) +
-  facet_nested(depth+startime~amendment+timepoint, switch="x") +
-  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
-  scale_fill_manual(values = c(splividis(19), "grey")) +
-  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
-  theme_bw() +
-  theme(
-    panel.spacing.x = unit(0, "lines"),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()
-  )
-
-# Summarized version
-env_concs %>%
-  mutate(mean_nm=mean(conc_nm_in_env), .by=compound_name) %>%
-  filter(!compound_name%in%c("Thymine", "Thymidine")) %>%
-  arrange(desc(mean_nm)) %>%
-  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", 
-                              "Glutamate", compound_name)) %>%
-  mutate(compound_name=str_remove_all(compound_name, "^L-|(?<=/)L-")) %>%
-  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:9], 
-                              compound_name, "Other")) %>%
-  mutate(compound_name=fct_inorder(compound_name)) %>%
-  summarise(conc_nm=sum(conc_nm_in_env), .by=c(compound_name, filename)) %>%
-  left_join(parametadata) %>%
-  summarise(conc_nm=mean(conc_nm), .by=c(compound_name, amendment, depth, timepoint)) %>%
-  ggplot(aes(x=timepoint, y=conc_nm, fill=compound_name)) +
-  geom_col(color="black", width=1) +
-  facet_grid(depth~amendment, scales="free") +
-  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
-  scale_fill_manual(values = c(splividis(9), "grey")) +
-  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
-  theme_bw()
-
-# In N moles
-env_concs %>%
-  left_join(stan_patched %>% distinct(compound_name, n_N)) %>%
-  mutate(nM_N=conc_nm_in_env*n_N) %>%
-  mutate(mean_nm=mean(nM_N), .by=compound_name) %>%
-  filter(!compound_name%in%c("Thymine", "Thymidine")) %>%
-  arrange(desc(mean_nm)) %>%
-  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", 
-                              "Glutamate", compound_name)) %>%
-  mutate(compound_name=str_remove_all(compound_name, "^L-|(?<=/)L-")) %>%
-  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:9], 
-                              compound_name, "Other")) %>%
-  mutate(compound_name=fct_inorder(compound_name)) %>%
-  summarise(conc_nm=sum(nM_N), .by=c(compound_name, filename)) %>%
-  left_join(parametadata) %>%
-  summarise(conc_nm=mean(conc_nm), .by=c(compound_name, amendment, depth, timepoint)) %>%
-  ggplot(aes(x=timepoint, y=conc_nm, fill=compound_name)) +
-  geom_col(color="black", width=1) +
-  facet_grid(depth~amendment, scales="free") +
-  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
-  scale_fill_manual(values = c(splividis(9), "grey")) +
-  labs(x=NULL, y="Nitrogen in unlabeled metabolites (nM)", fill=NULL) +
-  theme_bw()
-```
-
-```{r}
-peak_areas %>%
-  filter(str_detect(iso_name, "13C0, 15N0")) %>%
-  filter(!compound_name%in%c("Urea", "L-Valine", "Oxalic acid", "Malic acid", 
-                             "Acetylcholine", "Homarine")) %>%
-  left_join(parametadata %>% distinct(filename, samp_type, amendment)) %>%
-  filter(samp_type=="Smp") %>%
-  left_join(read_csv("targeted/rf_df.csv")) %>%
-  mutate(conc_nm=area/rf) %>%
-  filter(!is.na(conc_nm)) %>%
-  mutate(mean_nm=mean(conc_nm), .by=compound_name) %>%
-  arrange(desc(mean_nm)) %>%
-  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", 
-                              "Glutamate", compound_name)) %>%
-  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:19], 
-                              compound_name, "Other")) %>%
-  mutate(compound_name=fct_inorder(compound_name)) %>%
-  summarise(conc_nm=sum(conc_nm), .by=c(compound_name, filename)) %>%
-  left_join(parametadata) %>%
-  ggplot(aes(x=tripl, y=conc_nm, fill=compound_name)) +
-  geom_col(color="black", width=1) +
-  facet_nested(depth+startime~amendment+timepoint, switch="x") +
-  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
-  scale_fill_manual(values = c(splividis(19), "grey")) +
-  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
-  theme_bw() +
-  theme(
-    panel.spacing.x = unit(0, "lines"),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()
-  )
-```
-
-## Labeling traced into protein via total hydrolyzable amino acids
-
-```{r thaa data}
-thaa_quant <- read_csv("thaas/quant_data.csv")
-metathaa <- read_csv("thaas/metathaa.csv") %>%
-  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
-  mutate(startime=factor(startime, levels=c("Morn", "Eve")))
-
-all_thaa_plot <- thaa_quant %>%
-  filter(!str_detect(iso_name, "2H")) %>%
-  left_join(metathaa) %>%
-  filter(samp_type=="Smp") %>%
-  left_join(aa_ncs) %>%
-  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
-  mutate(n_N_lab=as.numeric(str_extract(iso_name, "(?<=15N)\\d+"))) %>%
-  mutate(n_N_unlab=n_N-n_N_lab) %>%
-  mutate(nM_N_lab=env_conc_nm*n_N_lab, nM_N_unlab=env_conc_nm*n_N_unlab) %>%
-  select(filename, compound_name, iso_name, nM_N_lab, nM_N_unlab) %>%
-  pivot_longer(c(nM_N_lab, nM_N_unlab)) %>%
-  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
-  left_join(metathaa) %>%
-  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
-  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
-  summarise(total_uM=mean(total_uM), .by=c(amendment, depth, startime, timepoint, name)) %>%
-  arrange(depth, startime) %>%
-  mutate(x_axis=paste(depth, startime)) %>%
-  mutate(x_axis=fct_inorder(x_axis)) %>%
-  ggplot(aes(x=x_axis, y=total_uM, fill=name)) +
-  geom_col(color="black", width=1) +
-  facet_nested(.~amendment, scales="free_y") +
-  theme_bw() +
-  scale_x_discrete(expand = expansion()) +
-  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
-  scale_fill_manual(breaks = c("nM_N_unlab", "nM_N_lab"), values=c("grey80", "grey20"), 
-                    labels=c("Unlabeled", "Labeled")) +
-  labs(x=NULL, y="Total uM N in THAAs", fill=NULL) +
-  theme(
-    axis.text.x=element_blank()
-  )
-
-thaa_plot_data <- thaa_quant %>%
-  left_join(metathaa) %>%
-  filter(samp_type=="Smp") %>%
-  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
-  summarise(avg_env_conc_nm=mean(env_conc_nm, na.rm=TRUE), .by=c(compound_name, iso_name, amendment, timepoint, startime, depth)) %>%
-  filter(compound_name%in%c("Glutamic acid", "Arginine")) %>%
-  mutate(iso_name=case_when(
-    str_detect(iso_name, "13C0, 15N0")~"Unlabeled",
-    str_detect(iso_name, "13C0, 15N1")~"Singly 15N labeled",
-    str_detect(iso_name, "13C1, 15N0")~"Singly 13C labeled",
-    str_detect(iso_name, "Glutamic acid, 13C5, 15N1")~"Fully C and N\nlabeled glutamate",
-    str_detect(iso_name, "Arginine, 13C0, 15N3")~"3x 15N\nlabeled arginine",
-    str_detect(iso_name, "Arginine, 13C0, 15N4")~"Fully N\nlabeled arginine",
-    str_detect(iso_name, "Arginine, 13C6, 15N4")~"Fully C and N\nlabeled arginine",
-    TRUE~"Other labeled"
-  )) %>%
-  mutate(iso_name=factor(iso_name, levels=c(
-    "Unlabeled", "Singly 15N labeled", "Singly 13C labeled", 
-    "3x 15N\nlabeled arginine", "Fully N\nlabeled arginine", 
-    "Fully C and N\nlabeled arginine", "Fully C and N\nlabeled glutamate", 
-    "Other labeled"))) %>%
-  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
-  arrange(depth, startime) %>%
-  mutate(x_axis=paste(depth, startime)) %>%
-  mutate(x_axis=fct_inorder(x_axis))
-cpal_breaks <- levels(thaa_plot_data$iso_name)
-cpal_cols <- c("grey80", viridis::viridis(length(cpal_breaks)-2), "grey20")
-arg_glut_thaa_plot <- ggplot(thaa_plot_data) +
-  geom_col(aes(x=x_axis, y=avg_env_conc_nm, fill=iso_name),
-           position="fill",
-           color="black", width=1) +
-  facet_wrap2(compound_name~amendment, ncol=5, strip = strip_nested(bleed = TRUE)) +
-  theme_bw() +
-  scale_x_discrete(expand = expansion()) +
-  scale_y_continuous(labels=scales::label_percent(), breaks=c(0, 0.5, 1),
-                     expand = expansion()) +
-  scale_fill_manual(breaks = cpal_breaks, values = cpal_cols) +
-  labs(x=NULL, y="Fraction labeled", fill=NULL) +
-  theme(
-    axis.text.x=element_text(angle=90, hjust=1, vjust = 0.5)
-  )
-total_thaa_gp <- all_thaa_plot / arg_glut_thaa_plot + plot_layout(heights = c(0.4, 0.6))
-ggsave("total_thaa_gp.png", total_thaa_gp, device = "png", path = "manuscript/figures",
-       width = 6.5, height = 5.5, units = "in", dpi = 300)
-```
-
-![](figures/total_thaa_gp.png)
-
-  - THAAs were 30-50% of total PC (~2 uM, need actual data) and 70-100% PN (~0.35uM)
-  - Maybe more labeling happening from inorganics in evening?
-  - 50% turnover of arginine in proteins in 3 days per +Arg data, less at surface than deep (??)
-  - 30-50% Glu/Gln turnover in proteins in one day per Amm data
-  - Glutamate, proline fully labeled labeled from arginine
-  - Phenylalanine, tyrosine, histidine(?) labeled from GMP
-  - How is there so much nitrate labeling when I don't see that in the metabolites?
+Upcoming.
 

### 2025-05-12 manuscript/manuscript.Rmd
Cleaning up data load things, placing in separate chunk

@@
 options(pillar.print_min = 100)
+```
 
+```{r load data}
 parametadata <- read_csv("metadata/parametadata.csv") %>%
@@
   select(-iso_name)
+
+thaa_quant <- read_csv("quant_data.csv")
+metathaa <- read_csv("metathaa.csv") %>%
+  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve")))
 ```
@@
 ```{r thaa data}
-thaa_quant <- read_csv("quant_data.csv")
-metathaa <- read_csv("metathaa.csv") %>%
-  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
-  mutate(startime=factor(startime, levels=c("Morn", "Eve")))
-
 all_thaa_plot <- thaa_quant %>%

### 2025-05-12 manuscript/manuscript.Rmd
Cleaning up old figures and making outputs clearer

@@
 ---
-
-```{r Run analysis, include=FALSE}
-# TBD
-```
 
@@
 
-```{r load data}
+```{r load data, eval=FALSE}
 parametadata <- read_csv("metadata/parametadata.csv") %>%
@@
   mutate(timepoint=fct_inorder(timepoint))
-peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv") %>%
-  left_join(parametadata %>% distinct(filename, shortname))
+
 all_stans <- read_csv("metadata/all_stans.csv")
@@
   mutate(compound_name=str_replace(compound_name, "Hypoxanthine", "Hypoxanthine/Allopurinol"))
-
 maa_names <- c("Also Palythene/Usujirene?", "Shinorine?", "Mycosporine-glycine?", 
@@
 
+peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv") %>%
+  left_join(parametadata %>% distinct(filename, shortname))
 samp_areas <- parametadata %>%
@@
 
-thaa_quant <- read_csv("quant_data.csv")
-metathaa <- read_csv("metathaa.csv") %>%
+thaa_quant <- read_csv("thaas/quant_data.csv") %>%
+  filter(iso_name!="Proline, 13C3, 15N1")
+metathaa <- read_csv("thaas/metathaa.csv") %>%
   mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
@@
   )
-ggsave("xamend_thaa_gp.png",plot = xamend_metab_gp, device = "png", path = "figures",
+ggsave("xamend_thaa_gp.png",plot = xamend_metab_gp, device = "png", path = "manuscript/figures",
        width = 6.5, height = 5.5, units = "in", dpi = 300)
@@
       - 30-50% arg in THAAs labeled @73hrs
-      - Clear reuse of arginine backbone as glutamate (fully labeled glutamate) and maybe proline?
+      - Clear reuse of arginine backbone as fully labeled glutamate and proline
   - Guanosine monophosphate
@@
 total_thaa_gp
-ggsave("total_thaa_gp.png", total_thaa_gp, device = "png", path = "figures",
+ggsave("total_thaa_gp.png", total_thaa_gp, device = "png", path = "manuscript/figures",
        width = 6.5, height = 5, units = "in", dpi = 300)
@@
 
-ggsave(filename = "MAA_curves.png", device = "png", path = "figures",
+ggsave(filename = "MAA_curves.png", device = "png", path = "manuscript/figures",
        width = 3, height = 6, units = "in", dpi = 300)
@@
     - Maybe also can do a search for fragments
-
-```{r}
-peak_areas %>%
-  filter(compound_name%in%chosen_maas) %>%
-  # filter(compound_name%in%maa_names) %>%
-  # filter(compound_name=="Shinorine?") %>%
-  left_join(parametadata, relationship = "many-to-many") %>%
-  filter(amendment%in%c("Nit", "Amm")) %>%
-  filter(depth=="Surf") %>%
-  mutate(n_N_lab=str_extract(iso_name, "(?<=15N)\\d+")) %>%
-  ggplot() +
-  geom_col(aes(x=tripl, y=area, fill=n_N_lab), color="black", width=1) +
-  facet_nested(compound_name~amendment+startime+timepoint, scales="free_y") +
-  theme_bw()
-```
 
@@
 
-ggsave(filename = "deep_vs_surf_labeling.png", surfdeep_gp, device = "png", path = "figures",
-       width = 6.5, height = 6.5, units = "in", dpi = 300)
+ggsave(filename = "deep_vs_surf_labeling.png", surfdeep_gp, device = "png", 
+       path = "manuscript/figures", width = 6.5, height = 6.5, units = "in", dpi = 300)
 ```
@@
   pull(compound_name) %>%
-  str_remove_all("^L-|(?<=)/L-") %>%
+  str_remove_all("^L-|(?<=/)L-") %>%
   c("5-Methylcytosine")
@@
   mutate(compound_name=str_remove_all(compound_name, "^L-|(?<=/)L-")) %>%
-  mutate(compound_name=factor(compound_name, cmpd_order)) %>%
+  mutate(order_name=factor(compound_name, cmpd_order)) %>%
   mutate(compound_class=case_when(
@@
 
-ggsave(plot = heatmap_gp, "manuscript/figures/supp_para_map.pdf", device = "pdf",
+ggsave(plot = heatmap_gp, "manuscript/supplement/para_heatmap.pdf", device = "pdf",
        width = 10, height = 12)
@@
 ```{r thaa supplement figures}
-thaa_quant %>%
-  left_join(metathaa) %>%
-  filter(samp_type=="Smp") %>%
-  ggplot() +
-  geom_col(aes(x=tripl, y=env_conc_nm, fill=compound_name), 
-           color="black", width=1, position="fill") +
-  facet_nested(depth+startime~amendment+timepoint) +
-  scale_x_discrete(expand = expansion()) +
-  scale_y_continuous(labels=scales::label_percent(), breaks=c(0, 0.5, 1),
-                     expand = expansion()) +
-  labs(x="Triplicate", y="Fraction labeled", fill=NULL) +
-  theme_bw() +
-  theme(
-    axis.text.y=element_markdown(vjust=c(0, 0.5, 1))
-  )
-
-THAA_mol_C_gp <- thaa_quant %>%
-  filter(!str_detect(iso_name, "2H")) %>%
-  left_join(metathaa) %>%
-  filter(samp_type=="Smp") %>%
-  left_join(aa_ncs) %>%
-  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
-  mutate(n_C_lab=as.numeric(str_extract(iso_name, "(?<=13C)\\d+"))/n_C) %>%
-  mutate(n_C_unlab=n_C-n_C_lab) %>%
-  mutate(nM_C_lab=env_conc_nm*n_C_lab, nM_C_unlab=env_conc_nm*n_C_unlab) %>%
-  select(filename, compound_name, iso_name, nM_C_lab, nM_C_unlab) %>%
-  pivot_longer(c(nM_C_lab, nM_C_unlab)) %>%
-  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
-  left_join(metathaa) %>%
-  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
-  ggplot(aes(x=tripl, y=total_uM, fill=name)) +
-  geom_col(color="black", width=1) +
-  facet_nested(depth+startime~amendment+timepoint) +
-  theme_bw() +
-  scale_x_discrete(expand = expansion()) +
-  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
-  scale_fill_manual(breaks = c("nM_C_unlab", "nM_C_lab"), values=c("grey80", "grey20"), 
-                    labels=c("Unlabeled", "Labeled")) +
-  labs(x="Triplicate", y="Total C in THAAs (uM)", fill=NULL) +
-  theme(
-    legend.position = "top"
-  )
-THAA_mol_N_gp <- thaa_quant %>%
-  filter(!str_detect(iso_name, "2H")) %>%
-  left_join(metathaa) %>%
-  filter(samp_type=="Smp") %>%
-  left_join(aa_ncs) %>%
-  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
-  mutate(n_N_lab=as.numeric(str_extract(iso_name, "(?<=15N)\\d+"))) %>%
-  mutate(n_N_unlab=n_N-n_N_lab) %>%
-  mutate(nM_N_lab=env_conc_nm*n_N_lab, nM_N_unlab=env_conc_nm*n_N_unlab) %>%
-  select(filename, compound_name, iso_name, nM_N_lab, nM_N_unlab) %>%
-  pivot_longer(c(nM_N_lab, nM_N_unlab)) %>%
-  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
-  left_join(metathaa) %>%
-  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"))) %>%
-  ggplot(aes(x=tripl, y=total_uM, fill=name)) +
-  geom_col(color="black", width=1) +
-  facet_nested(depth+startime~amendment+timepoint) +
-  theme_bw() +
-  scale_x_discrete(expand = expansion()) +
-  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
-  scale_fill_manual(breaks = c("nM_N_unlab", "nM_N_lab"), values=c("grey80", "grey20"), 
-                    labels=c("Unlabeled", "Labeled")) +
-  labs(x="Triplicate", y="Total N in THAAs (uM)", fill=NULL) +
-  theme(
-    legend.position = "top"
-  )
-all_thaa_NC_gp <- guide_area() + THAA_mol_C_gp + THAA_mol_N_gp + plot_layout(guides = "collect", heights = c(0.1, 0.45, 0.45), design = c("1\n2\n3"))
-ggsave("thaas/all_thaa_NC_gp.png", width=6.5, height=8)
-
-pdf("figures/all_cmpd_labels.pdf", width=9, height = 5)
+pdf("manuscript/supplement/THAA_labeling.pdf", width=9, height = 5)
 for(cmpd_i in setdiff(unique(thaa_quant$compound_name), c("Citrulline", "Ornithine", "Taurine"))){
@@
     filter(compound_name==cmpd_i) %>%
-    left_join(metathaa) %>%
+    left_join(metathaa, by = join_by(filename)) %>%
     drop_na() %>%
@@
     facet_nested(depth+startime~amendment+timepoint) +
-    ggtitle(cmpd_i)
+    scale_y_continuous(labels = scales::label_percent()) +
+    theme_bw() +
+    ggtitle(cmpd_i) +
+    labs(x="Biological triplicate", y="Fraction labeled", fill="Isotopologue")
   print(gp)
@@
 ```
-
-## Supp fig: Fraction of AA as DFAA vs THAA
-
-```{r fraction of aa pool in df vs th}
-thaa_pool_sizes <- thaa_quant %>%
-  filter(!compound_name%in%c("Citrulline", "Ornithine", "Taurine")) %>%
-  filter(str_detect(iso_name, "13C0, 15N0")) %>%
-  filter(str_detect(filename, "Smp.*T0")) %>%
-  select(filename, compound_name, env_conc_nm) %>%
-  mutate(compound_name=ifelse(compound_name=="Isoleucine", "Leucine", compound_name)) %>%
-  summarise(env_conc_nm=sum(env_conc_nm), .by=c(filename, compound_name)) %>%
-  mutate(compound_name=ifelse(compound_name=="Leucine", "Leu/Ile", compound_name)) %>%
-  mutate(compound_name=ifelse(compound_name=="Glutamic acid", "Glu/Gln", compound_name)) %>%
-  mutate(compound_name=ifelse(compound_name=="Aspartic acid", "Asp/Asn", compound_name)) %>%
-  left_join(metathaa) %>%
-  summarise(mean_thaa_conc=mean(env_conc_nm), .by=c(amendment, depth, startime, compound_name))
-metab_pool_sizes <- env_concs %>%
-  filter(str_detect(filename, "Smp.*T0")) %>%
-  mutate(compound_name=str_remove_all(compound_name, "L-")) %>%
-  mutate(compound_name=str_remove(compound_name, "/Homoserine")) %>%
-  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
-  mutate(compound_name=ifelse(compound_name=="Glutamine", "Glutamic acid", compound_name)) %>%
-  mutate(compound_name=ifelse(compound_name=="Asparagine", "Aspartic acid", compound_name)) %>%
-  summarise(conc_nm_in_env=sum(conc_nm_in_env), .by=c(filename, compound_name)) %>%
-  mutate(compound_name=ifelse(compound_name=="Glutamic acid", "Glu/Gln", compound_name)) %>%
-  mutate(compound_name=ifelse(compound_name=="Aspartic acid", "Asp/Asn", compound_name)) %>%
-  mutate(compound_name=ifelse(compound_name=="Leucine/Isoleucine", "Leu/Ile", compound_name)) %>%
-  filter(compound_name%in%thaa_pool_sizes$compound_name) %>%
-  left_join(parametadata) %>%
-  summarise(mean_metab_conc=mean(conc_nm_in_env), .by=c(amendment, depth, startime, compound_name))
-# Valine and phenylalanine failed QC
-thaa_pool_sizes %>%
-  left_join(metab_pool_sizes) %>%
-  ggplot() +
-  geom_point(aes(x=mean_thaa_conc, y=mean_metab_conc, color=compound_name)) +
-  facet_nested(depth+startime~amendment)
-thaa_pool_sizes %>%
-  left_join(metab_pool_sizes) %>%
-  ggplot() +
-  geom_point(aes(x=mean_thaa_conc, y=mean_metab_conc)) +
-  facet_wrap(~compound_name, scales="free")
-
-thaa_pool_sizes %>%
-  left_join(metab_pool_sizes) %>%
-  mutate(metab_frac=mean_metab_conc/mean_thaa_conc) %>%
-  ggplot() +
-  geom_density(aes(x=metab_frac, color=compound_name))
-thaa_pool_sizes %>%
-  left_join(metab_pool_sizes) %>%
-  mutate(metab_frac=mean_metab_conc/mean_thaa_conc) %>%
-  ggplot() +
-  geom_boxplot(aes(x=compound_name, y=metab_frac)) +
-  scale_y_continuous(labels = scales::label_percent()) +
-  labs(x=NULL, y='Fraction of AA in "free" pool: [metab] / [THAA]') +
-  coord_flip()
-```

### 2025-05-12 manuscript/manuscript.Rmd
Fixing THAA plot and adding proline + tyrosine

@@
 )
-# "#E41A1C" "#377EB8" "#4DAF4A" "#984EA3" "#FF7F00" "#FFFF33" "#A65628" "#F781BF" "#999999"
+thaa_spec_cmpds <- c("Glutamine + Glutamate", "Arginine", "Proline", "Histidine", "Phenylalanine", "Tyrosine")
 thaa_plot_data <- thaa_quant %>%
@@
   summarise(avg_env_conc_nm=mean(env_conc_nm, na.rm=TRUE), .by=c(compound_name, iso_name, amendment, timepoint)) %>%
-  filter(compound_name%in%c("Glutamic acid", "Arginine", "Histidine", "Phenylalanine")) %>%
   mutate(compound_name=ifelse(compound_name=="Glutamic acid", "Glutamine + Glutamate", compound_name)) %>%
+  mutate(compound_name=factor(compound_name, levels = thaa_spec_cmpds)) %>%
+  filter(!is.na(compound_name)) %>%
   mutate(plot_iso_name=case_when(
@@
     str_detect(iso_name, "13C1, 15N1")~"<sup>13</sup>C<sub>1</sub><sup>15</sup>N<sub>1</sub>",
-    str_detect(iso_name, "Glutamic acid, 13C5, 15N1")~"Entirely <sup>13</sup>C and <sup>15</sup>N",
+    str_detect(iso_name, "Glutamic acid, 13C5, 15N1")~"Entirely <sup>13</sup>C<br>and <sup>15</sup>N",
+    str_detect(iso_name, "Proline, 13C5, 15N1")~"Entirely <sup>13</sup>C<br>and <sup>15</sup>N",
     str_detect(iso_name, "Arginine, 13C6, 15N4")~"Entirely <sup>13</sup>C<br>and <sup>15</sup>N",
+    iso_name=="Tyrosine, 13C4, 15N1"~"<sup>13</sup>C<sub>4</sub><sup>15</sup>N<sub>1</sub>",
     iso_name=="Phenylalanine, 13C4, 15N1"~"<sup>13</sup>C<sub>4</sub><sup>15</sup>N<sub>1</sub>",

### 2025-05-16 untargeted/old/paruntargeted_control.Rmd
Moving new untargeted analysis over from ILC

--- 2023-11-07 untargeted/paruntargeted_control.Rmd
@@
 library(xcms)
+library(MSnbase)
 library(RaMS)
+library(squallms)
 
@@
 # polarity <- "neg"
-# amendment <- "Amm"
+amendment <- "Amm"
 # amendment <- "Nit"
-amendment <- "GMP"
+# amendment <- "GMP"
 # amendment <- "Arg"
@@
   mutate(samp_group=paste(amendment, depth, startime, timepoint, sep="-"))
-
-mz_group <- function(mz_vals, ppm){
-  group_vec <- numeric(length(mz_vals))
-  group_num <- 1L
-  init_vec <- mz_vals
-  names(init_vec) <- seq_along(init_vec)
-  while(length(init_vec)>0){
-    mz_i <- init_vec[1]
-    err <- mz_i*ppm/1000000
-    mz_idxs <- init_vec>mz_i-err & init_vec<mz_i+err
-    group_vec[as.numeric(names(mz_idxs)[mz_idxs])] <- group_num
-    init_vec <- init_vec[!mz_idxs]
-    group_num <- group_num+1L
-  }
-  group_vec
-}
-trapz <- function(x, y, na.rm=TRUE) {
-  y <- y[order(x)]
-  x <- x[order(x)]
-  m <- length(x)
-  xp <- c(x, x[m:1])
-  yp <- c(numeric(m), y[m:1])
-  n <- 2*m
-  p1 <- sum(xp[1:(n-1)]*yp[2:n], na.rm = na.rm) + xp[n]*yp[1]
-  p2 <- sum(xp[2:n]*yp[1:(n-1)], na.rm = na.rm) + xp[1]*yp[n]
-
-  return(0.5*(p1-p2))
-}
 ```
@@
 msnexp_filled <- readRDS(paste0(output_folder, "msnexp_filled.rds"))
-peak_data_long <- msnexp_filled %>%
-  chromPeaks() %>%
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  mutate(peakidx=row_number())
-peak_data <- msnexp_filled %>%
-  featureDefinitions() %>%
-  as.data.frame() %>%
-  select(mzmed, rtmed, npeaks, peakidx) %>%
-  rownames_to_column("id") %>%
-  unnest_longer(peakidx) %>%
-  rename_with(~paste0("feat_", .x), .cols = -peakidx) %>%
-  dplyr::rename(feature="feat_id") %>%
-  left_join(peak_data_long, by = join_by(peakidx)) %>%
-  mutate(filename=basename(fileNames(msnexp_filled))[sample]) %>%
-  select(feature, feat_mzmed, feat_rtmed, feat_npeaks, filename,
-         mz, mzmin, mzmax, rt, rtmin, rtmax, into, sn, beta_cor, beta_snr) %>%
+
+peak_data <- makeXcmsObjFlat(msnexp_filled, revert_rts = FALSE) %>%
+  group_by(filename, feature) %>%
   arrange(desc(into)) %>%
-  group_by(feature, filename) %>%
   slice(1) %>%
@@
 all_feat_data <- peak_data %>%
+  select(-filepath) %>%
   left_join(parametadata) %>%
@@
 output_csv_write(all_feat_data)
-# all_feat_data %>%
-#   # filter(med_cor>0.75) %>%
-#   mutate(n_group=cut(n, breaks = c(0, 2, 10, 40, 60, 62, 100, 130, Inf))) %>%
-#   ggplot(aes(label=feature)) +
-#   geom_point(aes(x=med_cor, y=med_snr, color=n_group))
 
@@
 
+```
+
+```{r}
+# best_feat_data <- all_feat_data %>%
+#   filter(med_cor>0.75 | feature%in%rescue_feats) %>%
+#   select(-med_cor, -med_snr)
+# output_csv_write(best_feat_data)
+
+library(squallms)
+pk_data <- makeXcmsObjFlat(msnexp_filled, revert_rts = FALSE) %>%
+  filter(str_detect(filename, "Poo"))
+msdata <- grabMSdata(unique(pk_data$filepath))
+feat_metrics <- extractChromMetrics(pk_data, recalc_betas = TRUE, verbosity = 2, 
+                                    ms1_data = msdata$MS1)
+class_labels <- labelFeatsLasso(pk_data, verbosity=1, ms1_data = msdata$MS1)
+good_cmpds <- logModelFeatQuality(feat_metrics, class_labels, likelihood_threshold=0.8)
+cmpds_keep <- c(good_cmpds, rescue_feats)
+
 best_feat_data <- all_feat_data %>%
-  filter(med_cor>0.75 | feature%in%rescue_feats) %>%
-  select(-med_cor, -med_snr)
+  filter(feature%in%cmpds_keep)
 output_csv_write(best_feat_data)
+```
+
+```{r}
+manual_labels <- labelFeatsManual(peak_data, existing_labels = class_labels, 
+                                  ms1_data = msdata$MS1, selection = "Labeled")
+table(manual_labels, class_labels)
+
 ```
@@
   ggplot() +
-  geom_point(aes(x=rtmed, y=mzmed, color=log10(n))) +
+  geom_point(aes(x=rtmed, y=mzmed, color=n)) +
   scale_color_viridis_c()
@@
   ggplot(aes(label=feature)) +
-  geom_point(aes(x=mz, y=mzmed-mz)) +
+  geom_point(aes(x=mz, y=mzmed)) +
   geom_abline(slope = 1, intercept = 0)

### 2025-05-16 untargeted/old/paruntargeted_sandbox.Rmd
Moving new untargeted analysis over from ILC

--- 2025-05-05 untargeted/paruntargeted_sandbox.Rmd
@@
   pol_i_areas <- map(c("Amm", "Nit", "GMP", "Arg"), function(amendment_i){
-    file_path <- paste0("G:/My Drive/PARAGON/untargeted/", amendment_i, "_", 
+    file_path <- paste0("C:/Users/willi/Documents/PARAGON/untargeted/", amendment_i, "_", 
                         pol_i, "_noIS_output/iso_areas.csv")
@@
 ```
+
+# Fresh start
+
+```{r}
+msnexp_filled <- readRDS("untargeted/Amm_pos_noIS_output/msnexp_filled.rds")
+summarized_feats <- msnexp_filled %>%
+  makeXcmsObjFlat(revert_rts = FALSE) %>%
+  summarise(feat_mzmed=unique(feat_mzmed), feat_rtmed=unique(feat_rtmed),
+            feat_into=mean(into, na.rm=TRUE), mean_cor=mean(beta_cor, na.rm=TRUE),
+            mean_snr=mean(beta_snr, na.rm=TRUE), n=n(), .by=feature)
+summarized_feats %>%
+  ggplot() +
+  geom_point(aes(x=feat_rtmed, y=feat_mzmed, color=log10(feat_into)))
+summarized_feats %>%
+  ggplot() +
+  geom_point(aes(x=mean_cor, y=mean_snr, color=log10(feat_into)))
+summarized_feats %>%
+  mutate(plot_col=feat_mzmed%between%pmppm(118.0865, 10)) %>%
+  ggplot() +
+  geom_point(aes(x=mean_cor, y=mean_snr, color=plot_col))
+```
+
+```{r}
+library(squallms)
+ms_peak_data <- msnexp_filled %>%
+  makeXcmsObjFlat(revert_rts = FALSE) %>%
+  filter(str_detect(filename, "Poo")) %>%
+  mutate(filepath=str_replace(filepath, "G:\\\\My Drive\\\\", "C:\\\\Users\\\\willi\\\\Documents\\\\"))
+labeled_feats <- squallms::labelFeatsLasso(ms_peak_data)
+
+good_feats <- msnexp_filled %>%
+  makeXcmsObjFlat(revert_rts = FALSE) %>%
+  summarise(med_cor=median(beta_cor, na.rm = TRUE), med_snr=median(beta_snr, na.rm = TRUE), .by=feature) %>%
+  squallms::logModelFeatQuality(feature_labels = labeled_feats, likelihood_threshold = 0.5)
+```
+
+```{r}
+msnexp_filled %>%
+  makeXcmsObjFlat(revert_rts = FALSE) %>%
+  summarise(feat_mzmed=unique(feat_mzmed), feat_rtmed=unique(feat_rtmed),
+            feat_into=mean(into, na.rm=TRUE), mean_cor=mean(beta_cor, na.rm=TRUE),
+            mean_snr=mean(beta_snr, na.rm=TRUE), n=n(), .by=feature) %>%
+  filter(feature%in%good_feats) %>%
+  ggplot() +
+  # geom_point(aes(x=feat_rtmed, y=feat_mzmed, color=log10(feat_into)))
+  geom_point(aes(x=mean_cor, y=mean_snr, color=log10(feat_into)))
+```

### 2025-05-16 untargeted/xcms_multicore.R

@@ -1,0 +1,338 @@
+
+library(tidyverse)
+# ms_files <- c(
+#   "Z:/1_QEdata/2021/2021_Will_PARAGON/210813_Will_NIT-Expt_PARAGON_HILIC/positive",
+#   "Z:/1_QEdata/2021/2021_Will_PARAGON/210916_Will_Amm-Expt_PARAGON_HILIC/positive",
+#   "Z:/1_QEdata/2022/220715_Will_GMP-Expt_PARAGON_HILIC/positive",
+#   "Z:/1_QEdata/2022/220729_Will_Arg-Expt_PARAGON_HILIC/positive"
+# ) %>%
+#   lapply(list.files, full.names=TRUE) %>%
+#   unlist() %>%
+#   str_subset("noIS") %>%
+#   str_subset("-neg", negate = TRUE) %>%
+#   # str_subset("Poo.*Full") %>%
+#   str_subset("DDA", negate = TRUE)
+# file.copy(ms_files, paste0("mzMLs/"))
+ms_files <- list.files("mzMLs", full.names = TRUE)
+
+
+# library(RaMS)
+# msdata <- grabMSdata(ms_files %>% str_subset("Poo.*Full"))
+# msdata$TIC %>%
+#   mutate(amendment=str_extract(filename, "Amm|Arg|GMP|Nit")) %>%
+#   qplotMS1data(color_col="filename") +
+#   facet_wrap(~amendment, ncol=1)
+# msdata$MS1[mz%between%pmppm(118.0862, 10)] %>%
+#   slice_max(int, by=c(filename, rt)) %>%
+#   mutate(amendment=str_extract(filename, "Amm|Arg|GMP|Nit")) %>%
+#   qplotMS1data(color_col = "amendment")
+
+
+library(xcms)
+register(BPPARAM = SnowParam(progressbar = TRUE, workers = 10, tasks = length(ms_files)))
+msnexp <- readMSData(ms_files, msLevel. = 1, mode = "onDisk")
+
+cwp <- CentWaveParam(
+  ppm=5,
+  peakwidth = c(20, 80),
+  prefilter=c(5, 1e4),
+  snthresh=5,
+  verboseColumns=TRUE,
+  extendLengthMSW=TRUE,
+  integrate=2
+)
+msnexp_withpeaks <- findChromPeaks(msnexp, cwp)
+saveRDS(msnexp_withpeaks, "msnexp_withpeaks.rds")
+
+
+
[…]

### 2025-05-16 manuscript/manuscript.Rmd
Moving analysis into manuscript from xcms script

@@
 
-Upcoming.
+```{r load clean_combined and make feat_shape_df}
+cleaned_combined_peaks <- readRDS("untargeted/cleaned_combined_peaks.rds")
+lmtest_output <- cleaned_combined_peaks %>%
+  arrange(feature) %>%
+  select(feature, filename, into) %>%
+  left_join(parametadata) %>%
+  filter(samp_type=="Smp") %>%
+  mutate(timepoint_num=ifelse(timepoint_num==0, 0.3, timepoint_num)) %>%
+  nest(data=-feature) %>%
+  filter(sapply(data, nrow)>2) %>%
+  mutate(lm_test=map(data, ~broom::tidy(lm(log10(.x$into)~log10(.x$timepoint_num))), .progress = TRUE)) %>%
+  unnest(lm_test) %>%
+  filter(term=="log10(.x$timepoint_num)") %>%
+  mutate(p_adj=p.adjust(p.value, method = "fdr"))
+
+feat_shape_df <- cleaned_combined_peaks %>%
+  summarise(mzmed=unique(mzmed),
+            rtmed=unique(rtmed),
+            med_iso_cor=median(iso_cor), 
+            med_snr=median(beta_snr),
+            med_cor=median(beta_cor),
+            .by=feature) %>%
+  left_join(lmtest_output %>% select(feature, estimate, p_adj)) %>%
+  mutate(peak_qual=ifelse(med_snr>-50*med_cor+50, "Good", "Bad")) %>%
+  mutate(is_iso=ifelse(med_iso_cor>0.8, "Is iso", "Not iso")) %>%
+  mutate(time_trend=ifelse(p_adj<0.001& estimate>0, "Increases", "No change"))
+feat_shape_df %>%
+  filter(!is.na(time_trend)) %>%
+  count(peak_qual, is_iso, time_trend)
+```
+
+```{r}
+best_labeled <- feat_shape_df %>%
+  filter(peak_qual=="Good" & is_iso=="Is iso" & time_trend=="Increases") %>%
+  arrange(feature) %>%
+  select(feature, mzmed, rtmed, med_iso_cor, med_snr, med_cor, p_adj)
+psb_stans <- all_stans %>%
+  filter(polarity=="pos") %>%
+  summarise(compound_name=paste(compound_name, collapse = "/"), .by = mz) %>%
+  rowwise() %>%
+  mutate(N15_mz_lower=pmppm(mz+0.997035, 10)[1]) %>%
+  mutate(N15_mz_upper=pmppm(mz+0.997035, 10)[2]) %>%
+  select(compound_name, N15_mz_lower, N15_mz_upper)
+best_labeled %>%
+  left_join(chosen_stans, join_by(between(x$mzmed, y$N15_mz_lower, y$N15_mz_upper))) %>%
+  select(-N15_mz_lower, -N15_mz_upper) %>%
+  mutate(compound_name=ifelse(duplicated(compound_name)&!is.na(compound_name), "Isocytosine", compound_name)) %>%
+  mutate(compound_name=ifelse((mzmed-lag(mzmed))%between%c(0.996, 0.998), "iso_prev?", compound_name)) %>%
+  mutate(rtmed=rtmed/60) %>%
+  select(feature, mzmed, rtmed, compound_name) %>%
+  arrange(mzmed) %>%
+  write_csv("untargeted/best_labeled_anno_2.csv")
+```
 

### 2025-05-16 untargeted/xcms_multicore.R

@@ -13,7 +13,7 @@ library(tidyverse)
 #   # str_subset("Poo.*Full") %>%
 #   str_subset("DDA", negate = TRUE)
 # file.copy(ms_files, paste0("mzMLs/"))
-ms_files <- list.files("mzMLs", full.names = TRUE)
+ms_files <- list.files("mzMLs/pos", full.names = TRUE, pattern="noIS")
 
 
 # library(RaMS)
@@ -42,11 +42,11 @@ cwp <- CentWaveParam(
   integrate=2
 )
 msnexp_withpeaks <- findChromPeaks(msnexp, cwp)
-saveRDS(msnexp_withpeaks, "msnexp_withpeaks.rds")
+saveRDS(msnexp_withpeaks, "untargeted/msnexp_withpeaks.rds")
 
 
 
-msnexp_withpeaks <- readRDS("msnexp_withpeaks.rds")
+msnexp_withpeaks <- readRDS("untargeted/msnexp_withpeaks.rds")
 register(BPPARAM = SerialParam(progressbar = TRUE))
 pdp <- PeakDensityParam(
   sampleGroups = str_remove(ms_files, "-[A-C](?=-)|_\\d(?=\\.mzML)"), 
@@ -58,12 +58,12 @@ msnexp_grouped <- groupChromPeaks(msnexp_withpeaks, pdp)
 
 register(BPPARAM = SnowParam(progressbar = TRUE, workers = 5, tasks = length(ms_files)))
 msnexp_filled <- fillChromPeaks(msnexp_grouped, FillChromPeaksParam())
-saveRDS(msnexp_filled, "msnexp_filled.rds")
+saveRDS(msnexp_filled, "untargeted/msnexp_filled.rds")
 
 
 
 
-msnexp_filled <- readRDS("msnexp_filled.rds")
+msnexp_filled <- readRDS("untargeted/msnexp_filled.rds")
 peak_df <- msnexp_filled %>%
   chromPeaks() %>%
   as.data.frame() %>%
@@ -89,11 +89,11 @@ oaoo_df <- feature_df %>%
 oaoo_df %>%
   count(feature) %>%
   distinct(n)
-saveRDS(oaoo_df, "one_and_only_one_peak_per_feature_file.rds")
+saveRDS(oaoo_df, "untargeted/one_and_only_one_peak_per_feature_file.rds")
 
 
 
-# oaoo_df <- readRDS("one_and_only_one_peak_per_feature_file.rds")
[…]

### 2025-05-16 manuscript/manuscript.Rmd
xcms_multicore.R now clean of all analysis

@@
 
+```{r source multicore}
+source("untargeted/xcms_multicore.R")
+```
+
 ```{r load clean_combined and make feat_shape_df}
@@
 
-```{r}
+```{r get counts per amendment instead of as a whole}
+# Repeat analysis across amendments to get additional numbers
+lmtest_xamend <- cleaned_combined_peaks %>%
+  arrange(feature) %>%
+  select(feature, filename, into) %>%
+  mutate(samp_type=str_extract(filename, "Smp|Blk|Poo")) %>%
+  filter(samp_type=="Smp") %>%
+  mutate(timepoint=as.numeric(str_extract(filename, "(?<=T)\\d+"))) %>%
+  mutate(timepoint=ifelse(timepoint==0, 0.3, timepoint)) %>%
+  mutate(amendment=str_extract(filename, "Amm|GMP|Arg|Nit")) %>%
+  mutate(depth=str_extract(filename, "Surf|Deep")) %>%
+  nest(data=-c(feature, amendment)) %>%
+  filter(sapply(data, nrow)>2) %>%
+  mutate(lm_test=map(data, ~broom::tidy(lm(log10(.x$into)~log10(.x$timepoint))), .progress = TRUE)) %>%
+  unnest(lm_test) %>%
+  filter(term=="log10(.x$timepoint)") %>%
+  mutate(p_adj=p.adjust(p.value, method = "fdr"))
+feat_shape_df_xamend <- cleaned_combined_peaks %>%
+  mutate(amendment=str_extract(filename, "Amm|GMP|Arg|Nit")) %>%
+  summarise(mzmed=unique(mzmed),
+            rtmed=unique(rtmed),
+            med_iso_cor=median(iso_cor), 
+            med_snr=median(beta_snr),
+            med_cor=median(beta_cor),
+            .by=c(feature, amendment)) %>%
+  left_join(lmtest_xamend %>% select(feature, amendment, estimate, p_adj)) %>%
+  mutate(peak_qual=ifelse(med_snr>-50*med_cor+50, "Good", "Bad")) %>%
+  mutate(is_iso=ifelse(med_iso_cor>0.8, "Is iso", "Not iso")) %>%
+  mutate(time_trend=ifelse(p_adj<0.001& estimate>0, "Increases", "No change"))
+feat_shape_df_xamend %>%
+  bind_rows(iso_df %>% mutate(amendment="Overall")) %>%
+  filter(!is.na(time_trend)) %>%
+  count(peak_qual, is_iso, time_trend, amendment) %>%
+  pivot_wider(names_from = amendment, values_from=n, values_fill = 0) %>%
+  print()
+feat_shape_df_xamend %>%
+  bind_rows(feat_shape_df %>% mutate(amendment="Overall")) %>%
+  filter(!is.na(time_trend)) %>%
+  ggplot() +
+  geom_bar(aes(x=amendment)) +
+  facet_wrap(~peak_qual+is_iso+time_trend, scales="free_y", ncol=4)
+```
+
+```{r write out the best labels for manual annotation}
 best_labeled <- feat_shape_df %>%
@@
 best_labeled %>%
-  left_join(chosen_stans, join_by(between(x$mzmed, y$N15_mz_lower, y$N15_mz_upper))) %>%
+  left_join(psb_stans, join_by(between(x$mzmed, y$N15_mz_lower, y$N15_mz_upper))) %>%
   select(-N15_mz_lower, -N15_mz_upper) %>%
@@
   select(feature, mzmed, rtmed, compound_name) %>%
-  arrange(mzmed) %>%
-  write_csv("untargeted/best_labeled_anno_2.csv")
+  write_csv("untargeted/best_labeled_anno.csv")
+```
+
+```{r create a heatmap and cluster all high-quality peaks}
+labeled_fracs <- cleaned_combined_peaks %>%
+  summarise(med_iso_cor=median(iso_cor), 
+            med_snr=median(beta_snr),
+            med_cor=median(beta_cor),
+            .by=c(feature)) %>%
+  filter(med_snr>-50*med_cor+50) %>%
+  filter(med_iso_cor>0.8) %>%
+  select(feature) %>%
+  left_join(cleaned_combined_peaks %>% select(feature, filename, lab_area, unlab_area), multiple = "all") %>%
+  mutate(labeled_frac=lab_area/(lab_area+unlab_area)) %>%
+  select(-lab_area, -unlab_area)
+
+cmpd_order <- labeled_fracs %>%
+  left_join(parametadata) %>%
+  summarise(mean_label=mean(labeled_frac, na.rm=TRUE), .by=feature) %>%
+  arrange(desc(mean_label)) %>%
+  pull(feature)
+
+colorpal <- viridis::viridis(10, alpha=seq(0, 1, length.out=10), direction = -1)
+untarg_heatmap <- labeled_fracs %>%
+  left_join(parametadata) %>%
+  filter(samp_type=="Smp") %>%
+  mutate(feature=factor(feature, levels=cmpd_order)) %>%
+  ggplot() +
+  geom_raster(aes(x=timepoint, y=tripl, fill=labeled_frac)) +
+  scale_fill_gradientn(colours = colorpal, limits = c(0, 1)) +
+  facet_wrap(feature~amendment+depth+startime) +
+  scale_y_discrete(expand=expansion()) +
+  scale_x_discrete(guide = guide_axis(n.dodge = 5), expand=expansion()) +
+  facet_nested(feature~amendment+depth+startime, scales = "free_x", switch="y", strip=strip_nested(size = "variable")) +
+  labs(x="Timepoint", y="Metabolite", fill="Fraction\nlabeled") +
+  theme_bw() +
+  theme(panel.spacing.x=unit(0.2, "lines"), 
+        panel.spacing.y = unit(0, "lines"),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        # axis.ticks.length.x=unit(1:5, "lines"),
+        # axis.text.x = element_markdown(vjust = -5:-1),
+        strip.text.y.left = element_text(angle=0, hjust = 1),
+        # strip.background.y = element_rect(fill=NA, color=NA),
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        legend.position = "top", 
+        legend.key.width = unit(1, "null")) +
+  guides(
+    fill=guide_colorbar(theme=theme(
+      legend.frame = element_rect(color = "black"),
+      legend.ticks = element_line(color="black"),
+      legend.text = ggtext::element_markdown(hjust=c(0, 0.5, 0.5, 0.5, 1)))
+    )
+  )
+ggsave("untargeted/untarg_heatmap.pdf", width = 10, height=36, units = "in")
 ```

### 2025-05-16 untargeted/xcms_multicore.R

@@ -186,129 +186,3 @@ cleaned_combined_peaks <- peak_quality_scores %>%
   left_join(peak_ids) %>%
   select(feature, mzmed, rtmed, filename, peak_id, mz, rt, into, everything())
 saveRDS(cleaned_combined_peaks, "untargeted/cleaned_combined_peaks.rds")
-
-
-
-
-
-
-
-iso_df <- cleaned_combined_peaks %>%
-  summarise(mzmed=unique(mzmed),
-            rtmed=unique(rtmed),
-            med_iso_cor=median(iso_cor), 
-            med_snr=median(beta_snr),
-            med_cor=median(beta_cor),
-            .by=feature) %>%
-  left_join(lmtest_output %>% select(feature, estimate, p_adj)) %>%
-  mutate(peak_qual=ifelse(med_snr>-50*med_cor+50, "Good", "Bad")) %>%
-  mutate(is_iso=ifelse(med_iso_cor>0.8, "Is iso", "Not iso")) %>%
-  mutate(time_trend=ifelse(p_adj<0.001& estimate>0, "Increases", "No change"))
-iso_df %>%
-  filter(!is.na(time_trend)) %>%
-  count(peak_qual, is_iso, time_trend)
-best_labeled <- iso_df %>%
-  filter(peak_qual=="Good" & is_iso=="Is iso" & time_trend=="Increases") %>%
-  arrange(feature) %>%
-  select(feature, mzmed, rtmed, med_iso_cor, med_snr, med_cor, p_adj)
-
-
-# Annotate best ones quickly with standards list, to be finessed later manually
-chosen_stans <- read_csv("https://github.com/IngallsLabUW/Ingalls_Standards/raw/refs/heads/master/Ingalls_Lab_Standards.csv") %>%
-  filter(z>0) %>%
-  filter(Column=="HILIC") %>%
-  summarise(compound_name=paste(Compound_Name, collapse = "/"), .by = mz) %>%
-  rowwise() %>%
-  mutate(N15_mz_lower=pmppm(mz+0.997035, 10)[1]) %>%
-  mutate(N15_mz_upper=pmppm(mz+0.997035, 10)[2]) %>%
-  select(compound_name, N15_mz_lower, N15_mz_upper)
-best_labeled %>%
-  left_join(chosen_stans, join_by(between(x$mzmed, y$N15_mz_lower, y$N15_mz_upper))) %>%
[…]

### 2025-05-17 manuscript/manuscript.Rmd
Fixing typo in heatmap plot

@@
   )
-ggsave("xamend_thaa_gp.png",plot = xamend_metab_gp, device = "png", path = "manuscript/figures",
+ggsave("xamend_metab_gp.png",plot = xamend_metab_gp, device = "png", path = "manuscript/figures",
        width = 6.5, height = 5.5, units = "in", dpi = 300)
@@
 
-![](figures/xamend_thaa_gp.png)
+![](figures/xamend_metab_gp.png)
 

### 2025-05-17 metadata/parametadata_control.Rmd
Dropping wonky nitrate data and correcting Arg timepoint

@@
   mutate(timepoint=str_extract(shortname, "T\\d+")) %>%
+  mutate(timepoint=case_when(
+    amendment=="Arg" & depth=="Surf" & startime=="Eve" & timepoint=="T3"~"T18"
+  )) %>%
   mutate(timepoint_num=as.numeric(str_remove(timepoint, "T"))) %>%
   mutate(tripl=str_extract(shortname, "[A-C]$")) %>%
-  mutate(untripl=str_remove(shortname, "-[A-C]$"))
+  mutate(untripl=str_remove(shortname, "-[A-C]$")) %>%
+  mutate(samp_type=case_when(
+    amendment=="Nit" & depth=="Deep" & startime=="Eve" & timepoint=="T10"~"Bad"
+  ))
 

### 2025-05-17 metadata/parametadata_control.Rmd
Forgot how case_when works :upsidedownface:

--- 2025-05-09 metadata/parametadata_control.Rmd
@@
   mutate(timepoint=str_extract(shortname, "T\\d+")) %>%
+  mutate(timepoint=case_when(
+    amendment=="Arg" & depth=="Surf" & startime=="Eve" & timepoint=="T3"~"T18",
+    TRUE~timepoint
+  )) %>%
   mutate(timepoint_num=as.numeric(str_remove(timepoint, "T"))) %>%
   mutate(tripl=str_extract(shortname, "[A-C]$")) %>%
-  mutate(untripl=str_remove(shortname, "-[A-C]$"))
+  mutate(untripl=str_remove(shortname, "-[A-C]$")) %>%
+  mutate(samp_type=case_when(
+    amendment=="Nit" & depth=="Deep" & startime=="Eve" & timepoint=="T10"~"Bad",
+    TRUE~samp_type
+  ))
 

### 2025-05-17 manuscript/manuscript.Rmd
Adding kmeans clustering code and figures

@@
   bind_rows(iso_df %>% mutate(amendment="Overall")) %>%
+  # filter(!rtmed%between%(c(14, 14.5)*60)) %>%
+  # filter(!rtmed<(2*60)) %>%
   filter(!is.na(time_trend)) %>%
@@
 
-```{r create a heatmap and cluster all high-quality peaks}
-labeled_fracs <- cleaned_combined_peaks %>%
+```{r heatmap of all high-quality feats}
+high_qual_feats <- cleaned_combined_peaks %>%
+  filter(rtmed>120 & !rtmed%between%c(14*60, 14.5*60)) %>%
+  left_join(parametadata %>% distinct(filename, samp_type)) %>%
+  filter(samp_type=="Smp") %>%
   summarise(med_iso_cor=median(iso_cor), 
@@
   filter(med_snr>-50*med_cor+50) %>%
-  filter(med_iso_cor>0.8) %>%
-  select(feature) %>%
-  left_join(cleaned_combined_peaks %>% select(feature, filename, lab_area, unlab_area), multiple = "all") %>%
+  filter(med_iso_cor>0.8)
+untarg_lab_fracs <- cleaned_combined_peaks %>%
+  filter(feature%in%high_qual_feats$feature) %>%
+  select(feature, filename, lab_area, unlab_area) %>%
+  left_join(parametadata %>% distinct(filename, samp_type)) %>%
+  filter(samp_type=="Smp") %>%
+  slice_max(unlab_area, by=c(feature, filename), with_ties = FALSE) %>%
   mutate(labeled_frac=lab_area/(lab_area+unlab_area)) %>%
-  select(-lab_area, -unlab_area)
+  left_join(parametadata %>% distinct(filename, shortname)) %>%
+  select(feature, shortname, labeled_frac)
 
-cmpd_order <- labeled_fracs %>%
-  left_join(parametadata) %>%
+cmpd_order <- untarg_lab_fracs %>%
   summarise(mean_label=mean(labeled_frac, na.rm=TRUE), .by=feature) %>%
@@
 colorpal <- viridis::viridis(10, alpha=seq(0, 1, length.out=10), direction = -1)
-untarg_heatmap <- labeled_fracs %>%
-  left_join(parametadata) %>%
-  filter(samp_type=="Smp") %>%
+untarg_heatmap <- untarg_lab_fracs %>%
+  complete(feature, shortname, fill=list(labeled_frac=0)) %>%
   mutate(feature=factor(feature, levels=cmpd_order)) %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
   ggplot() +
@@
   scale_fill_gradientn(colours = colorpal, limits = c(0, 1)) +
-  facet_wrap(feature~amendment+depth+startime) +
   scale_y_discrete(expand=expansion()) +
@@
 ggsave("untargeted/untarg_heatmap.pdf", width = 10, height=36, units = "in")
+```
+
+```{r run kmeans}
+untarg_lab_mat <- untarg_lab_fracs %>%
+  bind_rows(labeled_fracs %>% rename(feature=compound_name)) %>%
+  pivot_wider(names_from = "feature", values_from = "labeled_frac", values_fill=0) %>%
+  column_to_rownames("shortname") %>%
+  data.matrix()
+
+set.seed(124)
+
+# sapply(1:20, function(n_clust){
+#   replicate(10, {
+#     kmeans_out <- kmeans(t(untarg_lab_mat), centers = n_clust)
+#     kmeans_out$betweenss / kmeans_out$totss * 100
+#   })
+# }) %>%
+#   boxplot()
+
+n_clust <- 10
+untarg_kmeans <- untarg_lab_mat %>%
+  t() %>%
+  kmeans(centers = n_clust)
+kmeans_df <- data.frame(cluster=1:n_clust, n=untarg_kmeans$size) %>%
+  arrange(desc(n)) %>%
+  mutate(cluster_name=paste0("Cluster ", row_number(), " n=", n)) %>%
+  mutate(cluster_name=fct_inorder(cluster_name)) %>%
+  left_join(data.frame(cluster=untarg_kmeans$cluster, feature=names(untarg_kmeans$cluster))) %>%
+  select(cluster_name, feature)
+
+# Plot clusters as heatmap
+untarg_lab_fracs %>%
+  left_join(kmeans_df) %>%
+  summarise(labeled_frac=mean(labeled_frac), .by = c(cluster_name, shortname)) %>%
+  left_join(parametadata) %>%
+  ggplot() +
+  geom_raster(aes(x=timepoint, y=tripl, fill=labeled_frac)) +
+  scale_fill_gradientn(colours = colorpal, limits = c(0, 1)) +
+  facet_wrap(cluster_name~amendment+depth+startime) +
+  scale_y_discrete(expand=expansion()) +
+  scale_x_discrete(guide = guide_axis(n.dodge = 5), expand=expansion()) +
+  facet_nested(cluster_name~amendment+depth+startime, scales = "free_x", switch="y", 
+               strip=strip_nested(size = "variable")) +
+  labs(x="Timepoint", y=NULL, fill="Fraction\nlabeled") +
+  theme_bw() +
+  theme(panel.spacing.x=unit(0.2, "lines"), 
+        panel.spacing.y = unit(0, "lines"),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        # axis.ticks.length.x=unit(1:5, "lines"),
+        # axis.text.x = element_markdown(vjust = -5:-1),
+        strip.text.y.left = element_text(angle=0, hjust = 1),
+        # strip.background.y = element_rect(fill=NA, color=NA),
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        legend.position = "top", 
+        legend.key.width = unit(1, "null")) +
+  guides(
+    fill=guide_colorbar(theme=theme(
+      legend.frame = element_rect(color = "black"),
+      legend.ticks = element_line(color="black"),
+      legend.text = ggtext::element_markdown(hjust=c(0, 0.5, 0.5, 0.5, 1)))
+    )
+  )
+
+# Export as CSV
+kmeans_df %>%
+  filter(!startsWith(feature, "FT")) %>%
+  arrange(cluster_name, feature) %>%
+  split(.$cluster_name) %>%
+  map(function(df_i){
+    df_i %>% add_row(cluster_name="", feature="")
+  }) %>%
+  bind_rows() %>%
+  write_csv("untargeted/cluster_assignments.csv")
 ```

### 2025-05-17 manuscript/manuscript.Rmd
Rerunning manuscript figures with fixed metadata

@@
   filter(polarity=="pos") %>%
-  filter(samp_type=="Smp") %>%
+  filter(samp_type=="Smp" | samp_type=="Bad") %>%
   mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
@@
   arrange(amendment, depth, timepoint_num, startime) %>%
-  mutate(timepoint=fct_inorder(timepoint))
+  mutate(timepoint=factor(timepoint, levels=c("T0", "T1", "T3", "T10", "T18", "T26", "T73")))
 
@@
 xamend_metab_gp <- labeled_fracs %>%
-  left_join(parametadata %>% distinct(shortname, timepoint, timepoint_num, depth, startime, amendment, tripl)) %>%
+  left_join(parametadata %>% distinct(shortname, samp_type, timepoint, timepoint_num, depth, startime, amendment, tripl)) %>%
+  filter(samp_type!="Bad") %>%
   filter(depth=="Surf") %>%
@@
   left_join(parametadata %>% distinct(shortname, timepoint_num, depth, startime, amendment, tripl)) %>%
+  filter(timepoint_num!=18) %>%
   filter(compound_name=="Glutamic acid" & amendment=="Amm" |
@@
 surfdeep_all_cmpd_gp <- labeled_fracs %>%
-  left_join(parametadata %>% distinct(shortname, timepoint_num, depth, startime, amendment, tripl)) %>%
+  left_join(parametadata %>% distinct(shortname, timepoint, timepoint_num, depth, startime, amendment, tripl)) %>%
+  filter(timepoint_num!=18) %>%
   filter(!compound_name%in%maa_names) %>%
@@
   left_join(parametadata %>% distinct(shortname, timepoint_num, depth, startime, amendment, tripl)) %>%
+  filter(timepoint_num!=18) %>%
   filter(!compound_name%in%maa_names) %>%

### 2025-05-19 manuscript/manuscript.Rmd
Adding solvent artifact separation

@@
   mutate(peak_qual=ifelse(med_snr>-50*med_cor+50, "Good", "Bad")) %>%
+  mutate(peak_qual=ifelse(rtmed<2*60 | between(rtmed, 14*60, 14.5*60), "Solvent artifact", peak_qual)) %>%
   mutate(is_iso=ifelse(med_iso_cor>0.8, "Is iso", "Not iso")) %>%

### 2025-05-19 manuscript/manuscript.Rmd
Adding additional grey separator with new T18 timepoint

@@
   geom_raster(aes(x=timepoint, y=tripl, fill=avg_frac)) +
-  geom_vline(xintercept = c(1.5:4.5), color="grey80") +
+  geom_vline(xintercept = c(1.5:5.5), color="grey80") +
   facet_nested(compound_class~amendment, scales = "free_x", switch="y") +

### 2025-05-20 manuscript/manuscript.Rmd
Shortening fig 1

@@
       legend.ticks = element_line(color="black"),
-      legend.text = ggtext::element_markdown(hjust=c(0, 0, 0.5, 1, 1)))
+      legend.text = ggtext::element_markdown(hjust=c(0, 0, 0.5, 1, 1))),
+      text=element_text(size=12)
     )
@@
 ggsave("xamend_metab_gp.png",plot = xamend_metab_gp, device = "png", path = "manuscript/figures",
-       width = 6.5, height = 5.5, units = "in", dpi = 300)
+       width = 6.5, height = 4, units = "in", dpi = 300)
 ```

### 2025-05-20 manuscript/manuscript.Rmd
First day of writing results section

--- 2025-05-19 manuscript/manuscript.Rmd
@@
 library(minpack.lm)
+library(RaMS)
 options(pillar.sigfig = 7)
 options(pillar.print_min = 100)
+
+wc <- function(
+  comment, 
+  highlight = "", 
+  author = "Will", 
+  time = Sys.time(),
+  id = "0"
+) {
+  if (isTRUE(knitr:::pandoc_to() == "docx")) {
+    return(
+      sprintf(
+        '[%s]{.comment-start id="%s" author="%s" date="%s"} %s []{.comment-end id="%s"}',
+        comment,
+        id,
+        author,
+        time,
+        highlight,
+        id
+      )
+    )
+  } else {
+    return(
+      sprintf(
+        "*%s* **[Comment id %s by %s at time %s: %s]**", 
+        highlight,
+        id,
+        author, 
+        time, 
+        comment
+      )
+    )
+  }
+}
 ```
@@
 
+env_concs <- read_csv("targeted/final_env_concs.csv") %>%
+  select(-conc_um_in_vial) %>%
+  left_join(parametadata %>% distinct(filename, shortname)) %>%
+  select(-filename)
+nM_labeled <- samp_areas %>%
+  left_join(env_concs) %>%
+  mutate(labeled_frac=area/sum(area), .by = c("shortname", "compound_name")) %>%
+  mutate(rel_label=labeled_frac/labeled_frac[str_detect(iso_name, "13C0, 15N0")], .by = c("shortname", "compound_name")) %>%
+  mutate(nM_labeled=conc_nm_in_env*rel_label) %>%
+  select(filename, compound_name, iso_name, nM_labeled)
+
+cleaned_combined_peaks <- readRDS("untargeted/cleaned_combined_peaks.rds")
+
 thaa_quant <- read_csv("thaas/quant_data.csv") %>%
@@
 
-```{r}
+```{r gln numbers}
+labeled_fracs %>%
+  filter(compound_name=="L-Glutamine") %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(timepoint%in%c("T0", "T26")) %>%
+  filter(depth=="Surf") %>%
+  filter(amendment=="Amm") %>%
+  summarise(frac=mean(labeled_frac), .by=c(compound_name, timepoint))
+peak_areas %>%
+  filter(compound_name=="L-Glutamine") %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(depth=="Surf") %>%
+  filter(amendment=="Amm") %>%
+  ggplot(aes(x=tripl, y=area, fill=iso_name)) +
+  geom_col(color="black") +
+  facet_grid(startime~timepoint)
+nM_labeled %>%
+  filter(compound_name%in%c("L-Glutamine")) %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(depth=="Surf") %>%
+  filter(amendment=="Amm") %>%
+  summarise(conc=mean(nM_labeled), .by=c(compound_name, timepoint)) %>%
+  arrange(timepoint)
+```
+
+We observed significant differences in ^15^N labeling patterns between the different treatment incubations performed with ALOHA surface seawater (Figure 1).
+Ammonia incubations resulted in very rapid labeling, with glutamine more than 75% labeled even at the T0 timepoint, corresponding to the amount of time required to spike and immediately filter the 2 liters of seawater (~15 minutes).
+The dominant labeling pattern at T0 was a single ^15^N, though all later timepoints showed more doubly-^15^N labeled glutamine than singly-labeled.
+After 24 hours, glutamine was more than 95% labeled and had approximately doubled in pool size (initially ~0.3 nM becoming 0.8 nM).
+
+```{r fig1 grouped metab heatmap}
 xamend_metab_gp <- labeled_fracs %>%
@@
 
-*Figure 1: Fraction of each metabolite or metabolite group containing at least one 15N label during the surface seawater amended various substrates after the number of hours shown on the x-axis.
+*Figure 1: Fraction of each metabolite or metabolite group containing at least one ^15^N label during the surface seawater amendments with various substrates after the number of hours shown on the x-axis.
 Darker colors indicate a larger percentage of the compound signal was isotopically labeled.
@@
 
-  - Ammonia
-    - Really rapid labeling of glutamine/glutamate that aligns nicely with prior work
-      - Aspartate labeled faster than glutamate likely due to small aspartate pool size
-    - Surprisingly rapid labeling of citrulline that isn't reflected in arginine
-      - Pool size problem (arginine >> citrulline)?
-      - Indicates that urea cycle is only half-active?
-    - Putative annotation of theanine based on labeling patterns
-    - Rapid and heavy guanine labeling
-      - Potentially being used as a storage compound
-    - 10-40% of the N in the THAA (protein) pool ends up labeled by T26 (below)
-      - THAAs were 30-50% of total PC (~2 uM, need actual data) and 70-100% PN (~0.35uM)
-  - Nitrate
-    - Very little labeling, even in glutamate/glutamine
-      - Surface community not "expecting" nitrate?
-      - Redox problems with surface energy?
-    - Guanine and MAA labeling exceeds glutamate labeling
-      - Clear sign of a community *subset* using nitrate
-        - Specifically the MAA synthesizers (euks) and those storing N in guanine (diatoms/haptos/dinos)
-    - More extensive labeling (10-50% of the N) in the THAAs
-      - Must be a really high flux through a small AA pool subset, e.g. euks specifically
+```{r gluaspcit numbers}
+labeled_fracs %>%
+  filter(compound_name%in%c("L-Glutamic acid/beta-Glutamic acid", "L-Aspartic acid",
+                            "Citrulline", "L-Arginine", "Theanine?")) %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(timepoint=="T1") %>%
+  filter(depth=="Surf") %>%
+  filter(amendment=="Amm") %>%
+  summarise(frac=mean(labeled_frac), .by=c(compound_name))
+env_concs %>%
+  filter(compound_name%in%c("L-Glutamic acid/beta-Glutamic acid", "L-Aspartic acid",
+                            "Citrulline", "L-Arginine")) %>%
+  left_join(parametadata %>% filter(IS_type=="wIS")) %>%
+  filter(timepoint=="T1") %>%
+  filter(depth=="Surf") %>%
+  filter(amendment=="Amm") %>%
+  summarise(conc=mean(conc_nm_in_env), .by=c(compound_name))
+```
+
+```{r theanine annotation check}
+# Check theanine annotation and numbers
+cleaned_combined_peaks %>% 
+  filter(mzmed%between%pmppm(175.1077, 10)) %>%
+  left_join(parametadata) %>%
+  filter(amendment=="Amm") %>%
+  filter(depth=="Surf") %>%
+  summarise(mzmed=unique(mzmed), rtmed=unique(rtmed)/60, into=mean(into),
+            beta_snr=median(beta_snr), beta_cor=median(beta_cor), 
+            .by=feature)
+arrow::open_dataset("tmzMLs/pqds") %>%
+  filter(amendment=="Amm") %>%
+  filter(samp_type=="Smp") %>%
+  filter(polarity=="pos") %>%
+  filter(mz%between%pmppm(175.1077, 10) | mz%between%pmppm(176.1047, 10)) %>%
+  filter(rt%between%c(7.2, 8.5)) %>%
+  dplyr::collect() %>%
+  mutate(mz=round(mz)) %>%
+  slice_max(int, by=c(filename, mz, rt)) %>%
+  arrange(rt) %>%
+  summarise(area=trapz(rt, int), .by=c(filename, mz)) %>%
+  pivot_wider(names_from=mz, values_from=area) %>%
+  mutate(labeled_frac=`176`/(`175`+`176`)) %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(depth=="Surf") %>%
+  ggplot() +
+  geom_boxplot(aes(x=timepoint, y=labeled_frac, color=startime))
+```
+
+The ^15^N was then quickly propagated to glutamate and aspartate, creating a metabolite pool 40-50% labeled for each of these after a single hour.
+Aspartate was labeled slightly faster than glutamate, likely as a result of its smaller pool size (0.2 nM Asp versus 0.5 nM Glu).
+Also rapidly labeled was citrulline, reaching 20% labeled after a single hour, though arginine was not labeled until much later despite its location immediately downstream of this metabolite.
+Untargeted detection of other molecular features revealed an additional metabolite that was 30% labeled at the one hour timepoint with an *m/z* value of 175.1077 and a retention time of 8.2 minutes which we have putatively identified as theanine (N5-ethyl glutamine) based on *m/z* match and labeling pattern.
+
+```{r betaine and nucleo numbers}
+labeled_fracs %>%
+  filter(compound_name%in%nucleo_names) %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(timepoint=="T26") %>%
+  filter(amendment=="Amm") %>%
+  summarise(frac=mean(labeled_frac), .by=c(compound_name)) %>%
+  mutate(nuc_type=case_when(
+    compound_name%in%c("Cytosine", "Adenine", "Guanine", "Hypoxanthine/Allopurinol")~"Nucleobase",
+    compound_name%in%c("Cytidine", "Adenosine", "Guanosine")~"Nucleoside",
+    compound_name%in%c("Deoxycytidine?", "Deoxyadenosine", "Deoxyguanosine?")~"Deoxy"
+  )) %>%
+  mutate(nuc_shortname=str_extract(tolower(compound_name), tolower("Cy|Ad|Gu|Hy"))) %>%
+  select(nuc_shortname, nuc_type, frac) %>%
+  pivot_wider(names_from=nuc_type, values_from=frac) %>%
+  select(nuc_shortname, Nucleoside, Nucleobase, Deoxy)
+
+labeled_fracs %>%
+  filter(compound_name%in%betaine_names) %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(timepoint=="T26") %>%
+  filter(amendment=="Amm") %>%
+  summarise(frac=mean(labeled_frac), .by=c(compound_name))
+
+peak_areas %>%
+  filter(compound_name%in%nucleo_names) %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(timepoint=="T26") %>%
+  filter(amendment=="Amm") %>%
+  filter(depth=="Surf") %>%
+  mutate(iso_name=str_remove(iso_name, ".*?, ")) %>%
+  filter(iso_name!="13C0, 15N0") %>%
+  ggplot() +
+  geom_col(aes(x=tripl, y=area, fill=iso_name), position="fill", color="black") +
+  facet_wrap(~compound_name)
+```
+
+Other amino acids and nucleobases/nucleosides had trace amounts of labeling after 3 hours and were extensively labeled (>25%) after 24 hours (Figure 1).
+Cytidine (42.2% labeled) and adenosine (59.9%) were more heavily ^15^N labeled than their nucleobase constituents at T26 (22.9% and 42.1%, respectively), corresponding to the expected synthesis pathway.
+Guanine, however, had the inverse trend and at 44.9% labeled was 15% more labeled than guanosine (Supplemental figure XX).
+The deoxyribose forms were in all cases less completely labeled than their ribose counterparts (deoxycytidine = 18.9%, deoxyadenosine = 20.3%, and deoxyguanosine = 17.2%) `r wc("Maybe better as a small table?")` `r wc("Consider trying harder to get thymine/uracil/inosine values?")`.
+In each case, the fully ^15^N labeled nucleobase was the dominant constituent of the isotope envelope (^15^N_3_ for cytosine and ^15^N_5_ for adenine and guanine).
+Other compounds, including betaines (here restricted to quaternary amines with a carboxylic acid group) and choline derivatives showed very little labeling at all, with only homarine (15.0% labeled) and O-acetylcarnitine (9.9% labeled) exceeding 5% labeling after 24 hours`r wc("Expand betaine discussion and turn it into a separate paragraph?
+Not sure what more to say...")`.
+
+Amendments with ^15^NO_3_ showed a significantly different labeling pattern with very few metabolites incorporating the ^15^N label at the surface even after 24 hours.
+Glutamine and glutamate, which were labeled within 20 minutes from ammonia and are the main way in which inorganic nitrogen becomes organic, here saw only trace amounts of production from the added nitrate even after 24 hours.
+The most heavily labeled compound was guanine, which saw trace amounts of labeling at the T10 timepoint and reached ~40% labeling at T26.
+This labeling was not reflected in the other nucleobases.
+The only group of compounds that did see compelling degrees of label were the mycosporine-like amino acids (MAAs), a group of compounds produced under UV exposure (see below).
+These compounds, like guanine, were labeled in excess of the bulk glutamate/amino acid pool from which their nitrogen is sourced.
+This is only possible if the nitrate is assimilated by a subset of the community that is able to reduce it and prevents it from entering the environment.
+Metabolites specific to those organisms can then be made from the small pool of heavily-labeled organic material without affecting the fraction of glutamate or glutamine labeled in bulk.
+
+```{r arg numbers}
+labeled_fracs %>%
+  filter(compound_name=="L-Arginine") %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(depth=="Surf") %>%
+  filter(amendment=="Arg") %>%
+  summarise(frac=mean(labeled_frac), .by=c(compound_name, timepoint)) %>%
+  arrange(timepoint)
+```
+
+The organic nitrogen amendments also showed labeling patterns distinct from the ammonia incubations, though a large fraction of the nitrogen was clearly remineralized into ammonia/glutamate and used to label a similar suite of compounds (Figure 1).
+Fully C and N labeled (^13^C_6_, ^15^N_4_) arginine was visibly taken up at the T0 timepoint (within 20 minutes) and 80-90% of the pool was fully labeled after 12 hours.
+Surprisingly, the fraction of arginine labeled decreased from this point, with the T26 and T73 timepoints showing
+
   - Arginine
@@
     - Maybe less storage in guanine specifically, labeling is on par with other NAs unlike ammonia
-    - THAAs show clear usage of arginine in proteins
-      - 30-50% arg in THAAs labeled @73hrs
-      - Clear reuse of arginine backbone as fully labeled glutamate and proline
   - Guanosine monophosphate
@@
       - Implies mostly converted into ammonia as well and used that way
-    - Reuse of carbon backbone for THAAs though
+
+We also explored the degree of labeling incorporation into proteins by measuring the total hydrolyzable amino acids (THAAs) at both the start of each incubation and the end (Figure 2).
+Proteins were 30-50% of total particulate carbon pool (~2 uM) and 70-100% of the total particulate nitrogen (~0.25 uM), with free amino acids composing between 0.15% and 17% of the total.
+Glutamate + glutamine had the largest fraction in the free amino acid pool, varying between 3.4% and 17% while threonine had the largest fraction in protein (0.15% - 0.3% free).
+[Discuss additional variance here - see below tables(?)]
+
+  - Nitrate
+    - More extensive labeling (10-50% of the N) in the THAAs
+      - Must be a really high flux through a small AA pool subset, e.g. euks specifically
+  - Ammonia
+    - THAAs show clear usage of arginine in proteins
+    - 30-50% arg in THAAs labeled @73hrs
+    - Clear reuse of arginine backbone as fully labeled glutamate and proline
+  - GMP
+    - Reuse of carbon backbone
     - Histidine 13C5, 15N3
@@
 
-```{r thaa data}
+```{r thaa tables}
+thaa_avgs <- thaa_quant %>%
+  left_join(metathaa) %>%
+  filter(samp_type=="Smp") %>%
+  filter(depth=="Surf") %>%
+  select(compound_name, iso_name, amendment, timepoint, depth, startime, env_conc_nm) %>%
+  summarise(thaa_nM=mean(env_conc_nm), .by = c(compound_name, iso_name, amendment, timepoint)) %>%
+  mutate(compound_name=case_when(
+    compound_name%in%c("Aspartic acid")~"Aspartate + asparagine",
+    compound_name%in%c("Glutamic acid")~"Glutamate + glutamine",
+    compound_name%in%c("Leucine", "Isoleucine")~"Leucine + isoleucine",
+    TRUE~compound_name
+  )) %>%
+  mutate(iso_name=str_remove(iso_name, ".*(?=13)")) %>%
+  summarise(thaa_nM=sum(thaa_nM), .by=c(compound_name, iso_name, amendment, timepoint))
+metab_avgs <- nM_labeled %>%
+  mutate(compound_name=str_remove_all(compound_name, "L-")) %>%
+  mutate(iso_name=str_remove(iso_name, ".*(?=13)")) %>%
+  filter(compound_name%in%aa_names) %>%
+  filter(!compound_name%in%c("Methionine", "Valine", "Lysine")) %>%
+  mutate(compound_name=case_when(
+    compound_name%in%c("Asparagine", "Aspartic acid")~"Aspartate + asparagine",
+    compound_name%in%c("Glutamine", "Glutamic acid/beta-Glutamic acid")~"Glutamate + glutamine",
+    compound_name=="Threonine/Homoserine"~"Threonine",
+    compound_name=="Leucine/Isoleucine"~"Leucine + isoleucine",
+    TRUE~compound_name
+  )) %>%
+  summarise(nM_labeled=sum(nM_labeled), .by = c(compound_name, iso_name, filename)) %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  mutate(final_timepoint=ifelse(amendment%in%c("Amm", "Nit"), "T26", "T73")) %>%
+  filter(timepoint%in%c("T0", final_timepoint)) %>%
+  select(compound_name, iso_name, amendment, timepoint, depth, startime, nM_labeled) %>%
+  summarise(metab_nM=mean(nM_labeled), .by = c(compound_name, iso_name, amendment, timepoint))
+
+# Fraction of THAA that's DFAA at T0 and T26
+left_join(metab_avgs, thaa_avgs) %>%
+  summarise(metab_nM=sum(metab_nM, na.rm=TRUE), 
+            thaa_nM=sum(thaa_nM, na.rm=TRUE),
+            .by=c(compound_name, timepoint, amendment)) %>%
+  mutate(perc_in_free=metab_nM/(metab_nM+thaa_nM)*100) %>%
+  select(amendment, compound_name, timepoint, perc_in_free) %>%
+  pivot_wider(names_from=timepoint, values_from=perc_in_free) %>%
+  mutate(T26=ifelse(amendment%in%c("Arg", "GMP"), NA, T26)) %>%
+  arrange(amendment, compound_name)
+
+# Breakdown of THAA isotope labeling
+left_join(metab_avgs, thaa_avgs) %>%
+  mutate(perc_in_free=metab_nM/(metab_nM+thaa_nM)*100) %>%
+  arrange(amendment, compound_name) %>%
+  filter(perc_in_free<100) %>%
+  arrange(timepoint, compound_name)
+```
+
+```{r fig2 thaa data}
 all_thaa_plot <- thaa_quant %>%
@@
   mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"),
-                          labels=c("Initial", "+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
+                          labels=c("Initial (T0)", "+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
   summarise(total_uM=mean(total_uM), .by=c(amendment, timepoint, name)) %>%
@@
   ggplot(aes(x=amendment, y=total_uM, fill=name)) +
-  geom_col(color="black", width=1) +
+  geom_col(color="black", width=0.9) +
   theme_bw() +
-  scale_x_discrete(expand = expansion()) +
+  # scale_x_discrete(expand = expansion(mult = c(0.05))) +
   scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
@@
 ```{r load clean_combined and make feat_shape_df}
-cleaned_combined_peaks <- readRDS("untargeted/cleaned_combined_peaks.rds")
 lmtest_output <- cleaned_combined_peaks %>%
@@
 
+ - Putative annotation of theanine based on labeling patterns
+ - Annotation of pyrroline and friends too
+
 # Discussion
+
+[DISCUSSION: surface community not "expecting" nitrate? problems with redox at surface?]
+
+[DISCUSSION: not expecting to see any C labeling from the organics via CO2 because inorganic C pool is so much larger, unlike the N pool]
 

### 2025-05-20 manuscript/manuscript.Rmd
Fixing subscripts

@@
 The deoxyribose forms were in all cases less completely labeled than their ribose counterparts (deoxycytidine = 18.9%, deoxyadenosine = 20.3%, and deoxyguanosine = 17.2%) `r wc("Maybe better as a small table?")` `r wc("Consider trying harder to get thymine/uracil/inosine values?")`.
-In each case, the fully ^15^N labeled nucleobase was the dominant constituent of the isotope envelope (^15^N_3_ for cytosine and ^15^N_5_ for adenine and guanine).
+In each case, the fully ^15^N labeled nucleobase was the dominant constituent of the isotope envelope (^15^N~3~ for cytosine and ^15^N~5~ for adenine and guanine).
 Other compounds, including betaines (here restricted to quaternary amines with a carboxylic acid group) and choline derivatives showed very little labeling at all, with only homarine (15.0% labeled) and O-acetylcarnitine (9.9% labeled) exceeding 5% labeling after 24 hours`r wc("Expand betaine discussion and turn it into a separate paragraph?
@@
 
-Amendments with ^15^NO_3_ showed a significantly different labeling pattern with very few metabolites incorporating the ^15^N label at the surface even after 24 hours.
+Amendments with ^15^NO~3~ showed a significantly different labeling pattern with very few metabolites incorporating the ^15^N label at the surface even after 24 hours.
 Glutamine and glutamate, which were labeled within 20 minutes from ammonia and are the main way in which inorganic nitrogen becomes organic, here saw only trace amounts of production from the added nitrate even after 24 hours.
@@
 The organic nitrogen amendments also showed labeling patterns distinct from the ammonia incubations, though a large fraction of the nitrogen was clearly remineralized into ammonia/glutamate and used to label a similar suite of compounds (Figure 1).
-Fully C and N labeled (^13^C_6_, ^15^N_4_) arginine was visibly taken up at the T0 timepoint (within 20 minutes) and 80-90% of the pool was fully labeled after 12 hours.
+Fully C and N labeled (^13^C~6~, ^15^N~4~) arginine was visibly taken up at the T0 timepoint (within 20 minutes) and 80-90% of the pool was fully labeled after 12 hours.
 Surprisingly, the fraction of arginine labeled decreased from this point, with the T26 and T73 timepoints showing

### 2025-05-20 manuscript/manuscript.Rmd
Completed first draft of first manuscript results

@@
 
-```{r setup, echo=FALSE}
+```{r setup, echo=FALSE, include=FALSE}
 knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE)
@@
 Other compounds, including betaines (here restricted to quaternary amines with a carboxylic acid group) and choline derivatives showed very little labeling at all, with only homarine (15.0% labeled) and O-acetylcarnitine (9.9% labeled) exceeding 5% labeling after 24 hours`r wc("Expand betaine discussion and turn it into a separate paragraph?
-Not sure what more to say...")`.
+Not sure what more to say...
+fluxes?")`.
 
@@
 Fully C and N labeled (^13^C~6~, ^15^N~4~) arginine was visibly taken up at the T0 timepoint (within 20 minutes) and 80-90% of the pool was fully labeled after 12 hours.
-Surprisingly, the fraction of arginine labeled decreased from this point, with the T26 and T73 timepoints showing
-
-  - Arginine
-    - Clear uptake at T0 (20 minutes), 80-90% of the pool taken up at T10
-      - Labeling signal decreases after 24 hours
-    - Label obviously making it into citrulline so urea cycle is active
-      - Don't really have good enough urea data to see labeling change there?
-    - Big jump in labeling of other metabolites at T26 not seen at T10
-      - No hypotheses why this is happening but it occurs in both morn/eve so not light signal
-    - Maybe less storage in guanine specifically, labeling is on par with other NAs unlike ammonia
-  - Guanosine monophosphate
-    - Unable to measure GMP well enough directly?
-    - Guanosine labeling visible at T3, definitely T10, maybe some decrease at T73 like Arg?
-      - Guanosine pool never gets super heavily labeled
-    - Has a VERY similar pattern to arginine
-      - Also a big jump at T26 with basically no labeling at T10
-      - Implies mostly converted into ammonia as well and used that way
-
-We also explored the degree of labeling incorporation into proteins by measuring the total hydrolyzable amino acids (THAAs) at both the start of each incubation and the end (Figure 2).
-Proteins were 30-50% of total particulate carbon pool (~2 uM) and 70-100% of the total particulate nitrogen (~0.25 uM), with free amino acids composing between 0.15% and 17% of the total.
-Glutamate + glutamine had the largest fraction in the free amino acid pool, varying between 3.4% and 17% while threonine had the largest fraction in protein (0.15% - 0.3% free).
-[Discuss additional variance here - see below tables(?)]
-
-  - Nitrate
-    - More extensive labeling (10-50% of the N) in the THAAs
-      - Must be a really high flux through a small AA pool subset, e.g. euks specifically
-  - Ammonia
-    - THAAs show clear usage of arginine in proteins
-    - 30-50% arg in THAAs labeled @73hrs
-    - Clear reuse of arginine backbone as fully labeled glutamate and proline
-  - GMP
-    - Reuse of carbon backbone
-    - Histidine 13C5, 15N3
-    - Tyrosine 13C4, 15N1
-    - Phenylalanine 13C4, 15N1
+Surprisingly, the fraction of arginine labeled decreased from this point, with the T26 and T73 timepoints dropping to 73% and 46% labeled, respectively.
+This labeling rapidly propagated into citrulline via the urea cycle, indicating that labeled urea (^13^C~1~, ^15^N~2~) was created and available in the system as well though we were unable to measure this directly `r wc("I could try harder to get urea values for this but it's pretty bad.")`.
+Glutamate and glutamine were labeled much more slowly, reaching 10% labeled after 10 hours and a maximum of 66% labeled after 24 hours.
+Much of this nitrogen was then cycled very quickly among other compounds which also saw a jump in labeling at 24 hours.
+This was particularly true for other amino acids and the nucleobases/nucleosides and here, guanine was not labeled more than other nucleobases.
+Mycosporine-like amino acids and betaines were unlabeled above background even after 3 days of incubation.
+
+```{r gmp numbers}
+labeled_fracs %>%
+  filter(compound_name%in%c("Guanosine", "Guanine")) %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(depth=="Surf") %>%
+  filter(amendment=="GMP") %>%
+  summarise(frac=mean(labeled_frac), .by=c(compound_name, timepoint)) %>%
+  pivot_wider(names_from=compound_name, values_from=frac) %>%
+  arrange(timepoint)
+peak_areas %>%
+  filter(compound_name%in%c("Guanosine", "Guanine", "Adenine", "Adenosine", "Hypoxanthine/Allopurinol")) %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(amendment=="GMP") %>%
+  filter(depth=="Surf") %>%
+  mutate(iso_name=str_remove(iso_name, ".*?, ")) %>%
+  # filter(iso_name!="13C0, 15N0") %>%
+  ggplot() +
+  geom_col(aes(x=tripl, y=area, fill=iso_name), position="fill", color="black") +
+  facet_grid(compound_name~timepoint)
+```
+
+The guanosine monophosphate (GMP) incubations showed a pattern very similar to the arginine incubations in the identity, timing, and magnitude in which metabolites were labeled (Figure 1).
+While we were unable to measure GMP directly `r wc("I can try harder here but the data is pretty bad :(")` we saw clear labeling of fully C and N labeled guanosine (^13^C~10~, ^15^N~5~) after 3 hours that increased to a maximum of 75% after 24 hours before decreasing slightly to 64% labeled at the end of the three-day incubation.
+A small part of this signal was reflected in guanine, with ^13^C~5~, ^15^N~5~ guanine representing about 10% of the signal at T3 and reaching a maximum of 40% at T26.
+At T26 we also saw extensive labeling in adenine, adenosine, and hypoxanthine though their dominant labeling patterns (^13^C~0~ ^15^N~5~, ^13^C~5~ ^15^N~5~, and ^13^C~0~ ^15^N~4~ respectively, Supplemental figure XX) suggested that the adenine and hypoxanthine were synthesized from glutamate and that the ribose pool had been swamped by guanosine cleavage.
+Here, unlike the arginine incubations, we also saw some MAA labeling at the end of the 3 day incubation.
 
@@
   summarise(thaa_nM=sum(thaa_nM), .by=c(compound_name, iso_name, amendment, timepoint))
+
 metab_avgs <- nM_labeled %>%
@@
             .by=c(compound_name, timepoint, amendment)) %>%
+  filter(!(timepoint=="T26" & amendment%in%c("Arg", "GMP"))) %>%
   mutate(perc_in_free=metab_nM/(metab_nM+thaa_nM)*100) %>%
@@
   pivot_wider(names_from=timepoint, values_from=perc_in_free) %>%
-  mutate(T26=ifelse(amendment%in%c("Arg", "GMP"), NA, T26)) %>%
   arrange(amendment, compound_name)
@@
 ```
+
+We also explored the degree of labeling incorporation into proteins by measuring the total hydrolyzable amino acids (THAAs) at both the start of each incubation and the end (Figure 2).
+Proteins were 30-50% of total particulate carbon pool (~2 uM) and 70-100% of the total particulate nitrogen (~0.25 uM), with free amino acids composing between 0.15% and 17% of the total`r wc("I'm not really sure what to say here or how to dive into the variance that's of interest to us.
+Report the compound with the most/least in protein?
+Break it down by labeled vs unlabeled?")``.
+
+There were differences in THAA labeling patterns between the amendments (Figure 2).
+Despite their shorter incubation time (26 hours vs 73 hours), the inorganic nitrogen amendments generally higher fractions of labeled protein.
+The nitrate amendments in particular were surprisingly heavily labeled given the minimal signal seen in the free metabolites, with a degree of labeling exceeding the ammonia incubation.
+This indicates that most of the assimilation was done by the small subset of organisms able to reduce nitrate and use it for growth and that these organisms did not share the reduced nitrogen with the rest of the community`r wc("Is this something for the discussion instead?")`.
 
@@
 Colors denote specific labeling patterns detected in the THAAs.*
+
+We believe the organic nitrogen amendments largely resulted in metabolite labeling via remineralization to ammonia or glutamate based on the absence of a labeled carbon backbone in most of the labeled metabolites, but in the THAAs we were able to see some transformations that used the carbon backbone as-is.
+Proline and glutamate both had detectable ^13^C~5~ ^15^N~1~ labels in the arginine treatments.
+We also saw histidine ^13^C~5~ ^15^N~3~ synthesized during the guanosine monophosphate treatment, confirming again that the ribose pool must have been swamped with fully ^13^C labeled ribose cleaved from guanosine in order to enter the PRPP pathway.
+The ^15^N~3~ pattern also indicates that for these organisms, glutamate, glutamine, and adenine were also heavily labeled.
+Finally, both phenylalanine and tyrosine had significant ^13^C~4~ ^15^N~1~ labeling, likely from erythrose 4-phosphate via the pentose phosphate pathway indicating new synthesis through shikimate with nitrogen addition from a heavily labeled glutamate pool again.
 
@@
 
+[DISCUSSION: What does the decrease in arginine labeling in +Arg at T26 and T73 indicate? Pulse-chase behavior? How is the labeled pool specifically decreasing?]
+
+[DISCUSSION: Contrasting the THAA +Nit with the metabolites hints at a small but rapidly churning amino acid pool that's deeply segregated]
+
 # Conclusion

### 2025-05-20 manuscript/manuscript.Rmd
Adding MAA results section

@@
                "Also Porphyra-334?", "Asterina-330?", "Palythene/Usujirene?",
-               "Mycosporine-2-glycine?")
+               "Mycosporine-2-glycine?", "Mycosporine-alanine/Mycosporine-beta alanine?",
+               "Mycosporine-methylamine:serine?", "Mycosporine-glutamine?",
+               "Mycosporine glutaminol/Palythinol/Mycosporine-methylamine:threonine?", 
+               "Palythenic acid?")
 nucleo_names <- c("5-Methylcytosine", "Adenosine", "Cytidine", "Adenine", "Guanosine",
@@
 
+```{r maa numbers}
+labeled_fracs %>%
+  filter(compound_name%in%chosen_maas) %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(timepoint%in%c("T0", "T10")) %>%
+  filter(depth=="Surf") %>%
+  filter(startime=="Morn") %>%
+  filter(amendment=="Amm") %>%
+  summarise(frac=mean(labeled_frac), .by=c(compound_name, timepoint))
+```
+
+We performed the four nitrogen amendments at both 6AM and at 6PM to test for any diel effects in labeling patterns.
+Very few compounds showed a significant difference between the 6AM start time and the 6PM start time with the exception being the mycosporine-like amino acids (MAAs, Figure 3).
+Shinorine and palythine were confirmed via MS^2^ fragments while the other annotations remain putative, though their presence exclusively in the surface is compelling.
+During the ammonia incubations, six `r wc("double-check this after rerun")` MAAs showed clear ^15^N incorporation while exposed to sunlight, with the labeled fraction increasing sharply over the first 10 hours in the incubations started at 6AM.
+Incubations started at 6PM showed very little labeling from ammonia until a full 24 hours had passed.
+Shinorine, mycosporine-glycine, palythine, porphyra-334 were all clearly labeled (>20% ^15^N) as well as two of three molecular features detected at an *m/z* of 285.1445 possibly corresponding to palythene and usujirene.
+
 ```{r fig MAA labeling new}
+all_maa_data <- labeled_fracs %>%
+  filter(compound_name%in%maa_names) %>%
+  mutate(compound_name=factor(compound_name, levels=maa_names)) %>%
+  arrange(compound_name) %>%
+  mutate(compound_name=str_remove(compound_name, "\\?$")) %>%
+  mutate(compound_name=str_replace(compound_name, "Also Palythene/Usujirene", "Palythene or\n usujirene")) %>%
+  mutate(compound_name=str_replace(compound_name, "Palythene/Usujirene", "Usujirene or\n palythene")) %>%
+  mutate(compound_name=str_remove(compound_name, "Also ")) %>%
+  mutate(compound_name=str_replace(compound_name, "-glycine", "-\nglycine")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  left_join(parametadata, relationship = "many-to-many") %>%
+  filter(samp_type=="Smp") %>%
+  filter(IS_type=="noIS") %>%
+  filter(amendment%in%c("Amm", "Nit", "Arg", "GMP")) %>%
+  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"), labels=c("+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
+  mutate(clock_time=ifelse(startime=="Morn", 6+timepoint_num, 18+timepoint_num)) %>%
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start")))
+all_maa_gp <- ggplot(all_maa_data) +
+  annotate("rect", xmin=-Inf, xmax = 6, ymin=-Inf, ymax=Inf, fill="grey80") +
+  annotate("rect", xmin=18, xmax = 30, ymin=-Inf, ymax=Inf, fill="grey80") +
+  annotate("rect", xmin=42, xmax = Inf, ymin=-Inf, ymax=Inf, fill="grey80") +
+  geom_hline(yintercept = c(0.25, 0.5, 0.75, 1), color="grey90") +
+  geom_smooth(aes(x=clock_time, y=labeled_frac, color=startime),
+              method=nlsLM, formula="y~a*x^b+d", se = FALSE,
+              method.args=list(start=list(a=1, b=2, d=1),
+                               control=nls.lm.control(maxiter = 500))) +
+  # geom_smooth(aes(x=clock_time, y=labeled_frac, group=startime),
+  #             method=glm, formula = "y~x", method.args=list(family="quasibinomial"), 
+  #             color="grey10", se=FALSE) +
+  # geom_point(aes(x=clock_time, y=labeled_frac, shape=startime), color="black", 
+  #            fill="white", size=3) +
+  geom_point(aes(x=clock_time, y=labeled_frac, shape=startime, fill=startime),
+             color="black", size=2, alpha=0.7) +
+  facet_nested(compound_name~amendment, scales="free_x") +
+  scale_x_continuous(breaks = c(6, 18, 30, 42), labels = c("6AM", "6PM", "6AM", "6PM")) +
+  scale_y_continuous(labels = scales::label_percent(),
+                     limits = c(0, 1), breaks = c(0, 0.5, 1)) +
+  scale_shape_manual(values=c(24, 25), breaks = c("6AM start", "6PM start")) +
+  scale_fill_manual(values=c("#f4bb23", "#0b505c"), breaks = c("6AM start", "6PM start"),
+                    aesthetics = c("color", "fill")) +
+  labs(x=NULL, color=NULL, shape=NULL, fill=NULL, y="Fraction of MAA containing at least one 15N atom") +
+  theme_bw() +
+  theme(axis.text.y = element_markdown(vjust=c(0, 0.5, 1)),
+        axis.title.x = element_blank(),
+        legend.position = "inside",
+        legend.margin = margin(0.1,0.25,0.1,0.1, unit="lines"),
+        legend.position.inside = c(1, 1),
+        legend.justification = c(1, 1), 
+        legend.background = element_rect(fill="#FFFFFFAA", color="black", linewidth = 0)) +
+  guides(shape=guide_legend(override.aes = list(size=3, stroke=1.2, alpha=1, linetype=0)))
+ggsave(plot = all_maa_gp, filename = "supp_MAA_curves.png", device = "png", 
+       path = "manuscript/supplement", width = 6.5, height = 9, units = "in", dpi = 300)
+
 chosen_maas <- c("Mycosporine-glycine?", "Shinorine?", "Palythine?",
@@
                  "Also Porphyra-334?")
-labeled_fracs %>%
+chosen_maa_data <- labeled_fracs %>%
   filter(compound_name%in%chosen_maas) %>%
@@
   mutate(clock_time=ifelse(startime=="Morn", 6+timepoint_num, 18+timepoint_num)) %>%
-  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start"))) %>%
-  ggplot() +
+  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start")))
+chosen_maa_gp <- ggplot(chosen_maa_data) +
   annotate("rect", xmin=-Inf, xmax = 6, ymin=-Inf, ymax=Inf, fill="grey80") +
-  # annotate("rect", xmin=6, xmax = 18, ymin=-Inf, ymax=Inf, fill="#d0bd65") +
   annotate("rect", xmin=18, xmax = 30, ymin=-Inf, ymax=Inf, fill="grey80") +
-  # annotate("rect", xmin=30, xmax = 42, ymin=-Inf, ymax=Inf, fill="#d0bd65") +
   annotate("rect", xmin=42, xmax = Inf, ymin=-Inf, ymax=Inf, fill="grey80") +
   geom_hline(yintercept = c(0.25, 0.5, 0.75, 1), color="grey90") +
-  # annotate("rect", xmin=-Inf, xmax=Inf, ymin=-Inf, ymax = Inf, fill="white", alpha=0.3) +
   geom_smooth(aes(x=clock_time, y=labeled_frac, color=startime),
@@
                                control=nls.lm.control(maxiter = 500))) +
-  # geom_smooth(aes(x=clock_time, y=labeled_frac, group=startime),
-  #             method=glm, formula = "y~x", method.args=list(family="quasibinomial"), 
-  #             color="grey10", se=FALSE) +
-  # geom_point(aes(x=clock_time, y=labeled_frac, shape=startime), color="black", 
-  #            fill="white", size=3) +
   geom_point(aes(x=clock_time, y=labeled_frac, shape=startime, fill=startime),
@@
   guides(shape=guide_legend(override.aes = list(size=3, stroke=1.2, alpha=1, linetype=0)))
-
-ggsave(filename = "MAA_curves.png", device = "png", path = "manuscript/figures",
-       width = 3, height = 6, units = "in", dpi = 300)
+ggsave(plot = chosen_maa_gp, filename = "MAA_curves.png", device = "png", 
+       path = "manuscript/figures", width = 3, height = 6, units = "in", dpi = 300)
 ```
@@
 
-  - MAAs clearly show daytime synthesis during the day and not at night
-    - Very clear in +Ammonia, less so in +Nitrate
-    - Need to show this statistically somehow?
-  - High levels of labeling despite minimal glycine labeling and synthesis being 4DG+Gly for these
-  - Putative annotations based on labeling pattern (and only found at surface)
-    - Maybe also can do a search for fragments
-
-## Inorganics more bioavailable at surface, organics equally bioavailable at 25m and 175m
+Nitrate data showed similar degrees of labeling which was again surprising given the reduced amount of bulk metabolite labeling in those amendments, though the synthesis of MAAs from nitrate appeared less clearly diel.
+Shinorine and porphyra-334 were particularly well-labeled under nitrate, approaching the values achieved in the ammonia incubations.
+Very little MAA labeling was observed during the arginine and GMP amendments until the 73 hour timepoint, though at that point the fraction labeled was approximately equivalent to the inorganic amendments and slightly higher for GMP than arginine.
+
+## Ammonia more bioavailable at surface, organics equally bioavailable at 25m and 175m
 
@@
 [DISCUSSION: Contrasting the THAA +Nit with the metabolites hints at a small but rapidly churning amino acid pool that's deeply segregated]
+
+[DISCUSSION: MAAs are mostly made from the 4DG core plus N-containing amino acids (glycine). How does the labeling of these compare to their substituent AAs? (it's much larger)]
 

### 2025-05-22 manuscript/manuscript.Rmd
Results section finished and sent to Anitra
Needs review, esp. untargeted section.

@@
 # Results
+
+`r wc("Do I need a 'here's what the cruise looked like' results section?")`
 
@@
 
+```{r surfdeep nums}
+labeled_fracs %>%
+  # filter(compound_name%in%c("L-Glutamic acid/beta-Glutamic acid", "L-Aspartic acid", "L-Asparagine", "Theanine?")) %>%
+    # filter(compound_name%in%c("Guanine", "Guanosine", "Cytosine", "Cytidine", "Adenine", "Adenosine")) %>%
+  filter(compound_name%in%c("Glycine betaine", "Homarine", "Trigonelline", "Choline")) %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  filter(amendment=="Amm") %>%
+  summarise(frac=mean(labeled_frac), .by=c(compound_name, timepoint, depth)) %>%
+  pivot_wider(names_from=depth, values_from=frac) %>%
+  summarise(Surf/Deep, .by=c(compound_name, timepoint)) %>%
+  arrange(compound_name)
+```
+
+We also performed the four nitrogen amendments at both the surface (25 meters) and below the euphotic zone (175 meters) to investigate the effect of depth on organic and inorganic nitrogen use `r wc("Maybe find and include the light level at 175m?")`.
+We found that ammonia was used much more readily at the surface as measured by the fraction of metabolite labeled, with ^15^N incorporation for amino acids 5-25% greater in the surface than at depth (Figure 4).
+Glutamate, theanine, aspartate, and asparagine were among the most different with glutamate and theanine surface values about 10 times those at depth while ^15^N aspartate was undetectable until T3 and ^15^N asparagine never detected.
+In contrast, glutamine labeling was similar between the two depths, with the fraction labeled 10-20 percentage points higher at the surface across all timepoints; a trend which was clearly not propagated to glutamate or other amino acids and instead must have represented accumulation or protein incorporation.
+`r wc("Not super happy with the numeric language here - feels like I'm flipping between 'times greater' and 'percentage points higher'.")`
+
 ```{r fig surface vs deep labeling rates}
@@
 
-  - Ammonia incubations result in much faster labeling at 25 meters than 175m
-    - Presumed to be a phototrophic signal
-    - Major exception is ectoine, a bacterial biosignal (McParland 2021)
-  - Organic N incubations result in equal amounts of labeling at surface and depth
-    - Ectoine again predominantly labeled at depth
-    - Labeling of guanosine at T0 in the +GMP treatment implies rapid uptake and P-cleavage
-  - Nitrate data unavailable due to T10 and T26 deep samples contaminated
-    - See supplemental figure 1 for the data anyway
-
-## Other results from untargeted analysis
-
-```{r source multicore}
-source("untargeted/xcms_multicore.R")
-```
+Other metabolites varied in their degree of labeling (Figure 4).
+Nucleobases and nucleosides were much more heavily labeled from ammonia at the surface, with maximum labeling 3-10 times more extensive around T3 and T10 if the isotope was measurable at both depths and infinitely more so if not.
+Betaines and other nitrogen-containing metabolites, however, were almost entirely unlabeled at 175 meters until the 24 hour mark, at which the fraction labeled in both was 5-20 times higher at the surface.
+An important exception to this rule is the osmolyte ectoine, an aspartate derivative that nonetheless was synthesized with nitrogen from ammonia at depth at rates equal to or exceeding those of the surface.
+
+We were unfortunately unable to present the data collected from the 175 meter labeled nitrate incubations here due to missed timepoints and confusion during sample collection but the raw numbers are provided in the supplement (Table XX and Supplemental Figure XX).
+
+^15^N use from the organic nitrogen substrates was again strikingly similar between the 25 meter and 175 meter samples aside from the direct products of the amended compounds.
+In both cases, uptake and use of the labeled substrate was equal or slightly higher at depth for a majority of the metabolites measured.
+Exceptions to this were theanine, which was 20-30 percentage points more heavily labeled at the surface by the end of the incubations, and ectoine, which was synthesized more at depth even to a larger degree than it was during the ammonia treatments, reaching nearly 50 percentage points more labeled at 175 meters than it was at 25m.
+
+## Untargeted methods reveal widespread labeling throughout the metabolome
 
 ```{r load clean_combined and make feat_shape_df}
+cleaned_combined_peaks <- readRDS("untargeted/cleaned_combined_peaks.rds")
+
 lmtest_output <- cleaned_combined_peaks %>%
@@
   filter(sapply(data, nrow)>2) %>%
-  mutate(lm_test=map(data, ~broom::tidy(lm(log10(.x$into)~log10(.x$timepoint_num))), .progress = TRUE)) %>%
+  mutate(lm_test=map(data, ~broom::tidy(lm(log10(.x$into)~log10(.x$timepoint_num))),
+                     .progress = TRUE)) %>%
   unnest(lm_test) %>%
@@
   mutate(p_adj=p.adjust(p.value, method = "fdr"))
-
 feat_shape_df <- cleaned_combined_peaks %>%
@@
   mutate(peak_qual=ifelse(med_snr>-50*med_cor+50, "Good", "Bad")) %>%
-  mutate(peak_qual=ifelse(rtmed<2*60 | between(rtmed, 14*60, 14.5*60), "Solvent artifact", peak_qual)) %>%
+  mutate(peak_qual=ifelse(rtmed<2*60 | between(rtmed, 14*60, 14.5*60), 
+                          "Solvent artifact", peak_qual)) %>%
   mutate(is_iso=ifelse(med_iso_cor>0.8, "Is iso", "Not iso")) %>%
   mutate(time_trend=ifelse(p_adj<0.001& estimate>0, "Increases", "No change"))
-feat_shape_df %>%
-  filter(!is.na(time_trend)) %>%
-  count(peak_qual, is_iso, time_trend)
-```
-
-```{r get counts per amendment instead of as a whole}
+
 # Repeat analysis across amendments to get additional numbers
@@
   filter(sapply(data, nrow)>2) %>%
-  mutate(lm_test=map(data, ~broom::tidy(lm(log10(.x$into)~log10(.x$timepoint))), .progress = TRUE)) %>%
+  mutate(lm_test=map(data, ~broom::tidy(lm(log10(.x$into)~log10(.x$timepoint))), 
+                     .progress = TRUE)) %>%
   unnest(lm_test) %>%
@@
   mutate(peak_qual=ifelse(med_snr>-50*med_cor+50, "Good", "Bad")) %>%
+  mutate(peak_qual=ifelse(rtmed<2*60 | between(rtmed, 14*60, 14.5*60), 
+                          "Solvent artifact", peak_qual)) %>%
   mutate(is_iso=ifelse(med_iso_cor>0.8, "Is iso", "Not iso")) %>%
   mutate(time_trend=ifelse(p_adj<0.001& estimate>0, "Increases", "No change"))
-feat_shape_df_xamend %>%
-  bind_rows(iso_df %>% mutate(amendment="Overall")) %>%
-  # filter(!rtmed%between%(c(14, 14.5)*60)) %>%
-  # filter(!rtmed<(2*60)) %>%
-  filter(!is.na(time_trend)) %>%
-  count(peak_qual, is_iso, time_trend, amendment) %>%
-  pivot_wider(names_from = amendment, values_from=n, values_fill = 0) %>%
-  print()
+```
+
+```{r count categories as a whole and by amendment}
 feat_shape_df_xamend %>%
@@
   filter(!is.na(time_trend)) %>%
-  ggplot() +
-  geom_bar(aes(x=amendment)) +
-  facet_wrap(~peak_qual+is_iso+time_trend, scales="free_y", ncol=4)
-```
+  count(peak_qual, is_iso, time_trend, amendment) %>%
+  pivot_wider(names_from = amendment, values_from=n, values_fill = 0)
+```
+
+After a heuristic compound discovery approach revealed the putatively identified theanine, deoxyguanosine, deoxycytidine, and MAAs had interesting labeling patterns, we decided to perform a fully untargeted analysis to detect additional compounds that were labeled over the course of the incubations.
+Fairly lax peakpicking settings for xcms returned nearly 10,000 chromatographic peaks in each of the ~280 samples across the four LC runs used, one for each amendment sample set.
+At this point we noted that the nitrate amendment samples had an unusually large number of peaks detected due to the inclusion of a contaminant suite, likely plastic, which fortunately eluted before many of our compounds of interest.
+Correspondence requiring the presence of a peak in all three biological triplicates for a given treatment/timepoint combination reduced that number to approximately 5,000 unique features.
+Random inspection revealed that the majority of these features were noise so we calculated additional peak quality metrics as well as reintegrated each peak and its predicted isotope (see Methods).
+Of the initial 5,602 features, over half (2,956) failed quality control and an additional 1,690 were determined to be likely solvent artifacts due to their elution in less than two minutes or `r wc("when a large proportion of junk was coming off the column (median retention times between 14.0 and 14.5 minutes)?
+during the retention time when sulfate eluted?
+based on their peak shape and retention time?
+due to the fact that they showed no change across any of the treatments or timepoints?
+The darn things just look bad and I hate them but I can't really explain why.")` and were removed from downstream analysis.
 
@@
 ```
+
+Of the 219 features that remained, the majority (150) had a ^15^N isotope with the correct *m/z* value and a chromatographic similarity score above 0.8 (see Methods).
+80 of these 150 also showed linear increases in peak area across the timepoints, denoting likely ^15^N-labeled metabolites.
+We manually inspected each of these 80 metabolites and were able to identify the majority as either compounds in our list of authentic standards (19/80), an isotope of a known compound (39/80), or an in-source fragment of a known compound (12/80, with fragments from proline (70.0658), glutamic acid/glutamate (84.0449, 102.0550, and 130.0499), and the loss of ribose from nucleosides).
+Five compounds were putatively identified using isotope labeling patterns and literature searches: FT1362 with an *m/z* of 121.07798 and a retention time (RT) of 4.5 minutes we flagged as indoline, FT1564 with an *m/z* of 124.0524 and a 2.7 minute RT we flagged as the ^15^N isotope of nicotinamide, FT4529 with an *m/z* of 228.0979 and a 6.7 minute RT we flagged as deoxycytidine, FT5860 with an *m/z* of 282.1182 and an 11.1 minute RT we flagged as pangamic acid, and FT6622 with an *m/z* of 333.1292 and an 11.2 minute RT we flagged as shinorine.
+Four additional compounds could not be putatively identified but had plausible molecular formulae: FT3699 with an *m/z* of 192.0866 and 10.3 RT we assigned C~7~H~14~NO~5~, FT4598 with an *m/z* of 231.1703	and 1.9 RT we assigned C~11~H~23~N~2~O~3~, FT4733 with an *m/z* of 236.1492 and 10.0 RT we assigned C~10~H~22~NO~5~, and FT5185 with an *m/z* of 252.1441 and 10.7 RT we assigned C~10~H~22~NO~6~.
+One feature we were unable to annotate at all (FT5076, *m/z* 248.1258 and RT 12.0).
 
@@
 
+To derive biological meaning from the untargeted analysis, we performed a k-means clustering of the features across all the treatments and timepoints and included our targeted data in the clustering.
+Clear patterns of isotope incorporation emerged in 6 of the 10 clusters we used (Figure 5).
+Cluster 1, the largest with 64 molecular features, contained many of the unlabeled compounds including those without any nitrogen and all of the betaines.
+Cluster 2 contained most of the amino acids, the nucleobases cytosine and guanine, the and the deoxynucleosides as well as ectoine and citrulline.
+Cluster 3 contained homarine, beta-alanine, and O-acetylcarnitine as well as the remaining amino acids aside from aspartate, glutamate, and glutamine which fell into cluster 4.
+Known compounds in cluster 5 consisted of adenine, adenosine, cytidine, and hypoxanthine while guanosine fell alone into cluster 6.
+The MAAs were distributed across clusters 8 and 9, while clusters 7 and 10 had no known compounds.
+
 ```{r run kmeans}
 untarg_lab_mat <- untarg_lab_fracs %>%
-  bind_rows(labeled_fracs %>% rename(feature=compound_name)) %>%
+  bind_rows(labeled_fracs %>% rename(feature=compound_name) %>% select(-filename)) %>%
   pivot_wider(names_from = "feature", values_from = "labeled_frac", values_fill=0) %>%
@@
 # Plot clusters as heatmap
-untarg_lab_fracs %>%
+untarg_heatmap <- untarg_lab_fracs %>%
   left_join(kmeans_df) %>%
@@
   )
+ggsave("untarg_heatmap.png", untarg_heatmap, path="manuscript/figures",
+       width=6.5, height=4, units = "in")
 
@@
 
- - Putative annotation of theanine based on labeling patterns
- - Annotation of pyrroline and friends too
+![](figures/untarg_heatmap.png)
+
+*Figure 5: Fraction of each cluster containing at least one ^15^N label during the surface seawater amendments with various substrates after the number of hours shown on the x-axis.
+Darker colors indicate a larger percentage of the feature signal was isotopically labeled.
+Three biological replicates are stacked vertically at each timepoint and the labeling within each clusters has been averaged across all the composing features.
+The number of features in each cluster is specified by the n in the facet name.
+GMP = guanosine monophosphate.*
 
@@
 # Methods
+
+[explain chromatographic similarity scoring as pearson's correlation coefficient]
 

### 2025-05-22 manuscript/manuscript.Rmd
Finished annotation of high-quality untargeted features
Cleaning up figures and adding color palette

@@
                   "Guanosine", "Other nucleic acids", "MAAs", "Betaines", "All others")
+class_colors <- c("#08519C", "#2171B5", "#4292C6", "#6BAED6", "#8EBAD1", 
+                  "#08306B", "#006D2C", "#319B4D", "#AB0022", "#6A51A3", 
+                  "#AE9AD8", "#4E1F88", "#E0C174", "#4D2B01", "grey40")
 aa_ncs <- all_stans %>%
@@
   geom_vline(xintercept = c(1.5:5.5), color="grey80") +
-  facet_nested(compound_class~amendment, scales = "free_x", switch="y") +
+  facet_nested(compound_class~amendment, scales = "free_x", switch="y", 
+               strip=strip_nested(text_y = elem_list_text(color=class_colors, 
+                                                          face=rep("bold", length(class_colors))))) +
   scale_fill_gradientn(colours = viridis::viridis(10, alpha = seq(0, 1, length.out=10), direction = -1),
@@
         legend.position = "top", 
-        legend.key.width = unit(1, "null")) +
+        legend.key.width = unit(1, "null"),
+        text=element_text(size=15)) +
   guides(
@@
   )
+xamend_metab_gp
 ggsave("xamend_metab_gp.png",plot = xamend_metab_gp, device = "png", path = "manuscript/figures",
-       width = 6.5, height = 4, units = "in", dpi = 300)
+       width = 6.5, height = 4.5, units = "in", dpi = 300)
 ```
@@
   geom_col(color="black", width=0.9) +
+  geom_hline(yintercept = 0) +
   theme_bw() +
@@
   scale_fill_manual(breaks = c("Unlabeled", "Labeled"), values=c("grey80", "grey20")) +
-  labs(x=NULL, y="Total uM N in THAAs", fill=NULL)
+  labs(x=NULL, y="Total uM N in THAAs      ", fill=NULL)
 
@@
 plot_iso_colors <- c(
-  "grey80",
-  "#e0ecf6",
-  "#97bfe0",
-  "#377EB8",
-  "#285b84",
-  "#ffb4b5",
-  "#d7b6dc",
-  "#FFFF33",
-  "#4DAF4A",
-  "#984EA3",
-  "grey20"
+  "grey80","#e0ecf6","#97bfe0","#377EB8","#285b84","#ffb4b5","#AE9AD8",
+  "#E0C144","#319B4D","#5A4193","grey30"
 )
@@
            position="fill",
-           color="black", width=1) +
+           color="black", width=1, lwd=0.2) +
   facet_wrap(~compound_name) +
@@
 ggsave("total_thaa_gp.png", total_thaa_gp, device = "png", path = "manuscript/figures",
-       width = 6.5, height = 5, units = "in", dpi = 300)
+       width = 6.5, height = 4.5, units = "in", dpi = 300)
 ```
@@
   facet_wrap(~amendment, nrow=1) +
-  scale_color_discrete(drop=FALSE) +
+  scale_color_manual(breaks=class_levels, values = class_colors) +
   scale_x_log10(breaks=c(0.3, 3, 30), labels=c("0.3", "3", "30"), expand=expansion(0.15)) +
@@
   scale_shape_manual(breaks=setdiff(class_levels, "MAAs"), values = rep(21:25, length.out=14)) +
+  scale_color_manual(breaks=class_levels, values = class_colors, aesthetics = c("color", "fill")) +
   labs(x="Hours after <sup>15</sup>N spike", color=NULL, fill=NULL, shape=NULL,
@@
 
+surfdeep_gp
 ggsave(filename = "deep_vs_surf_labeling.png", surfdeep_gp, device = "png", 
@@
 Four additional compounds could not be putatively identified but had plausible molecular formulae: FT3699 with an *m/z* of 192.0866 and 10.3 RT we assigned C~7~H~14~NO~5~, FT4598 with an *m/z* of 231.1703	and 1.9 RT we assigned C~11~H~23~N~2~O~3~, FT4733 with an *m/z* of 236.1492 and 10.0 RT we assigned C~10~H~22~NO~5~, and FT5185 with an *m/z* of 252.1441 and 10.7 RT we assigned C~10~H~22~NO~6~.
-One feature we were unable to annotate at all (FT5076, *m/z* 248.1258 and RT 12.0).
+One feature we were unable to annotate at all (FT5076, *m/z* 248.1258 and RT 12.0).`r wc("Is this information useful?
+Or should I just refer people to a supplemental table after giving the general breakdown?")`
 

### 2025-05-22 manuscript/manuscript.Rmd
Adding untargeted results

@@
 cleaned_combined_peaks <- readRDS("untargeted/cleaned_combined_peaks.rds")
+high_qual_feats <- cleaned_combined_peaks %>%
+  filter(rtmed>120 & !rtmed%between%c(14*60, 14.5*60)) %>%
+  left_join(parametadata %>% distinct(filename, samp_type)) %>%
+  filter(samp_type=="Smp") %>%
+  summarise(med_iso_cor=median(iso_cor), 
+            med_snr=median(beta_snr),
+            med_cor=median(beta_cor),
+            .by=c(feature)) %>%
+  filter(med_snr>-50*med_cor+50) %>%
+  filter(med_iso_cor>0.8)
+untarg_lab_fracs <- cleaned_combined_peaks %>%
+  filter(feature%in%high_qual_feats$feature) %>%
+  select(feature, filename, lab_area, unlab_area) %>%
+  left_join(parametadata %>% distinct(filename, samp_type)) %>%
+  filter(samp_type=="Smp") %>%
+  slice_max(unlab_area, by=c(feature, filename), with_ties = FALSE) %>%
+  mutate(labeled_frac=lab_area/(lab_area+unlab_area)) %>%
+  left_join(parametadata %>% distinct(filename, shortname)) %>%
+  select(feature, shortname, labeled_frac)
 
@@
 
-```{r heatmap of all high-quality feats}
-high_qual_feats <- cleaned_combined_peaks %>%
-  filter(rtmed>120 & !rtmed%between%c(14*60, 14.5*60)) %>%
-  left_join(parametadata %>% distinct(filename, samp_type)) %>%
-  filter(samp_type=="Smp") %>%
-  summarise(med_iso_cor=median(iso_cor), 
-            med_snr=median(beta_snr),
-            med_cor=median(beta_cor),
-            .by=c(feature)) %>%
-  filter(med_snr>-50*med_cor+50) %>%
-  filter(med_iso_cor>0.8)
-untarg_lab_fracs <- cleaned_combined_peaks %>%
-  filter(feature%in%high_qual_feats$feature) %>%
-  select(feature, filename, lab_area, unlab_area) %>%
-  left_join(parametadata %>% distinct(filename, samp_type)) %>%
-  filter(samp_type=="Smp") %>%
-  slice_max(unlab_area, by=c(feature, filename), with_ties = FALSE) %>%
-  mutate(labeled_frac=lab_area/(lab_area+unlab_area)) %>%
-  left_join(parametadata %>% distinct(filename, shortname)) %>%
-  select(feature, shortname, labeled_frac)
-
-cmpd_order <- untarg_lab_fracs %>%
-  summarise(mean_label=mean(labeled_frac, na.rm=TRUE), .by=feature) %>%
-  arrange(desc(mean_label)) %>%
-  pull(feature)
-
-colorpal <- viridis::viridis(10, alpha=seq(0, 1, length.out=10), direction = -1)
-untarg_heatmap <- untarg_lab_fracs %>%
-  complete(feature, shortname, fill=list(labeled_frac=0)) %>%
-  mutate(feature=factor(feature, levels=cmpd_order)) %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  ggplot() +
-  geom_raster(aes(x=timepoint, y=tripl, fill=labeled_frac)) +
-  scale_fill_gradientn(colours = colorpal, limits = c(0, 1)) +
-  scale_y_discrete(expand=expansion()) +
-  scale_x_discrete(guide = guide_axis(n.dodge = 5), expand=expansion()) +
-  facet_nested(feature~amendment+depth+startime, scales = "free_x", switch="y", strip=strip_nested(size = "variable")) +
-  labs(x="Timepoint", y="Metabolite", fill="Fraction\nlabeled") +
-  theme_bw() +
-  theme(panel.spacing.x=unit(0.2, "lines"), 
-        panel.spacing.y = unit(0, "lines"),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        # axis.ticks.length.x=unit(1:5, "lines"),
-        # axis.text.x = element_markdown(vjust = -5:-1),
-        strip.text.y.left = element_text(angle=0, hjust = 1),
-        # strip.background.y = element_rect(fill=NA, color=NA),
-        panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank(),
-        legend.position = "top", 
-        legend.key.width = unit(1, "null")) +
-  guides(
-    fill=guide_colorbar(theme=theme(
-      legend.frame = element_rect(color = "black"),
-      legend.ticks = element_line(color="black"),
-      legend.text = ggtext::element_markdown(hjust=c(0, 0.5, 0.5, 0.5, 1)))
-    )
-  )
-ggsave("untargeted/untarg_heatmap.pdf", width = 10, height=36, units = "in")
-```
-
 To derive biological meaning from the untargeted analysis, we performed a k-means clustering of the features across all the treatments and timepoints and included our targeted data in the clustering.
@@
 
-```{r run kmeans}
+```{r run kmeans and save heatmap and save clusters}
 untarg_lab_mat <- untarg_lab_fracs %>%
@@
 
-set.seed(124)
+set.seed(123)
 
@@
 untarg_heatmap <- untarg_lab_fracs %>%
+  bind_rows(labeled_fracs %>% rename(feature=compound_name) %>% select(-filename)) %>%
   left_join(kmeans_df) %>%
@@
         axis.ticks.y = element_blank(),
-        # axis.ticks.length.x=unit(1:5, "lines"),
-        # axis.text.x = element_markdown(vjust = -5:-1),
         strip.text.y.left = element_text(angle=0, hjust = 1),
-        # strip.background.y = element_rect(fill=NA, color=NA),
         panel.grid.major = element_blank(), 
@@
 GMP = guanosine monophosphate.*
+
+```{r supp fig heatmap of all high-quality feats}
+colorpal <- viridis::viridis(10, alpha=seq(0, 1, length.out=10), direction = -1)
+untarg_heatmap <- untarg_lab_fracs %>%
+  bind_rows(labeled_fracs %>% rename(feature=compound_name) %>% select(-filename)) %>%
+  complete(feature, shortname, fill=list(labeled_frac=0)) %>%
+  left_join(kmeans_df) %>%
+  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
+  ggplot() +
+  geom_raster(aes(x=timepoint, y=tripl, fill=labeled_frac)) +
+  scale_fill_gradientn(colours = colorpal, limits = c(0, 1)) +
+  scale_y_discrete(expand=expansion()) +
+  scale_x_discrete(guide = guide_axis(n.dodge = 5), expand=expansion()) +
+  facet_nested(cluster_name~feature~amendment+depth+startime, 
+               scales = "free_x", switch="y", strip=strip_nested(size = "variable")) +
+  labs(x="Timepoint", y=NULL, fill="Fraction\nlabeled") +
+  theme_bw() +
+  theme(panel.spacing.x=unit(0.2, "lines"), 
+        panel.spacing.y = unit(0, "lines"),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        strip.text.y.left = element_text(angle=0, hjust = 1),
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        legend.position = "top", 
+        legend.key.width = unit(1, "null")) +
+  guides(
+    fill=guide_colorbar(theme=theme(
+      legend.frame = element_rect(color = "black"),
+      legend.ticks = element_line(color="black"),
+      legend.text = ggtext::element_markdown(hjust=c(0, 0.5, 0.5, 0.5, 1)))
+    )
+  )
+ggsave("manuscript/supplement/untarg_heatmap.pdf", width = 10, height=36, units = "in")
+```
 

### 2025-05-22 README.Rmd
Adding project README

--- 2025-05-22 manuscript/manuscript.Rmd
@@
 ---
-title: "Nitrogen's form determines its fate in the North Pacific Subtropical Gyre during PARAGON I"
-author:
-  - William Kumler
-  - Laura T. Carlson
-  - Anitra E. Ingalls
-  - Other SCOPE collaborators
-output: word_document
+title: "Nitrogen's form determines its fate in the North Pacific Subtropical Gyre"
+output: github_document
 ---
 
-```{r setup, echo=FALSE, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE)
-library(tidyverse)
-library(ggh4x)
-library(patchwork)
-library(ggtext)
-library(minpack.lm)
-library(RaMS)
-options(pillar.sigfig = 7)
-options(pillar.print_min = 100)
-
-wc <- function(
-  comment, 
-  highlight = "", 
-  author = "Will", 
-  time = Sys.time(),
-  id = "0"
-) {
-  if (isTRUE(knitr:::pandoc_to() == "docx")) {
-    return(
-      sprintf(
-        '[%s]{.comment-start id="%s" author="%s" date="%s"} %s []{.comment-end id="%s"}',
-        comment,
-        id,
-        author,
-        time,
-        highlight,
-        id
-      )
-    )
-  } else {
-    return(
-      sprintf(
-        "*%s* **[Comment id %s by %s at time %s: %s]**", 
-        highlight,
-        id,
-        author, 
-        time, 
-        comment
-      )
-    )
-  }
-}
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-```{r load data, eval=FALSE}
-parametadata <- read_csv("metadata/parametadata.csv") %>%
-  # filter(IS_type=="noIS") %>%
-  filter(polarity=="pos") %>%
-  filter(samp_type=="Smp" | samp_type=="Bad") %>%
-  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
-  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
-  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"))) %>%
-  arrange(amendment, depth, timepoint_num, startime) %>%
-  mutate(timepoint=factor(timepoint, levels=c("T0", "T1", "T3", "T10", "T18", "T26", "T73")))
+Nitrogen availability controls productivity in much of the surface ocean and 
+the form of this nitrogen controls who it is available to and how it is used. 
+Here, we explored how four different forms of nitrogen (ammonia, nitrate, 
+arginine, and guanosine monophosphate) were taken up and transformed in the 
+North Pacific Subtropical Gyre using stable isotope labeling metabolomics. We 
+repeated these experiments both at the surface and at the base of the euphotic 
+zone as well as in the morning and evening. We found that ammonia is converted 
+into glutamate in less than 20 minutes and results in labeling throughout the 
+metabolome, though osmolytes like betaines had slow turnover rates. This process 
+was slower at depth but largely independent of time of day. Nitrate amendments 
+resulted in the labeling of only a small subset of molecules such as 
+mycosporine-like amino acids, whose incorporation patterns suggest that this 
+form of nitrogen was available to a subset of the community, likely eukaryotes. 
+The two organic nitrogen amendments labeled the metabolome more slowly than the
+inorganic forms of nitrogen and showed similar usage at depth and at the 
+surface, hinting at the major role that heterotrophic bacteria likely play in 
+their uptake and use. A large fraction of the organic nitrogen added was 
+remineralized into ammonia and used as glutamate rather than as intact 
+substrates, though labeled arginine was incorporated into proteins directly. 
+These measurements reveal the forms and fluxes of organic nitrogen in the 
+natural environment to provide constraints on biogeochemical models and better 
+characterize the transformations between organic and inorganic nitrogen.
 
-all_stans <- read_csv("metadata/all_stans.csv")
-stan_patched <- all_stans %>%
-  filter(!compound_name=="beta-Glutamic acid") %>%
-  mutate(compound_name=str_replace(compound_name, "L-Glutamic acid", "L-Glutamic acid/beta-Glutamic acid")) %>%
-  filter(!compound_name=="L-Isoleucine") %>%
-  mutate(compound_name=str_replace(compound_name, "L-Leucine", "L-Leucine/L-Isoleucine")) %>%
-  filter(!compound_name=="L-Homoserine") %>%
-  mutate(compound_name=str_replace(compound_name, "L-Threonine", "L-Threonine/L-Homoserine")) %>%
-  filter(!compound_name=="Allopurinol") %>%
-  mutate(compound_name=str_replace(compound_name, "Hypoxanthine", "Hypoxanthine/Allopurinol"))
-maa_names <- c("Also Palythene/Usujirene?", "Shinorine?", "Mycosporine-glycine?", 
-               "Palythine?", "Another Palythene/Usujirene?", "Porphyra-334?", 
-               "Also Porphyra-334?", "Asterina-330?", "Palythene/Usujirene?",
-               "Mycosporine-2-glycine?", "Mycosporine-alanine/Mycosporine-beta alanine?",
-               "Mycosporine-methylamine:serine?", "Mycosporine-glutamine?",
-               "Mycosporine glutaminol/Palythinol/Mycosporine-methylamine:threonine?", 
-               "Palythenic acid?")
-nucleo_names <- c("5-Methylcytosine", "Adenosine", "Cytidine", "Adenine", "Guanosine",
-                  "Cytosine", "Hypoxanthine/Allopurinol", "Guanine", "Deoxycytidine?",
-                  "Deoxyadenosine", "Deoxyguanosine?")
-betaine_names <- c("beta-Alaninebetaine", "Glycine betaine", "Homarine", "Trigonelline",
-                   "Choline", "Glycerophosphocholine", "Homoserine betaine?", 
-                   "Threonine betaine?", "Proline betaine",
-                   "(3-Carboxypropyl)trimethylammonium", "O-Acetylcarnitine", 
-                   "Betonicine", "Acetylcholine", "Phosphocholine")
-non_n_names <- c("Dimethylsulfonioacetate", "Dimethylsulfoniopropionate",
-                 "Arsenobetaine", "Dimethylsulfoxonium propionate?", "Gonyol",
-                 "2-O-alpha-D-Glucosylglycerol")
-aa_names <- c("Glutamine", "Aspartic acid", "Glutamic acid/beta-Glutamic acid",
-              "Asparagine", "Alanine", "Methionine", "Arginine", "Proline",
-              "Lysine", "Threonine/Homoserine", "Serine", "Glycine", "Histidine",
-              "Leucine/Isoleucine", "Valine", "Arginine")
-bad_cmpds <- c("Melamine", "Uridine", "Itaconic acid?", "Inosine", "Dexpanthenol", 
-               "Choline sulfate", "5-Hydroxyectoine", "L-Tyrosine")
-class_levels <- c("Glutamine", "Glutamic acid", "Aspartic acid", "Asparagine", "Arginine", 
-                  "Other amino acids", "Theanine?", "Citrulline", "Ectoine", "Guanine", 
-                  "Guanosine", "Other nucleic acids", "MAAs", "Betaines", "All others")
-aa_ncs <- all_stans %>%
-  filter(compound_type=="Amino Acid") %>%
-  mutate(compound_name=str_remove(compound_name, "^L-")) %>%
-  distinct(compound_name, n_C, n_N) %>%
-  add_row(compound_name="Phenylalanine", n_C=9, n_N=1)
+## Figure gallery
 
-peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv") %>%
-  left_join(parametadata %>% distinct(filename, shortname))
-samp_areas <- parametadata %>%
-  filter(samp_type=="Smp") %>%
-  distinct(shortname, depth) %>%
-  left_join(peak_areas) %>%
-  mutate(shortname=fct_inorder(shortname)) %>%
-  filter(!compound_name%in%bad_cmpds) %>%
-  filter(!(depth=="Deep" & compound_name%in%maa_names)) %>%
-  select(-depth)
-labeled_fracs <- samp_areas %>%
-  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
-  select(compound_name, iso_name, shortname, labeled_frac) %>%
-  filter(str_detect(iso_name, "13C0, 15N0")) %>%
-  select(-iso_name)
+![](manuscript/figures/xamend_metab_gp.png)
 
-env_concs <- read_csv("targeted/final_env_concs.csv") %>%
-  select(-conc_um_in_vial) %>%
-  left_join(parametadata %>% distinct(filename, shortname)) %>%
-  select(-filename)
-nM_labeled <- samp_areas %>%
-  left_join(env_concs) %>%
-  mutate(labeled_frac=area/sum(area), .by = c("shortname", "compound_name")) %>%
-  mutate(rel_label=labeled_frac/labeled_frac[str_detect(iso_name, "13C0, 15N0")], .by = c("shortname", "compound_name")) %>%
-  mutate(nM_labeled=conc_nm_in_env*rel_label) %>%
-  select(filename, compound_name, iso_name, nM_labeled)
+Figure 1: Fraction of each metabolite or metabolite group containing at least 
+one ^15^N label during the surface seawater amendments with various substrates 
+after the number of hours shown on the x-axis. Darker colors indicate a larger 
+percentage of the compound signal was isotopically labeled. Three biological 
+replicates are stacked vertically at each timepoint and the categories with 
+multiple entries have been averaged. MAA = mycosporine-like amino acids, 
+GMP = guanosine monophosphate.
 
-cleaned_combined_peaks <- readRDS("untargeted/cleaned_combined_peaks.rds")
+![](manuscript/figures/total_thaa_gp.png)
 
-thaa_quant <- read_csv("thaas/quant_data.csv") %>%
-  filter(iso_name!="Proline, 13C3, 15N1")
-metathaa <- read_csv("thaas/metathaa.csv") %>%
-  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
-  mutate(startime=factor(startime, levels=c("Morn", "Eve")))
-```
+Figure 2: Stacked bar plots showing the fraction of various total hydrolyzable 
+amino acids (THAAs) labeled at the end of the incubations across the four 
+amendments. Total uM nitrogen (N) in the protein pool is shown in the top plot 
+while the fraction of individual amino acids labeled is shown in the rows below. 
+Colors denote specific labeling patterns detected in the THAAs.
 
-# Abstract
+<div style="display: flex; align-items: center;">
+  <img src="manuscript/figures/MAA_curves.png" style="width: 50%; margin-right: 10px;" />
+  <div style="text-align: center; flex: 1;">
+    <p>Figure 3: Fraction of UV-reactive mycosporine-like amino acids (MAAs) containing at least one ^15^N label over one day during the treatments amended with ^15^N labeled ammonia or nitrate.
+Incubations started at 6AM are shown in yellow while incubations started at 6PM are shown in blue, with grey background regions highlighting nighttime.
+Three replicates are included at each of the 5 timepoints and best-fit power law curves have been fit behind the data.</p>
+  </div>
+</div>
 
-Nitrogen availability controls productivity in much of the surface ocean and the form of this nitrogen controls who it is available to and how it is used.
-Here, we explored how four different forms of nitrogen (ammonia, nitrate, arginine, and guanosine monophosphate) were taken up and transformed in the North Pacific Subtropical Gyre using stable isotope labeling metabolomics.
-We repeated these experiments both at the surface and at the base of the euphotic zone as well as in the morning and evening.
-We found that ammonia is converted into glutamate in less than 20 minutes and results in labeling throughout the metabolome, though osmolytes like betaines had slow turnover rates.
-This process was slower at depth but largely independent of time of day.
-Nitrate amendments resulted in the labeling of only a small subset of molecules such as mycosporine-like amino acids, whose incorporation patterns suggest that this form of nitrogen was available to a subset of the community, likely eukaryotes.
-The two organic nitrogen amendments labeled the metabolome more slowly than the inorganic forms of nitrogen and showed similar usage at depth and at the surface, hinting at the major role that heterotrophic bacteria likely play in their uptake and use.
-A large fraction of the organic nitrogen added was remineralized into ammonia and used as glutamate rather than as intact substrates, though labeled arginine was incorporated into proteins directly.
-These measurements reveal the forms and fluxes of organic nitrogen in the natural environment to provide constraints on biogeochemical models and better characterize the transformations between organic and inorganic nitrogen.
+<div clear="left"></div>
 
-# Introduction
+![](manuscript/figures/deep_vs_surf_labeling.png)
 
-Marine carbon is fixed into biomass very quickly.
-In most of the surface ocean this flux is controlled by the availability of nitrogen, an element with a complex and incompletely characterized biogeochemical cycle [@Capone2008; @Moore2013; @Hutchins2022].
-Organic nitrogen in particular is often treated as a black box despite substantial variability in its bioavailability and chemical nature.
-While extensive work has been done to characterize the forms and fluxes of nitrogen within the inorganic pool, the equivalent for the organic pool is woefully underdeveloped [@ref, prob NME2008 or Hansell2024, maybe @Hutchins2022 for visual?].
-This is largely due to the difficulty of comprehensively measuring the many organic molecules that contain nitrogen in the marine environment [@Boysen2018; @Moran2022].
-The bioavailability of the nitrogen atom is in the environment is largely determined by the form of the organic material it composes.
-
-The mechanisms by which dissolved nitrogen becomes biomass and vice versa are major factors in our ability to predict ecosystem productivity and therefore carbon fixation and export.
-Traditionally, the conversion of inorganic material into organic substrate was thought to be limited to autotrophs while remineralization was performed by the heterotrophic community.
-Today, it’s known that many photosynthetic organisms are able to take up and use organic nitrogen [@Anita1991; @Morando2018; @Hugo2021] and that only a select subset of phototrophs are capable of nitrate reduction or nitrogen fixation [@ref], while inorganic nitrogen is directly accessible to heterotrophs as well.
-This means that the transformations possible for a given N-containing compound are a function of the microbial community and its environment while the microbial community is in turn also expected to be a function of the N-containing molecules available.
-This creates a recursive network that’s difficult to untangle and cannot necessarily be extrapolated from axenic culture studies.
-
-One way to reveal the pathways and transformations that marine N undergoes in a natural community is via 15N labeling paired with metabolomics.
-Metabolites that incorporate the labeled nitrogen can then be separated on the mass spectrometer and quantified separately from the unlabeled pool.
-Here, we trace the uptake and use of labeled ammonia, nitrate, arginine, and guanosine monophosphate in a natural community from the NPSG at multiple depths and diel conditions.
-Our application of metabolomics to these samples allows us to map out the pathways and restructuring that organic nitrogen experiences in the largest biomes on the planet, quantifying the pool sizes and turnover rates for various low molecular weight compounds that serve as building blocks and intermediates of cell biology.
-
-## Other things to maybe discuss:
-
-  -	mention the expected differences with depth and diel effects
-  -	detail which compounds we chose and why?
-  -	expand on DON-PON interactions and importance?
-  - Extracellular enzymes, organic use as ammonia vs as-is?
-  -	expand on the importance of biogeochemical models and why we need pathways mapped out?
-  -	mention basic biosynthetic pathways (maybe better suited to discussion?)
-  -	new (nitrate) vs regenerated (ammonia and DON) production (maybe also discussion?)
-
-## Other soundbites:
-  - It is assumed that most fixed nitrogen becomes protein, presumably based on the large fraction of cell nitrogen in protein and previous research showing how quickly it enters the (free?) amino acid pool.
-However, 1) important intermediates molecules 2) doesn’t explain where biounavailable N comes from 3) makes gene-based molecular modeling really difficult.
-  - In each case, assimilation into organic matter is typically done via the glutamine synthetase (GS)/glutamate synthase (GOGAT) or glutamate dehydrogenase (GDH) pathways to produce glutamate from α-ketoglutarate [@ref, see bronk chapter].
-Glutamate then serves as a nitrogen supply in an enormous number of biochemical pathways in both primary and secondary metabolism [@Walker2016].
-  - nitrate amendments typically resulting in diatom blooms while ammonia incubations favor cyanobacteria [@Glibert2016].
-  - Communities to distinguish are 1) surface (largely regenerated production) and 2) 175m (largely new production(??))
-  - use of organics vs synthesis of organics
-  - Compounds such as non-proteinogenic amino acids (e.g., ornithine, citrulline, creatine, MAAs), osmolytes like betaines, nucleobases, and sulfur-containing molecules like taurine represent significant but poorly quantified pools.
-Understanding the rates and pathways by which these compounds are utilized is crucial, as they form the fundamental unit linking genetic potential to elemental cycling in biogeochemical models.
-  - Nitrogen use varies significantly between the sunlit surface where productivity is largely regenerated through recycling of ammonia and organic nitrogen, and deeper waters near the base of the euphotic zone, where nitrate supports “new” production.
-Quantifying rates of nitrogen assimilation into ammonia and downstream products across these gradients is therefore essential for predicting marine productivity and nutrient cycling.
-  - Nonetheless, how the community uses a substrate is arguably more important than whether it can.
-
-# Results
-
-`r wc("Do I need a 'here's what the cruise looked like' results section?")`
-
-## Differences in nitrogen labeling at the surface between amendments
-
-```{r gln numbers}
-labeled_fracs %>%
-  filter(compound_name=="L-Glutamine") %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(timepoint%in%c("T0", "T26")) %>%
-  filter(depth=="Surf") %>%
-  filter(amendment=="Amm") %>%
-  summarise(frac=mean(labeled_frac), .by=c(compound_name, timepoint))
-peak_areas %>%
-  filter(compound_name=="L-Glutamine") %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(depth=="Surf") %>%
-  filter(amendment=="Amm") %>%
-  ggplot(aes(x=tripl, y=area, fill=iso_name)) +
-  geom_col(color="black") +
-  facet_grid(startime~timepoint)
-nM_labeled %>%
-  filter(compound_name%in%c("L-Glutamine")) %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(depth=="Surf") %>%
-  filter(amendment=="Amm") %>%
-  summarise(conc=mean(nM_labeled), .by=c(compound_name, timepoint)) %>%
-  arrange(timepoint)
-```
-
-We observed significant differences in ^15^N labeling patterns between the different treatment incubations performed with ALOHA surface seawater (Figure 1).
-Ammonia incubations resulted in very rapid labeling, with glutamine more than 75% labeled even at the T0 timepoint, corresponding to the amount of time required to spike and immediately filter the 2 liters of seawater (~15 minutes).
-The dominant labeling pattern at T0 was a single ^15^N, though all later timepoints showed more doubly-^15^N labeled glutamine than singly-labeled.
-After 24 hours, glutamine was more than 95% labeled and had approximately doubled in pool size (initially ~0.3 nM becoming 0.8 nM).
-
-```{r fig1 grouped metab heatmap}
-xamend_metab_gp <- labeled_fracs %>%
-  left_join(parametadata %>% distinct(shortname, samp_type, timepoint, timepoint_num, depth, startime, amendment, tripl)) %>%
-  filter(samp_type!="Bad") %>%
-  filter(depth=="Surf") %>%
-  mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
-  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
-  mutate(compound_class=case_when(
-    compound_name%in%c("Glutamine", "Aspartic acid", "Asparagine", "Glutamic acid", "Ectoine", "Theanine?", "Citrulline", "Guanine", "Guanosine", "Arginine") ~ compound_name,
-    compound_name%in%maa_names~"MAAs",
-    compound_name%in%aa_names~"Other amino acids",
-    compound_name%in%nucleo_names~"Other nucleic acids",
-    compound_name%in%betaine_names~"Betaines",
-    TRUE~"All others"
-  )) %>%
-  mutate(compound_class=factor(compound_class, levels=class_levels)) %>%
-  summarise(avg_frac=mean(labeled_frac), .by = c(compound_class, amendment, timepoint, tripl)) %>%
-  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"),
-                          labels=c("+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
-  ggplot() +
-  geom_raster(aes(x=timepoint, y=tripl, fill=avg_frac)) +
-  geom_vline(xintercept = c(1.5:5.5), color="grey80") +
-  facet_nested(compound_class~amendment, scales = "free_x", switch="y") +
-  scale_fill_gradientn(colours = viridis::viridis(10, alpha = seq(0, 1, length.out=10), direction = -1),
-                       limits=c(0, 1), breaks=c(0, 0.25, 0.5, 0.75, 1),
-                       labels=c("Fully unlabeled", "", "50% labeled", "", "Fully labeled")) +
-  scale_x_discrete(expand = expansion()) +
-  scale_y_discrete(expand = expansion()) +
-  labs(x=NULL, y=NULL, fill=NULL) +
-  theme_bw() +
-  theme(panel.spacing.x=unit(0.5, "lines"), 
-        panel.spacing.y = unit(0, "lines"),
-        axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        strip.text.y.left = element_text(angle=0, hjust = 1),
-        strip.background.y = element_rect(fill=NA, color=NA),
-        panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank(),
-        legend.position = "top", 
-        legend.key.width = unit(1, "null")) +
-  guides(
-    fill=guide_colorbar(theme=theme(
-      legend.frame = element_rect(color = "black"),
-      legend.ticks = element_line(color="black"),
-      legend.text = ggtext::element_markdown(hjust=c(0, 0, 0.5, 1, 1))),
-      text=element_text(size=12)
-    )
-  )
-ggsave("xamend_metab_gp.png",plot = xamend_metab_gp, device = "png", path = "manuscript/figures",
-       width = 6.5, height = 4, units = "in", dpi = 300)
-```
-
-![](figures/xamend_metab_gp.png)
-
-*Figure 1: Fraction of each metabolite or metabolite group containing at least one ^15^N label during the surface seawater amendments with various substrates after the number of hours shown on the x-axis.
-Darker colors indicate a larger percentage of the compound signal was isotopically labeled.
-Three biological replicates are stacked vertically at each timepoint and the categories with multiple entries have been averaged.
-MAA = mycosporine-like amino acids, GMP = guanosine monophosphate.*
-
-```{r gluaspcit numbers}
-labeled_fracs %>%
-  filter(compound_name%in%c("L-Glutamic acid/beta-Glutamic acid", "L-Aspartic acid",
-                            "Citrulline", "L-Arginine", "Theanine?")) %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(timepoint=="T1") %>%
-  filter(depth=="Surf") %>%
-  filter(amendment=="Amm") %>%
-  summarise(frac=mean(labeled_frac), .by=c(compound_name))
-env_concs %>%
-  filter(compound_name%in%c("L-Glutamic acid/beta-Glutamic acid", "L-Aspartic acid",
-                            "Citrulline", "L-Arginine")) %>%
-  left_join(parametadata %>% filter(IS_type=="wIS")) %>%
-  filter(timepoint=="T1") %>%
-  filter(depth=="Surf") %>%
-  filter(amendment=="Amm") %>%
-  summarise(conc=mean(conc_nm_in_env), .by=c(compound_name))
-```
-
-```{r theanine annotation check}
-# Check theanine annotation and numbers
-cleaned_combined_peaks %>% 
-  filter(mzmed%between%pmppm(175.1077, 10)) %>%
-  left_join(parametadata) %>%
-  filter(amendment=="Amm") %>%
-  filter(depth=="Surf") %>%
-  summarise(mzmed=unique(mzmed), rtmed=unique(rtmed)/60, into=mean(into),
-            beta_snr=median(beta_snr), beta_cor=median(beta_cor), 
-            .by=feature)
-arrow::open_dataset("tmzMLs/pqds") %>%
-  filter(amendment=="Amm") %>%
-  filter(samp_type=="Smp") %>%
-  filter(polarity=="pos") %>%
-  filter(mz%between%pmppm(175.1077, 10) | mz%between%pmppm(176.1047, 10)) %>%
-  filter(rt%between%c(7.2, 8.5)) %>%
-  dplyr::collect() %>%
-  mutate(mz=round(mz)) %>%
-  slice_max(int, by=c(filename, mz, rt)) %>%
-  arrange(rt) %>%
-  summarise(area=trapz(rt, int), .by=c(filename, mz)) %>%
-  pivot_wider(names_from=mz, values_from=area) %>%
-  mutate(labeled_frac=`176`/(`175`+`176`)) %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(depth=="Surf") %>%
-  ggplot() +
-  geom_boxplot(aes(x=timepoint, y=labeled_frac, color=startime))
-```
-
-The ^15^N was then quickly propagated to glutamate and aspartate, creating a metabolite pool 40-50% labeled for each of these after a single hour.
-Aspartate was labeled slightly faster than glutamate, likely as a result of its smaller pool size (0.2 nM Asp versus 0.5 nM Glu).
-Also rapidly labeled was citrulline, reaching 20% labeled after a single hour, though arginine was not labeled until much later despite its location immediately downstream of this metabolite.
-Untargeted detection of other molecular features revealed an additional metabolite that was 30% labeled at the one hour timepoint with an *m/z* value of 175.1077 and a retention time of 8.2 minutes which we have putatively identified as theanine (N5-ethyl glutamine) based on *m/z* match and labeling pattern.
-
-```{r betaine and nucleo numbers}
-labeled_fracs %>%
-  filter(compound_name%in%nucleo_names) %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(timepoint=="T26") %>%
-  filter(amendment=="Amm") %>%
-  summarise(frac=mean(labeled_frac), .by=c(compound_name)) %>%
-  mutate(nuc_type=case_when(
-    compound_name%in%c("Cytosine", "Adenine", "Guanine", "Hypoxanthine/Allopurinol")~"Nucleobase",
-    compound_name%in%c("Cytidine", "Adenosine", "Guanosine")~"Nucleoside",
-    compound_name%in%c("Deoxycytidine?", "Deoxyadenosine", "Deoxyguanosine?")~"Deoxy"
-  )) %>%
-  mutate(nuc_shortname=str_extract(tolower(compound_name), tolower("Cy|Ad|Gu|Hy"))) %>%
-  select(nuc_shortname, nuc_type, frac) %>%
-  pivot_wider(names_from=nuc_type, values_from=frac) %>%
-  select(nuc_shortname, Nucleoside, Nucleobase, Deoxy)
-
-labeled_fracs %>%
-  filter(compound_name%in%betaine_names) %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(timepoint=="T26") %>%
-  filter(amendment=="Amm") %>%
-  summarise(frac=mean(labeled_frac), .by=c(compound_name))
-
-peak_areas %>%
-  filter(compound_name%in%nucleo_names) %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(timepoint=="T26") %>%
-  filter(amendment=="Amm") %>%
-  filter(depth=="Surf") %>%
-  mutate(iso_name=str_remove(iso_name, ".*?, ")) %>%
-  filter(iso_name!="13C0, 15N0") %>%
-  ggplot() +
-  geom_col(aes(x=tripl, y=area, fill=iso_name), position="fill", color="black") +
-  facet_wrap(~compound_name)
-```
-
-Other amino acids and nucleobases/nucleosides had trace amounts of labeling after 3 hours and were extensively labeled (>25%) after 24 hours (Figure 1).
-Cytidine (42.2% labeled) and adenosine (59.9%) were more heavily ^15^N labeled than their nucleobase constituents at T26 (22.9% and 42.1%, respectively), corresponding to the expected synthesis pathway.
-Guanine, however, had the inverse trend and at 44.9% labeled was 15% more labeled than guanosine (Supplemental figure XX).
-The deoxyribose forms were in all cases less completely labeled than their ribose counterparts (deoxycytidine = 18.9%, deoxyadenosine = 20.3%, and deoxyguanosine = 17.2%) `r wc("Maybe better as a small table?")` `r wc("Consider trying harder to get thymine/uracil/inosine values?")`.
-In each case, the fully ^15^N labeled nucleobase was the dominant constituent of the isotope envelope (^15^N~3~ for cytosine and ^15^N~5~ for adenine and guanine).
-Other compounds, including betaines (here restricted to quaternary amines with a carboxylic acid group) and choline derivatives showed very little labeling at all, with only homarine (15.0% labeled) and O-acetylcarnitine (9.9% labeled) exceeding 5% labeling after 24 hours`r wc("Expand betaine discussion and turn it into a separate paragraph?
-Not sure what more to say...
-fluxes?")`.
-
-Amendments with ^15^NO~3~ showed a significantly different labeling pattern with very few metabolites incorporating the ^15^N label at the surface even after 24 hours.
-Glutamine and glutamate, which were labeled within 20 minutes from ammonia and are the main way in which inorganic nitrogen becomes organic, here saw only trace amounts of production from the added nitrate even after 24 hours.
-The most heavily labeled compound was guanine, which saw trace amounts of labeling at the T10 timepoint and reached ~40% labeling at T26.
-This labeling was not reflected in the other nucleobases.
-The only group of compounds that did see compelling degrees of label were the mycosporine-like amino acids (MAAs), a group of compounds produced under UV exposure (see below).
-These compounds, like guanine, were labeled in excess of the bulk glutamate/amino acid pool from which their nitrogen is sourced.
-This is only possible if the nitrate is assimilated by a subset of the community that is able to reduce it and prevents it from entering the environment.
-Metabolites specific to those organisms can then be made from the small pool of heavily-labeled organic material without affecting the fraction of glutamate or glutamine labeled in bulk.
-
-```{r arg numbers}
-labeled_fracs %>%
-  filter(compound_name=="L-Arginine") %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(depth=="Surf") %>%
-  filter(amendment=="Arg") %>%
-  summarise(frac=mean(labeled_frac), .by=c(compound_name, timepoint)) %>%
-  arrange(timepoint)
-```
-
-The organic nitrogen amendments also showed labeling patterns distinct from the ammonia incubations, though a large fraction of the nitrogen was clearly remineralized into ammonia/glutamate and used to label a similar suite of compounds (Figure 1).
-Fully C and N labeled (^13^C~6~, ^15^N~4~) arginine was visibly taken up at the T0 timepoint (within 20 minutes) and 80-90% of the pool was fully labeled after 12 hours.
-Surprisingly, the fraction of arginine labeled decreased from this point, with the T26 and T73 timepoints dropping to 73% and 46% labeled, respectively.
-This labeling rapidly propagated into citrulline via the urea cycle, indicating that labeled urea (^13^C~1~, ^15^N~2~) was created and available in the system as well though we were unable to measure this directly `r wc("I could try harder to get urea values for this but it's pretty bad.")`.
-Glutamate and glutamine were labeled much more slowly, reaching 10% labeled after 10 hours and a maximum of 66% labeled after 24 hours.
-Much of this nitrogen was then cycled very quickly among other compounds which also saw a jump in labeling at 24 hours.
-This was particularly true for other amino acids and the nucleobases/nucleosides and here, guanine was not labeled more than other nucleobases.
-Mycosporine-like amino acids and betaines were unlabeled above background even after 3 days of incubation.
-
-```{r gmp numbers}
-labeled_fracs %>%
-  filter(compound_name%in%c("Guanosine", "Guanine")) %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(depth=="Surf") %>%
-  filter(amendment=="GMP") %>%
-  summarise(frac=mean(labeled_frac), .by=c(compound_name, timepoint)) %>%
-  pivot_wider(names_from=compound_name, values_from=frac) %>%
-  arrange(timepoint)
-peak_areas %>%
-  filter(compound_name%in%c("Guanosine", "Guanine", "Adenine", "Adenosine", "Hypoxanthine/Allopurinol")) %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(amendment=="GMP") %>%
-  filter(depth=="Surf") %>%
-  mutate(iso_name=str_remove(iso_name, ".*?, ")) %>%
-  # filter(iso_name!="13C0, 15N0") %>%
-  ggplot() +
-  geom_col(aes(x=tripl, y=area, fill=iso_name), position="fill", color="black") +
-  facet_grid(compound_name~timepoint)
-```
-
-The guanosine monophosphate (GMP) incubations showed a pattern very similar to the arginine incubations in the identity, timing, and magnitude in which metabolites were labeled (Figure 1).
-While we were unable to measure GMP directly `r wc("I can try harder here but the data is pretty bad :(")` we saw clear labeling of fully C and N labeled guanosine (^13^C~10~, ^15^N~5~) after 3 hours that increased to a maximum of 75% after 24 hours before decreasing slightly to 64% labeled at the end of the three-day incubation.
-A small part of this signal was reflected in guanine, with ^13^C~5~, ^15^N~5~ guanine representing about 10% of the signal at T3 and reaching a maximum of 40% at T26.
-At T26 we also saw extensive labeling in adenine, adenosine, and hypoxanthine though their dominant labeling patterns (^13^C~0~ ^15^N~5~, ^13^C~5~ ^15^N~5~, and ^13^C~0~ ^15^N~4~ respectively, Supplemental figure XX) suggested that the adenine and hypoxanthine were synthesized from glutamate and that the ribose pool had been swamped by guanosine cleavage.
-Here, unlike the arginine incubations, we also saw some MAA labeling at the end of the 3 day incubation.
-
-```{r thaa tables}
-thaa_avgs <- thaa_quant %>%
-  left_join(metathaa) %>%
-  filter(samp_type=="Smp") %>%
-  filter(depth=="Surf") %>%
-  select(compound_name, iso_name, amendment, timepoint, depth, startime, env_conc_nm) %>%
-  summarise(thaa_nM=mean(env_conc_nm), .by = c(compound_name, iso_name, amendment, timepoint)) %>%
-  mutate(compound_name=case_when(
-    compound_name%in%c("Aspartic acid")~"Aspartate + asparagine",
-    compound_name%in%c("Glutamic acid")~"Glutamate + glutamine",
-    compound_name%in%c("Leucine", "Isoleucine")~"Leucine + isoleucine",
-    TRUE~compound_name
-  )) %>%
-  mutate(iso_name=str_remove(iso_name, ".*(?=13)")) %>%
-  summarise(thaa_nM=sum(thaa_nM), .by=c(compound_name, iso_name, amendment, timepoint))
-
-metab_avgs <- nM_labeled %>%
-  mutate(compound_name=str_remove_all(compound_name, "L-")) %>%
-  mutate(iso_name=str_remove(iso_name, ".*(?=13)")) %>%
-  filter(compound_name%in%aa_names) %>%
-  filter(!compound_name%in%c("Methionine", "Valine", "Lysine")) %>%
-  mutate(compound_name=case_when(
-    compound_name%in%c("Asparagine", "Aspartic acid")~"Aspartate + asparagine",
-    compound_name%in%c("Glutamine", "Glutamic acid/beta-Glutamic acid")~"Glutamate + glutamine",
-    compound_name=="Threonine/Homoserine"~"Threonine",
-    compound_name=="Leucine/Isoleucine"~"Leucine + isoleucine",
-    TRUE~compound_name
-  )) %>%
-  summarise(nM_labeled=sum(nM_labeled), .by = c(compound_name, iso_name, filename)) %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  mutate(final_timepoint=ifelse(amendment%in%c("Amm", "Nit"), "T26", "T73")) %>%
-  filter(timepoint%in%c("T0", final_timepoint)) %>%
-  select(compound_name, iso_name, amendment, timepoint, depth, startime, nM_labeled) %>%
-  summarise(metab_nM=mean(nM_labeled), .by = c(compound_name, iso_name, amendment, timepoint))
-
-# Fraction of THAA that's DFAA at T0 and T26
-left_join(metab_avgs, thaa_avgs) %>%
-  summarise(metab_nM=sum(metab_nM, na.rm=TRUE), 
-            thaa_nM=sum(thaa_nM, na.rm=TRUE),
-            .by=c(compound_name, timepoint, amendment)) %>%
-  filter(!(timepoint=="T26" & amendment%in%c("Arg", "GMP"))) %>%
-  mutate(perc_in_free=metab_nM/(metab_nM+thaa_nM)*100) %>%
-  select(amendment, compound_name, timepoint, perc_in_free) %>%
-  pivot_wider(names_from=timepoint, values_from=perc_in_free) %>%
-  arrange(amendment, compound_name)
-
-# Breakdown of THAA isotope labeling
-left_join(metab_avgs, thaa_avgs) %>%
-  mutate(perc_in_free=metab_nM/(metab_nM+thaa_nM)*100) %>%
-  arrange(amendment, compound_name) %>%
-  filter(perc_in_free<100) %>%
-  arrange(timepoint, compound_name)
-```
-
-We also explored the degree of labeling incorporation into proteins by measuring the total hydrolyzable amino acids (THAAs) at both the start of each incubation and the end (Figure 2).
-Proteins were 30-50% of total particulate carbon pool (~2 uM) and 70-100% of the total particulate nitrogen (~0.25 uM), with free amino acids composing between 0.15% and 17% of the total`r wc("I'm not really sure what to say here or how to dive into the variance that's of interest to us.
-Report the compound with the most/least in protein?
-Break it down by labeled vs unlabeled?")``.
-
-There were differences in THAA labeling patterns between the amendments (Figure 2).
-Despite their shorter incubation time (26 hours vs 73 hours), the inorganic nitrogen amendments generally higher fractions of labeled protein.
-The nitrate amendments in particular were surprisingly heavily labeled given the minimal signal seen in the free metabolites, with a degree of labeling exceeding the ammonia incubation.
-This indicates that most of the assimilation was done by the small subset of organisms able to reduce nitrate and use it for growth and that these organisms did not share the reduced nitrogen with the rest of the community`r wc("Is this something for the discussion instead?")`.
-
-```{r fig2 thaa data}
-all_thaa_plot <- thaa_quant %>%
-  filter(!str_detect(iso_name, "2H")) %>%
-  left_join(metathaa) %>%
-  filter(samp_type=="Smp") %>%
-  left_join(aa_ncs) %>%
-  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
-  mutate(n_N_lab=as.numeric(str_extract(iso_name, "(?<=15N)\\d+"))) %>%
-  mutate(n_N_unlab=n_N-n_N_lab) %>%
-  mutate(nM_N_lab=env_conc_nm*n_N_lab, nM_N_unlab=env_conc_nm*n_N_unlab) %>%
-  select(filename, compound_name, iso_name, nM_N_lab, nM_N_unlab) %>%
-  pivot_longer(c(nM_N_lab, nM_N_unlab)) %>%
-  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
-  left_join(metathaa) %>%
-  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
-  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"),
-                          labels=c("Initial (T0)", "+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
-  summarise(total_uM=mean(total_uM), .by=c(amendment, timepoint, name)) %>%
-  mutate(name=factor(name, levels=c("nM_N_unlab", "nM_N_lab"), labels = c("Unlabeled", "Labeled"))) %>%
-  ggplot(aes(x=amendment, y=total_uM, fill=name)) +
-  geom_col(color="black", width=0.9) +
-  theme_bw() +
-  # scale_x_discrete(expand = expansion(mult = c(0.05))) +
-  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
-  scale_fill_manual(breaks = c("Unlabeled", "Labeled"), values=c("grey80", "grey20")) +
-  labs(x=NULL, y="Total uM N in THAAs", fill=NULL)
-
-plot_iso_levels <- c(
-  "Unlabeled", 
-  "<sup>15</sup>N<sub>1</sub>", 
-  "<sup>15</sup>N<sub>2</sub>", 
-  "<sup>15</sup>N<sub>3</sub>", 
-  "Entirely <sup>15</sup>N",
-  "<sup>13</sup>C<sub>1</sub>", 
-  "<sup>13</sup>C<sub>1</sub><sup>15</sup>N<sub>1</sub>",
-  "<sup>13</sup>C<sub>4</sub><sup>15</sup>N<sub>1</sub>", 
-  "<sup>13</sup>C<sub>5</sub><sup>15</sup>N<sub>3</sub>",
-  "Entirely <sup>13</sup>C<br>and <sup>15</sup>N", 
-  "Other labels")
-plot_iso_colors <- c(
-  "grey80",
-  "#e0ecf6",
-  "#97bfe0",
-  "#377EB8",
-  "#285b84",
-  "#ffb4b5",
-  "#d7b6dc",
-  "#FFFF33",
-  "#4DAF4A",
-  "#984EA3",
-  "grey20"
-)
-thaa_spec_cmpds <- c("Glutamine + Glutamate", "Arginine", "Proline", "Histidine", "Phenylalanine", "Tyrosine")
-thaa_plot_data <- thaa_quant %>%
-  left_join(metathaa) %>%
-  filter(samp_type=="Smp") %>%
-  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
-  summarise(avg_env_conc_nm=mean(env_conc_nm, na.rm=TRUE), .by=c(compound_name, iso_name, amendment, timepoint)) %>%
-  mutate(compound_name=ifelse(compound_name=="Glutamic acid", "Glutamine + Glutamate", compound_name)) %>%
-  mutate(compound_name=factor(compound_name, levels = thaa_spec_cmpds)) %>%
-  filter(!is.na(compound_name)) %>%
-  mutate(plot_iso_name=case_when(
-    str_detect(iso_name, "13C0, 15N0")~"Unlabeled",
-    str_detect(iso_name, "13C0, 15N1")~"<sup>15</sup>N<sub>1</sub>",
-    str_detect(iso_name, "13C0, 15N2")~"<sup>15</sup>N<sub>2</sub>",
-    str_detect(iso_name, "Histidine, 13C0, 15N3")~"Entirely <sup>15</sup>N",
-    str_detect(iso_name, "13C0, 15N3")~"<sup>15</sup>N<sub>3</sub>",
-    str_detect(iso_name, "Arginine, 13C0, 15N4")~"Entirely <sup>15</sup>N",
-    str_detect(iso_name, "13C1, 15N0")~"<sup>13</sup>C<sub>1</sub>",
-    str_detect(iso_name, "13C1, 15N1")~"<sup>13</sup>C<sub>1</sub><sup>15</sup>N<sub>1</sub>",
-    str_detect(iso_name, "Glutamic acid, 13C5, 15N1")~"Entirely <sup>13</sup>C<br>and <sup>15</sup>N",
-    str_detect(iso_name, "Proline, 13C5, 15N1")~"Entirely <sup>13</sup>C<br>and <sup>15</sup>N",
-    str_detect(iso_name, "Arginine, 13C6, 15N4")~"Entirely <sup>13</sup>C<br>and <sup>15</sup>N",
-    iso_name=="Tyrosine, 13C4, 15N1"~"<sup>13</sup>C<sub>4</sub><sup>15</sup>N<sub>1</sub>",
-    iso_name=="Phenylalanine, 13C4, 15N1"~"<sup>13</sup>C<sub>4</sub><sup>15</sup>N<sub>1</sub>",
-    iso_name=="Histidine, 13C5, 15N3"~"<sup>13</sup>C<sub>5</sub><sup>15</sup>N<sub>3</sub>",
-    TRUE~"Other labels"
-  )) %>%
-  mutate(plot_iso_name=factor(plot_iso_name, levels=plot_iso_levels)) %>%
-  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP")))
-arg_glut_thaa_plot <- ggplot(thaa_plot_data) +
-  geom_col(aes(x=amendment, y=avg_env_conc_nm, fill=plot_iso_name, label=iso_name),
-           position="fill",
-           color="black", width=1) +
-  facet_wrap(~compound_name) +
-  theme_bw() +
-  scale_x_discrete(expand = expansion()) +
-  scale_y_continuous(labels=scales::label_percent(), breaks=c(0, 0.5, 1),
-                     expand = expansion()) +
-  scale_fill_manual(breaks = plot_iso_levels, values = plot_iso_colors) +
-  labs(x=NULL, y="Fraction of THAA labeled", fill=NULL) +
-  theme(
-    # axis.text.x=element_text(angle=90, hjust=1, vjust = 0.5),
-    legend.text = element_markdown()
-  )
-total_thaa_gp <- all_thaa_plot / arg_glut_thaa_plot + plot_layout(heights = c(0.3, 0.7))
-total_thaa_gp
-ggsave("total_thaa_gp.png", total_thaa_gp, device = "png", path = "manuscript/figures",
-       width = 6.5, height = 5, units = "in", dpi = 300)
-```
-
-![](figures/total_thaa_gp.png)
-
-*Figure 2: Stacked bar plots showing the fraction of various total hydrolyzable amino acids (THAAs) labeled at the end of the incubations across the four amendments.
-Total uM nitrogen (N) in the protein pool is shown in the top plot while the fraction of individual amino acids labeled is shown in the rows below.
-Colors denote specific labeling patterns detected in the THAAs.*
-
-We believe the organic nitrogen amendments largely resulted in metabolite labeling via remineralization to ammonia or glutamate based on the absence of a labeled carbon backbone in most of the labeled metabolites, but in the THAAs we were able to see some transformations that used the carbon backbone as-is.
-Proline and glutamate both had detectable ^13^C~5~ ^15^N~1~ labels in the arginine treatments.
-We also saw histidine ^13^C~5~ ^15^N~3~ synthesized during the guanosine monophosphate treatment, confirming again that the ribose pool must have been swamped with fully ^13^C labeled ribose cleaved from guanosine in order to enter the PRPP pathway.
-The ^15^N~3~ pattern also indicates that for these organisms, glutamate, glutamine, and adenine were also heavily labeled.
-Finally, both phenylalanine and tyrosine had significant ^13^C~4~ ^15^N~1~ labeling, likely from erythrose 4-phosphate via the pentose phosphate pathway indicating new synthesis through shikimate with nitrogen addition from a heavily labeled glutamate pool again.
-
-## Minimal diel signal observed except in the UV-reactive mycosporine-like amino acids
-
-```{r maa numbers}
-labeled_fracs %>%
-  filter(compound_name%in%chosen_maas) %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(timepoint%in%c("T0", "T10")) %>%
-  filter(depth=="Surf") %>%
-  filter(startime=="Morn") %>%
-  filter(amendment=="Amm") %>%
-  summarise(frac=mean(labeled_frac), .by=c(compound_name, timepoint))
-```
-
-We performed the four nitrogen amendments at both 6AM and at 6PM to test for any diel effects in labeling patterns.
-Very few compounds showed a significant difference between the 6AM start time and the 6PM start time with the exception being the mycosporine-like amino acids (MAAs, Figure 3).
-Shinorine and palythine were confirmed via MS^2^ fragments while the other annotations remain putative, though their presence exclusively in the surface is compelling.
-During the ammonia incubations, six `r wc("double-check this after rerun")` MAAs showed clear ^15^N incorporation while exposed to sunlight, with the labeled fraction increasing sharply over the first 10 hours in the incubations started at 6AM.
-Incubations started at 6PM showed very little labeling from ammonia until a full 24 hours had passed.
-Shinorine, mycosporine-glycine, palythine, porphyra-334 were all clearly labeled (>20% ^15^N) as well as two of three molecular features detected at an *m/z* of 285.1445 possibly corresponding to palythene and usujirene.
-
-```{r fig MAA labeling new}
-all_maa_data <- labeled_fracs %>%
-  filter(compound_name%in%maa_names) %>%
-  mutate(compound_name=factor(compound_name, levels=maa_names)) %>%
-  arrange(compound_name) %>%
-  mutate(compound_name=str_remove(compound_name, "\\?$")) %>%
-  mutate(compound_name=str_replace(compound_name, "Also Palythene/Usujirene", "Palythene or\n usujirene")) %>%
-  mutate(compound_name=str_replace(compound_name, "Palythene/Usujirene", "Usujirene or\n palythene")) %>%
-  mutate(compound_name=str_remove(compound_name, "Also ")) %>%
-  mutate(compound_name=str_replace(compound_name, "-glycine", "-\nglycine")) %>%
-  mutate(compound_name=fct_inorder(compound_name)) %>%
-  left_join(parametadata, relationship = "many-to-many") %>%
-  filter(samp_type=="Smp") %>%
-  filter(IS_type=="noIS") %>%
-  filter(amendment%in%c("Amm", "Nit", "Arg", "GMP")) %>%
-  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"), labels=c("+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
-  mutate(clock_time=ifelse(startime=="Morn", 6+timepoint_num, 18+timepoint_num)) %>%
-  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start")))
-all_maa_gp <- ggplot(all_maa_data) +
-  annotate("rect", xmin=-Inf, xmax = 6, ymin=-Inf, ymax=Inf, fill="grey80") +
-  annotate("rect", xmin=18, xmax = 30, ymin=-Inf, ymax=Inf, fill="grey80") +
-  annotate("rect", xmin=42, xmax = Inf, ymin=-Inf, ymax=Inf, fill="grey80") +
-  geom_hline(yintercept = c(0.25, 0.5, 0.75, 1), color="grey90") +
-  geom_smooth(aes(x=clock_time, y=labeled_frac, color=startime),
-              method=nlsLM, formula="y~a*x^b+d", se = FALSE,
-              method.args=list(start=list(a=1, b=2, d=1),
-                               control=nls.lm.control(maxiter = 500))) +
-  # geom_smooth(aes(x=clock_time, y=labeled_frac, group=startime),
-  #             method=glm, formula = "y~x", method.args=list(family="quasibinomial"), 
-  #             color="grey10", se=FALSE) +
-  # geom_point(aes(x=clock_time, y=labeled_frac, shape=startime), color="black", 
-  #            fill="white", size=3) +
-  geom_point(aes(x=clock_time, y=labeled_frac, shape=startime, fill=startime),
-             color="black", size=2, alpha=0.7) +
-  facet_nested(compound_name~amendment, scales="free_x") +
-  scale_x_continuous(breaks = c(6, 18, 30, 42), labels = c("6AM", "6PM", "6AM", "6PM")) +
-  scale_y_continuous(labels = scales::label_percent(),
-                     limits = c(0, 1), breaks = c(0, 0.5, 1)) +
-  scale_shape_manual(values=c(24, 25), breaks = c("6AM start", "6PM start")) +
-  scale_fill_manual(values=c("#f4bb23", "#0b505c"), breaks = c("6AM start", "6PM start"),
-                    aesthetics = c("color", "fill")) +
-  labs(x=NULL, color=NULL, shape=NULL, fill=NULL, y="Fraction of MAA containing at least one 15N atom") +
-  theme_bw() +
-  theme(axis.text.y = element_markdown(vjust=c(0, 0.5, 1)),
-        axis.title.x = element_blank(),
-        legend.position = "inside",
-        legend.margin = margin(0.1,0.25,0.1,0.1, unit="lines"),
-        legend.position.inside = c(1, 1),
-        legend.justification = c(1, 1), 
-        legend.background = element_rect(fill="#FFFFFFAA", color="black", linewidth = 0)) +
-  guides(shape=guide_legend(override.aes = list(size=3, stroke=1.2, alpha=1, linetype=0)))
-ggsave(plot = all_maa_gp, filename = "supp_MAA_curves.png", device = "png", 
-       path = "manuscript/supplement", width = 6.5, height = 9, units = "in", dpi = 300)
-
-chosen_maas <- c("Mycosporine-glycine?", "Shinorine?", "Palythine?",
-                 "Palythene/Usujirene?", "Also Palythene/Usujirene?", 
-                 "Also Porphyra-334?")
-chosen_maa_data <- labeled_fracs %>%
-  filter(compound_name%in%chosen_maas) %>%
-  mutate(compound_name=factor(compound_name, levels=chosen_maas)) %>%
-  arrange(compound_name) %>%
-  mutate(compound_name=str_remove(compound_name, "\\?$")) %>%
-  mutate(compound_name=str_replace(compound_name, "Also Palythene/Usujirene", "Palythene or\n usujirene")) %>%
-  mutate(compound_name=str_replace(compound_name, "Palythene/Usujirene", "Usujirene or\n palythene")) %>%
-  mutate(compound_name=str_remove(compound_name, "Also ")) %>%
-  mutate(compound_name=str_replace(compound_name, "-glycine", "-\nglycine")) %>%
-  mutate(compound_name=fct_inorder(compound_name)) %>%
-  left_join(parametadata, relationship = "many-to-many") %>%
-  filter(samp_type=="Smp") %>%
-  filter(IS_type=="noIS") %>%
-  filter(amendment%in%c("Amm", "Nit")) %>%
-  mutate(amendment=factor(amendment, levels=c("Amm", "Nit"), labels=c("+Ammonia", "+Nitrate"))) %>%
-  mutate(clock_time=ifelse(startime=="Morn", 6+timepoint_num, 18+timepoint_num)) %>%
-  mutate(startime=factor(startime, levels=c("Morn", "Eve"), labels=c("6AM start", "6PM start")))
-chosen_maa_gp <- ggplot(chosen_maa_data) +
-  annotate("rect", xmin=-Inf, xmax = 6, ymin=-Inf, ymax=Inf, fill="grey80") +
-  annotate("rect", xmin=18, xmax = 30, ymin=-Inf, ymax=Inf, fill="grey80") +
-  annotate("rect", xmin=42, xmax = Inf, ymin=-Inf, ymax=Inf, fill="grey80") +
-  geom_hline(yintercept = c(0.25, 0.5, 0.75, 1), color="grey90") +
-  geom_smooth(aes(x=clock_time, y=labeled_frac, color=startime),
-              method=nlsLM, formula="y~a*x^b+d", se = FALSE,
-              method.args=list(start=list(a=1, b=2, d=1),
-                               control=nls.lm.control(maxiter = 500))) +
-  geom_point(aes(x=clock_time, y=labeled_frac, shape=startime, fill=startime),
-             color="black", size=2, alpha=0.7) +
-  facet_nested(compound_name~amendment) +
-  scale_x_continuous(breaks = c(6, 18, 30, 42), labels = c("6AM", "6PM", "6AM", "")) +
-  scale_y_continuous(labels = scales::label_percent(),
-                     limits = c(0, 1), breaks = c(0, 0.5, 1)) +
-  scale_shape_manual(values=c(24, 25), breaks = c("6AM start", "6PM start")) +
-  scale_fill_manual(values=c("#f4bb23", "#0b505c"), breaks = c("6AM start", "6PM start"),
-                    aesthetics = c("color", "fill")) +
-  labs(x=NULL, color=NULL, shape=NULL, fill=NULL, y="Fraction of MAA containing at least one 15N atom") +
-  theme_bw() +
-  theme(axis.text.y = element_markdown(vjust=c(0, 0.5, 1)),
-        axis.title.x = element_blank(),
-        legend.position = "inside",
-        legend.margin = margin(0.1,0.25,0.1,0.1, unit="lines"),
-        legend.position.inside = c(1, 1),
-        legend.justification = c(1, 1), 
-        legend.background = element_rect(fill="#FFFFFFAA", color="black", linewidth = 0)) +
-  guides(shape=guide_legend(override.aes = list(size=3, stroke=1.2, alpha=1, linetype=0)))
-ggsave(plot = chosen_maa_gp, filename = "MAA_curves.png", device = "png", 
-       path = "manuscript/figures", width = 3, height = 6, units = "in", dpi = 300)
-```
-
-![](figures/MAA_curves.png)
-
-*Figure 3: Fraction of UV-reactive mycosporine-like amino acids (MAAs) containing at least one 15N label over one day during the treatments amended with 15N labeled ammonia or nitrate.
-Incubations started at 6AM are shown in yellow while incubations started at 6PM are shown in blue, with grey background regions highlighting nighttime.
-Three replicates are included at each of the 5 timepoints and best-fit power law curves have been fit behind the data.*
-
-Nitrate data showed similar degrees of labeling which was again surprising given the reduced amount of bulk metabolite labeling in those amendments, though the synthesis of MAAs from nitrate appeared less clearly diel.
-Shinorine and porphyra-334 were particularly well-labeled under nitrate, approaching the values achieved in the ammonia incubations.
-Very little MAA labeling was observed during the arginine and GMP amendments until the 73 hour timepoint, though at that point the fraction labeled was approximately equivalent to the inorganic amendments and slightly higher for GMP than arginine.
-
-## Ammonia more bioavailable at surface, organics equally bioavailable at 25m and 175m
-
-```{r surfdeep nums}
-labeled_fracs %>%
-  # filter(compound_name%in%c("L-Glutamic acid/beta-Glutamic acid", "L-Aspartic acid", "L-Asparagine", "Theanine?")) %>%
-    # filter(compound_name%in%c("Guanine", "Guanosine", "Cytosine", "Cytidine", "Adenine", "Adenosine")) %>%
-  filter(compound_name%in%c("Glycine betaine", "Homarine", "Trigonelline", "Choline")) %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  filter(amendment=="Amm") %>%
-  summarise(frac=mean(labeled_frac), .by=c(compound_name, timepoint, depth)) %>%
-  pivot_wider(names_from=depth, values_from=frac) %>%
-  summarise(Surf/Deep, .by=c(compound_name, timepoint)) %>%
-  arrange(compound_name)
-```
-
-We also performed the four nitrogen amendments at both the surface (25 meters) and below the euphotic zone (175 meters) to investigate the effect of depth on organic and inorganic nitrogen use `r wc("Maybe find and include the light level at 175m?")`.
-We found that ammonia was used much more readily at the surface as measured by the fraction of metabolite labeled, with ^15^N incorporation for amino acids 5-25% greater in the surface than at depth (Figure 4).
-Glutamate, theanine, aspartate, and asparagine were among the most different with glutamate and theanine surface values about 10 times those at depth while ^15^N aspartate was undetectable until T3 and ^15^N asparagine never detected.
-In contrast, glutamine labeling was similar between the two depths, with the fraction labeled 10-20 percentage points higher at the surface across all timepoints; a trend which was clearly not propagated to glutamate or other amino acids and instead must have represented accumulation or protein incorporation.
-`r wc("Not super happy with the numeric language here - feels like I'm flipping between 'times greater' and 'percentage points higher'.")`
-
-```{r fig surface vs deep labeling rates}
-surfdeep_demo_cmpd_gp <- labeled_fracs %>%
-  mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
-  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
-  mutate(compound_class=case_when(
-    compound_name%in%c("Glutamine", "Aspartic acid", "Asparagine", "Glutamic acid", "Ectoine", "Theanine?", "Citrulline", "Guanine", "Guanosine", "Arginine") ~ compound_name,
-    compound_name%in%aa_names~"Other amino acids",
-    compound_name%in%nucleo_names~"Other nucleic acids",
-    compound_name%in%betaine_names~"Betaines",
-    TRUE~"All others"
-  )) %>%
-  mutate(compound_class=factor(compound_class, levels=setdiff(class_levels, "MAAs"))) %>%
-  left_join(parametadata %>% distinct(shortname, timepoint_num, depth, startime, amendment, tripl)) %>%
-  filter(timepoint_num!=18) %>%
-  filter(compound_name=="Glutamic acid" & amendment=="Amm" |
-           compound_name=="Arginine" & amendment=="Arg" |
-           compound_name=="Guanosine" & amendment=="GMP") %>%
-  mutate(timepoint_num=ifelse(timepoint_num==0, 0.3, timepoint_num)) %>%
-  mutate(amendment=factor(amendment, levels=c("Amm", "Arg", "GMP"),
-                          labels=c("+Ammonia", "+Arginine", "+GMP"))) %>%
-  select(-shortname) %>%
-  summarise(mean_label=mean(labeled_frac), .by=c(compound_class, amendment, timepoint_num, depth)) %>%
-  pivot_wider(names_from=depth, values_from=mean_label) %>%
-  ggplot(aes(x=timepoint_num, color=compound_class, shape=compound_class)) +
-  geom_segment(aes(xend=timepoint_num, y=Surf, yend=Deep), lwd=1) +
-  geom_point(aes(y=Surf), fill="grey80", size=3, stroke=1) +
-  geom_point(aes(y=Deep), fill="grey20", size=3, stroke=1) +
-  stat_summary(aes(y=Surf, label=paste0(round(Surf-Deep, 2)*100, "%"), vjust=-sign(Surf-Deep)*1.5+0.75), 
-               geom = "text", fun = mean) +
-  facet_wrap(~amendment, nrow=1) +
-  scale_color_discrete(drop=FALSE) +
-  scale_x_log10(breaks=c(0.3, 3, 30), labels=c("0.3", "3", "30"), expand=expansion(0.15)) +
-  scale_y_continuous(labels=scales::label_percent(), expand=expansion(mult=c(0, 0.1)),
-                     breaks=c(0, 0.5, 1), limits = c(0, 1)) +
-  scale_shape_manual(breaks=c("Glutamic acid", "Arginine", "Guanosine"), values=c(22, 25, 21)) +
-  labs(x="Hours after <sup>15</sup>N spike", color=NULL, fill=NULL, shape=NULL,
-       y="Percent metabolite\nlabeled") +
-  theme_bw() +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    legend.position = "none"
-  )
-
-surfdeep_all_cmpd_gp <- labeled_fracs %>%
-  left_join(parametadata %>% distinct(shortname, timepoint, timepoint_num, depth, startime, amendment, tripl)) %>%
-  filter(timepoint_num!=18) %>%
-  filter(!compound_name%in%maa_names) %>%
-  mutate(compound_name=str_remove_all(compound_name, "^L-")) %>%
-  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
-  mutate(compound_class=case_when(
-    compound_name%in%c("Glutamine", "Aspartic acid", "Asparagine", "Glutamic acid", "Ectoine", "Theanine?", "Citrulline", "Guanine", "Guanosine", "Arginine") ~ compound_name,
-    compound_name%in%aa_names~"Other amino acids",
-    compound_name%in%nucleo_names~"Other nucleic acids",
-    compound_name%in%betaine_names~"Betaines",
-    TRUE~"All others"
-  )) %>%
-  mutate(compound_class=factor(compound_class, levels=class_levels)) %>%
-  summarise(avg_frac=mean(labeled_frac), .by = c(compound_class, amendment, timepoint_num, depth, tripl)) %>%
-  filter(amendment!="Nit") %>%
-  summarise(avg_label=mean(avg_frac), .by=c(compound_class, amendment, depth, timepoint_num)) %>%
-  pivot_wider(names_from = depth, values_from=avg_label) %>%
-  mutate(timepoint_num=ifelse(timepoint_num==0, 0.3, timepoint_num)) %>%
-  mutate(amendment=factor(amendment, levels=c("Amm", "Arg", "GMP"),
-                          labels=c("+Ammonia", "+Arginine", "+GMP"))) %>%
-  ggplot(aes(x=timepoint_num, y=Surf-Deep, color=compound_class, 
-             fill=compound_class, shape=compound_class)) +
-  geom_line(lwd=0.8) +
-  geom_point(color="black", size=3) +
-  geom_hline(yintercept = 0) +
-  facet_wrap(~amendment, nrow=1) +
-  scale_x_log10(breaks=c(0.3, 3, 30), labels=c("0.3", "3", "30"), expand=expansion(0.15)) +
-  scale_y_continuous(labels=scales::label_percent()) +
-  scale_shape_manual(breaks=setdiff(class_levels, "MAAs"), values = rep(21:25, length.out=14)) +
-  labs(x="Hours after <sup>15</sup>N spike", color=NULL, fill=NULL, shape=NULL,
-       y="Percent metabolite labeled at 25m\nminus percent labeled at 175m") +
-  theme_bw() +
-  theme(
-    # axis.title.x = element_markdown(),
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    strip.text = element_blank(),
-    strip.background = element_blank()
-  )
-
-surfdeep_all_cmpd_boxplot_gp <- labeled_fracs %>%
-  left_join(parametadata %>% distinct(shortname, timepoint_num, depth, startime, amendment, tripl)) %>%
-  filter(timepoint_num!=18) %>%
-  filter(!compound_name%in%maa_names) %>%
-  filter(amendment!="Nit") %>%
-  mutate(timepoint_num=ifelse(timepoint_num==0, 0.3, timepoint_num)) %>%
-  summarise(labeled_frac=mean(labeled_frac), .by=c(timepoint_num, compound_name, depth, amendment)) %>%
-  ggplot(aes(x=timepoint_num, y=labeled_frac, group=interaction(timepoint_num, depth), fill=depth)) +
-  geom_boxplot(color="black", outlier.shape=21) +
-  facet_wrap(~amendment, nrow=1) +
-  scale_x_log10(breaks=c(0.3, 3, 30), labels=c("0.3", "3", "30"), expand=expansion(0.15/2)) +
-  scale_y_continuous(labels = scales::label_percent(), breaks=c(0, 0.5, 1)) +
-  scale_fill_manual(breaks=c("Surf", "Deep"), values=c("grey90", "grey30"), labels=c("25 meters", "175 meters")) +
-  labs(x="Hours after <sup>15</sup>N spike", color=NULL, fill=NULL, shape=NULL,
-       y="Percent metabolite\nlabeled") +
-  theme_bw() +
-  theme(
-    axis.title.x = element_markdown(),
-    strip.text = element_blank(),
-    strip.background = element_blank()
-  )
-
-surfdeep_gp <- surfdeep_demo_cmpd_gp / surfdeep_all_cmpd_gp / surfdeep_all_cmpd_boxplot_gp + plot_layout(guides = "collect", heights = c(0.3, 0.5, 0.2))
-
-ggsave(filename = "deep_vs_surf_labeling.png", surfdeep_gp, device = "png", 
-       path = "manuscript/figures", width = 6.5, height = 6.5, units = "in", dpi = 300)
-```
-
-![](figures/deep_vs_surf_labeling.png)
-
-*Figure 4: Plots of the metabolite fraction labeled at the surface (25 meters) and the fraction labeled at 175 meters.
+Figure 4: Plots of the metabolite fraction labeled at the surface (25 meters) and the fraction labeled at 175 meters.
 Incubations with added ammonia, arginine, and guanosine monophosphate (GMP) are shown.
@@
 Finally, the bottom row of plots shows the relative difference between the surface and deep as boxplots for all compounds without aggregating.
-Points in the first two rows are the average of six biological replicates (incubations started in the morning and the evening are combined here for additional power) while points in the bottom row denote metabolites with labeled fractions falling outside 1.5 times the interquartile range.*
+Points in the first two rows are the average of six biological replicates (incubations started in the morning and the evening are combined here for additional power) while points in the bottom row denote metabolites with labeled fractions falling outside 1.5 times the interquartile range.
 
-Other metabolites varied in their degree of labeling (Figure 4).
-Nucleobases and nucleosides were much more heavily labeled from ammonia at the surface, with maximum labeling 3-10 times more extensive around T3 and T10 if the isotope was measurable at both depths and infinitely more so if not.
-Betaines and other nitrogen-containing metabolites, however, were almost entirely unlabeled at 175 meters until the 24 hour mark, at which the fraction labeled in both was 5-20 times higher at the surface.
-An important exception to this rule is the osmolyte ectoine, an aspartate derivative that nonetheless was synthesized with nitrogen from ammonia at depth at rates equal to or exceeding those of the surface.
+![](manuscript/figures/untarg_heatmap.png)
 
-We were unfortunately unable to present the data collected from the 175 meter labeled nitrate incubations here due to missed timepoints and confusion during sample collection but the raw numbers are provided in the supplement (Table XX and Supplemental Figure XX).
-
-^15^N use from the organic nitrogen substrates was again strikingly similar between the 25 meter and 175 meter samples aside from the direct products of the amended compounds.
-In both cases, uptake and use of the labeled substrate was equal or slightly higher at depth for a majority of the metabolites measured.
-Exceptions to this were theanine, which was 20-30 percentage points more heavily labeled at the surface by the end of the incubations, and ectoine, which was synthesized more at depth even to a larger degree than it was during the ammonia treatments, reaching nearly 50 percentage points more labeled at 175 meters than it was at 25m.
-
-## Untargeted methods reveal widespread labeling throughout the metabolome
-
-```{r load clean_combined and make feat_shape_df}
-cleaned_combined_peaks <- readRDS("untargeted/cleaned_combined_peaks.rds")
-
-lmtest_output <- cleaned_combined_peaks %>%
-  arrange(feature) %>%
-  select(feature, filename, into) %>%
-  left_join(parametadata) %>%
-  filter(samp_type=="Smp") %>%
-  mutate(timepoint_num=ifelse(timepoint_num==0, 0.3, timepoint_num)) %>%
-  nest(data=-feature) %>%
-  filter(sapply(data, nrow)>2) %>%
-  mutate(lm_test=map(data, ~broom::tidy(lm(log10(.x$into)~log10(.x$timepoint_num))),
-                     .progress = TRUE)) %>%
-  unnest(lm_test) %>%
-  filter(term=="log10(.x$timepoint_num)") %>%
-  mutate(p_adj=p.adjust(p.value, method = "fdr"))
-feat_shape_df <- cleaned_combined_peaks %>%
-  summarise(mzmed=unique(mzmed),
-            rtmed=unique(rtmed),
-            med_iso_cor=median(iso_cor), 
-            med_snr=median(beta_snr),
-            med_cor=median(beta_cor),
-            .by=feature) %>%
-  left_join(lmtest_output %>% select(feature, estimate, p_adj)) %>%
-  mutate(peak_qual=ifelse(med_snr>-50*med_cor+50, "Good", "Bad")) %>%
-  mutate(peak_qual=ifelse(rtmed<2*60 | between(rtmed, 14*60, 14.5*60), 
-                          "Solvent artifact", peak_qual)) %>%
-  mutate(is_iso=ifelse(med_iso_cor>0.8, "Is iso", "Not iso")) %>%
-  mutate(time_trend=ifelse(p_adj<0.001& estimate>0, "Increases", "No change"))
-
-# Repeat analysis across amendments to get additional numbers
-lmtest_xamend <- cleaned_combined_peaks %>%
-  arrange(feature) %>%
-  select(feature, filename, into) %>%
-  mutate(samp_type=str_extract(filename, "Smp|Blk|Poo")) %>%
-  filter(samp_type=="Smp") %>%
-  mutate(timepoint=as.numeric(str_extract(filename, "(?<=T)\\d+"))) %>%
-  mutate(timepoint=ifelse(timepoint==0, 0.3, timepoint)) %>%
-  mutate(amendment=str_extract(filename, "Amm|GMP|Arg|Nit")) %>%
-  mutate(depth=str_extract(filename, "Surf|Deep")) %>%
-  nest(data=-c(feature, amendment)) %>%
-  filter(sapply(data, nrow)>2) %>%
-  mutate(lm_test=map(data, ~broom::tidy(lm(log10(.x$into)~log10(.x$timepoint))), 
-                     .progress = TRUE)) %>%
-  unnest(lm_test) %>%
-  filter(term=="log10(.x$timepoint)") %>%
-  mutate(p_adj=p.adjust(p.value, method = "fdr"))
-feat_shape_df_xamend <- cleaned_combined_peaks %>%
-  mutate(amendment=str_extract(filename, "Amm|GMP|Arg|Nit")) %>%
-  summarise(mzmed=unique(mzmed),
-            rtmed=unique(rtmed),
-            med_iso_cor=median(iso_cor), 
-            med_snr=median(beta_snr),
-            med_cor=median(beta_cor),
-            .by=c(feature, amendment)) %>%
-  left_join(lmtest_xamend %>% select(feature, amendment, estimate, p_adj)) %>%
-  mutate(peak_qual=ifelse(med_snr>-50*med_cor+50, "Good", "Bad")) %>%
-  mutate(peak_qual=ifelse(rtmed<2*60 | between(rtmed, 14*60, 14.5*60), 
-                          "Solvent artifact", peak_qual)) %>%
-  mutate(is_iso=ifelse(med_iso_cor>0.8, "Is iso", "Not iso")) %>%
-  mutate(time_trend=ifelse(p_adj<0.001& estimate>0, "Increases", "No change"))
-```
-
-```{r count categories as a whole and by amendment}
-feat_shape_df_xamend %>%
-  bind_rows(feat_shape_df %>% mutate(amendment="Overall")) %>%
-  filter(!is.na(time_trend)) %>%
-  count(peak_qual, is_iso, time_trend, amendment) %>%
-  pivot_wider(names_from = amendment, values_from=n, values_fill = 0)
-```
-
-After a heuristic compound discovery approach revealed the putatively identified theanine, deoxyguanosine, deoxycytidine, and MAAs had interesting labeling patterns, we decided to perform a fully untargeted analysis to detect additional compounds that were labeled over the course of the incubations.
-Fairly lax peakpicking settings for xcms returned nearly 10,000 chromatographic peaks in each of the ~280 samples across the four LC runs used, one for each amendment sample set.
-At this point we noted that the nitrate amendment samples had an unusually large number of peaks detected due to the inclusion of a contaminant suite, likely plastic, which fortunately eluted before many of our compounds of interest.
-Correspondence requiring the presence of a peak in all three biological triplicates for a given treatment/timepoint combination reduced that number to approximately 5,000 unique features.
-Random inspection revealed that the majority of these features were noise so we calculated additional peak quality metrics as well as reintegrated each peak and its predicted isotope (see Methods).
-Of the initial 5,602 features, over half (2,956) failed quality control and an additional 1,690 were determined to be likely solvent artifacts due to their elution in less than two minutes or `r wc("when a large proportion of junk was coming off the column (median retention times between 14.0 and 14.5 minutes)?
-during the retention time when sulfate eluted?
-based on their peak shape and retention time?
-due to the fact that they showed no change across any of the treatments or timepoints?
-The darn things just look bad and I hate them but I can't really explain why.")` and were removed from downstream analysis.
-
-```{r write out the best labels for manual annotation}
-best_labeled <- feat_shape_df %>%
-  filter(peak_qual=="Good" & is_iso=="Is iso" & time_trend=="Increases") %>%
-  arrange(feature) %>%
-  select(feature, mzmed, rtmed, med_iso_cor, med_snr, med_cor, p_adj)
-psb_stans <- all_stans %>%
-  filter(polarity=="pos") %>%
-  summarise(compound_name=paste(compound_name, collapse = "/"), .by = mz) %>%
-  rowwise() %>%
-  mutate(N15_mz_lower=pmppm(mz+0.997035, 10)[1]) %>%
-  mutate(N15_mz_upper=pmppm(mz+0.997035, 10)[2]) %>%
-  select(compound_name, N15_mz_lower, N15_mz_upper)
-best_labeled %>%
-  left_join(psb_stans, join_by(between(x$mzmed, y$N15_mz_lower, y$N15_mz_upper))) %>%
-  select(-N15_mz_lower, -N15_mz_upper) %>%
-  mutate(compound_name=ifelse(duplicated(compound_name)&!is.na(compound_name), "Isocytosine", compound_name)) %>%
-  mutate(compound_name=ifelse((mzmed-lag(mzmed))%between%c(0.996, 0.998), "iso_prev?", compound_name)) %>%
-  mutate(rtmed=rtmed/60) %>%
-  select(feature, mzmed, rtmed, compound_name) %>%
-  write_csv("untargeted/best_labeled_anno.csv")
-```
-
-Of the 219 features that remained, the majority (150) had a ^15^N isotope with the correct *m/z* value and a chromatographic similarity score above 0.8 (see Methods).
-80 of these 150 also showed linear increases in peak area across the timepoints, denoting likely ^15^N-labeled metabolites.
-We manually inspected each of these 80 metabolites and were able to identify the majority as either compounds in our list of authentic standards (19/80), an isotope of a known compound (39/80), or an in-source fragment of a known compound (12/80, with fragments from proline (70.0658), glutamic acid/glutamate (84.0449, 102.0550, and 130.0499), and the loss of ribose from nucleosides).
-Five compounds were putatively identified using isotope labeling patterns and literature searches: FT1362 with an *m/z* of 121.07798 and a retention time (RT) of 4.5 minutes we flagged as indoline, FT1564 with an *m/z* of 124.0524 and a 2.7 minute RT we flagged as the ^15^N isotope of nicotinamide, FT4529 with an *m/z* of 228.0979 and a 6.7 minute RT we flagged as deoxycytidine, FT5860 with an *m/z* of 282.1182 and an 11.1 minute RT we flagged as pangamic acid, and FT6622 with an *m/z* of 333.1292 and an 11.2 minute RT we flagged as shinorine.
-Four additional compounds could not be putatively identified but had plausible molecular formulae: FT3699 with an *m/z* of 192.0866 and 10.3 RT we assigned C~7~H~14~NO~5~, FT4598 with an *m/z* of 231.1703	and 1.9 RT we assigned C~11~H~23~N~2~O~3~, FT4733 with an *m/z* of 236.1492 and 10.0 RT we assigned C~10~H~22~NO~5~, and FT5185 with an *m/z* of 252.1441 and 10.7 RT we assigned C~10~H~22~NO~6~.
-One feature we were unable to annotate at all (FT5076, *m/z* 248.1258 and RT 12.0).
-
-```{r heatmap of all high-quality feats}
-high_qual_feats <- cleaned_combined_peaks %>%
-  filter(rtmed>120 & !rtmed%between%c(14*60, 14.5*60)) %>%
-  left_join(parametadata %>% distinct(filename, samp_type)) %>%
-  filter(samp_type=="Smp") %>%
-  summarise(med_iso_cor=median(iso_cor), 
-            med_snr=median(beta_snr),
-            med_cor=median(beta_cor),
-            .by=c(feature)) %>%
-  filter(med_snr>-50*med_cor+50) %>%
-  filter(med_iso_cor>0.8)
-untarg_lab_fracs <- cleaned_combined_peaks %>%
-  filter(feature%in%high_qual_feats$feature) %>%
-  select(feature, filename, lab_area, unlab_area) %>%
-  left_join(parametadata %>% distinct(filename, samp_type)) %>%
-  filter(samp_type=="Smp") %>%
-  slice_max(unlab_area, by=c(feature, filename), with_ties = FALSE) %>%
-  mutate(labeled_frac=lab_area/(lab_area+unlab_area)) %>%
-  left_join(parametadata %>% distinct(filename, shortname)) %>%
-  select(feature, shortname, labeled_frac)
-
-cmpd_order <- untarg_lab_fracs %>%
-  summarise(mean_label=mean(labeled_frac, na.rm=TRUE), .by=feature) %>%
-  arrange(desc(mean_label)) %>%
-  pull(feature)
-
-colorpal <- viridis::viridis(10, alpha=seq(0, 1, length.out=10), direction = -1)
-untarg_heatmap <- untarg_lab_fracs %>%
-  complete(feature, shortname, fill=list(labeled_frac=0)) %>%
-  mutate(feature=factor(feature, levels=cmpd_order)) %>%
-  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
-  ggplot() +
-  geom_raster(aes(x=timepoint, y=tripl, fill=labeled_frac)) +
-  scale_fill_gradientn(colours = colorpal, limits = c(0, 1)) +
-  scale_y_discrete(expand=expansion()) +
-  scale_x_discrete(guide = guide_axis(n.dodge = 5), expand=expansion()) +
-  facet_nested(feature~amendment+depth+startime, scales = "free_x", switch="y", strip=strip_nested(size = "variable")) +
-  labs(x="Timepoint", y="Metabolite", fill="Fraction\nlabeled") +
-  theme_bw() +
-  theme(panel.spacing.x=unit(0.2, "lines"), 
-        panel.spacing.y = unit(0, "lines"),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        # axis.ticks.length.x=unit(1:5, "lines"),
-        # axis.text.x = element_markdown(vjust = -5:-1),
-        strip.text.y.left = element_text(angle=0, hjust = 1),
-        # strip.background.y = element_rect(fill=NA, color=NA),
-        panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank(),
-        legend.position = "top", 
-        legend.key.width = unit(1, "null")) +
-  guides(
-    fill=guide_colorbar(theme=theme(
-      legend.frame = element_rect(color = "black"),
-      legend.ticks = element_line(color="black"),
-      legend.text = ggtext::element_markdown(hjust=c(0, 0.5, 0.5, 0.5, 1)))
-    )
-  )
-ggsave("untargeted/untarg_heatmap.pdf", width = 10, height=36, units = "in")
-```
-
-To derive biological meaning from the untargeted analysis, we performed a k-means clustering of the features across all the treatments and timepoints and included our targeted data in the clustering.
-Clear patterns of isotope incorporation emerged in 6 of the 10 clusters we used (Figure 5).
-Cluster 1, the largest with 64 molecular features, contained many of the unlabeled compounds including those without any nitrogen and all of the betaines.
-Cluster 2 contained most of the amino acids, the nucleobases cytosine and guanine, the and the deoxynucleosides as well as ectoine and citrulline.
-Cluster 3 contained homarine, beta-alanine, and O-acetylcarnitine as well as the remaining amino acids aside from aspartate, glutamate, and glutamine which fell into cluster 4.
-Known compounds in cluster 5 consisted of adenine, adenosine, cytidine, and hypoxanthine while guanosine fell alone into cluster 6.
-The MAAs were distributed across clusters 8 and 9, while clusters 7 and 10 had no known compounds.
-
-```{r run kmeans}
-untarg_lab_mat <- untarg_lab_fracs %>%
-  bind_rows(labeled_fracs %>% rename(feature=compound_name) %>% select(-filename)) %>%
-  pivot_wider(names_from = "feature", values_from = "labeled_frac", values_fill=0) %>%
-  column_to_rownames("shortname") %>%
-  data.matrix()
-
-set.seed(124)
-
-# sapply(1:20, function(n_clust){
-#   replicate(10, {
-#     kmeans_out <- kmeans(t(untarg_lab_mat), centers = n_clust)
-#     kmeans_out$betweenss / kmeans_out$totss * 100
-#   })
-# }) %>%
-#   boxplot()
-
-n_clust <- 10
-untarg_kmeans <- untarg_lab_mat %>%
-  t() %>%
-  kmeans(centers = n_clust)
-kmeans_df <- data.frame(cluster=1:n_clust, n=untarg_kmeans$size) %>%
-  arrange(desc(n)) %>%
-  mutate(cluster_name=paste0("Cluster ", row_number(), " n=", n)) %>%
-  mutate(cluster_name=fct_inorder(cluster_name)) %>%
-  left_join(data.frame(cluster=untarg_kmeans$cluster, feature=names(untarg_kmeans$cluster))) %>%
-  select(cluster_name, feature)
-
-# Plot clusters as heatmap
-untarg_heatmap <- untarg_lab_fracs %>%
-  left_join(kmeans_df) %>%
-  summarise(labeled_frac=mean(labeled_frac), .by = c(cluster_name, shortname)) %>%
-  left_join(parametadata) %>%
-  ggplot() +
-  geom_raster(aes(x=timepoint, y=tripl, fill=labeled_frac)) +
-  scale_fill_gradientn(colours = colorpal, limits = c(0, 1)) +
-  facet_wrap(cluster_name~amendment+depth+startime) +
-  scale_y_discrete(expand=expansion()) +
-  scale_x_discrete(guide = guide_axis(n.dodge = 5), expand=expansion()) +
-  facet_nested(cluster_name~amendment+depth+startime, scales = "free_x", switch="y", 
-               strip=strip_nested(size = "variable")) +
-  labs(x="Timepoint", y=NULL, fill="Fraction\nlabeled") +
-  theme_bw() +
-  theme(panel.spacing.x=unit(0.2, "lines"), 
-        panel.spacing.y = unit(0, "lines"),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        # axis.ticks.length.x=unit(1:5, "lines"),
-        # axis.text.x = element_markdown(vjust = -5:-1),
-        strip.text.y.left = element_text(angle=0, hjust = 1),
-        # strip.background.y = element_rect(fill=NA, color=NA),
-        panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank(),
-        legend.position = "top", 
-        legend.key.width = unit(1, "null")) +
-  guides(
-    fill=guide_colorbar(theme=theme(
-      legend.frame = element_rect(color = "black"),
-      legend.ticks = element_line(color="black"),
-      legend.text = ggtext::element_markdown(hjust=c(0, 0.5, 0.5, 0.5, 1)))
-    )
-  )
-ggsave("untarg_heatmap.png", untarg_heatmap, path="manuscript/figures",
-       width=6.5, height=4, units = "in")
-
-# Export as CSV
-kmeans_df %>%
-  filter(!startsWith(feature, "FT")) %>%
-  arrange(cluster_name, feature) %>%
-  split(.$cluster_name) %>%
-  map(function(df_i){
-    df_i %>% add_row(cluster_name="", feature="")
-  }) %>%
-  bind_rows() %>%
-  write_csv("untargeted/cluster_assignments.csv")
-```
-
-![](figures/untarg_heatmap.png)
-
-*Figure 5: Fraction of each cluster containing at least one ^15^N label during the surface seawater amendments with various substrates after the number of hours shown on the x-axis.
+Figure 5: Fraction of each cluster containing at least one ^15^N label during the surface seawater amendments with various substrates after the number of hours shown on the x-axis.
 Darker colors indicate a larger percentage of the feature signal was isotopically labeled.
@@
 The number of features in each cluster is specified by the n in the facet name.
-GMP = guanosine monophosphate.*
+GMP = guanosine monophosphate.
 
-# Discussion
+## Data products
 
-[DISCUSSION: surface community not "expecting" nitrate? problems with redox at surface?]
+[Metadata for each of the sample files](metadata/parametadata.csv)
 
-[DISCUSSION: not expecting to see any C labeling from the organics via CO2 because inorganic C pool is so much larger, unlike the N pool]
+[Environmental concentrations for XX metabolites](targeted/final_env_concs.csv)
 
-[DISCUSSION: What does the decrease in arginine labeling in +Arg at T26 and T73 indicate? Pulse-chase behavior? How is the labeled pool specifically decreasing?]
+[Relative abundances and isotope labeling patterns for XX metabolites](targeted/all_peak_areas_w_isos.csv)
 
-[DISCUSSION: Contrasting the THAA +Nit with the metabolites hints at a small but rapidly churning amino acid pool that's deeply segregated]
+[Environmental concentrations for THAAs and their isotopes](thaas/quant_data.csv)
 
-[DISCUSSION: MAAs are mostly made from the 4DG core plus N-containing amino acids (glycine). How does the labeling of these compare to their substituent AAs? (it's much larger)]
+[Relative abundance (peak area) and peak quality for 500,000 xcms peaks](untargeted/cleaned_combined_peaks.rds)
 
-# Conclusion
+[Annotation of the 80 best features in the untargeted analysis](untargeted/best_labeled_anno.xlsx)
 
-# Methods
+---
 
-[explain chromatographic similarity scoring as pearson's correlation coefficient]
-
-# Acknowledgements
-
-# Data availability
-
-# References
-
-```{r packages, eval=FALSE}
-packages_used <- list.files(pattern = "*\\.R$|*\\.Rmd$", recursive = TRUE, full.names = TRUE) %>%
-  str_subset("old|scripts", negate=TRUE) %>%
-  map(function(file_name){
-    v <- readLines(file_name)
-    libs <- unique(str_extract(v, "(?<=library\\().*?(?=\\)|,)"))
-    refs <- unique(str_extract(v, "[[:alnum:]]*(?=::)"))
-    pkgs <- unique(c(na.omit(libs), na.omit(refs)))
-  }, .progress = TRUE) %>% unlist() %>% unique()
-knitr::write_bib(packages_used[nchar(packages_used)>0], file = "manuscript/r-packages.bib")
-```
-
-<div id="refs"></div>
-
-# Supplement
-
-## Supp fig: full heatmap of all compounds and timepoints
-
-```{r suppfig full heatmap}
-cmpd_order <- labeled_fracs %>%
-  left_join(parametadata, relationship = "many-to-many") %>%
-  filter(amendment=="Amm") %>%
-  filter(depth=="Surf") %>%
-  filter(startime=="Morn") %>%
-  summarise(mean_label=mean(labeled_frac, na.rm=TRUE), .by=compound_name) %>%
-  arrange(desc(mean_label)) %>%
-  pull(compound_name) %>%
-  str_remove_all("^L-|(?<=/)L-") %>%
-  c("5-Methylcytosine")
-
-heatmap_gp <- labeled_fracs %>%
-  left_join(parametadata, relationship = "many-to-many") %>%
-  arrange(amendment, depth, timepoint_num, startime) %>%
-  mutate(shortname=fct_inorder(shortname)) %>%
-  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"),
-                          labels = c("+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
-  mutate(depth=factor(depth, levels=c("Surf", "Deep"), labels=c("25m", "175m"))) %>%
-  mutate(compound_name=str_remove_all(compound_name, "^L-|(?<=/)L-")) %>%
-  mutate(order_name=factor(compound_name, cmpd_order)) %>%
-  mutate(compound_class=case_when(
-    compound_name%in%maa_names~"MAA",
-    compound_name%in%aa_names~"Amino\nacid",
-    compound_name%in%nucleo_names~"Nucleic\nacid",
-    compound_name%in%non_n_names~"N-free",
-    compound_name%in%betaine_names~"Betaine",
-    TRUE~"Other"
-  )) %>%
-  mutate(compound_class=factor(compound_class, levels=c("Amino\nacid", "Nucleic\nacid", "MAA", "Betaine", "N-free", "Other"))) %>%
-  # filter(compound_class=="Other") %>%
-  ggplot() +
-  geom_raster(aes(x=timepoint, y=tripl, fill=labeled_frac)) +
-  scale_fill_gradientn(colours = viridis::viridis(10, alpha = seq(0, 1, length.out=10), direction = -1), limits=c(0, 1), breaks=c(0, 0.25, 0.5, 0.75, 1), labels=c("0%<br>(Unlabeled)", "", "50%", "", "100%<br>(Fully labeled)")) +
-  scale_y_discrete(expand=expansion()) +
-  scale_x_discrete(guide = guide_axis(n.dodge = 5), expand=expansion()) +
-  facet_nested(compound_class+compound_name~amendment+depth+startime, scales = "free_x", switch="y", strip=strip_nested(size = "variable")) +
-  labs(x="Timepoint", y="Metabolite", fill="Fraction\nlabeled") +
-  theme_bw() +
-  theme(panel.spacing.x=unit(0.2, "lines"), 
-        panel.spacing.y = unit(0, "lines"),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        # axis.ticks.length.x=unit(1:5, "lines"),
-        # axis.text.x = element_markdown(vjust = -5:-1),
-        strip.text.y.left = element_text(angle=0, hjust = 1),
-        # strip.background.y = element_rect(fill=NA, color=NA),
-        panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank(),
-        legend.position = "top", 
-        legend.key.width = unit(1, "null")) +
-  guides(
-    fill=guide_colorbar(theme=theme(
-      legend.frame = element_rect(color = "black"),
-      legend.ticks = element_line(color="black"),
-      legend.text = ggtext::element_markdown(hjust=c(0, 0.5, 0.5, 0.5, 1)))
-    )
-  )
-
-ggsave(plot = heatmap_gp, "manuscript/supplement/para_heatmap.pdf", device = "pdf",
-       width = 10, height = 12)
-```
-
-## Supp fig: THAA labeling broken down by triplicate
-
-```{r thaa supplement figures}
-pdf("manuscript/supplement/THAA_labeling.pdf", width=9, height = 5)
-for(cmpd_i in setdiff(unique(thaa_quant$compound_name), c("Citrulline", "Ornithine", "Taurine"))){
-  print(cmpd_i)
-  gp <- thaa_quant %>%
-    filter(compound_name==cmpd_i) %>%
-    left_join(metathaa, by = join_by(filename)) %>%
-    drop_na() %>%
-    ggplot() +
-    geom_col(aes(x=tripl, y=env_conc_nm, fill=iso_name), 
-             position="fill",
-             color="black", width=1) +
-    facet_nested(depth+startime~amendment+timepoint) +
-    scale_y_continuous(labels = scales::label_percent()) +
-    theme_bw() +
-    ggtitle(cmpd_i) +
-    labs(x="Biological triplicate", y="Fraction labeled", fill="Isotopologue")
-  print(gp)
-}
-dev.off()
-```
+README last built on "`r Sys.Date()`"

### 2025-05-22 README.Rmd
Fixing image alignment

@@
 
-<div style="display: flex; align-items: center;">
-  <img src="manuscript/figures/MAA_curves.png" style="width: 50%; margin-right: 10px;" />
-  <div style="text-align: center; flex: 1;">
-    <p>Figure 3: Fraction of UV-reactive mycosporine-like amino acids (MAAs) containing at least one ^15^N label over one day during the treatments amended with ^15^N labeled ammonia or nitrate.
+<table>
+  <tr>
+    <td>
+      <img src="manuscript/figures/MAA_curves.png" width="4500"/>
+    </td>
+    <td style="vertical-align: middle; text-align: center;">
+      Figure 3: Fraction of UV-reactive mycosporine-like amino acids (MAAs) containing at least one ^15^N label over one day during the treatments amended with ^15^N labeled ammonia or nitrate.
 Incubations started at 6AM are shown in yellow while incubations started at 6PM are shown in blue, with grey background regions highlighting nighttime.
-Three replicates are included at each of the 5 timepoints and best-fit power law curves have been fit behind the data.</p>
-  </div>
-</div>
-
-<div clear="left"></div>
+Three replicates are included at each of the 5 timepoints and best-fit power law curves have been fit behind the data.
+    </td>
+  </tr>
+</table>
 

### 2025-05-22 README.Rmd
Tidying README

@@
 
+---
+
 ![](manuscript/figures/total_thaa_gp.png)
@@
   <tr>
-    <td>
-      <img src="manuscript/figures/MAA_curves.png" width="4500"/>
-    </td>
     <td style="vertical-align: middle; text-align: center;">
@@
     </td>
+    <td>
+      <img src="manuscript/figures/MAA_curves.png" width="4500"/>
+    </td>
   </tr>
 </table>
+
+---
 
@@
 Points in the first two rows are the average of six biological replicates (incubations started in the morning and the evening are combined here for additional power) while points in the bottom row denote metabolites with labeled fractions falling outside 1.5 times the interquartile range.
+
+---
 

### 2025-05-23 tmzMLs/arrowscript.R
Adding separate wIS handling

@@ -26,6 +26,30 @@ for(samp_type_i in unique(parametadata$samp_type)){
   }
 }
 
+parametadata <- fread("metadata/parametadata.csv")[
+  IS_type=="noIS"|is.na(IS_type),c("filepath", "filename", "samp_type", "amendment", "polarity")
+]
+parametadata[, .N, by=c("samp_type", "amendment", "polarity")]
+for(samp_type_i in unique(parametadata$samp_type)){
+  for(amendment_i in unique(parametadata$amendment)){
+    for(polarity_i in unique(parametadata$polarity)){
+      ms_files <- parametadata[samp_type==samp_type_i][amendment==amendment_i][polarity==polarity_i]$filepath
+      # ms_files <- head(ms_files, 3)
+      if(length(ms_files)==0)next
+      msdata <- grabMSdata(ms_files, grab_what = "MS1")
+      parq_dir <- paste0("tmzMLs/pqds",
+                         "/samp_type=", samp_type_i, 
+                         "/amendment=", amendment_i, 
+                         "/polarity=", polarity_i)
+      if(!dir.exists(parq_dir))dir.create(parq_dir, recursive = TRUE)
+      parq_name <- paste0(parq_dir, "/part-0.parquet")
+      print(parq_name)
+      write_parquet(msdata$MS1[order(mz)], parq_name)
+    }
+  }
+}
+
+
 # library(tidyverse)
 # open_dataset("tmzMLs/pqds") %>%
 #   filter(amendment=="Amm") %>%

### 2025-05-23 manuscript/manuscript.Rmd
Moving MAA selection higher

@@
                "Palythenic acid?")
+chosen_maas <- c("Mycosporine-glycine?", "Shinorine?", "Palythine?",
+                 "Palythene/Usujirene?", "Also Palythene/Usujirene?", 
+                 "Also Porphyra-334?")
 nucleo_names <- c("5-Methylcytosine", "Adenosine", "Cytidine", "Adenine", "Guanosine",
@@
 
-chosen_maas <- c("Mycosporine-glycine?", "Shinorine?", "Palythine?",
-                 "Palythene/Usujirene?", "Also Palythene/Usujirene?", 
-                 "Also Porphyra-334?")
 chosen_maa_data <- labeled_fracs %>%

### 2025-05-23 manuscript/manuscript.Rmd
Cleaning up color palette use

@@
                   "#AE9AD8", "#4E1F88", "#E0C174", "#4D2B01", "grey40")
+
+heatmap_colorpal <- viridis::viridis(10, alpha=seq(0, 1, length.out=10), direction = -1)
+
 aa_ncs <- all_stans %>%
@@
                                                           face=rep("bold", length(class_colors))))) +
-  scale_fill_gradientn(colours = viridis::viridis(10, alpha = seq(0, 1, length.out=10), direction = -1),
+  scale_fill_gradientn(colours = heatmap_colorpal,
                        limits=c(0, 1), breaks=c(0, 0.25, 0.5, 0.75, 1),
@@
 
-```{r load clean_combined and make feat_shape_df}
-cleaned_combined_peaks <- readRDS("untargeted/cleaned_combined_peaks.rds")
-
+```{r make feat_shape_df}
 lmtest_output <- cleaned_combined_peaks %>%
@@
 untarg_lab_mat <- untarg_lab_fracs %>%
-  bind_rows(labeled_fracs %>% rename(feature=compound_name) %>% select(-filename)) %>%
+  bind_rows(labeled_fracs %>% rename(feature=compound_name)) %>%
   pivot_wider(names_from = "feature", values_from = "labeled_frac", values_fill=0) %>%
@@
 untarg_heatmap <- untarg_lab_fracs %>%
-  bind_rows(labeled_fracs %>% rename(feature=compound_name) %>% select(-filename)) %>%
+  bind_rows(labeled_fracs %>% rename(feature=compound_name)) %>%
   left_join(kmeans_df) %>%
@@
   geom_raster(aes(x=timepoint, y=tripl, fill=labeled_frac)) +
-  scale_fill_gradientn(colours = colorpal, limits = c(0, 1)) +
+  scale_fill_gradientn(colours = heatmap_colorpal, limits = c(0, 1)) +
   facet_wrap(cluster_name~amendment+depth+startime) +
@@
 ```{r supp fig heatmap of all high-quality feats}
-colorpal <- viridis::viridis(10, alpha=seq(0, 1, length.out=10), direction = -1)
 untarg_heatmap <- untarg_lab_fracs %>%
-  bind_rows(labeled_fracs %>% rename(feature=compound_name) %>% select(-filename)) %>%
+  bind_rows(labeled_fracs %>% rename(feature=compound_name)) %>%
   complete(feature, shortname, fill=list(labeled_frac=0)) %>%
@@
   geom_raster(aes(x=timepoint, y=tripl, fill=labeled_frac)) +
-  scale_fill_gradientn(colours = colorpal, limits = c(0, 1)) +
+  scale_fill_gradientn(colours = heatmap_colorpal, limits = c(0, 1)) +
   scale_y_discrete(expand=expansion()) +
@@
   geom_raster(aes(x=timepoint, y=tripl, fill=labeled_frac)) +
-  scale_fill_gradientn(colours = viridis::viridis(10, alpha = seq(0, 1, length.out=10), direction = -1), limits=c(0, 1), breaks=c(0, 0.25, 0.5, 0.75, 1), labels=c("0%<br>(Unlabeled)", "", "50%", "", "100%<br>(Fully labeled)")) +
+  scale_fill_gradientn(colours = heatmap_colorpal, limits=c(0, 1), breaks=c(0, 0.25, 0.5, 0.75, 1), labels=c("0%<br>(Unlabeled)", "", "50%", "", "100%<br>(Fully labeled)")) +
   scale_y_discrete(expand=expansion()) +

### 2025-05-23 manuscript/manuscript.Rmd
New introduction drafted

@@
 
-Marine carbon is fixed into biomass very quickly.
-In most of the surface ocean this flux is controlled by the availability of nitrogen, an element with a complex and incompletely characterized biogeochemical cycle [@Capone2008; @Moore2013; @Hutchins2022].
+In most of the surface ocean the flux of carbon is controlled by the availability of nitrogen, an element with a complex and incompletely characterized biogeochemical cycle [@Capone2008; @Moore2013; @Hutchins2022].
 Organic nitrogen in particular is often treated as a black box despite substantial variability in its bioavailability and chemical nature.
-While extensive work has been done to characterize the forms and fluxes of nitrogen within the inorganic pool, the equivalent for the organic pool is woefully underdeveloped [@ref, prob NME2008 or Hansell2024, maybe @Hutchins2022 for visual?].
+Turnover times for labile organic compounds like urea and amino acids can be orders of magnitude shorter than high molecular weight DON [@Aluwihare2005; @Bronk2024].
+Additionally, the availability of the precious nitrogen atom is determined not only by the form of the organic material it composes but also the environment in which it's found.
+
+While extensive work has been done to characterize the forms and fluxes of nitrogen within the inorganic pool, the equivalent for the organic pool is woefully underdeveloped [@Zehr2011; @NME].
 This is largely due to the difficulty of comprehensively measuring the many organic molecules that contain nitrogen in the marine environment [@Boysen2018; @Moran2022].
-The bioavailability of the nitrogen atom is in the environment is largely determined by the form of the organic material it composes.
-
-The mechanisms by which dissolved nitrogen becomes biomass and vice versa are major factors in our ability to predict ecosystem productivity and therefore carbon fixation and export.
-Traditionally, the conversion of inorganic material into organic substrate was thought to be limited to autotrophs while remineralization was performed by the heterotrophic community.
-Today, it’s known that many photosynthetic organisms are able to take up and use organic nitrogen [@Anita1991; @Morando2018; @Hugo2021] and that only a select subset of phototrophs are capable of nitrate reduction or nitrogen fixation [@ref], while inorganic nitrogen is directly accessible to heterotrophs as well.
-This means that the transformations possible for a given N-containing compound are a function of the microbial community and its environment while the microbial community is in turn also expected to be a function of the N-containing molecules available.
-This creates a recursive network that’s difficult to untangle and cannot necessarily be extrapolated from axenic culture studies.
-
-One way to reveal the pathways and transformations that marine N undergoes in a natural community is via 15N labeling paired with metabolomics.
-Metabolites that incorporate the labeled nitrogen can then be separated on the mass spectrometer and quantified separately from the unlabeled pool.
-Here, we trace the uptake and use of labeled ammonia, nitrate, arginine, and guanosine monophosphate in a natural community from the NPSG at multiple depths and diel conditions.
-Our application of metabolomics to these samples allows us to map out the pathways and restructuring that organic nitrogen experiences in the largest biomes on the planet, quantifying the pool sizes and turnover rates for various low molecular weight compounds that serve as building blocks and intermediates of cell biology.
+Recently, however, we have improved our understanding of community composition via genomics and transcriptomics and made progress in our ability to characterize organic matter via proteomics and metabolomics.
+These advancements, paired with the increased availability of pure and isotope-labeled substrates, mean that we can now apply metabolically resolved isotope tracing to resolve how organic nitrogen is transformed in the marketplace of metabolites that empower the surface ocean.
+
+Nitrogen assimilation is a well-documented process in the ocean after decades of research and demonstrates how biochemistry derived from non-marine organisms can be applied to the ocean.
+Central to the process is ammonia, which can be directly taken up from the environment by all organisms and combined with glutamate.
+This forms glutamine, which is then combined with alpha-ketoglutarate to form two glutamate molecules that can then be used as a nitrogen source throughout the cell [@Walker2016].
+Nitrate, on the other hand, must first be reduced to ammonia in an energetically expensive process that some organisms have stripped from their genomes [@Partensky2010; @Zehr2011].
+
+Unlike the relatively well-resolved pathways for inorganic nitrogen, the use of organic nitrogen remains underexplored in the marine environment.
+Organic nitrogen is generally thought to be accessible to most organisms but the high levels of DON make it clear that this is highly substrate-specific [@Antia1991; @Bronk2007; @Aluwihare2008; @Zehr2011; @Bronk2024].
+Additionally, many of the transformations possible are only active in specific communities or times of day [@Bender2012].
+Linking the transformations that nitrogen undergoes to known molecular pathways documented in other organisms or in culture is a particularly promising way to understand what we observe on the global scale.
+
+The lack of a comprehensive understanding of nitrogen's fate after it is assimilated makes molecular modeling difficult and undermines our ability to connect the community to the organic nitrogen it produces.
+Here, we trace the uptake and use of labeled ammonia, nitrate, arginine, and guanosine monophosphate in a natural community from the North Pacific Subtropical Gyre at multiple depths and diel conditions.
+Our application of metabolomics to these samples allows us to map the pathways and restructuring that organic nitrogen experiences in the largest biomes on the planet, quantifying the pool sizes and turnover rates for various low molecular weight compounds that serve as building blocks and intermediates of cell biology.
 
@@
 
+Incubations are imperfect metrics of the actual environment because amendments with *anything* will heavily skew the community - over what time scale though? [@Eppley1977; @Shilova2017; @Rii2018]
+
 # Conclusion

### 2025-05-24 manuscript/manuscript.Rmd
Abstract edits from Anitra

@@
 We repeated these experiments both at the surface and at the base of the euphotic zone as well as in the morning and evening.
-We found that ammonia is converted into glutamate in less than 20 minutes and results in labeling throughout the metabolome, though osmolytes like betaines had slow turnover rates.
-This process was slower at depth but largely independent of time of day.
-Nitrate amendments resulted in the labeling of only a small subset of molecules such as mycosporine-like amino acids, whose incorporation patterns suggest that this form of nitrogen was available to a subset of the community, likely eukaryotes.
-The two organic nitrogen amendments labeled the metabolome more slowly than the inorganic forms of nitrogen and showed similar usage at depth and at the surface, hinting at the major role that heterotrophic bacteria likely play in their uptake and use.
-A large fraction of the organic nitrogen added was remineralized into ammonia and used as glutamate rather than as intact substrates, though labeled arginine was incorporated into proteins directly.
-These measurements reveal the forms and fluxes of organic nitrogen in the natural environment to provide constraints on biogeochemical models and better characterize the transformations between organic and inorganic nitrogen.
+We found that ammonia is converted into glutamate in less than 20 minutes and results in labeling throughout the metabolome after 24 hours, though osmolytes like betaines had slow turnover rates.
+Nitrate amendments resulted in the labeling of only a small subset of molecules such as mycosporine-like amino acids (MAAs), whose incorporation patterns suggest that this form of nitrogen was available to a subset of the community, likely eukaryotes.
+The two organic nitrogen amendments labeled the metabolome more slowly than the inorganic forms of nitrogen and showed similar usage at 25 meters and 175 meters, hinting at the major role that heterotrophic bacteria play in cycling organic matter.
+A majority of the organic nitrogen added was converted into glutamate after 24 hours rather than used as intact substrates, though labeled arginine was incorporated into proteins directly and we were able to trace the carbon backbone from both arginine and GMP into several additional metabolites.
+We also show that nitrogen assimilation and use was largely the same between morning and evening with the exception of UV-sensitive compounds such as MAAs.
+These measurements reveal the forms and fluxes of nitrogen in the natural environment and provide constraints on biogeochemical models that require characterizing the transformations between organic and inorganic nitrogen.
+
+```{r}
+length(strsplit("", split = " ")[[1]]) # ~286 words
+length(strsplit("", split = "")[[1]])  # ~1851 characters
+```
 

### 2025-05-24 manuscript/manuscript.Rmd
Removing soundbites from intro and tidying output

@@
 Our application of metabolomics to these samples allows us to map the pathways and restructuring that organic nitrogen experiences in the largest biomes on the planet, quantifying the pool sizes and turnover rates for various low molecular weight compounds that serve as building blocks and intermediates of cell biology.
-
-## Other things to maybe discuss:
-
-  -	mention the expected differences with depth and diel effects
-  -	detail which compounds we chose and why?
-  -	expand on DON-PON interactions and importance?
-  - Extracellular enzymes, organic use as ammonia vs as-is?
-  -	expand on the importance of biogeochemical models and why we need pathways mapped out?
-  -	mention basic biosynthetic pathways (maybe better suited to discussion?)
-  -	new (nitrate) vs regenerated (ammonia and DON) production (maybe also discussion?)
-
-## Other soundbites:
-  - It is assumed that most fixed nitrogen becomes protein, presumably based on the large fraction of cell nitrogen in protein and previous research showing how quickly it enters the (free?) amino acid pool.
-However, 1) important intermediates molecules 2) doesn’t explain where biounavailable N comes from 3) makes gene-based molecular modeling really difficult.
-  - In each case, assimilation into organic matter is typically done via the glutamine synthetase (GS)/glutamate synthase (GOGAT) or glutamate dehydrogenase (GDH) pathways to produce glutamate from α-ketoglutarate [@ref, see bronk chapter].
-Glutamate then serves as a nitrogen supply in an enormous number of biochemical pathways in both primary and secondary metabolism [@Walker2016].
-  - nitrate amendments typically resulting in diatom blooms while ammonia incubations favor cyanobacteria [@Glibert2016].
-  - Communities to distinguish are 1) surface (largely regenerated production) and 2) 175m (largely new production(??))
-  - use of organics vs synthesis of organics
-  - Compounds such as non-proteinogenic amino acids (e.g., ornithine, citrulline, creatine, MAAs), osmolytes like betaines, nucleobases, and sulfur-containing molecules like taurine represent significant but poorly quantified pools.
-Understanding the rates and pathways by which these compounds are utilized is crucial, as they form the fundamental unit linking genetic potential to elemental cycling in biogeochemical models.
-  - Nitrogen use varies significantly between the sunlit surface where productivity is largely regenerated through recycling of ammonia and organic nitrogen, and deeper waters near the base of the euphotic zone, where nitrate supports “new” production.
-Quantifying rates of nitrogen assimilation into ammonia and downstream products across these gradients is therefore essential for predicting marine productivity and nutrient cycling.
-  - Nonetheless, how the community uses a substrate is arguably more important than whether it can.
 

### 2025-05-24 manuscript/manuscript.Rmd
Adding refs and refining intro a little bit

@@
 ---
-title: "Nitrogen's form determines its fate in the North Pacific Subtropical Gyre during PARAGON I"
+title: "Nitrogen's form determines its fate in the North Pacific Subtropical Gyre"
 author:
@@
 
-In most of the surface ocean the flux of carbon is controlled by the availability of nitrogen, an element with a complex and incompletely characterized biogeochemical cycle [@Capone2008; @Moore2013; @Hutchins2022].
+In most of the surface ocean the flux of carbon is controlled by the availability of nitrogen, an element with a complex and incompletely characterized biogeochemical cycle [@Bronk2008; @Moore2013; @Hutchins2022].
 Organic nitrogen in particular is often treated as a black box despite substantial variability in its bioavailability and chemical nature.
@@
 Recently, however, we have improved our understanding of community composition via genomics and transcriptomics and made progress in our ability to characterize organic matter via proteomics and metabolomics.
-These advancements, paired with the increased availability of pure and isotope-labeled substrates, mean that we can now apply metabolically resolved isotope tracing to resolve how organic nitrogen is transformed in the marketplace of metabolites that empower the surface ocean.
-
-Nitrogen assimilation is a well-documented process in the ocean after decades of research and demonstrates how biochemistry derived from non-marine organisms can be applied to the ocean.
+These advancements, paired with the increased availability of pure and isotope-labeled substrates, mean that we can now apply metabolically resolved isotope tracing to resolve how organic nitrogen is transformed in the marketplace of metabolites that empower the surface ocean [@Heal2021; @McParland2021; @Moran2022].
+
+Nitrogen assimilation is a well-documented process in the ocean after decades of research and demonstrates how biochemistry derived from non-marine organisms can be applied to the ocean [@Mulholland2008].
 Central to the process is ammonia, which can be directly taken up from the environment by all organisms and combined with glutamate.
@@
 Additionally, many of the transformations possible are only active in specific communities or times of day [@Bender2012].
-Linking the transformations that nitrogen undergoes to known molecular pathways documented in other organisms or in culture is a particularly promising way to understand what we observe on the global scale.
-
-The lack of a comprehensive understanding of nitrogen's fate after it is assimilated makes molecular modeling difficult and undermines our ability to connect the community to the organic nitrogen it produces.
+Linking the transformations that nitrogen undergoes to known molecular pathways from other environments or axenic cultures is a particularly promising way to understand what we observe on the global scale.
+
+The lack of a comprehensive understanding of nitrogen's fate after it is assimilated makes molecular modeling difficult and undermines our ability to connect the community to the organic nitrogen it produces [@Jones2024].
 Here, we trace the uptake and use of labeled ammonia, nitrate, arginine, and guanosine monophosphate in a natural community from the North Pacific Subtropical Gyre at multiple depths and diel conditions.
@@
 
+Comments from Anitra:
+
+0.3nM glutamine becoming 0.8nM means that 95% of the 0.8nM was produced, not counting all the N that went through glutamine into other things.
+Do you know what typically happens to biomass during ALOHA + DIN incubation?
+How does this compare to measures of regenerated production?
+(Was this increase a result of more biomass or just higher gln concentrations?)
+
 # Conclusion
